### 2023-11-08 article.tex
Initial commit

\documentclass{article}
\usepackage[T1]{fontenc}
\usepackage[style=ieee, backend=bibtex8]{biblatex}
\usepackage{beramono}
\usepackage{courier}
\usepackage{filecontents}
\usepackage{graphicx} % Required for inserting images
\usepackage{listings}
\usepackage{realboxes}
\usepackage{titlesec}
\usepackage{xcolor}
\usepackage{xpatch}

\titleformat{\paragraph}
{\normalfont\normalsize\bfseries}{\theparagraph}{1em}{}
\titlespacing*{\paragraph}
{0pt}{3.25ex plus 1ex minus .2ex}{1.5ex plus .2ex}

\lstdefinelanguage{none}{
  identifierstyle=
}

\definecolor{mygreen}{rgb}{0,0.6,0}
\definecolor{mygray}{rgb}{0.5,0.5,0.5}
\definecolor{mymauve}{rgb}{0.58,0,0.82}
\definecolor{mygray}{rgb}{0.94,0.94,0.94}

\addbibresource{\jobname.bib}
\AtEveryBibitem{\printfield{note}\clearfield{note}\item}

\lstset{ 
  backgroundcolor=\color{mygray},   % choose the background color; you must add \usepackage{color} or \usepackage{xcolor}; should come as last argument
  breakatwhitespace=false,         % sets if automatic breaks should only happen at whitespace
  basicstyle=\ttfamily,
  breaklines=true,                 % sets automatic line breaking
  captionpos=b,                    % sets the caption-position to bottom
  commentstyle=\color{mygreen},    % comment style
  deletekeywords={...},            % if you want to delete keywords from the given language
  escapeinside={\%*}{*)},          % if you want to add LaTeX within your code
  extendedchars=true,              % lets you use non-ASCII characters; for 8-bits encodings only, does not work with UTF-8
  firstnumber=1000,                % start line enumeration with line 1000
  frame=none,	                   % adds a frame around the code
  keepspaces=true,                 % keeps spaces in text, useful for keeping indentation of code (possibly needs columns=flexible)
  keywordstyle=\color{blue},       % keyword style
  language=Octave,                 % the language of the code
  morekeywords={*,...},            % if you want to add more keywords to the set
  numbers=none,                    % where to put the line-numbers; possible values are (none, left, right)
  numbersep=5pt,                   % how far the line-numbers are from the code
  numberstyle=\tiny\color{red}, % the style that is used for the line-numbers
  rulecolor=\color{black},         % if not set, the frame-color may be changed on line-breaks within not-black text (e.g. comments (green here))
  showspaces=false,                % show spaces everywhere adding particular underscores; it overrides 'showstringspaces'
  showstringspaces=false,          % underline spaces within strings only
  showtabs=false,                  % show tabs within strings adding particular underscores
  stepnumber=2,                    % the step between two line-numbers. If it's 1, each line will be numbered
  stringstyle=\color{mymauve},     % string literal style
  tabsize=2,	                   % sets default tabsize to 2 spaces
  title=\lstname                   % show the filename of files included with \lstinputlisting; also try caption instead of title
}

\title{Implementation of register allocation to translate \lstinline!LLVM--! to x86}
\author{William Welle Tange}
\date{2023}

\makeatletter
\xpretocmd\lstinline{\Colorbox{mygray}\bgroup\appto\lst@DeInit{\egroup}}{}{}
\makeatother

\begin{document}

\maketitle

\section{Introduction}

\section{Review of literature}

%* todo: ssa/llvm vs appel analysis
%
%* henvis til blog posts som archive
%
%    * så tæller det som litteratur
%    
%* tag essensen ud af hvad der bliver sagt i teksten
%
%    * som er relevant for mit arbejde
%    
%* HENVIS TIL KAPITEL 19 OM SSA FORM
%
%* linear scan approach
%
%* snak om hvilke dele af bogen i hvilken rækkefølge
%
%Appel (208p):
%
%* Translation To Intermediate Code (153-179, 26p)
%
%* Basic Blocks And Traces (179-193, 14p)
%
%*Instruction Selection (193-219, 26p)
%
%* Liveness Analysis (219-237, 18p)
%
%* Register Allocation (237-267, 30p)
%
%* Dataflow Analysis (387-415, 28p)
%
%* Loop Optimizations (415-439, 24p)
%
%* Static Single Assignment Form (439-481, 42p)
%
%*using IR when only translating from one language to another:
%
%>Even when only one front end and one back end are being built, a good IR can modularize the task, so that the front end is not complicated with machine-specific details, and the back end is not bothered with information specific to one source language.

\section{Limited instruction set of \lstinline!LLVM--!}

The intermediate representation emitted by the frontend of a compiler serves as a stepping stone independent of the target architecture. The \lstinline!LLVM! infrastructure is the industry standard in terms of bridging this gap and was consequently the library used to translate the semantically annotated abstract syntax tree to executable machine code in the 2022 compilers course. As this project is an expansion on this, it follows naturally to build on this.

The instruction set used in this paper will be a union of the sets used in the 2022 and 2023 compilers courses in order to work as a drop-in replacement of LLVM for either of the two respective source languages:  Tiger and Dolphin. This instruction set is a subset of the one used in practice, as neither of the languages implemented support exception handling, floating point operations and so on, and instead only strive for a seemingly comfortanle middle ground.

The instructions included are
\begin{enumerate}
    \item binary operations of \lstinline!add!, \lstinline!and!, \lstinline!ashr!, \lstinline!lshr!, \lstinline!mul!, \lstinline!or!, \lstinline!sdiv!, \lstinline!srem!, \lstinline!shl!, \lstinline!sub! and \lstinline!xor!
    \item logic/arithmetic comparison of \lstinline!icmp! (conditions being \lstinline!eq!, \lstinline!ne!, \lstinline!sge!, \lstinline!sgt!, \lstinline!sle! and \lstinline!slt!)
    \item memory/address operations of \lstinline!alloca!, \lstinline!gep!, \lstinline!load! and \lstinline!store!
    \item trivial/non-trivial \lstinline!mov! operations \lstinline!gep!, \lstinline!phi!, \lstinline!ptrtoint! and \lstinline!zext!
    \item control flow operations of \lstinline!call!
\end{enumerate}
with only three possible block terminators 
\begin{enumerate}
    \item \lstinline!br! with an optional condition
    \item \lstinline!ret!
    \item \lstinline!unreachable!
\end{enumerate}

\section{LLVM -> CFG}

%* description of blocks, how they connect etc.

\section{CFG -> INTERF}

%* describe the dataflow algorithm from the book
%
%* describe the interference criteria from the book

\section{INTERF -> DOT}

%* ez ???

\section{INTERF -> DOT/Assignment}

%* list of methods for coalescing/allocating:
%
%    * ocamlgraph builtin
%
%    * greedy 
%
%    * briggs
%
%    * george
%
%    * welsh-powell?

\section{Naive Codegen}

%* safest possible x86 translation
%
%    * moving sys-v abi arguments to assigned registers is treacherous

\section{Linear scan approach}

%Live ranges for each of the temporaries are derived by visiting each instruction of each control flow block and examining its \textit{uses} and \textit{defs}.

\section{Benchmarking}
Benchmarks are performed on Apple Silicon (M1 Pro, i.e. ARMv8.5-A \cite{AM1}) despite the target instruction set being x86-64. This is possible because of Rosetta 2, which  translates x86-64 to ARMv8 before execution \cite{A2012}.  Time is measured using the \lstinline!time! function of the bash shell as it is native to the chosen Docker image.  As all execution relevant to the performance of register allocation is done in user mode, only the \lstinline!user! metric is noted.

%* fib
%* ack
%* \lstinline!/usr/bin/time -al! counts cpu cycles

\section{Evaluation}
%Benchmarks are performed in the environment described by the Dockerfile in the repository and executed on an Apple Silicon/M1 chip. Time is measured using the \lstinline!time! function of the bash shell as it is native to the chosen Docker image.  As all execution relevant to the performance of register allocation is done in user mode, only the \lstinline!user! metric is noted.

%TODO:
%
%* should I use shell time, or /usr/bin/time?
%* how do I measure cpu cycles?

\section{Comparison to other work and ideas for future work}

\section{Conclusion}

%\quickcharcount{main}

%\defbibnote{myprenote}{This thesis is based on the following original publications:}
%\printbibliography[prenote=myprenote,title={References}]
\printbibliography[title={References}]

\appendix

\newpage
\section{Benchmarks}

\subsection{\lstinline!benches/fib.ll!}
\begin{lstlisting}[language=LLVM]
declare i32 @atoi(i8*)
define i32 @fib(i32 %n0) {
  %cn =  icmp sle i32 %n0, 2
  br i1 %cn, label %base, label %rec
base:
  ret i32 1
rec:
  %n1 = sub i32 %n0, 1
  %v0 = call i32 @fib(i32 %n1)
  %n2 = sub i32 %n0, 2
  %v1 = call i32 @fib(i32 %n2)
  %v2 = add i32 %v1, %v2
  ret i32 %v2
}
define i32 @main(i32 %argc, i8** %argv) {
  %arg1ptr = getelementptr i8*, i8** %argv, i64 1
  %arg1 = load i8*, i8** %arg1ptr
  %n = call i32 @atoi(i8* %arg1)
  call i32 @fib(i32 %n)
  ret i32 0
}

\end{lstlisting}

\subsubsection{\lstinline!make fib-clang!}
\begin{lstlisting}[language=none]
$ make fib-clang
clang -O0 -target x86_64-unknown-darwin benches/fib.ll -o fib-clang
\end{lstlisting}
\paragraph{\lstinline!make bench fib-clang 42!}
\begin{lstlisting}[language=none]
$ make bench fib-clang 42
/usr/bin/time -al ./fib-clang 42
        1.58 real         1.56 user         0.00 sys
             2678784  maximum resident set size
                   0  average shared memory size
                   0  average unshared data size
                   0  average unshared stack size
                 765  page reclaims
                   0  page faults
                   0  swaps
                   0  block input operations
                   0  block output operations
                   0  messages sent
                   0  messages received
                   0  signals received
                   0  voluntary context switches
                  29  involuntary context switches
         11000239251  instructions retired
          4962802652  cycles elapsed
             1708928  peak memory footprint
\end{lstlisting}
\paragraph{\lstinline!make bench fib-clang 43!}
\begin{lstlisting}[language=none]
make bench fib-clang 43
$ /usr/bin/time -al ./fib-clang 43
        2.71 real         2.49 user         0.00 sys
             2678784  maximum resident set size
                   0  average shared memory size
                   0  average unshared data size
                   0  average unshared stack size
                 700  page reclaims
                  66  page faults
                   0  swaps
                   0  block input operations
                   0  block output operations
                   0  messages sent
                   0  messages received
                   0  signals received
                  13  voluntary context switches
                  32  involuntary context switches
         17797124508  instructions retired
          8035601598  cycles elapsed
             1708928  peak memory footprint
\end{lstlisting}

\paragraph{\lstinline!make bench fib-clang 44!}
\begin{lstlisting}[language=none]
$ make bench fib-clang 44
/usr/bin/time -al ./fib-clang 44
        4.06 real         4.04 user         0.00 sys
             2678784  maximum resident set size
                   0  average shared memory size
                   0  average unshared data size
                   0  average unshared stack size
                 700  page reclaims
                  66  page faults
                   0  swaps
                   0  block input operations
                   0  block output operations
                   0  messages sent
                   0  messages received
                   0  signals received
                   5  voluntary context switches
                  29  involuntary context switches
         28781328564  instructions retired
         12988681489  cycles elapsed
             1708928  peak memory footprint
\end{lstlisting}

\paragraph{\lstinline!make bench fib-clang 45!}
\begin{lstlisting}[language=none]
$ make bench fib-clang 45
/usr/bin/time -al ./fib-clang 45
        6.54 real         6.51 user         0.00 sys
             2678784  maximum resident set size
                   0  average shared memory size
                   0  average unshared data size
                   0  average unshared stack size
                 700  page reclaims
                  66  page faults
                   0  swaps
                   0  block input operations
                   0  block output operations
                   0  messages sent
                   0  messages received
                   0  signals received
                   6  voluntary context switches
                  39  involuntary context switches
         46555259189  instructions retired
         20968319385  cycles elapsed
             1708928  peak memory footprint
\end{lstlisting}

\paragraph{\lstinline!make bench fib-clang 46!}
\begin{lstlisting}[language=none]
$ make bench fib-clang 46
/usr/bin/time -al ./fib-clang 46
       10.58 real        10.54 user         0.00 sys
             2678784  maximum resident set size
                   0  average shared memory size
                   0  average unshared data size
                   0  average unshared stack size
                 700  page reclaims
                  66  page faults
                   0  swaps
                   0  block input operations
                   0  block output operations
                   0  messages sent
                   0  messages received
                   0  signals received
                   5  voluntary context switches
                 107  involuntary context switches
         75314385847  instructions retired
         33939361073  cycles elapsed
             1725376  peak memory footprint
\end{lstlisting}

\paragraph{\lstinline!make bench fib-clang 47!}
\begin{lstlisting}[language=none]
$ make bench fib-clang 47
/usr/bin/time -al ./fib-clang 47
       17.09 real        17.06 user         0.00 sys
             2678784  maximum resident set size
                   0  average shared memory size
                   0  average unshared data size
                   0  average unshared stack size
                 766  page reclaims
                   0  page faults
                   0  swaps
                   0  block input operations
                   0  block output operations
                   0  messages sent
                   0  messages received
                   0  signals received
                   0  voluntary context switches
                 100  involuntary context switches
        121838577947  instructions retired
         54901966523  cycles elapsed
             1708928  peak memory footprint
\end{lstlisting}

\paragraph{\lstinline!make bench fib-clang 48!}
\begin{lstlisting}[language=none]
$ make bench fib-clang 48
/usr/bin/time -al ./fib-clang 48
       27.65 real        27.59 user         0.00 sys
             2678784  maximum resident set size
                   0  average shared memory size
                   0  average unshared data size
                   0  average unshared stack size
                 766  page reclaims
                   0  page faults
                   0  swaps
                   0  block input operations
                   0  block output operations
                   0  messages sent
                   0  messages received
                   0  signals received
                   0  voluntary context switches
                 161  involuntary context switches
        197130073055  instructions retired
         88835375488  cycles elapsed
             1708928  peak memory footprint
\end{lstlisting}

\end{document}

### 2023-11-08 article.tex
Some work on intro etc.

@@
 \documentclass{article}
+\usepackage[style=ieee]{biblatex}
+
+\addbibresource{\jobname.bib}
+%\AtEveryBibitem{\printfield{note}\clearfield{note}\item}
+
+\usepackage{tikz}
+\usetikzlibrary{chains}
+\usepackage{graphicx} % Required for inserting images
 \usepackage[T1]{fontenc}
-\usepackage[style=ieee, backend=bibtex8]{biblatex}
 \usepackage{beramono}
 \usepackage{courier}
-\usepackage{filecontents}
-\usepackage{graphicx} % Required for inserting images
 \usepackage{listings}
+\usepackage{xpatch}
+\usepackage{xcolor}
 \usepackage{realboxes}
 \usepackage{titlesec}
-\usepackage{xcolor}
-\usepackage{xpatch}
+
+\setcounter{secnumdepth}{4}
 
@@
 
-\addbibresource{\jobname.bib}
-\AtEveryBibitem{\printfield{note}\clearfield{note}\item}
+%%% Wrap colorbox around lstinline %%%
+\makeatletter
+\xpretocmd\lstinline{\Colorbox{mygray}\bgroup\appto\lst@DeInit{\egroup}}{}{}
+\makeatother
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-\title{Implementation of register allocation to translate \lstinline!LLVM--! to x86}
+\title{Implementation and comparison of register allocation to translate \lstinline!LLVM--! to x86}
 \author{William Welle Tange}
 \date{2023}
-
-\makeatletter
-\xpretocmd\lstinline{\Colorbox{mygray}\bgroup\appto\lst@DeInit{\egroup}}{}{}
-\makeatother
 
@@
 
+\tableofcontents
+\newpage
+
+%% TODO: tjek hvad LLVM gør %%
+
+%\texttt{LLVM-{}-2}
+%* skab så mange underpunkter som muligt
+    %* jo mindre man skal overskue af gangen jo lettere er det at komme i gang med at skrive
+
 \section{Introduction}
+
+Compilers can be divided into several distinct phases, with the two halfs of which colloquially referred to as the \textit{frontend} and \textit{backend} stages.
+With the former serving the purpose of translating a high-level programming language to an \textit{intermediate representation} (IR) layer and the latter translating said IR to executable machine code on a target platform.
+
+\begin{figure}[h]
+  \centering
+\includegraphics[width=0.8\textwidth]{phases.png}
+\caption{Compiler phases}
+\end{figure}
+
+%* overfladisk intro til hvad helvede register allokering er og hvorfor
+    %* hovedtræk
+%* rudimentær del af compilation/oversættelse fra high-level sprog til x86 er register allokering
+    %* projektet er skridtet videre fra compilers kurset
+    %* siger hvad jeg står ovenpå/bygger på
+%snak om hvorfor jeg valgte llvm/ssa frem for x86 til global register llocation
 
@@
 
+Write about Appel book.
+
+Write about SSA form liveness analysis.
+
+Keywords: program points, dataflow equations, fixed point
+
 %* todo: ssa/llvm vs appel analysis
-%
 %* henvis til blog posts som archive
-%
 %    * så tæller det som litteratur
-%    
 %* tag essensen ud af hvad der bliver sagt i teksten
-%
-%    * som er relevant for mit arbejde
+    %* som er relevant for mit arbejde
 %
@@
 %Appel (208p):
-%
 %*   Translation To Intermediate Code (153-179, 26p)
-%
 %*   Basic Blocks And Traces (179-193, 14p)
-%
-%*Instruction Selection (193-219, 26p)
-%
+%*   Instruction Selection (193-219, 26p)
 %*   Liveness Analysis (219-237, 18p)
-%
 %*   Register Allocation (237-267, 30p)
-%
 %*   Dataflow Analysis (387-415, 28p)
-%
 %*   Loop Optimizations (415-439, 24p)
-%
 %*   Static Single Assignment Form (439-481, 42p)
@@
 
-\section{Limited instruction set of \lstinline!LLVM--!}
+\section{\lstinline!LLVM--! instruction set}
 
@@
 The instruction set used in this paper will be a union of the sets used in the 2022 and 2023 compilers courses in order to work as a drop-in replacement of LLVM for either of the two respective source languages:  Tiger and Dolphin.
-This instruction set is a subset of the one used in practice, as neither of the languages implemented support exception handling, floating point operations and so on, and instead only strive for a seemingly comfortanle middle ground.
+This instruction set is a subset of the one used in practice, as, for instance, neither of the languages implemented support exception handling, floating point operations and so on, and instead only strive to cover the basics of compilers.
 
@@
 \begin{enumerate}
-    \item binary operations of \lstinline!add!, \lstinline!and!, \lstinline!ashr!, \lstinline!lshr!, \lstinline!mul!, \lstinline!or!, \lstinline!sdiv!, \lstinline!srem!, \lstinline!shl!, \lstinline!sub!
+    \item binary operations \lstinline!add!, \lstinline!and!, \lstinline!ashr!, \lstinline!lshr!, \lstinline!mul!, \lstinline!or!, \lstinline!sdiv!, \lstinline!srem!, \lstinline!shl!, \lstinline!sub!
 and \lstinline!xor!
-    \item logic/arithmetic comparison of \lstinline!icmp! (conditions being \lstinline!eq!, \lstinline!ne!, \lstinline!sge!, \lstinline!sgt!, \lstinline!sle! and \lstinline!slt!)
-    \item memory/address operations of \lstinline!alloca!, \lstinline!gep!, \lstinline!load! and \lstinline!store!
-    \item trivial/non-trivial \lstinline!mov! operations \lstinline!gep!, \lstinline!phi!, \lstinline!ptrtoint! and \lstinline!zext!
-    \item control flow operations of \lstinline!call!
+    \item integer comparison \lstinline!icmp! (conditions being \lstinline!eq!, \lstinline!ne!, \lstinline!sge!, \lstinline!sgt!, \lstinline!sle! and \lstinline!slt!)
+    \item memory/address operations \lstinline!alloca!, \lstinline!gep!, \lstinline!load! and \lstinline!store!
+    \item  \lstinline!mov! operations \lstinline!gep!, \lstinline!phi!, \lstinline!ptrtoint! and \lstinline!zext!
+    \item control flow operations \lstinline!br!, \lstinline!call! and \lstinline!ret!
+    \item a  nullary terminator \lstinline!unreachable!
 \end{enumerate}
-with only three possible block terminators 
-\begin{enumerate}
-    \item \lstinline!br! with an optional condition
-    \item \lstinline!ret!
-    \item \lstinline!unreachable!
-\end{enumerate}
+%with only four distinct block terminators 
+%\begin{enumerate}
+%    \item Unconditional \lstinline!br! followed by the label to branch to
+%    \item Conditional \lstinline!br!  followed by an \lstinline!i1! condition and two labels
+%    \item \lstinline!ret! followed by an optional return value
+%    \item \lstinline!unreachable!
+%\end{enumerate}
+The branching for most of these is trivial: after successful execution the flow of all but \lstinline!br!, \lstinline!ret!
+and \lstinline!unreachable!
+will unconditionally attempt to execute the next instruction in memory (i.e.
+increment the instruction pointer by instruction length).
 
-\section{LLVM -> CFG}
+Because of this, these instructions are referred to as \textit{terminators}, as their purpose is to disrupt what had otherwise been a linear flow from the beginning of this continus sequence of instructions, henceforth referred to as a \textit{basic block}.
+
+\section{Control Flow Analysis}
+
+For a given program, it's representation as linear sequence of instructions from the above set isn't immediately useful for liveness analysis.
+(FIXME: why not?)  hence a \textit{control flow graph} is derived from these.
+A terminator will branch elsewhere to continue execution.
+The destination of this has to be an annotated by a label, hence labels \textit{initiate} blocks.
+
+With an input stream of instructions
 
@@
 
-\section{CFG -> INTERF}
+Each function defined in an LLVM program is constructed with the help of
+
+\section{Liveness Analysis}
+
+%\section{CFG -> INTERF}
 
 %* describe the dataflow algorithm from the book
-%
 %* describe the interference criteria from the book
 
-\section{INTERF -> DOT}
+%\section{INTERF -> DOT}
 
@@
 
-\section{INTERF -> DOT/Assignment}
+%\section{INTERF -> DOT/Assignment}
 
 %* list of methods for coalescing/allocating:
-%
 %   * ocamlgraph builtin
-%
 %   * greedy 
-%
 %   * briggs
-%
 %   * george
-%
 %   * welsh-powell?
+\section{Graph Coloring}
+The graph coloring by simplification as described by Appel
 
-\section{Naive Codegen}
+%% TODO
+%* optimistic coloring
+
+\section{Linear Scan}
+
+%Live ranges for each of the temporaries are derived by visiting each instruction of each control flow block and examining its \textit{uses} and \textit{defs}.
+
+\section{Codegen}
 
@@
 %
-%    * moving sys-v abi arguments to assigned registers is treacherous
+    %* moving sys-v abi arguments to assigned registers is treacherous
 
-\section{Linear scan approach}
+\section{Optimization}
 
-%Live ranges for each of the temporaries are derived by visiting each instruction of each control flow block and examining its \textit{uses} and \textit{defs}.
+% TODO: quote computability theory infinitely many optimizations
 
-\section{Benchmarking}
+\section{Evaluation}
+
+Relevant metrics by which to compare these variations are naturally the performance of the code generated at runtime, but also the time efficiency at compile-time.
+While the code written isn't expected to outperform LLVM, due to the time complexity of the linear scan and simplification algorithms implemented their respective runtime performances are expected to outperform the builtin graph coloring algorithm.
+
+% TODO: cpu cycles
+
+Another factor worth considering would be memory usage, caches misses, garbage collection (although not relevant to this as it doesn't use GC), vectorizatiion (loop optimizations by SIMD) etc.
+
+\subsection{Benchmarking}
+
 Benchmarks are performed on Apple Silicon (M1 Pro, i.e.
@@
 
-\section{Evaluation}
 %Benchmarks are performed in the environment described by the Dockerfile in the repository and executed on an Apple Silicon/M1 chip.
@@
 
-\section{Comparison to other work and ideas for future work}
+\subsection{Comparison to other work and ideas for future work}
 
@@
 
-%\quickcharcount{main}
-
 %\defbibnote{myprenote}{This thesis is based on the following original publications:}
 %\printbibliography[prenote=myprenote,title={References}]
-\printbibliography[title={References}]
+%\printbibliography[title={References}]
+
+\printbibliography
 
@@
 \newpage
+
+%%% NOTE: Don't include subsections of tthe benchmarks section
+\addtocontents{toc}{\protect\setcounter{tocdepth}{1}}
+
 \section{Benchmarks}

### 2023-11-22 article.tex
More work on intro etc.

@@
 
+\usepackage{amsmath}
 \usepackage{tikz}
 \usetikzlibrary{chains}
+\usetikzlibrary{angles, quotes}
+\usetikzlibrary{decorations.pathreplacing}
 \usepackage{graphicx} % Required for inserting images
@@
 
-Compilers can be divided into several distinct phases, with the two halfs of which colloquially referred to as the \textit{frontend} and \textit{backend} stages.
-With the former serving the purpose of translating a high-level programming language to an \textit{intermediate representation} (IR) layer and the latter translating said IR to executable machine code on a target platform.
+Compilation refers to the process of translating from one language to another, most often from a programming language intended for humans to work with, to machine-/bytecode intended to be executed on a target platform.
+This process can be divided into several distinct phases, which are grouped into one of two stages, colloquially referred to as the \textit{frontend} and \textit{backend}.
+The former translating a high-level programming language to an \textit{intermediate representation} (IR) layer and the latter translating IR to executable machine code on a target platform (e.g.
+x86, ARM) or bytecode for a target VM (e.g.
+JVM, CLR).
 
@@
   \centering
-\includegraphics[width=0.8\textwidth]{phases.png}
-\caption{Compiler phases}
+  \resizebox{0.8\textwidth}{!}{
+\begin{tikzpicture}
+  \tikzset{
+box/.style={draw,
+  inner sep=0.3cm,
+    minimum height=1.75cm,
+    align=center},
+}
+
+  % Nodes
+  \node [box, left, align=center] (incg) at (0,2.25) {Intermediate\\Codegen};
+  \node [box, left=0.7 of incg, align=center] (semant) {Semantic\\Analysis};
+  \node [box, left=0.7 of semant, align=center] (parse) {Parser};
+  \node [box, left=0.7 of parse, align=center] (lex) {Lexical\\Analysis};
+  \node [ left=0.9 of lex, align=center] (entry) {};
+  \node [box, left, align=center] (cfa) at (0,0) {Control-\\Flow\\Analysis};
+  \node [box, left=0.7 of cfa, align=center] (lva)  {Liveness/\\Dataflow\\Analysis};
+  \node [box, left=0.7 of lva, align=center] (alloc)  {Register\\Allocation};
+  \node [box, left=0.7 of alloc, align=center] (x86)  {Instruction\\Selection};
+  \node [ left=0.70 of x86, align=center] (exit) {};
+
+  % Edge with right-angle corners
+  \draw [-latex] (entry.east) -- (lex.west)  node [near start,pos=0.325,rotate=90,fill=white,opacity=0.9] {\small Source};
+  \draw [-latex] (lex.east) -- (parse.west)  node [midway,pos=0.4,rotate=90,fill=white,opacity=0.9] {\small Tokens};
+  \draw [-latex] (parse.east) -- (semant.west)  node [midway,pos=0.4,rotate=90,fill=white,opacity=0.9] {\small AST};
+  \draw [-latex] (semant.east) -- (incg.west)  node [midway,pos=0.4,rotate=90,fill=white,opacity=0.9] {\small \(t\)-AST};
+  \draw [-latex] (incg.east) -- +(0.4,0) |- node[pos=0.25,fill=white,opacity=0.9] {IR} (cfa.east);
+  \draw [-latex] (cfa.west) -- (lva.east)  node [midway,pos=0.4,rotate=-90,fill=white,opacity=0.9] {\small Flow Graph};
+  \draw [-latex] (lva.west) -- (alloc.east)  node [midway,pos=0.4,rotate=-90,fill=white,opacity=0.9] {\small Int. Graph};
+  \draw [-latex] (alloc.west) -- (x86.east)  node [midway,pos=0.4,rotate=-90,fill=white,opacity=0.9] {\small Reg. Asn.};
+  \draw [-latex] (x86.west) -- (exit.east)  node [midway,pos=0.4,rotate=-90,fill=white,opacity=0.9] {\small x86};
+
+  % backend outline
+  \fill [opacity=0.1] ([shift={(-0.15,-0.15)}]x86.south west) rectangle ([shift={(0.8,0.15)}]cfa.north east);
+
+  \draw [decorate,decoration={brace,amplitude=5pt,mirror,raise=4ex}]
+    (0.3,-1) -- (0.3,1) node[midway,xshift=3em,rotate=-90]{Backend};
+  \draw [decorate,decoration={brace,amplitude=5pt,mirror,raise=4ex}]
+    (0.3,1.25) -- (0.3,3.25) node[midway,xshift=3em,rotate=-90]{Frontend};
+\end{tikzpicture}}
+\caption{Compiler phases, backend highlighted}
 \end{figure}
+A critical part of the backend stage is assigning each variable (also referred to as \textit{virtual}) of the source program to a physical register on the target processor.
 
@@
 
-\section{Review of literature}
+\section{Review of Literature}
 
-Write about Appel book.
+The two primary sources are the Tiger- \cite{AppelAndrewW} and Dragonbook \cite{alma99122774374105763}, in addition to scientific publications concerning the linear scan approach.
 
-Write about SSA form liveness analysis.
+\subsection{Control Flow Analysis}
 
-Keywords: program points, dataflow equations, fixed point
+The backend of a compiler takes some form of IR as input, usually a linear sequence of instructions for each separate function.
+This representation is close to the level of an actual processor by design, but it isn't immediately useful for the further analysis steps needed to generate optimized code for the target architecture.
+%%(FIXME: why not?)
+The control flow of a given program refers to the order in which instructions are executed.
+While the flow of most instructions is linear, in the sense that the next instruction to be executed is located immediately after, some send the execution elsewhere or outright terminate it.
+%% FIXME: why is this important?
+
+These instructions are referred to as \textit{terminators}, as they potentially terminate a contiguous flow of execution.
+This subset of instructions can be further divided into \textit{branch instructions} and those concerned with handling the callstack in function/subroutine management.
+
+These so-called \textit{branch instructions} that alter the flow of execution are further c
+
+hence a \textit{control flow graph} is derived from these.
+A terminator will branch elsewhere to continue execution.
+The destination of this has to be an annotated by a label, hence labels \textit{initiate} blocks.
+
+\subsection{Liveness Analysis}
+
+Liveness analysis is the process of finding the program points at which a variable is live.
+This serves the function of determining which variables are live simultaneously, i.e.
+in interference/conflict with one another, so cannot be assigned the same physical register without overwriting the value of one another in execution.
+These can be derived recursively by iteratively applying so-called \textit{dataflow equations} to each node in a \textit{control-flow graph} (CFG) until a stable state/fixed point is reached.
+
+The algorithm described in \cite{AppelAndrewW} and \cite{alma99122774374105763} is based on the following equations for the live-in and live-out variables respectively:
+\begin{align}
+  in\left[n\right] &= use\left[n\right] \cup (out\left[n\right] - def\left[n\right])\\
+  out\left[n\right] &= \bigcup_{s\in succ\left[n\right]} in\left[s\right]
+\end{align}
+where \(use\left[n\right]\) is the set of all dependent variables, \(def\left[n\right]\) is the set of all variables defines and \(succ\left[n\right]\) is the set of all immediate successor nodes of \(n\).
+
+The structure of the CFG is inherently vague as both  \cite{AppelAndrewW} and \cite{alma99122774374105763}, while based on the same unde
+
+rlying conceptions of dataflow, have different approaches to nodes of a CFG, with the former taking each individual instruction into consideration and the latter each basic block.
+Either of these approaches are sound
+
+% TODO: try int set instead of symbols
+
+\subsection{Graph Coloring}
+
+\subsection{Linear Scan}
+
+%Write about Appel book.
+
+%Write about SSA form liveness analysis.
+
+%Keywords: program points, dataflow equations, fixed point
 
@@
 The instructions included are
-\begin{enumerate}
-    \item binary operations \lstinline!add!, \lstinline!and!, \lstinline!ashr!, \lstinline!lshr!, \lstinline!mul!, \lstinline!or!, \lstinline!sdiv!, \lstinline!srem!, \lstinline!shl!, \lstinline!sub!
-and \lstinline!xor!
-    \item integer comparison \lstinline!icmp! (conditions being \lstinline!eq!, \lstinline!ne!, \lstinline!sge!, \lstinline!sgt!, \lstinline!sle! and \lstinline!slt!)
-    \item memory/address operations \lstinline!alloca!, \lstinline!gep!, \lstinline!load! and \lstinline!store!
-    \item  \lstinline!mov! operations \lstinline!gep!, \lstinline!phi!, \lstinline!ptrtoint! and \lstinline!zext!
-    \item control flow operations \lstinline!br!, \lstinline!call! and \lstinline!ret!
-    \item a  nullary terminator \lstinline!unreachable!
-\end{enumerate}
+\lstinline!trunc! has only been added in order to help cover more generated LLVM.
 %with only four distinct block terminators 
@@
 
-For a given program, it's representation as linear sequence of instructions from the above set isn't immediately useful for liveness analysis.
-(FIXME: why not?)  hence a \textit{control flow graph} is derived from these.
-A terminator will branch elsewhere to continue execution.
-The destination of this has to be an annotated by a label, hence labels \textit{initiate} blocks.
-
 With an input stream of instructions
@@
 \section{Liveness Analysis}
+
+Dataflow analysis
 
@@
 %* optimistic coloring
+\subsection{Coalescing}
 
@@
 
-\section{Codegen}
+\section{Instruction Selection}
 
@@
 
-\section{Optimization}
+\section{Further Optimization}
 
@@
 Benchmarks are performed on Apple Silicon (M1 Pro, i.e.
-ARMv8.5-A \cite{AM1}) despite the target instruction set being x86-64.
-This is possible because of Rosetta 2, which  translates x86-64 to ARMv8 before execution \cite{A2012}.
+ARMv8.5-A \cite{article}) despite the target instruction set being x86-64.
+This is possible because of Rosetta 2, which  translates x86-64 to ARMv8 before execution \cite{PolettoMassimiliano1999Lsra}.
 Time is measured using the \lstinline!time!
@@
 %* how do I measure cpu cycles?
+\begin{center}
+\begin{tabular}{ c c c }
+ cell1 & cell2 & cell3 \\ 
+ cell4 & cell5 & cell6 \\  
+ cell7 & cell8 & cell9    
+\end{tabular}
+\end{center}
+
+\begin{center}
+\begin{tabular}{|c||c|c|c|c|} 
+ \hline
+ sha256 & -O0 & -O1 & -O2 & -O3 \\ [0.5ex] 
+ \hline
+ 1000000 & 1 & 0.73 & 0.73 & 0.73 \\ [1ex] 
+ \hline
+\end{tabular}
+\end{center}
 
@@
 \addtocontents{toc}{\protect\setcounter{tocdepth}{1}}
+
+\section{LLVM-- instruction set}
+\label{appendix:llvmis}
+\begin{enumerate}
+    \item binary operations \lstinline!add!, \lstinline!and!, \lstinline!ashr!, \lstinline!lshr!, \lstinline!mul!, \lstinline!or!, \lstinline!sdiv!, \lstinline!srem!, \lstinline!shl!, \lstinline!sub!
+and \lstinline!xor!
+    \item integer comparison \lstinline!icmp! (conditions being \lstinline!eq!, \lstinline!ne!, \lstinline!sge!, \lstinline!sgt!, \lstinline!sle! and \lstinline!slt!)
+    \item memory/address operations \lstinline!alloca!, \lstinline!gep!, \lstinline!load! and \lstinline!store!
+    \item  \lstinline!mov! operations \lstinline!gep!, \lstinline!phi!, \lstinline!ptrtoint!, \lstinline!trunc!, and \lstinline!zext!
+    \item control flow operations \lstinline!br!, \lstinline!call! and \lstinline!ret!
+    \item a  nullary terminator \lstinline!unreachable!
+\end{enumerate}
 

### 2023-11-24 article.tex
Looks good -Jane

@@
 
+% mention basic block early on, they're a contiguous sequence of instructions that do not transfer control flow
+
 These instructions are referred to as \textit{terminators}, as they potentially terminate a contiguous flow of execution.
@@
 
-These so-called \textit{branch instructions} that alter the flow of execution are further c
+%These so-called \textit{branch instructions} that alter the flow of execution are further c
 
@@
 This serves the function of determining which variables are live simultaneously, i.e.
-in interference/conflict with one another, so cannot be assigned the same physical register without overwriting the value of one another in execution.
+in interference/conflict with one another, which cannot be assigned the same physical register without overwriting the value of one another in execution.
 These can be derived recursively by iteratively applying so-called \textit{dataflow equations} to each node in a \textit{control-flow graph} (CFG) until a stable state/fixed point is reached.
@@
 \begin{align}
-  in\left[n\right] &= use\left[n\right] \cup (out\left[n\right] - def\left[n\right])\\
-  out\left[n\right] &= \bigcup_{s\in succ\left[n\right]} in\left[s\right]
+  in\left[n\right] &= use\left[n\right] \cup (out\left[n\right] - \mathit{def}\left[n\right])\\
+  out\left[n\right] &= \bigcup_{s\in \mathit{succ}\left[n\right]} in\left[s\right]
 \end{align}
@@
 
-The structure of the CFG is inherently vague as both  \cite{AppelAndrewW} and \cite{alma99122774374105763}, while based on the same unde
-
-rlying conceptions of dataflow, have different approaches to nodes of a CFG, with the former taking each individual instruction into consideration and the latter each basic block.
+The structure of the CFG is inherently vague as both  \textit{Modern Compiler Implementation in ML} \cite{AppelAndrewW} and \textit{Compilers : Principles, Techniques, and Tools} \cite{alma99122774374105763}, while based on the same underlying % TODO: fix references
+conceptions of dataflow, have different approaches to nodes of a CFG, with the former taking each individual instruction into consideration and the latter each basic block.
 Either of these approaches are sound
@@
 \section{Instruction Selection}
+
+Translating each IR instruction to x86 correctly is a matter of eliminating unintended side-effects. Each LLVM-- instruction is defined to have only one purpose, %% FIXME is purpose the right word?
+as concepts such as calling conventions, stack frames, or a \lstinline!FLAGS! register are completely abstracted over in order to remain platform independent.
+
+In contrast, the x86 \textit{instruction set architecture} (ISA) is targeting a  \textit{complex instruction set computer} (CISC) family of processors, the instructions of which perform a much broader set of operations \cite[p.~190]{AppelAndrewW}.
+This is is in part due to pipelining, i.e.
+an abstraction over the concrete implementation of the actual processor, which in turn is made for the sake of performance.
+
+An example of this would be the division/remainder operation: since integer division is a non-trivial iterative process wherein both the quotient and remainder is needed throughout, the result of both of these is stored in the \lstinline!%rax!
+and \lstinline!%rdx!
+registers respectively.
+This means two operations are performed simultaneously regardless of which value is used, hence they need to be restored before executing the next instruction, as any variables assigned to \lstinline!%rax!
+or \lstinline!%rdx!
+will be overwritten.
+
+\begin{center}
+\begin{tabular}{|c||c|c|c|c|} 
+ \hline
+ sha256 & -O0 & -O1 & -O2 & -O3 \\ [0.5ex] 
+ \hline
+ 1000000 & 1 & 0.73 & 0.73 & 0.73 \\ [1ex] 
+ \hline
+\end{tabular}
+\end{center}
 

### 2023-11-27 article.tex
Add some benchmarks

@@
 
-The two primary sources are the Tiger- \cite{AppelAndrewW} and Dragonbook \cite{alma99122774374105763}, in addition to scientific publications concerning the linear scan approach.
+The two primary sources are the \textit{Modern Compiler Implementation in ML} \cite{AppelAndrewW} and Dragonbook \cite{alma99122774374105763}, in addition to scientific publications concerning the linear scan approach.
 
@@
 
-The algorithm described in \cite{AppelAndrewW} and \cite{alma99122774374105763} is based on the following equations for the live-in and live-out variables respectively:
+The algorithm described in \textit{Modern Compiler Implementation in ML} \cite{AppelAndrewW} and \cite{alma99122774374105763} is based on the following equations for the live-in and live-out variables respectively:
 \begin{align}
@@
 
-The structure of the CFG is inherently vague as both  \textit{Modern Compiler Implementation in ML} \cite{AppelAndrewW} and \textit{Compilers : Principles, Techniques, and Tools} \cite{alma99122774374105763}, while based on the same underlying % TODO: fix references
+The structure of the CFG is inherently vague as both   \cite{AppelAndrewW} and \textit{Compilers : Principles, Techniques, and Tools} \cite{alma99122774374105763}, while based on the same underlying % TODO: fix references
 conceptions of dataflow, have different approaches to nodes of a CFG, with the former taking each individual instruction into consideration and the latter each basic block.
@@
  \hline
- sha256 & -O0 & -O1 & -O2 & -O3 \\ [0.5ex] 
+ add & addx & \{\}\\
  \hline
- 1000000 & 1 & 0.73 & 0.73 & 0.73 \\ [1ex] 
+ mul & imul & \{ \%rax \} \\
+ \hline
+ sdiv & idivx & \{\%rax, \%rdx\} \\
+ \hline
+ srem & idivx & \{\%rax, \%rdx \\
+ \hline
+   call & callx & \{\%rax, \textit{caller saved registers}\} \\
+ \hline
  \hline
@@
  \hline
- sha256 & -O0 & -O1 & -O2 & -O3 \\ [0.5ex] 
+ fib & clang & greedy & briggs & linear \\ [0.5ex] 
+ \hline
+ 40 & 1.000000 & 1.438693 & 1.441270 & 0.73 \\ [1ex] 
+ 41 & 1.000000 & 1.441035 & 1.440626 & 0.73 \\ [1ex] 
+ \hline
+\end{tabular}
+\end{center}
+
+bench clang	benches/fib.ll 42 	min=5494929992 (1.000000)	avg=5545852373 (1.000000)	max=5669289413
+bench greedy	benches/fib.ll 42 	min=7951933724 (1.447140)	avg=7969947833 (1.437101)	max=8004638438
+bench briggs	benches/fib.ll 42 	min=7946186345 (1.446094)
+
+\begin{center}
+\begin{tabular}{|c||c|c|c|c|} 
+ \hline
+ sha256 & clang & greedy & briggs & linear \\ [0.5ex] 
  \hline

### 2023-11-27 article.tex
Expand intro

@@
 
-Compilation refers to the process of translating from one language to another, most often from a programming language intended for humans to work with, to machine-/bytecode intended to be executed on a target platform.
-This process can be divided into several distinct phases, which are grouped into one of two stages, colloquially referred to as the \textit{frontend} and \textit{backend}.
-The former translating a high-level programming language to an \textit{intermediate representation} (IR) layer and the latter translating IR to executable machine code on a target platform (e.g.
-x86, ARM) or bytecode for a target VM (e.g.
-JVM, CLR).
+Compilation refers to the process of translating from one language to another, most often from a high-level programming language intended for humans to work with, to machine- or bytecode intended to be executed on a target platform.
+This process can be divided into several distinct phases, which are grouped into one of two stages colloquially referred to as the \textit{frontend} and \textit{backend}.
+The former translating a high-level programming language to an \textit{intermediate representation} (IR) layer and the latter translating IR to executable machine code on a target platform or bytecode for a target \textit{virtual machine} (VM).
 
@@
 \end{figure}
-A critical part of the backend stage is assigning each variable (also referred to as \textit{virtual}) of the source program to a physical register on the target processor.
+Most operations of a general-purpose programming language are translated to a set of logic,  control, and arithmetic processor instructions, that are intended to be executed sequentially on a computer processor: a single circuit/chip, referred to as the \textit{central processing unit} (CPU).
+CPU design varies and has evolved over time, but most processors executing logic operations are \textit{register machines}, in that they use a limited set of \textit{general-purpose registers} (GPRs) to store working values in combination with \textit{random access memory} (RAM) for mid-term and other I/O peripherals for long-term storage.
+
+The reason for this can largely be attributed to performance, as register machines routinely outperform so-called stack machines \cite{ShiYunhe2008VmsS} that are often used in VMs.
+Due to this, and the limited amount of GPRs available, a crucial part of the backend stage for an optimizing compiler is assigning each variable  of the source program to a GPR in such a way that maximizes performance without sacrificing correctness.
+
+The process of assigning each variable to a GPR is generally referred to as \textit{register allocation}, and
 

### 2023-11-27 article.tex
More detail

@@
 \documentclass{article}
-\usepackage[style=ieee]{biblatex}
+\usepackage[style=apa]{biblatex}
 
@@
 \end{figure}
-Most operations of a general-purpose programming language are translated to a set of logic,  control, and arithmetic processor instructions, that are intended to be executed sequentially on a computer processor: a single circuit/chip, referred to as the \textit{central processing unit} (CPU).
+Most operations of a general-purpose programming language are translated to a set of logic,  control, and arithmetic instructions that are intended to be executed sequentially on a computer processor: a single circuit/chip, referred to as the \textit{central processing unit} (CPU).
 CPU design varies and has evolved over time, but most processors executing logic operations are \textit{register machines}, in that they use a limited set of \textit{general-purpose registers} (GPRs) to store working values in combination with \textit{random access memory} (RAM) for mid-term and other I/O peripherals for long-term storage.
 
-The reason for this can largely be attributed to performance, as register machines routinely outperform so-called stack machines \cite{ShiYunhe2008VmsS} that are often used in VMs.
-Due to this, and the limited amount of GPRs available, a crucial part of the backend stage for an optimizing compiler is assigning each variable  of the source program to a GPR in such a way that maximizes performance without sacrificing correctness.
+This can largely be attributed to performance, as register machines routinely outperform so-called \textit{stack machines} \parencite{ShiYunhe2008VmsS} that are often used in VMs.
+Because of the limited amount of GPRs available simultaneously, a crucial part of the backend stage for an optimizing compiler is assigning each variable  of the source program to a GPR in such a way that maximizes performance without sacrificing correctness.
 
-The process of assigning each variable to a GPR is generally referred to as \textit{register allocation}, and
+The process of assigning each variable to a GPR is referred to as \textit{register allocation}, and can be approached in several different ways.
+This paper will seek to implement graph coloring and linear scan and evaluate them in terms of runtime performance after compilation.
+%% TODO: compare with time-complexity/compile-time duration?
+The primary sources will be \textit{Modern Compiler Implementation in ML} \parencite{AppelAndrewW} and \textit{Compilers: Principles, techniques, and tools} by \parencite{alma99122774374105763}, in addition to scientific publications concerning the linear scan approach.
 

### 2023-11-27 article.tex
Fix hbox

@@
 \begin{enumerate}
-    \item binary operations \lstinline!add!, \lstinline!and!, \lstinline!ashr!, \lstinline!lshr!, \lstinline!mul!, \lstinline!or!, \lstinline!sdiv!, \lstinline!srem!, \lstinline!shl!, \lstinline!sub!
+    \item binary operations \lstinline!add!, \lstinline!and!, \lstinline!ashr!, \lstinline!lshr!, \lstinline!mul!, \lstinline!or!, \lstinline!sdiv!,\\\lstinline!srem!, \lstinline!shl!, \lstinline!sub!
 and \lstinline!xor!
@@
     \item control flow operations \lstinline!br!, \lstinline!call! and \lstinline!ret!
-    \item a  nullary terminator \lstinline!unreachable!
+    \item a  nullary block terminator \lstinline!unreachable!
 \end{enumerate}

### 2023-11-27 article.tex
Fix references

@@
 %% TODO: compare with time-complexity/compile-time duration?
-The primary sources will be \textit{Modern Compiler Implementation in ML} \parencite{AppelAndrewW} and \textit{Compilers: Principles, techniques, and tools} by \parencite{alma99122774374105763}, in addition to scientific publications concerning the linear scan approach.
+The primary sources will be \textit{Modern Compiler Implementation in ML} \parencite{tiger} and \textit{Compilers: Principles, techniques, and tools} \parencite{dragon}, in addition to publications concerning the linear scan approach.
 
@@
 
-The two primary sources are the \textit{Modern Compiler Implementation in ML} \cite{AppelAndrewW} and Dragonbook \cite{alma99122774374105763}, in addition to scientific publications concerning the linear scan approach.
+The two primary sources are the \textit{Modern Compiler Implementation in ML} \cite{tiger} and Dragonbook \cite{dragon}, in addition to scientific publications concerning the linear scan approach.
 
@@
 
-The algorithm described in \textit{Modern Compiler Implementation in ML} \cite{AppelAndrewW} and \cite{alma99122774374105763} is based on the following equations for the live-in and live-out variables respectively:
+The algorithm described in \textit{Modern Compiler Implementation in ML} \cite{tiger} and \cite{dragon} is based on the following equations for the live-in and live-out variables respectively:
 \begin{align}
@@
 
-The structure of the CFG is inherently vague as both   \cite{AppelAndrewW} and \textit{Compilers : Principles, Techniques, and Tools} \cite{alma99122774374105763}, while based on the same underlying % TODO: fix references
+The structure of the CFG is inherently vague as both   \cite{tiger} and \textit{Compilers : Principles, Techniques, and Tools} \cite{dragon}, while based on the same underlying % TODO: fix references
 conceptions of dataflow, have different approaches to nodes of a CFG, with the former taking each individual instruction into consideration and the latter each basic block.
@@
 
-In contrast, the x86 \textit{instruction set architecture} (ISA) is targeting a  \textit{complex instruction set computer} (CISC) family of processors, the instructions of which perform a much broader set of operations \cite[p.~190]{AppelAndrewW}.
+In contrast, the x86 \textit{instruction set architecture} (ISA) is targeting a  \textit{complex instruction set computer} (CISC) family of processors, the instructions of which perform a much broader set of operations \cite[p.~190]{tiger}.
 This is is in part due to pipelining, i.e.

### 2023-11-29 article.tex
Intro basic blocks before terms

@@
 % mention basic block early on, they're a contiguous sequence of instructions that do not transfer control flow
+
+These contiguous sequences of instructions are referred to as \textit{basic blocks}: a basic block is defined as a sequence of instructions with no branches in or out except for the first and last instructions, and can be collectively construed of as a node in a control flow graph, whose directed entry edges are from any block that directs control flow towards itself.
 

### 2023-12-01 article.tex
Expan benchmarks

@@
  \hline
- fib & clang & greedy & briggs & linear \\ [0.5ex] 
+ fib & clang & greedy & simple & linear \\ [0.5ex] 
  \hline
@@
  41 & 1.000000 & 1.441035 & 1.440626 & 0.73 \\ [1ex] 
+ 42 & 	1.000000 & 1.447140 & 1.446094 & 0\\
  \hline
@@
 \end{center}
-
-bench clang	benches/fib.ll 42 	min=5494929992 (1.000000)	avg=5545852373 (1.000000)	max=5669289413
-bench greedy	benches/fib.ll 42 	min=7951933724 (1.447140)	avg=7969947833 (1.437101)	max=8004638438
-bench briggs	benches/fib.ll 42 	min=7946186345 (1.446094)
 

### 2023-12-01 article.tex
After LSV meeting

@@
 \usepackage{xpatch}
-\usepackage{xcolor}
 \usepackage{realboxes}
 \usepackage{titlesec}
+\usepackage{xcolor}
 
@@
 
-\definecolor{mygreen}{rgb}{0,0.6,0}
+\definecolor{mygreen}{rgb}{0,0.8,0}
 \definecolor{mygray}{rgb}{0.5,0.5,0.5}
@@
 \end{figure}
-Most operations of a general-purpose programming language are translated to a set of logic,  control, and arithmetic instructions that are intended to be executed sequentially on a computer processor: a single circuit/chip, referred to as the \textit{central processing unit} (CPU).
-CPU design varies and has evolved over time, but most processors executing logic operations are \textit{register machines}, in that they use a limited set of \textit{general-purpose registers} (GPRs) to store working values in combination with \textit{random access memory} (RAM) for mid-term and other I/O peripherals for long-term storage.
+Most operations of a general-purpose programming language are translated to a set of logic,  control, and arithmetic instructions to be executed sequentially on a computer processor: a single circuit/chip, referred to as the \textit{central processing unit} (CPU), the design of which has varied and evolved over time.
+Most processors executing logic operations are \textit{register machines}, in that they use a limited set of \textit{general-purpose registers} (GPRs) to store working values in combination with \textit{random access memory} (RAM) for mid-term, and other I/O peripherals for long-term storage.
 
@@
 The control flow of a given program refers to the order in which instructions are executed.
-While the flow of most instructions is linear, in the sense that the next instruction to be executed is located immediately after, some send the execution elsewhere or outright terminate it.
+While the flow of most instructions is linear, in the sense that the next instruction to be executed is located immediately after, some transfer the flow of execution elsewhere or terminate it.
 %% FIXME: why is this important?
@@
 
-These contiguous sequences of instructions are referred to as \textit{basic blocks}: a basic block is defined as a sequence of instructions with no branches in or out except for the first and last instructions, and can be collectively construed of as a node in a control flow graph, whose directed entry edges are from any block that directs control flow towards itself.
+Such contiguous sequences of instructions are referred to as \textit{basic blocks}: a basic block is defined as a sequence of instructions with no branches in or out except for the first and last instructions, and can be collectively construed of as a node in a control flow graph, whose directed entry edges are from any block that directs control flow towards itself.
 
@@
 
-Benchmarks are performed on Apple Silicon (M1 Pro, i.e.
+%Runtime performance of generated code can be measured 
+%Measuring the performance of a compiler backend involves assessing how well it translates high-level programming code into machine code, considering factors like execution speed and generated code size.
+Here are several methods to evaluate the performance of a given compiler:
+
+%Benchmarks are performed on Apple Silicon (M1 Pro, i.e.
 ARMv8.5-A \cite{article}) despite the target instruction set being x86-64.
@@
 %* how do I measure cpu cycles?
-\begin{center}
-\begin{tabular}{ c c c }
- cell1 & cell2 & cell3 \\ 
- cell4 & cell5 & cell6 \\  
- cell7 & cell8 & cell9    
-\end{tabular}
-\end{center}
+%\begin{center}
+%\begin{tabular}{ c c c }
+ %cell1 & cell2 & cell3 \\ 
+ %cell4 & cell5 & cell6 \\  
+ %cell7 & cell8 & cell9    
+%\end{tabular}
+%\end{center}
 
@@
 
-\begin{center}
-\begin{tabular}{|c||c|c|c|c|} 
- \hline
- sha256 & clang & greedy & briggs & linear \\ [0.5ex] 
- \hline
- 1000000 & 1 & 0.73 & 0.73 & 0.73 \\ [1ex] 
- \hline
-\end{tabular}
-\end{center}
+%\begin{center}
+%\begin{tabular}{|c||c|c|c|c|} 
+ %\hline
+ %sha256 & clang & greedy & briggs & linear \\ [0.5ex] 
+ %\hline
+ %1000000 & 1 & 0.73 & 0.73 & 0.73 \\ [1ex] 
+ %\hline
+%\end{tabular}
+%\end{center}
 

### 2023-12-05 article.tex
Reorder

@@
 
-\section{\lstinline!LLVM--! instruction set}
+\section{Control Flow Analysis}
+
+\subsection{Building a graph}
+
+\subsection{Parameterized over individual instructions}
+
+\subsection{Parameterized over basic blocks}
+
+With an input stream of instructions
+
+%* description of blocks, how they connect etc.
+
+Each function defined in an LLVM program is constructed with the help of
+
+\section{Liveness Analysis}
+
+\subsection{Dataflow Analysis}
+
+Dataflow analysis
+
+\subsection{Constructing an interference graph}
+
+%\section{CFG -> INTERF}
+
+%* describe the dataflow algorithm from the book
+%* describe the interference criteria from the book
+
+%\section{INTERF -> DOT}
+
+%* ez ???
+
+%\section{INTERF -> DOT/Assignment}
+
+%* list of methods for coalescing/allocating:
+%   * ocamlgraph builtin
+%   * greedy 
+%   * briggs
+%   * george
+%   * welsh-powell?
+\section{Graph Coloring}
+The graph coloring by simplification as introduced in the Appel text takes an iterative approach.
+
+\subsection{Coloring by simplification}
+
+%% TODO
+%* optimistic coloring
+\subsection{Coalescing}
+
+Coalescing is the process of eliminating moves/copies of data from one GPR to another by combining their interference graph nodes.
+
+\section{Linear Scan}
+
+%Live ranges for each of the temporaries are derived by visiting each instruction of each control flow block and examining its \textit{uses} and \textit{defs}.
+
+\section{Instruction Selection}
+
+\subsection{\lstinline!LLVM--! instruction set}
 
@@
 
-\section{Control Flow Analysis}
-
-With an input stream of instructions
-
-%* description of blocks, how they connect etc.
-
-Each function defined in an LLVM program is constructed with the help of
-
-\section{Liveness Analysis}
-
-Dataflow analysis
-
-%\section{CFG -> INTERF}
-
-%* describe the dataflow algorithm from the book
-%* describe the interference criteria from the book
-
-%\section{INTERF -> DOT}
-
-%* ez ???
-
-%\section{INTERF -> DOT/Assignment}
-
-%* list of methods for coalescing/allocating:
-%   * ocamlgraph builtin
-%   * greedy 
-%   * briggs
-%   * george
-%   * welsh-powell?
-\section{Graph Coloring}
-The graph coloring by simplification as described by Appel
-
-%% TODO
-%* optimistic coloring
-\subsection{Coalescing}
-
-\section{Linear Scan}
-
-%Live ranges for each of the temporaries are derived by visiting each instruction of each control flow block and examining its \textit{uses} and \textit{defs}.
-
-\section{Instruction Selection}
+%%%
+\subsection{Translating to x86}
 
@@
 %Runtime performance of generated code can be measured
-%Measuring the performance of a compiler backend involves assessing how well it translates high-level programming code into machine code, considering factors like execution speed and generated code size.
-Here are several methods to evaluate the performance of a given compiler:
+
+There are different approaches to assessing how well a compiler backend translates IR to machine code targeting a specific architecture and platform.
+One of the primary means is to simply measure the time it takes to execute the code generated by it.
+
+Since the target architecture is x86-64, all benchmarks are executed natively on the fastest processor available to be used consistently at the time of writing.
+This is to maximize the sample size and in turn reduce uncertainty, but the clock speed at which the benchmarks are performed is irrelevant to how fast the generated code is.
+Because of this, the metric recorded is the amount of CPU cycles spent executing from start to finish instead of actual time in seconds, as the time measure only works as a more imprecise estimate of the amount of underlying work/execution steps performed on the processor.
+
+Needless to say, CPU cycles isn't a perfect measurement either, but it works to eliminate the clock speed as well as negate much of the time the scheduler allocates for other processes.
+Scheduling still has a negative impact on execution because of the cache misses caused by context switching, but outside of executing each program in immediate mode which isn't possible on Linux or macOS this is a sensible estimate.
+To further reduce context switching, benchmarks are made on a fresh restart with minimal background processes running.
+
+Additionally, all benchmarks are deterministic, so the value of each virtual variable will stay the same at every instruction step of execution when compared to another run with the same starting conditions.
+%all the way through the execution compared to another run of the same program with the same input.
+%% FIXME: not necessarily true as loops exist
+This means that the cycles measured over infinitely many runs will converge towards a 'perfect' run with no cache misses as caused by context switching.
+So across \(n\) runs, the run least affected by context switching will be the one with the least CPU cycles, so is the one most representative of the actual performance without noise.
+
+Measurements are made with the \lstinline!perf!
+utility, which relies on the \textit{hardware counters} (HC) registers of modern microprocessors, which are special purpose registers meant to record performance data live during execution.
+Because the analysis is integrated into the chip on which it is executing, very little overhead is incurred, and is therefore the go-to for estimating native performance.
+
+\subsubsection{Measurements}
+
+Each \lstinline!.ll! file present in the \lstinline!benches! directory is listed here in a table denoting the input paramater on the left column and type of allocator used.
+
+%for several reasons, one of which is to negate the clock speed as this holds no relevance to the compiler.
+
+%considering factors like execution speed and generated code size. Here are several methods to evaluate the performance of a given compiler:
 
@@
 
+\paragraph{fib.ll}
+The 'dumb' fib is a very naive approach to finding the \(n\)th number of the Fibonacci sequence by calling itself recursively twice (decrementing \(n\) before each), which causes an exponential growth in function calls.
 \begin{center}

### 2023-12-06 article.tex
Some notes

@@
 \section{Graph Coloring}
-The graph coloring by simplification as introduced in the Appel text takes an iterative approach.
+The heuristic for graph coloring as introduced in both the Appel and Aho et al. texts is a linear time approximation of an NP-complete problem. It is based on an iterative approach.
+
+Let \(k\) be the number of working registers available on the target architecture.
+After building the interference graph \(G\), a node \(n\) with fewer than \(k\) neighbors is chosen.
+As \(n\) has at most \(k-1\) neighbors, it can be removed safely by ensuring it is not assigned any color assigned to its neighbors, effectively simplifying \(G\).
+
+%  (i.e. \(\mathit{deg}(n) < k\))
+
+It must hold that \(G-\{n\}\) is \(k\)-colorable if \(G\) is \(k\)-colorable
 
@@
 \section{Instruction Selection}
+
+%% NOTE: special implementations
+%%
+%% * phi nodes
+%% * calling convention:
+%%   * save/restore callee saved registers in pro/epilogue
+%%   * save/restore caller saved registersk before call
+%% * jump flags for comparison
 

### 2023-12-07 article.tex
asd

@@
 
+\usepackage{subcaption}
 \usepackage{amsmath}
 \usepackage{tikz}
+\usetikzlibrary{angles}
+\usetikzlibrary{arrows.meta}
+\usetikzlibrary{automata}
+\usetikzlibrary{calc}
 \usetikzlibrary{chains}
-\usetikzlibrary{angles, quotes}
 \usetikzlibrary{decorations.pathreplacing}
+\usetikzlibrary{positioning}
+\usetikzlibrary{quotes}
 \usepackage{graphicx} % Required for inserting images
@@
 
+\tikzset{
+box/.style={draw,
+  inner sep=0.3cm,
+    minimum height=1.75cm,
+    align=center},
+cbox/.style={draw,
+  font=\ttfamily,},
+}
+\tikzset{invisible/.style={minimum width=0mm,inner sep=0mm,outer sep=0mm}}
+
 \title{Implementation and comparison of register allocation to translate \lstinline!LLVM--! to x86}
@@
 This process can be divided into several distinct phases, which are grouped into one of two stages colloquially referred to as the \textit{frontend} and \textit{backend}.
-The former translating a high-level programming language to an \textit{intermediate representation} (IR) layer and the latter translating IR to executable machine code on a target platform or bytecode for a target \textit{virtual machine} (VM).
+The former translating a high-level programming language to an \textit{intermediate representation} (IR) and the latter translating IR to executable machine code of a target architecture or bytecode of a target \textit{virtual machine} (VM).
 
@@
 \begin{tikzpicture}
-  \tikzset{
-box/.style={draw,
-  inner sep=0.3cm,
-    minimum height=1.75cm,
-    align=center},
-}
 
@@
 \end{figure}
-Most operations of a general-purpose programming language are translated to a set of logic,  control, and arithmetic instructions to be executed sequentially on a computer processor: a single circuit/chip, referred to as the \textit{central processing unit} (CPU), the design of which has varied and evolved over time.
-Most processors executing logic operations are \textit{register machines}, in that they use a limited set of \textit{general-purpose registers} (GPRs) to store working values in combination with \textit{random access memory} (RAM) for mid-term, and other I/O peripherals for long-term storage.
+Most operations of a general-purpose programming language are translated to a set of control, logic, and arithmetic instructions to be executed sequentially on a computer processor: a single circuit/chip, referred to as the \textit{central processing unit} (CPU), the design of which has varied and evolved over time.
 
-This can largely be attributed to performance, as register machines routinely outperform so-called \textit{stack machines} \parencite{ShiYunhe2008VmsS} that are often used in VMs.
+Most CPUs are \textit{register machines}, in that they use a limited set of \textit{general-purpose registers} (GPRs) to store working values in combination with \textit{random access memory} (RAM) for mid-term, and other I/O peripherals for long-term storage.
+This can largely be attributed to performance, as register machines routinely outperform \textit{stack machines} \parencite{ShiYunhe2008VmsS} that are often used in VMs.
 Because of the limited amount of GPRs available simultaneously, a crucial part of the backend stage for an optimizing compiler is assigning each variable  of the source program to a GPR in such a way that maximizes performance without sacrificing correctness.
@@
 
-\section{Review of Literature}
+%\section{Review of Literature}
 
-The two primary sources are the \textit{Modern Compiler Implementation in ML} \cite{tiger} and Dragonbook \cite{dragon}, in addition to scientific publications concerning the linear scan approach.
+%The two primary sources are the \textit{Modern Compiler Implementation in ML} \cite{tiger} and Dragonbook \cite{dragon}, in addition to scientific publications concerning the linear scan approach.
 
-\subsection{Control Flow Analysis}
+\section{Control Flow Analysis}
+
+% TODO: figure out what traces are
 
@@
 The control flow of a given program refers to the order in which instructions are executed.
-While the flow of most instructions is linear, in the sense that the next instruction to be executed is located immediately after, some transfer the flow of execution elsewhere or terminate it.
+While the flow of most instructions is linear, in the sense that the next instruction executed is located immediately after, some transfer the flow of execution elsewhere or even terminate it.
 %% FIXME: why is this important?
@@
 
-Such contiguous sequences of instructions are referred to as \textit{basic blocks}: a basic block is defined as a sequence of instructions with no branches in or out except for the first and last instructions, and can be collectively construed of as a node in a control flow graph, whose directed entry edges are from any block that directs control flow towards itself.
+Continuous sequences of instructions are referred to as \textit{basic blocks}.
+A basic block is defined as a sequence of instructions with no branches in or out except for the first instruction, which follows immediately either the function entry point or a label,  and the last instruction, which either terminates or transfers the flow of execution, hence these are often referred to as terminators and can be thought of as a different categoriy of instructions entirely.
 
-These instructions are referred to as \textit{terminators}, as they potentially terminate a contiguous flow of execution.
+Basic blocks represent a single node in a \textit{control flow graph} (CFG), which is a directed graph whose edges denote transfer of  control flow.
+The unconditional branch terminator always transfers control flow to the block labelled, meaning only one successor will follow, whereas conditional branching could transfer to either of the two, but because control flow analysis is not concerned with data, it simply adds both as successors.
+% TODO: rets
+While each block has 0-2 immediate successors, the amount of predecessors is unbounded.
+
+\begin{figure}[h]
+  \centering
+  \begin{subfigure}{0.48\textwidth}
+  \centering
+  \begin{tikzpicture}
+    \node [cbox, font=\ttfamily\tiny, align=left] (cfa) at (0,0) {define i32 @square(i32 \%0) \{\\1:\\\ \%2 = icmp sgt i32 \%0, 0\\\ br i1 \%2, label \%3, label \%9\ ; succs = \%3, \%9\\3:\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ ; preds = \%1, \%3\\\ \%4 = phi i32 [ \%0, \%1 ], [ \%7, \%3 ]\\\ \%5 = phi i32 [ 0, \%1 ], [ \%6, \%3 ]\\\ \%6 = add i32 \%5, \%0\\\ \%7 = sub i32 \%4, 1\\\ \%8 = icmp ugt i32 \%7, 0\\\ br i1 \%8, label \%3, label \%9\ ; succs = \%3, \%9\\9:\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ ; preds = \%3, \%1\\\ \%10 = phi i32 [ 0, \%1 ], [ \%6, \%3 ]\\\ ret i32 \%10\\\}};
+  \end{tikzpicture}
+  \end{subfigure}
+  \begin{subfigure}{0.48\textwidth}
+  \centering
+  \begin{tikzpicture}
+    \node [cbox, font=\ttfamily\tiny, align=left] (a) at (0,0) {1:\\\ \%2 = icmp sgt i32 \%0, 0\\\ br i1 \%2, label \%3, label \%9\ \ \ \ \ \ \ };
+    \node [cbox, font=\ttfamily\tiny, align=left, below=0.5cm of a] (b) {3:\\\ \%4 = phi i32 [ \%0, \%1 ], [ \%7, \%3 ]\\\ \%5 = phi i32 [ 0, \%1 ], [ \%6, \%3 ]\\\ \%6 = add i32 \%5, \%0\\\ \%7 = sub i32 \%4, 1\\\ \%8 = icmp ugt i32 \%7, 0\\\ br i1 \%8, label \%3, label \%9\\9:\\\ \%10 = phi i32 [ 0, \%1 ], [ \%6, \%3 ]\\\ ret i32 \%10\\\}};
+    \node [cbox, font=\ttfamily\tiny, align=left, below=0.5cm of b] (c) {9:\\\ \%10 = phi i32 [ 0, \%1 ], [ \%6, \%3 ]\\\ ret i32 \%10};
+    \node[right=2cm, invisible] (ab) at ($(a.south)!0.5!(b.north)$) {};
+    \node[right=2cm, invisible] (bc) at ($(b.south)!0.5!(c.north)$) {};
+  \draw [-latex] (a) -- (b);
+  %\draw [-latex] (a) to[out=-15,in=15] (c);
+  \draw [-latex] (b) -- (c);
+  %\draw [-latex] (c.north) to[bend right=1000] (b.north);
+  \draw  (a.south) to [in=0] (ab);
+  \draw  (ab) to [in=0] (bc);
+  \draw [-latex]  (bc) to [out=180,in=70] (c.north);
+  \end{tikzpicture}
+  \end{subfigure}
+  \hfill
+  \caption{Compiler phases, backend highlighted}
+\end{figure}
+
+%These instructions are referred to as \textit{terminators}, as they potentially terminate a contiguous flow of execution.
 This subset of instructions can be further divided into \textit{branch instructions} and those concerned with handling the callstack in function/subroutine management.
@@
 
-hence a \textit{control flow graph} is derived from these.
+%hence a \textit{control flow graph} is derived from these.
 A terminator will branch elsewhere to continue execution.
@@
 
-\subsection{Liveness Analysis}
+\subsection{Building a graph}
+
+\subsection{Parameterized over individual instructions}
+
+\subsection{Parameterized over basic blocks}
+
+With an input stream of instructions
+
+%* description of blocks, how they connect etc.
+
+Each function defined in an LLVM program is constructed with the help of
+
+\section{Liveness Analysis}
+
+% * Different approaches:
+%   * Naive/greedy 
+%   * Dataflow analysis 
+%     * forward must
+%     * backwards may
+%   * linear scan
 
@@
 
-\subsection{Graph Coloring}
+The purpose of liveness analysis is to determine which variables are live at which program point, which in turn is used to
 
-\subsection{Linear Scan}
+\subsection{Dataflow Analysis}
+
+Dataflow analysis
+
+\subsection{Constructing an interference graph}
+
+%\section{CFG -> INTERF}
+
+%* describe the dataflow algorithm from the book
+%* describe the interference criteria from the book
+
+%\section{INTERF -> DOT}
+
+%* ez ???
+
+%\section{INTERF -> DOT/Assignment}
+
+%* list of methods for coalescing/allocating:
+%   * ocamlgraph builtin
+%   * greedy 
+%   * briggs
+%   * george
+%   * welsh-powell?
+\section{Graph Coloring}
+The heuristic for graph coloring as introduced in Appel and Aho et al. texts is a linear time approximation of an NP-complete problem. It is based on an iterative approach.
+
+Let \(k\) be the number of working registers available on the target architecture.
+After building the interference graph \(G\), a node \(n\) with fewer than \(k\) neighbors is chosen.
+As \(n\) has at most \(k-1\) neighbors, it can be removed safely by ensuring it is not assigned any color assigned to its neighbors, effectively simplifying \(G\).
+
+%  (i.e. \(\mathit{deg}(n) < k\))
+
+It must hold that \(G-\{n\}\) is \(k\)-colorable if \(G\) is \(k\)-colorable
+
+\subsection{Coloring by simplification}
+
+%% TODO
+%* optimistic coloring
+\subsection{Coalescing}
+
+Coalescing is the process of eliminating moves/copies of data from one GPR to another by combining their interference graph nodes.
+
+\section{Linear Scan}
 
@@
 
-\section{Control Flow Analysis}
-
-\subsection{Building a graph}
-
-\subsection{Parameterized over individual instructions}
-
-\subsection{Parameterized over basic blocks}
-
-With an input stream of instructions
-
-%* description of blocks, how they connect etc.
-
-Each function defined in an LLVM program is constructed with the help of
-
-\section{Liveness Analysis}
-
-\subsection{Dataflow Analysis}
-
-Dataflow analysis
-
-\subsection{Constructing an interference graph}
-
-%\section{CFG -> INTERF}
-
-%* describe the dataflow algorithm from the book
-%* describe the interference criteria from the book
-
-%\section{INTERF -> DOT}
-
-%* ez ???
-
-%\section{INTERF -> DOT/Assignment}
-
-%* list of methods for coalescing/allocating:
-%   * ocamlgraph builtin
-%   * greedy 
-%   * briggs
-%   * george
-%   * welsh-powell?
-\section{Graph Coloring}
-The heuristic for graph coloring as introduced in both the Appel and Aho et al. texts is a linear time approximation of an NP-complete problem. It is based on an iterative approach.
-
-Let \(k\) be the number of working registers available on the target architecture.
-After building the interference graph \(G\), a node \(n\) with fewer than \(k\) neighbors is chosen.
-As \(n\) has at most \(k-1\) neighbors, it can be removed safely by ensuring it is not assigned any color assigned to its neighbors, effectively simplifying \(G\).
-
-%  (i.e. \(\mathit{deg}(n) < k\))
-
-It must hold that \(G-\{n\}\) is \(k\)-colorable if \(G\) is \(k\)-colorable
-
-\subsection{Coloring by simplification}
-
-%% TODO
-%* optimistic coloring
-\subsection{Coalescing}
-
-Coalescing is the process of eliminating moves/copies of data from one GPR to another by combining their interference graph nodes.
-
-\section{Linear Scan}
+As linear scan
 
@@
 
+\newpage
 \subsection{Benchmarking}
@@
 Because the analysis is integrated into the chip on which it is executing, very little overhead is incurred, and is therefore the go-to for estimating native performance.
+
+%% NOTE: måske noget om hvordan intel laver sin egen arkitektur?
 

### 2023-12-07 article.tex
Color changes for 8/12

@@
 This process can be divided into several distinct phases, which are grouped into one of two stages colloquially referred to as the \textit{frontend} and \textit{backend}.
-The former translating a high-level programming language to an \textit{intermediate representation} (IR) and the latter translating IR to executable machine code of a target architecture or bytecode of a target \textit{virtual machine} (VM).
+\textcolor{orange}{The former translating a high-level programming language to an \textit{intermediate representation} (IR) and the latter translating IR to executable machine code of a target architecture or bytecode of a target \textit{virtual machine} (VM).}
 
@@
 
+\textcolor{mygreen}{
 Continuous sequences of instructions are referred to as \textit{basic blocks}.
 A basic block is defined as a sequence of instructions with no branches in or out except for the first instruction, which follows immediately either the function entry point or a label,  and the last instruction, which either terminates or transfers the flow of execution, hence these are often referred to as terminators and can be thought of as a different categoriy of instructions entirely.
+}
 
+\textcolor{mygreen}{
 Basic blocks represent a single node in a \textit{control flow graph} (CFG), which is a directed graph whose edges denote transfer of  control flow.
@@
 While each block has 0-2 immediate successors, the amount of predecessors is unbounded.
+}
 
@@
 
+\color{mygreen}
 The purpose of liveness analysis is to determine which variables are live at which program point, which in turn is used to
@@
 Coalescing is the process of eliminating moves/copies of data from one GPR to another by combining their interference graph nodes.
+\color{black}
 
@@
 
-\newpage
+\color{mygreen}
 \subsection{Benchmarking}
@@
 The 'dumb' fib is a very naive approach to finding the \(n\)th number of the Fibonacci sequence by calling itself recursively twice (decrementing \(n\) before each), which causes an exponential growth in function calls.
+\color{black}
 \begin{center}

### 2023-12-08 article.tex
Jane approved

--- 2023-12-07 article.tex
@@
 
-Compilation refers to the process of translating from one language to another, most often from a high-level programming language intended for humans to work with, to machine- or bytecode intended to be executed on a target platform.
-This process can be divided into several distinct phases, which are grouped into one of two stages colloquially referred to as the \textit{frontend} and \textit{backend}.
-The former translating a high-level programming language to an \textit{intermediate representation} (IR) and the latter translating IR to executable machine code of a target architecture or bytecode of a target \textit{virtual machine} (VM).
+Compilation refers to the process of translating from one language to another, most often from a high-level programming language intended for humans to work with, to machine- or bytecode intended to be executed on a target architecture.
+This process can be divided into several distinct phases, which are grouped into one of two stages colloquially referred to as the \textit{frontend} and \textit{backend}, the former translating a high-level programming language to an \textit{intermediate representation} (IR) and the latter translating IR to executable machine code of a target architecture or bytecode of a target \textit{virtual machine} (VM).
 
@@
 
-Continuous sequences of instructions are referred to as \textit{basic blocks}.
-A basic block is defined as a sequence of instructions with no branches in or out except for the first instruction, which follows immediately either the function entry point or a label,  and the last instruction, which either terminates or transfers the flow of execution, hence these are often referred to as terminators and can be thought of as a different categoriy of instructions entirely.
+A continuous sequence of instructions is referred to as a \textit{basic block}, which is defined as a sequence with no branches in or out except for the first instruction, following immediately after the function entry or a label, and the last instruction, either terminates or transferring the flow of execution.
+These are often referred to as terminators and can be thought of as an entirely different categoriy of instructions.
+% TODO: måske drop 'can be thought of'
 
@@
 
-\newpage
 \subsection{Benchmarking}

### 2023-12-08 article.tex
Make cfg example look ok

@@
     \node [cbox, font=\ttfamily\tiny, align=left, below=0.5cm of b] (c) {9:\\\ \%10 = phi i32 [ 0, \%1 ], [ \%6, \%3 ]\\\ ret i32 \%10};
-    \node[right=2cm, invisible] (ab) at ($(a.south)!0.5!(b.north)$) {};
-    \node[right=2cm, invisible] (bc) at ($(b.south)!0.5!(c.north)$) {};
+
+    \node[right= 1cm of a.south, invisible] (ac1)  {};
+    \node[below= 0.166667cm of ac1, invisible] (ac2)  {};
+    \node[right= 1.333333cm of ac2, invisible] (ac3)  {};
+    \node[right= 1cm of c.north, invisible] (ac6)  {};
+    \node[above= 0.166667cm of ac6, invisible] (ac5)  {};
+    \node[right= 1.333333cm of ac5, invisible] (ac4)  {};
+
+    \node[right= 1cm of b.north, invisible] (bb1)  {};
+    \node[above= 0.166667cm of bb1, invisible] (bb2)  {};
+    \node[right= 1.166667cm of bb2, invisible] (bb3)  {};
+    \node[right= 1cm of b.south, invisible] (bb6)  {};
+    \node[below= 0.166667cm of bb6, invisible] (bb5)  {};
+    \node[right= 1.166667cm of bb5, invisible] (bb4)  {};
+
+    %\node[right=2.5cm of ab, invisible] (abr)  {};
+    %\node[right=2.5cm of bc, invisible] (bcr)  {};
   \draw [-latex] (a) -- (b);
@@
   \draw [-latex] (b) -- (c);
+
+  \draw          (ac1.center) -- (ac2.center) -- (ac3.center) -- (ac4.center) -- (ac5.center);
+  \draw [-latex] (ac5.center) -- (ac6.center);
+
+  \draw          (bb2.center) -- (bb3.center) -- (bb4.center) -- (bb5.center) -- (bb6.center);
+  \draw [-latex] (bb2.center) -- (bb1.center);
   %\draw [-latex] (c.north) to[bend right=1000] (b.north);
-  \draw  (a.south) to [in=0] (ab);
-  \draw  (ab) to [in=0] (bc);
-  \draw [-latex]  (bc) to [out=180,in=70] (c.north);
+  %\draw  (a.south) to  (ab);
+  %\draw  (ab) to  (abr);
+  %\draw  (abr) to  (bcr);
+  %\draw  (bcr) to  (bc);
+  %\draw  (ab) to  (bc);
+  %\draw [-latex]  (bc) to [out=180,in=70] (c.north);
   \end{tikzpicture}
@@
   \hfill
-  \caption{Compiler phases, backend highlighted}
+  \caption{CFG of \lstinline!@square! of \lstinline!square.ll!} % FIXME: not entirely correct (square.ll is @main)
 \end{figure}

### 2023-12-12 article.tex
Ok.

@@
   showtabs=false,                  % show tabs within strings adding particular underscores
-  stepnumber=2,                    % the step between two line-numbers. If it's 1, each line will be numbered
+  %stepnumber=2,                    % the step between two line-numbers. If it's 1, each line will be numbered
   stringstyle=\color{mymauve},     % string literal style
@@
 }
+
+\definecolor{codegreen}{rgb}{0,0.6,0}
+\definecolor{codegray}{rgb}{0.5,0.5,0.5}
+\definecolor{codepurple}{rgb}{0.58,0,0.82}
+\definecolor{backcolour}{rgb}{0.95,0.95,0.92}
+
+\lstdefinestyle{mystyle}{
+    backgroundcolor=\color{backcolour},   
+    commentstyle=\color{codegreen},
+    keywordstyle=\color{magenta},
+    numberstyle=\tiny\color{codegray},
+    stringstyle=\color{codepurple},
+    basicstyle=\ttfamily\footnotesize,
+    breakatwhitespace=false,         
+    %breaklines=true,                 
+    %captionpos=b,                    
+    keepspaces=true,                 
+    numbers=left,                    
+    numbersep=5pt,                  
+    showspaces=false,                
+    showstringspaces=false,
+    showtabs=false,                  
+    tabsize=2
+}
+
+\lstset{style=mystyle}
 
@@
 % TODO: figure out what traces are
+% TODO: mention dominators etc.
 
@@
 
-A continuous sequence of instructions is referred to as a \textit{basic block}, which is defined as a sequence with no branches in or out except for the first instruction, following immediately after the function entry or a label, and the last instruction, either terminates or transferring the flow of execution.
-These are often referred to as terminators and can be thought of as an entirely different categoriy of instructions.
+A continuous flow of instructions is referred to as a \textit{basic block}, defined as a sequence of instructions with no branches in or out except for the first instruction (referred to as a \textit{leader}, immediately following either the function entry or label) and the last (referred to as a \textit{terminator}, as it either terminates or transfers the flow of execution).
+%These are often referred to as terminators and can be thought of as an entirely different categoriy of instructions.
 % TODO: måske drop 'can be thought of'
@@
 % TODO: rets
-While each block has 0-2 immediate successors, the amount of predecessors is unbounded.
+While each block has 0-2 immediate successors, the amount of predecessors is unbounded as the amount of branches targeting a specific leader is unlimited.
 
@@
     \node [cbox, font=\ttfamily\tiny, align=left] (a) at (0,0) {1:\\\ \%2 = icmp sgt i32 \%0, 0\\\ br i1 \%2, label \%3, label \%9\ \ \ \ \ \ \ };
-    \node [cbox, font=\ttfamily\tiny, align=left, below=0.5cm of a] (b) {3:\\\ \%4 = phi i32 [ \%0, \%1 ], [ \%7, \%3 ]\\\ \%5 = phi i32 [ 0, \%1 ], [ \%6, \%3 ]\\\ \%6 = add i32 \%5, \%0\\\ \%7 = sub i32 \%4, 1\\\ \%8 = icmp ugt i32 \%7, 0\\\ br i1 \%8, label \%3, label \%9\\9:\\\ \%10 = phi i32 [ 0, \%1 ], [ \%6, \%3 ]\\\ ret i32 \%10\\\}};
+    \node [cbox, font=\ttfamily\tiny, align=left, below=0.5cm of a] (b) {3:\\\ \%4 = phi i32 [ \%0, \%1 ], [ \%7, \%3 ]\\\ \%5 = phi i32 [ 0, \%1 ], [ \%6, \%3 ]\\\ \%6 = add i32 \%5, \%0\\\ \%7 = sub i32 \%4, 1\\\ \%8 = icmp ugt i32 \%7, 0\\\ br i1 \%8, label \%3, label \%9};
     \node [cbox, font=\ttfamily\tiny, align=left, below=0.5cm of b] (c) {9:\\\ \%10 = phi i32 [ 0, \%1 ], [ \%6, \%3 ]\\\ ret i32 \%10};
@@
 %   * linear scan
+%
+%  TODO: comment on interrupts
 
-Liveness analysis is the process of finding the program points at which a variable is live.
+Translating IR with an unbounded number of variables to a CPU with a bounded number of registers involves the process of assigning each variable a register s.t.
+no value that can be used in the future is overwritten.
+%Therefore, any variable that may be used in the future is considered \textit{live}, and said to be in interference with any other variable live at any intersecting point.
+Variables that are in use at a given program point are considered \textit{'live'}, and although variables can be assigned the same register, %if they are not in interference
+variables that are live at the same time (i.e. at intersecting program points) cannot, in which case they are also said to be in interference with one another.
+Although variables that are not in interference can be assigned the same register, and finding the precise points at which any variable is live is trivial for linear sequences of instructions, when conditional branching is introduced, deriving the path of execution becomes undecidable because of the halting problem.
+
+Suppose a function that calls another:
+
+\begin{lstlisting}[language=LLVM,firstnumber=1]
+define i32 @countcall(i32 %x0) {
+  %x1 = add i32 %x0, 1
+  call ptr @subproc()
+  ret i32 %x1
+}
+\end{lstlisting}
+Because of the halting problem, static analysis cannot determine if the call to \lstinline!@subproc!
+will return for every possible implementation.
+%which may or may not return control to the calling function:
+ So when assigning \lstinline!%x1!
+a register it is undecidable whether the variable must be live in the last return instruction, i.e.
+needs to live across the call to another function which can overwrite several registers depending on calling conventions.
+Because of this, any sound approach to liveness analysis will be an approximation.
+
+ % TODO: maybe find a better exampple with a better assignment that computer can't find?
+ % TODO: maybe mention exponential brute force algorithms?
+
+A sound albeit very naive approach is to consider every variable live at every program point, s.t.
+every variable is in interference with one another, producing a fully connected interference graph.
+% FIXME: find a place for: This generally isn't a very performant approach in terms of execution
+Such a heuristic is greedy in the sense that it picks an assignment known to be safe for the least amount of preprocessing work possible.
+Although this can possibly be a very inefficient assignment at runtime, as a number of \(n\)  variables greater than \(k\) working registers causes \(n-k\) variables to be spilled to memory, which can greatly reduce performance, but  allows for translation in constant time.
+%FIXME: is this even true?
+
+\subsection{Dataflow Analysis}
+
+Another approach, which is a much more precise approximation, is a specific variant of the dataflow analysis as described in \textit{Modern Compiler Implementation in ML} (\cite{tiger}) and \textit{Compilers: Principles, Techniques, and Tools} (\cite{dragon}).
+In general, dataflow analysis is the process of finding the possible paths in which data may propagate through different branches of execution.
+While several applications of this exist (like constant propagation, reaching defnitions, available expressions etc.), one that is immediately beneficial in the case of liveness analysis is one that traverses a CFG in the reverse order of execution (i.e.
+\textit{backwards} flow), and tagging any instructions in which a variable \textit{may} be used: also referred to as \textit{backwards may} analysis.
+This algorithm calculates which program points each variable may be accessed from with some conservative constraints known to maintain correctness.
+Specifically, these are the transfer and control-flow constraints.
+%, the former of which must satisfy a transfer function a
+
+The transfer constraint is based on a \textit{transfer function} that describe how liveness is affected across instructions.
+For each instruction, there is a transfer function that describe how liveness changes from one point to the one immediately after.
+For example, as an arithmetic operation needs to be assigned a new temporary variable, the liveness of a new variable is propagated to all  instructions  executed subsequently.
+This is done by applying the transfer function to the current \textit{live-out} set to stop further propagation of variables defined by the currently visited instruction.
+% TODO: In fact, because the IR is in SSA form, the only time
+\begin{align}
+  \mathit{in}\left[n\right] &= \mathit{use}\left[n\right] \cup (\mathit{out}\left[n\right] - \mathit{def}\left[n\right])
+\end{align}
+Control-flow constraints on the other hand propagate the use of variables to previously executed instructions, expecting these to be defined somewhere further up the CFG.
+This is done by propagating the union of the \textit{live-in} set associated with all immediate successor nodes.
+This is also referred to as the \textit{meet operator}, whose operator depends on the type of dataflow alanysis as well, but for liveness analysis a union is performed on the previos \textit{live-in} variables.
+\begin{align}
+  \mathit{out}\left[n\right] &= \bigcup_{s\in \mathit{succ}\left[n\right]} \mathit{in}\left[s\right]
+\end{align}
+%in order to maintain correctness by iteratively applying a set of constraints until a fixed point is reached.
+%These constraints are applied to the \(\textit{live-in}\left[i\right]\) and  \(\textit{live-out}\left[i\right]\) sets for every instruction \(i\). Both are initialized to the empty set \(\emptyset\).
+%The dataflow analysis approach builds the set of variables that 
+%Each type of instruction has some semantic attributes that are applied to the system state upon execution.
+This change is also referres to as a \textit{transfer function} as it transitions from one state to another.
+%can be thought of as a \textit{transfer function} in terms of liveness, denoting which variables may be live before and after execution.
+Initially, two sets are associated with each instruction: the \textit{live-in} and \textit{live-out} sets, which is the set of variables live before and after execution respectively.
+Then the following two equations are applied iteratively until a fixed point is reached, i.e.
+a point in which neither \(\mathit{in}[n]\) or \(\mathit{out}[n]\) is changed for all \(n\) instructions.
+
+%builds sets of \textit{live-in} and  \textit{live-out} variable for each instruction.
+
+% must be assigned in such a way that ensures its availability at every subsequent use throughout execution.
+
+%Variables can be assigned the same register if they are not in use at the same time, so variables that are in use at the same time are said to be in interference with one another.
+
+%Liveness analysis is the process of finding the program points at which a variable is live.
 This serves the function of determining which variables are live simultaneously, i.e.
-in interference/conflict with one another, which cannot be assigned the same physical register without overwriting the value of one another in execution.
-These can be derived recursively by iteratively applying so-called \textit{dataflow equations} to each node in a \textit{control-flow graph} (CFG) until a stable state/fixed point is reached.
+in conflict/interference with one another, so cannot be assigned the same physical register without overwriting the value of one another.
 
-The algorithm described in \textit{Modern Compiler Implementation in ML} \cite{tiger} and \cite{dragon} is based on the following equations for the live-in and live-out variables respectively:
-\begin{align}
-  in\left[n\right] &= use\left[n\right] \cup (out\left[n\right] - \mathit{def}\left[n\right])\\
-  out\left[n\right] &= \bigcup_{s\in \mathit{succ}\left[n\right]} in\left[s\right]
-\end{align}
-where \(use\left[n\right]\) is the set of all dependent variables, \(def\left[n\right]\) is the set of all variables defines and \(succ\left[n\right]\) is the set of all immediate successor nodes of \(n\).
+%Another approach that takes CFG into consideration is that of
 
-The structure of the CFG is inherently vague as both   \cite{tiger} and \textit{Compilers : Principles, Techniques, and Tools} \cite{dragon}, while based on the same underlying % TODO: fix references
-conceptions of dataflow, have different approaches to nodes of a CFG, with the former taking each individual instruction into consideration and the latter each basic block.
+%These can be derived recursively by iteratively applying so-called \textit{dataflow equations} to each node in a \textit{control-flow graph} (CFG) until a stable state/fixed point is reached.
+
+%The algorithm described in \textit{Modern Compiler Implementation in ML} \cite{tiger} and \cite{dragon} is based on the following equations for the live-in and live-out variables respectively:
+%where \(use\left[n\right]\) is the set of all dependent variables, \(def\left[n\right]\) is the set of all variables defines and \(succ\left[n\right]\) is the set of all immediate successor nodes of \(n\).
+
+%The structure of the CFG is inherently vague as both   \cite{tiger} and \textit{Compilers : Principles, Techniques, and Tools} \cite{dragon}, while based on the same underlying % TODO: fix references
+%conceptions of dataflow, have different approaches to nodes of a CFG, with the former taking each individual instruction into consideration and the latter each basic block.
 Either of these approaches are sound
@@
 
-The purpose of liveness analysis is to determine which variables are live at which program point, which in turn is used to
-
-\subsection{Dataflow Analysis}
-
-Dataflow analysis
+%The purpose of liveness analysis is to determine which variables are live at which program point, which in turn is used to
 

### 2023-12-13 article.tex
Jane approved

@@
 
-Translating IR with an unbounded number of variables to a CPU with a bounded number of registers involves the process of assigning each variable a register s.t.
-no value that can be used in the future is overwritten.
+Translating IR with an unbounded number of variables to a CPU with a bounded number of registers involves the process of assigning each variable a register such that no value that can be used in the future is overwritten.
 %Therefore, any variable that may be used in the future is considered \textit{live}, and said to be in interference with any other variable live at any intersecting point.
@@
 variables that are live at the same time (i.e. at intersecting program points) cannot, in which case they are also said to be in interference with one another.
-Although variables that are not in interference can be assigned the same register, and finding the precise points at which any variable is live is trivial for linear sequences of instructions, when conditional branching is introduced, deriving the path of execution becomes undecidable because of the halting problem.
+ Variables that are not in interference can be assigned the same register, and finding the precise points at which any variable is live is trivial for linear sequences of instructions.
+However, when conditional branching is introduced, deriving the path of execution becomes undecidable because of the halting problem.
+% TODO: finish this sentence
 
@@
 
-A sound albeit very naive approach is to consider every variable live at every program point, s.t.
-every variable is in interference with one another, producing a fully connected interference graph.
+A sound albeit very naive approach is to consider every variable live at every program point, such that every variable is in interference with one another, producing a fully connected interference graph.
 % FIXME: find a place for: This generally isn't a very performant approach in terms of execution
 Such a heuristic is greedy in the sense that it picks an assignment known to be safe for the least amount of preprocessing work possible.
-Although this can possibly be a very inefficient assignment at runtime, as a number of \(n\)  variables greater than \(k\) working registers causes \(n-k\) variables to be spilled to memory, which can greatly reduce performance, but  allows for translation in constant time.
-%FIXME: is this even true?
+However, this can be a very inefficient assignment at runtime: a number of \(n\)  variables greater than \(k\) working registers causes \(n-k\) variables to be spilled to memory, % TODO: skal der stå stack?
+which can greatly reduce performance, but  allows for translation in constant time. %FIXME: is this even true?
 
+\newpage
 \subsection{Dataflow Analysis}
@@
 While several applications of this exist (like constant propagation, reaching defnitions, available expressions etc.), one that is immediately beneficial in the case of liveness analysis is one that traverses a CFG in the reverse order of execution (i.e.
-\textit{backwards} flow), and tagging any instructions in which a variable \textit{may} be used: also referred to as \textit{backwards may} analysis.
+\textit{backwards} flow), and extracts any variable that \textit{may} be used in execution (also referred to as \textit{backwards may} analysis).
+
 This algorithm calculates which program points each variable may be accessed from with some conservative constraints known to maintain correctness.
@@
 
-The transfer constraint is based on a \textit{transfer function} that describe how liveness is affected across instructions.
-For each instruction, there is a transfer function that describe how liveness changes from one point to the one immediately after.
+The transfer constraint is based on a \textit{transfer function} that describes how liveness is affected across instructions.
+For each instruction, there is a transfer function that describes how liveness changes from one point to the one immediately after.
 For example, as an arithmetic operation needs to be assigned a new temporary variable, the liveness of a new variable is propagated to all  instructions  executed subsequently.
@@
 %can be thought of as a \textit{transfer function} in terms of liveness, denoting which variables may be live before and after execution.
-Initially, two sets are associated with each instruction: the \textit{live-in} and \textit{live-out} sets, which is the set of variables live before and after execution respectively.
+Initially, two sets are associated with each instruction: the \textit{live-in} and \textit{live-out} sets, which are the sets of variables that are live respectively before and after execution.
 Then the following two equations are applied iteratively until a fixed point is reached, i.e.

### 2023-12-17 article.tex
Start adding code

@@
 \usepackage[style=apa]{biblatex}
+
+\bibliography{article}
 
@@
 \usepackage{subcaption}
+\usepackage[margin=1in]{geometry}
 \usepackage{amsmath}
@@
 \usepackage{xcolor}
+\usepackage{minted}
+
+\definecolor{LightGray}{gray}{0.97}
 
@@
 % TODO: In fact, because the IR is in SSA form, the only time
-\begin{align}
+\begin{align}\label{flowin}
   \mathit{in}\left[n\right] &= \mathit{use}\left[n\right] \cup (\mathit{out}\left[n\right] - \mathit{def}\left[n\right])
@@
 This is also referred to as the \textit{meet operator}, whose operator depends on the type of dataflow alanysis as well, but for liveness analysis a union is performed on the previos \textit{live-in} variables.
-\begin{align}
+\begin{align}\label{flowout}
   \mathit{out}\left[n\right] &= \bigcup_{s\in \mathit{succ}\left[n\right]} \mathit{in}\left[s\right]
@@
 a point in which neither \(\mathit{in}[n]\) or \(\mathit{out}[n]\) is changed for all \(n\) instructions.
+
+% TODO: onwards!
 
@@
 
-\subsection{Constructing an interference graph}
+\subsubsection{Implementation}
+
+The \texttt{dataflow} function takes a list of instructions \texttt{insns}, an array of CFG vertices \texttt{ids} as well as the CFG \texttt{g}.
+The \(i\)th element of \texttt{ids} is the CFG vertex corresponding to the \(i\)th entry of \texttt{insns}.
+The decision to use arrays instead of pairing up each instruction with both \textit{live-in} and \textit{live-out} sets was made because of the \textit{live-out} equation \eqref{flowout} that needs to find all \textit{live-out} sets of the successor blocks.
+Hence random access arrays were chosen instead of linear search.
+\begin{minted}[linenos,bgcolor=LightGray]{ocaml}
+let dataflow (insns : Cfg.insn list) (ids : Cfg.G.V.t array) (g : Cfg.G.t) =
+  let insns = List.mapi (fun i v -> (i, v)) insns |> List.rev in
+  let in_ = Array.init (List.length insns) (fun _ -> S.SS.empty) in
+  let out = Array.init (List.length insns) (fun _ -> S.SS.empty) in
+  let rec dataflow () =
+    let flowout = (* ... *)
+    let flowin = (* ... *)
+    let flow changed insn = changed || flowout insn || flowin insn in
+    if List.fold_left flow false insns then dataflow () else (in_, out)
+  in
+  dataflow ()
+\end{minted}
+The \texttt{flowin} function correponds to the \textit{live-in} equation \eqref{flowin} and is implemented as follows:
+\begin{minted}[linenos,bgcolor=LightGray]{ocaml}
+let flowin (i, insn) =
+  let newin = S.SS.union (use insn) (S.SS.diff out.(i) (def insn)) in
+  let changed = not (S.SS.equal newin in_.(i)) in
+  if changed then in_.(i) <- newin;
+  changed
+\end{minted}
+Where \lstinline!S.SS! is a \lstinline!Set.S! module built over the \lstinline!symbol! type found in \lstinline!lib/symbol.ml!:
+
+\begin{minted}[linenos,bgcolor=LightGray]{ocaml}
+type symbol = string * int
+(* ... *)
+module SS = Set.Make (struct
+  type t = symbol
+  let compare (_, n1) (_, n2) = compare n1 n2
+end)
+type set = SS.t
+\end{minted}
+
+\begin{minted}[linenos,bgcolor=LightGray]{ocaml}
+let flowout (i, _) =
+  let newout =
+    let succ = Cfg.G.succ g ids.(i) in
+    List.fold_left
+      (fun s v -> S.SS.union s in_.(Cfg.G.V.label v))
+      S.SS.empty succ
+  in
+  let changed = not (S.SS.equal newout out.(i)) in
+  if changed then out.(i) <- newout;
+  changed
+\end{minted}
+
+\subsection{Interference Graph}
+
+The purpose for conducting dataflow analysis as above is finding variables that may be assigned the same register.
+This is done by building an interference graph, which is an undirected graph, whose nodes represent variables and edges signify interference between them, i.e.
+variables \(a\) and \(b\) live at overlapping program points is represented with an edge \((a,b)\).
+
+Constructing an interference graph only depends on the \textit{live-out} set and type of instruction.
+If the instruction defines a variable, said variable is in interference with all variables in the \textit{live-out} set.
+There is one exception however: according to the Appel text, move instructions (i.e.
+phi nodes in the case of SSA form) are given special consideration.
+The purpose of phi nodes is to copy/move a certain value from a certain predecessor, so they are not necessarily in conflict for being live at the same time.
+Rather it would often benefit if they were assigned the same register to spare unnecessary moves.
+% TODO: uddyb hvorfor de ikke er i interference?
+
+Because of this, for any phi node of the form 
+\[
+  a = \Phi (b_1, ..., b_n)
+\]
+add edges to all \textit{live-out} variables not in \(B\)
+\[
+  \forall b_j \in (\mathit{out}[i] \setminus B), \mathit{add\_edge}(a, b_j)
+\]
+For any other instruction the defines a variable \(a\)
+\[
+  \forall b_j \in B, \mathit{add\_edge}(a, b_j)
+\]
+Although the Appel text notes interference with concrete registers as well as overlapping variables, this isn't considered in this implementation.
 
@@
 \section{Graph Coloring}
-The heuristic for graph coloring as introduced in Appel and Aho et al. texts is a linear time approximation of an NP-complete problem. It is based on an iterative approach.
+Once an interference graph is constructed, the actual assignments can be found using graph coloring.
+Although this has long been known to be NP-complete, the heuristic as introduced in both the Appel and Aho et al.
+texts is a linear time approximation to this problem.
+It is based on an iterative approach wherein nodes known to be colorable are removed until either an empty graph remains in which case the original graph \(G\) is \(k\)-colorable  or nodes with more than \(k\) neighbours remain, in which case a node is chosen to be spilled to the stack and removed.
+This is then repeated.
 
@@
 After building the interference graph \(G\), a node \(n\) with fewer than \(k\) neighbors is chosen.
-As \(n\) has at most \(k-1\) neighbors, it can be removed safely by ensuring it is not assigned any color assigned to its neighbors, effectively simplifying \(G\).
-
+As \(n\) has at most \(k-1\) neighbors, it can be removed safely, effectively simplifying \(G\).
+It is pushed to a stack in order to preserve the order in which they are removed so that \(G\) can be rebuilt and the corresponding registers can be assigned correctly once a \(k\)-colorable assignment is found.
 %  (i.e. \(\mathit{deg}(n) < k\))
 
-It must hold that \(G-\{n\}\) is \(k\)-colorable if \(G\) is \(k\)-colorable
+%It must hold that \(G-\{n\}\) is \(k\)-colorable if \(G\) is \(k\)-colorable
 
@@
 Coalescing is the process of eliminating moves/copies of data from one GPR to another by combining their interference graph nodes.
+This is similar to but  not the same as  the lack of interference edges between variables subject to move operations.
+This is because variables \(a\) and \(b\) may still be assigned different registers or even spilled if,  for instance, either of them are of significant degree.
+Coalescing joins nodes \(a\) and \(b\) to node \(ab\) preserving the edges of both to maintain soundness.
+
+Since all edges are preserved, the resulting node \(an\) may be of a much higher degree.
+Because to this, only strategies that produce a \(k\)-colorable graph are worth considering as worth considering as additional spills negate the purpose entirely.
 
@@
     %* moving sys-v abi arguments to assigned registers is treacherous
+
+\subsection{Asserting correctness}
+
+Validating the correctness of each instruction translation is a complicated task given how low-level it is.
+One approach is attaching a debugger and examining the state before and after the translated sequence of x86 instructions is executed.
+The result of applying a binary operation should only affect the assigned register, for instance.
 

### 2023-12-20 article.tex
Explain B
More code and some sven notes

@@
 
+% FIXME: introducer phi nodes og nævn dominators i CFA afsnittet
+
 %hence a \textit{control flow graph} is derived from these.
@@
 
+% FIXME: beskriv use, def osv. sets tidligere!
+
 Another approach, which is a much more precise approximation, is a specific variant of the dataflow analysis as described in \textit{Modern Compiler Implementation in ML} (\cite{tiger}) and \textit{Compilers: Principles, Techniques, and Tools} (\cite{dragon}).
@@
 \textit{backwards} flow), and extracts any variable that \textit{may} be used in execution (also referred to as \textit{backwards may} analysis).
+% FIXME: forklar hvorfor List.rev ikke udgør backwards flow
 
@@
 
-\subsubsection{Implementation}
+%\subsubsection{Implementation}
 
-The \texttt{dataflow} function takes a list of instructions \texttt{insns}, an array of CFG vertices \texttt{ids} as well as the CFG \texttt{g}.
-The \(i\)th element of \texttt{ids} is the CFG vertex corresponding to the \(i\)th entry of \texttt{insns}.
-The decision to use arrays instead of pairing up each instruction with both \textit{live-in} and \textit{live-out} sets was made because of the \textit{live-out} equation \eqref{flowout} that needs to find all \textit{live-out} sets of the successor blocks.
-Hence random access arrays were chosen instead of linear search.
+The simplest ones to implement are the \lstinline!def!
+and \lstinline!use!
+functions, as all of the values of interest are located immediately within the instruction itself and not hidden behind some layer of indirection:
+\begin{minted}[linenos,bgcolor=LightGray]{ocaml}
+let def (s : S.SS.t) (insn : Cfg.insn) =
+  match insn with Insn (Some dop, _) -> S.SS.add dop s | _ -> s
+
+let use (s : S.SS.t) (insn : Cfg.insn) =
+  let op o s = match o with Ll.Id i -> S.SS.add i s | _ -> s in
+  let po s o = op o s in
+  match insn with
+  | Insn (_, AllocaN (_, (_, o)))
+  | Insn (_, Bitcast (_, o, _))
+  | Insn (_, Load (_, o))
+  | Insn (_, Ptrtoint (_, o, _))
+  | Insn (_, Sext (_, o, _))
+  | Insn (_, Trunc (_, o, _))
+  | Insn (_, Zext (_, o, _)) ->
+      op o s
+  | Insn (_, Binop (_, _, l, r))
+  | Insn (_, Icmp (_, _, l, r))
+  | Insn (_, Store (_, l, r)) ->
+      op l s |> op r
+  | Insn (_, Call (_, _, args)) -> List.map snd args |> List.fold_left po s
+  | Insn (_, Gep (_, bop, ops)) -> List.fold_left po (op bop s) ops
+  | Insn (_, Select (c, (_, l), (_, r))) -> op c s |> op l |> op r
+  | Insn (_, PhiNode (_, ops)) -> List.map fst ops |> List.fold_left po s
+  | Term (Ret (_, Some o) | Cbr (o, _, _)) -> op o s
+  | _ -> s
+\end{minted}
+Where \lstinline!S.SS! is a \lstinline!Set.S! module built over the \lstinline!symbol! type found in \lstinline!lib/symbol.ml!:
+\begin{minted}[linenos,bgcolor=LightGray]{ocaml}
+type symbol = string * int
+(* ... *)
+module SS = Set.Make (struct
+  type t = symbol
+  let compare (_, n1) (_, n2) = compare n1 n2
+end)
+type set = SS.t
+\end{minted}
+With the actual fixed-point iteration performed as follows:
 \begin{minted}[linenos,bgcolor=LightGray]{ocaml}
@@
 \end{minted}
-Where \lstinline!S.SS! is a \lstinline!Set.S! module built over the \lstinline!symbol! type found in \lstinline!lib/symbol.ml!:
-
-\begin{minted}[linenos,bgcolor=LightGray]{ocaml}
-type symbol = string * int
-(* ... *)
-module SS = Set.Make (struct
-  type t = symbol
-  let compare (_, n1) (_, n2) = compare n1 n2
-end)
-type set = SS.t
-\end{minted}
-
+And the \texttt{flowout} function which correponds to the \textit{live-out} equation \eqref{flowout} implemented as follows:
 \begin{minted}[linenos,bgcolor=LightGray]{ocaml}
@@
 \[
-  \forall b_j \in (\mathit{out}[i] \setminus B), \mathit{add\_edge}(a, b_j)
+  \forall b_j \in (\mathit{out}[i] \setminus B), \mathit{add\_edge}(a, b_j) \text{ where } B = \{b_1, ..., b_n\}
 \]

### 2023-12-20 article.tex
Highlight changes

@@
 
+\color{mygreen}
 The simplest ones to implement are the \lstinline!def!
@@
 It is pushed to a stack in order to preserve the order in which they are removed so that \(G\) can be rebuilt and the corresponding registers can be assigned correctly once a \(k\)-colorable assignment is found.
+
+\color{black}
 %  (i.e. \(\mathit{deg}(n) < k\))
@@
 Coalescing is the process of eliminating moves/copies of data from one GPR to another by combining their interference graph nodes.
-This is similar to but  not the same as  the lack of interference edges between variables subject to move operations.
+\color{mygreen} This is similar to but  not the same as  the lack of interference edges between variables subject to move operations.
 This is because variables \(a\) and \(b\) may still be assigned different registers or even spilled if,  for instance, either of them are of significant degree.
@@
 Because to this, only strategies that produce a \(k\)-colorable graph are worth considering as worth considering as additional spills negate the purpose entirely.
+\color{black}
 
@@
 
-As linear scan
+%TODO: As linear scan
 
@@
 
+\color{mygreen}
 Validating the correctness of each instruction translation is a complicated task given how low-level it is.
@@
 The result of applying a binary operation should only affect the assigned register, for instance.
+\color{black}
 
@@
 %% NOTE: måske noget om hvordan intel laver sin egen arkitektur?
+
+% TODO: uddyb mere om instrumentering og måske mere om cache meisses osv.
+%
+% FIXME: husk at nævn hvordan størrelsen af kompileret kode har en indflydelse på performance
 

### 2023-12-20 article.tex
Jane approved

--- 2023-12-20 article.tex
@@
 
-The simplest ones to implement are the \lstinline!def!
+The simplest equations to implement are the \lstinline!def!
 and \lstinline!use!
@@
 The purpose for conducting dataflow analysis as above is finding variables that may be assigned the same register.
-This is done by building an interference graph, which is an undirected graph, whose nodes represent variables and edges signify interference between them, i.e.
+This is done by building an interference graph, which is an undirected graph, whose nodes represent variables and edges that signify interference between them, i.e.
 variables \(a\) and \(b\) live at overlapping program points is represented with an edge \((a,b)\).
@@
 \]
-add edges to all \textit{live-out} variables not in \(B\)
+add edges to all \textit{live-out} variables not in \(B\):
 \[
@@
 \]
-For any other instruction the defines a variable \(a\)
+For any other instruction that defines a variable \(a\)
 \[
-  \forall b_j \in B, \mathit{add\_edge}(a, b_j)
+  \forall b_j \in B, \mathit{add\_edge}(a, b_j).
 \]
-Although the Appel text notes interference with concrete registers as well as overlapping variables, this isn't considered in this implementation.
+Although the Appel text (\cite{tiger}) describes interference of variables with concrete registers as well as overlapping variables, this isn't considered in this implementation.
 
@@
 texts is a linear time approximation to this problem.
-It is based on an iterative approach wherein nodes known to be colorable are removed until either an empty graph remains in which case the original graph \(G\) is \(k\)-colorable  or nodes with more than \(k\) neighbours remain, in which case a node is chosen to be spilled to the stack and removed.
+It is based on an iterative approach wherein nodes known to be colorable are removed until either an empty graph remains, in which case the original graph \(G\) is \(k\)-colorable  or nodes with more than \(k\) neighbours remain.
+In this case a node is chosen to be spilled to the stack and removed.
 This is then repeated.
@@
 As \(n\) has at most \(k-1\) neighbors, it can be removed safely, effectively simplifying \(G\).
-It is pushed to a stack in order to preserve the order in which they are removed so that \(G\) can be rebuilt and the corresponding registers can be assigned correctly once a \(k\)-colorable assignment is found.
+It is pushed to a stack to preserve the order in which they are removed so that \(G\) can be rebuilt and the corresponding registers can be assigned correctly once a \(k\)-colorable assignment is found.
+
 %  (i.e. \(\mathit{deg}(n) < k\))
@@
 Coalescing is the process of eliminating moves/copies of data from one GPR to another by combining their interference graph nodes.
-This is similar to but  not the same as  the lack of interference edges between variables subject to move operations.
+This is similar to but  not the same as  the lack of interference edges between variables that are subject to move operations.
 This is because variables \(a\) and \(b\) may still be assigned different registers or even spilled if,  for instance, either of them are of significant degree.
@@
 
-Since all edges are preserved, the resulting node \(an\) may be of a much higher degree.
-Because to this, only strategies that produce a \(k\)-colorable graph are worth considering as worth considering as additional spills negate the purpose entirely.
+Since all edges are preserved, the resulting node \(ab\) may be of a much higher degree.
+Because of this, only strategies that produce a \(k\)-colorable graph are worth considering, as additional spills negate the purpose entirely.
 
@@
 %Keywords: program points, dataflow equations, fixed point
+
+% TODO: find ud af om man skal henvise til kurser på au på speciel form
 
@@
 One approach is attaching a debugger and examining the state before and after the translated sequence of x86 instructions is executed.
-The result of applying a binary operation should only affect the assigned register, for instance.
+The result of applying a binary operation,  for instance, should only affect the assigned register.
 

### 2023-12-22 article.tex
Not sure

@@
 
+\newpage
+
 \section{Control Flow Analysis}
@@
 The control flow of a given program refers to the order in which instructions are executed.
-While the flow of most instructions is linear, in the sense that the next instruction executed is located immediately after, some transfer the flow of execution elsewhere or even terminate it.
+While the flow of most instructions is linear, in the sense that the next instruction executed is located immediately after, some transfer the flow of execution elsewhere or terminate it.
 %% FIXME: why is this important?
@@
 
-A continuous flow of instructions is referred to as a \textit{basic block}, defined as a sequence of instructions with no branches in or out except for the first instruction (referred to as a \textit{leader}, immediately following either the function entry or label) and the last (referred to as a \textit{terminator}, as it either terminates or transfers the flow of execution).
+A continuous flow of instructions is referred to as a \textit{basic block}, defined as a sequence of instructions with no branches in or out except for the first instruction (referred to as a \textit{leader}, immediately following either the function entry or label within it) and the last (referred to as a \textit{terminator}, as it either terminates or transfers the flow of execution).
 %These are often referred to as terminators and can be thought of as an entirely different categoriy of instructions.
 % TODO: måske drop 'can be thought of'
-
-Basic blocks represent a single node in a \textit{control flow graph} (CFG), which is a directed graph whose edges denote transfer of  control flow.
-The unconditional branch terminator always transfers control flow to the block labelled, meaning only one successor will follow, whereas conditional branching could transfer to either of the two, but because control flow analysis is not concerned with data, it simply adds both as successors.
+In order to represent the order in which instructions are executed, a \textit{control flow graph} (CFG) is introduced.
+It is a directed graph whose edges represent transfer of control.
+The type of node varies over the source material, with CFGs of the Appel test constructed over individual instructions (\cite{tiger}) and the Aho text over basic blocks (\cite{dragon}).
+Either way leaders may have some predecessors and terminators some successors associated with them.
+
+Terminators have one or two successors in the case of branching or none at all in the case of function exit. % These can either be associated basic blocks or individual instructions that make it up.
+Unconditional branching always transfers control flow to the block labelled, meaning only one successor follows, whereas conditional branching could transfer to either of the two, but because control flow analysis is not concerned with data, both are considered as possible successors.
 % TODO: rets
-While each block has 0-2 immediate successors, the amount of predecessors is unbounded as the amount of branches targeting a specific leader is unlimited.
-
-\begin{figure}[h]
-  \centering
-  \begin{subfigure}{0.48\textwidth}
-  \centering
-  \begin{tikzpicture}
-    \node [cbox, font=\ttfamily\tiny, align=left] (cfa) at (0,0) {define i32 @square(i32 \%0) \{\\1:\\\ \%2 = icmp sgt i32 \%0, 0\\\ br i1 \%2, label \%3, label \%9\ ; succs = \%3, \%9\\3:\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ ; preds = \%1, \%3\\\ \%4 = phi i32 [ \%0, \%1 ], [ \%7, \%3 ]\\\ \%5 = phi i32 [ 0, \%1 ], [ \%6, \%3 ]\\\ \%6 = add i32 \%5, \%0\\\ \%7 = sub i32 \%4, 1\\\ \%8 = icmp ugt i32 \%7, 0\\\ br i1 \%8, label \%3, label \%9\ ; succs = \%3, \%9\\9:\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ ; preds = \%3, \%1\\\ \%10 = phi i32 [ 0, \%1 ], [ \%6, \%3 ]\\\ ret i32 \%10\\\}};
-  \end{tikzpicture}
-  \end{subfigure}
-  \begin{subfigure}{0.48\textwidth}
-  \centering
-  \begin{tikzpicture}
-    \node [cbox, font=\ttfamily\tiny, align=left] (a) at (0,0) {1:\\\ \%2 = icmp sgt i32 \%0, 0\\\ br i1 \%2, label \%3, label \%9\ \ \ \ \ \ \ };
-    \node [cbox, font=\ttfamily\tiny, align=left, below=0.5cm of a] (b) {3:\\\ \%4 = phi i32 [ \%0, \%1 ], [ \%7, \%3 ]\\\ \%5 = phi i32 [ 0, \%1 ], [ \%6, \%3 ]\\\ \%6 = add i32 \%5, \%0\\\ \%7 = sub i32 \%4, 1\\\ \%8 = icmp ugt i32 \%7, 0\\\ br i1 \%8, label \%3, label \%9};
-    \node [cbox, font=\ttfamily\tiny, align=left, below=0.5cm of b] (c) {9:\\\ \%10 = phi i32 [ 0, \%1 ], [ \%6, \%3 ]\\\ ret i32 \%10};
-
-    \node[right= 1cm of a.south, invisible] (ac1)  {};
-    \node[below= 0.166667cm of ac1, invisible] (ac2)  {};
-    \node[right= 1.333333cm of ac2, invisible] (ac3)  {};
-    \node[right= 1cm of c.north, invisible] (ac6)  {};
-    \node[above= 0.166667cm of ac6, invisible] (ac5)  {};
-    \node[right= 1.333333cm of ac5, invisible] (ac4)  {};
-
-    \node[right= 1cm of b.north, invisible] (bb1)  {};
-    \node[above= 0.166667cm of bb1, invisible] (bb2)  {};
-    \node[right= 1.166667cm of bb2, invisible] (bb3)  {};
-    \node[right= 1cm of b.south, invisible] (bb6)  {};
-    \node[below= 0.166667cm of bb6, invisible] (bb5)  {};
-    \node[right= 1.166667cm of bb5, invisible] (bb4)  {};
-
-    %\node[right=2.5cm of ab, invisible] (abr)  {};
-    %\node[right=2.5cm of bc, invisible] (bcr)  {};
-  \draw [-latex] (a) -- (b);
-  %\draw [-latex] (a) to[out=-15,in=15] (c);
-  \draw [-latex] (b) -- (c);
-
-  \draw          (ac1.center) -- (ac2.center) -- (ac3.center) -- (ac4.center) -- (ac5.center);
-  \draw [-latex] (ac5.center) -- (ac6.center);
-
-  \draw          (bb2.center) -- (bb3.center) -- (bb4.center) -- (bb5.center) -- (bb6.center);
-  \draw [-latex] (bb2.center) -- (bb1.center);
-  %\draw [-latex] (c.north) to[bend right=1000] (b.north);
-  %\draw  (a.south) to  (ab);
-  %\draw  (ab) to  (abr);
-  %\draw  (abr) to  (bcr);
-  %\draw  (bcr) to  (bc);
-  %\draw  (ab) to  (bc);
-  %\draw [-latex]  (bc) to [out=180,in=70] (c.north);
-  \end{tikzpicture}
-  \end{subfigure}
-  \hfill
-  \caption{CFG of \lstinline!@square! of \lstinline!square.ll!} % FIXME: not entirely correct (square.ll is @main)
-\end{figure}
+
+Successors of node \(n\) are denoted
+\[
+  \mathit{succ}[n],
+\]
+and while each node also has predecessors associated, which consist of an  unbounded number of nodes from which cotrol may be transferred, it isn't particularly useful in the following analysis.
+
+%Basic blocks represent a single node in a \textit{control flow graph} (CFG), which is a directed graph whose edges denote transfer of  control flow.
+%The unconditional branch terminator always transfers control flow to the block labelled, meaning only one successor will follow, whereas conditional branching could transfer to either of the two, but because control flow analysis is not concerned with data, it simply adds both as successors.
+% TODO: rets
 
@@
 
-\subsection{Parameterized over individual instructions}
-
-\subsection{Parameterized over basic blocks}
+%\subsection{Parameterized over individual instructions}
+
+%\subsection{Parameterized over basic blocks}
 
@@
 
-Translating IR with an unbounded number of variables to a CPU with a bounded number of registers involves the process of assigning each variable a register such that no value that can be used in the future is overwritten.
+Translating IR with an unbounded number of variables to a CPU with a bounded number of registers involves the process of assigning each variable a register such that no value that may be needed in the future is overwritten.
 %Therefore, any variable that may be used in the future is considered \textit{live}, and said to be in interference with any other variable live at any intersecting point.
-Variables that are in use at a given program point are considered \textit{'live'}, and although variables can be assigned the same register, %if they are not in interference
+Variables that are in use at a given program point are considered \textit{live}, and although variables can be assigned the same register, %if they are not in interference
 variables that are live at the same time (i.e. at intersecting program points) cannot, in which case they are also said to be in interference with one another.
@@
 
-Suppose a function that calls another:
-
-\begin{lstlisting}[language=LLVM,firstnumber=1]
+For instance, suppose a function that calls another:
+
+\begin{minted}[linenos,bgcolor=LightGray]{llvm}
 define i32 @countcall(i32 %x0) {
@@
 }
-\end{lstlisting}
+\end{minted}
 Because of the halting problem, static analysis cannot determine if the call to \lstinline!@subproc!
@@
 However, this can be a very inefficient assignment at runtime: a number of \(n\)  variables greater than \(k\) working registers causes \(n-k\) variables to be spilled to memory, % TODO: skal der stå stack?
-which can greatly reduce performance, but  allows for translation in constant time. %FIXME: is this even true?
-
-\newpage
+which can greatly reduce performance, but  allows for translation in linear time. %FIXME: is this even true?
+
 \subsection{Dataflow Analysis}
@@
 Another approach, which is a much more precise approximation, is a specific variant of the dataflow analysis as described in \textit{Modern Compiler Implementation in ML} (\cite{tiger}) and \textit{Compilers: Principles, Techniques, and Tools} (\cite{dragon}).
-In general, dataflow analysis is the process of finding the possible paths in which data may propagate through different branches of execution.
+In general, dataflow analysis is the process of finding the possible paths in which data may propagate through various branches of execution.
 While several applications of this exist (like constant propagation, reaching defnitions, available expressions etc.), one that is immediately beneficial in the case of liveness analysis is one that traverses a CFG in the reverse order of execution (i.e.
@@
 %, the former of which must satisfy a transfer function a
-
 The transfer constraint is based on a \textit{transfer function} that describes how liveness is affected across instructions.
@@
 \end{align}
+with the \(\mathit{use}[n]\) set being defined as any variables that may be used and the \(\mathit{def}[n]\) as the set of variables defined in node \(n\).
+Because the chosen IR is in SSA form the \(\mathit{def}[n]\) either consists of one variable or equal to \(\emptyset\).
+The  \(\mathit{use}[n]\) set is effectively unbounded as some instructions take any \(m\) variables as parameters.
+
 Control-flow constraints on the other hand propagate the use of variables to previously executed instructions, expecting these to be defined somewhere further up the CFG.
 This is done by propagating the union of the \textit{live-in} set associated with all immediate successor nodes.
-This is also referred to as the \textit{meet operator}, whose operator depends on the type of dataflow alanysis as well, but for liveness analysis a union is performed on the previos \textit{live-in} variables.
+This is also referred to as the \textit{meet operator}, whose operator depends on the type of dataflow analysis as well, but for liveness analysis a union is performed on the previos \textit{live-in} variables.
 \begin{align}\label{flowout}
@@
 Initially, two sets are associated with each instruction: the \textit{live-in} and \textit{live-out} sets, which are the sets of variables that are live respectively before and after execution.
-Then the following two equations are applied iteratively until a fixed point is reached, i.e.
-a point in which neither \(\mathit{in}[n]\) or \(\mathit{out}[n]\) is changed for all \(n\) instructions.
+Then the two equations above are applied iteratively until a fixed point is reached, i.e.
+an invariant point where neither \(\mathit{in}[n]\) or \(\mathit{out}[n]\) is changed for all instructions \(n\).
 
@@
 %%%
-\subsection{Translating to x86}
+\subsection{Translating \lstinline!LLVM--! to x86}
 

### 2023-12-27 article.tex
some

@@
 \usepackage{amsmath}
+\usepackage[table]{xcolor}% http://ctan.org/pkg/xcolor
 \usepackage{tikz}
@@
 \usepackage{minted}
+\usepackage{nicematrix}
+\usepackage{collcell}
+\usepackage{varwidth}
 
@@
 This can largely be attributed to performance, as register machines routinely outperform \textit{stack machines} \parencite{ShiYunhe2008VmsS} that are often used in VMs.
-Because of the limited amount of GPRs available simultaneously, a crucial part of the backend stage for an optimizing compiler is assigning each variable  of the source program to a GPR in such a way that maximizes performance without sacrificing correctness.
+Although register machines generally also have a stack available, as being limited to mere bytes of storage is simply unfeasible for most large scale applications, using it compared to GPRs is several orders of magnitude slower.
+Because of this, a crucial part of the backend stage for an optimizing compiler is assigning each variable  of the source program to a GPR in such a way that maximizes performance without sacrificing correctness.
 
@@
 It is a directed graph whose edges represent transfer of control.
-The type of node varies over the source material, with CFGs of the Appel test constructed over individual instructions (\cite{tiger}) and the Aho text over basic blocks (\cite{dragon}).
-Either way leaders may have some predecessors and terminators some successors associated with them.
+The type of node varies over the source material, with CFGs of the Appel text constructed over individual instructions (\cite{tiger}) and the Aho text over basic blocks (\cite{dragon}).
+Either way, leaders have a set of predecessors and terminators a set of successors associated with them.
 
 Terminators have one or two successors in the case of branching or none at all in the case of function exit. % These can either be associated basic blocks or individual instructions that make it up.
-Unconditional branching always transfers control flow to the block labelled, meaning only one successor follows, whereas conditional branching could transfer to either of the two, but because control flow analysis is not concerned with data, both are considered as possible successors.
+Unconditional branching always transfers control to the block labelled, meaning only one successor follows, whereas conditional branching could transfer to either of the two, but because control flow analysis is not concerned with data, both are considered as possible successors.
 % TODO: rets
 
-Successors of node \(n\) are denoted
-\[
-  \mathit{succ}[n],
-\]
-and while each node also has predecessors associated, which consist of an  unbounded number of nodes from which cotrol may be transferred, it isn't particularly useful in the following analysis.
+Successors of node \(n\) are denoted \( \mathit{succ}[n], \)
+and while each node also has predecessors associated, which consist of an  unbounded number of nodes from which control may be transferred, it isn't particularly useful in the following analysis.
 
@@
  Variables that are not in interference can be assigned the same register, and finding the precise points at which any variable is live is trivial for linear sequences of instructions.
-However, when conditional branching is introduced, deriving the path of execution becomes undecidable because of the halting problem.
+However, when conditional branching is introduced, deriving the path of execution becomes undecidable.
 % TODO: finish this sentence
@@
 a register it is undecidable whether the variable must be live in the last return instruction, i.e.
-needs to live across the call to another function which can overwrite several registers depending on calling conventions.
+needs to live across the call to another.
 Because of this, any sound approach to liveness analysis will be an approximation.
+
+In some cases it is simply impossible to assign each variable its own register without conflict along interference edges, in which case one or more variables need to be assigned to memory instead.
+This is also referred to as \textit{spilling} the variable/register to the stack, or the variable itself is referred to as \textit{spilled}.
 
@@
 A sound albeit very naive approach is to consider every variable live at every program point, such that every variable is in interference with one another, producing a fully connected interference graph.
-% FIXME: find a place for: This generally isn't a very performant approach in terms of execution
-Such a heuristic is greedy in the sense that it picks an assignment known to be safe for the least amount of preprocessing work possible.
-However, this can be a very inefficient assignment at runtime: a number of \(n\)  variables greater than \(k\) working registers causes \(n-k\) variables to be spilled to memory, % TODO: skal der stå stack?
+This forces every variable to be assigned a different register, which is a viable approach for small programs with fewer variables than the set of assignable registers.
+%where every variable is assigned a different register. % FIXME: find a place for: This generally isn't a very performant approach in terms of execution 
+Such a heuristic is greedy in the sense that it picks an assignment known to be safe for the least amount of preprocessing possible. However, this can be a very inefficient assignment at runtime. 
+When the number of variables is greater than the number of assignable registers, the variable is spilled.
+A number of \(n\)  variables greater than \(k\) working registers causes \(n-k\) variables to be spilled to memory, % TODO: skal der stå stack?
 which can greatly reduce performance, but  allows for translation in linear time. %FIXME: is this even true?
@@
 This is done by propagating the union of the \textit{live-in} set associated with all immediate successor nodes.
-This is also referred to as the \textit{meet operator}, whose operator depends on the type of dataflow analysis as well, but for liveness analysis a union is performed on the previos \textit{live-in} variables.
+This is also referred to as the \textit{meet operator}, whose operator depends on the type of dataflow analysis as well, but for liveness analysis a union is performed on the \textit{live-in} sets of  any successive  nodes:
 \begin{align}\label{flowout}
@@
 
-\subsection{Asserting correctness}
+\subsection{Assessing Correctness}
+It is difficult to assert the correctness of a compiler.
+While writing simple unit tests is easy, guaranteeing the correctness of code generated for any given input is on a different order of magnitude in terms of complexity.
 
@@
 
-\section{Further Optimization}
+\subsubsection{Debugger Harness}
+A more elaborate way of testing correctness of translation is to assert that there are no effects outside of those intended.
+One way of achieving this is to check if a group of x86 instructions representing a single LLVM-- operation change any values except those expected.
+An approach to this is to attach a debugger, insert a breakpoint before every LLVM-- instruction, and ensure that only the intended registers are altered.
+
+To achieve this, only a few changes need to be made to the codebase.
+A debug flag is introduced, which inserts a label before every instruction by adding a \lstinline!Breakpoint!
+variant of the \lstinline!ins!
+type before the assembly emitted of that specific operation.
+In this a unique identifier as well as a list changed expected to be made is encoded.
+This includes a bitmask of GPRs that can possibly be changed (aside from the scrath registers of couse) as well as which parts of the stack may be modified. % TODO: implement the stack checker
+In addition to this, each of the labels are output with a \lstinline!.globl! directive in the file header. A file is built with these extra symbols using the \lstinline!-d! flag.
+
+To validate that this works as intended,
+
+%\section{Further Optimization}
 
@@
 
-Relevant metrics by which to compare these variations are naturally the performance of the code generated at runtime, but also the time efficiency at compile-time.
+While there are many approaches to assessing the quality of translation, one of the primary means is to simply measure the time taken to execute the code generated. 
+Assuming the translation is sound given the previous steps taken to validate correctness in translating \lstinline!LLVM--!
+to x86, it is a solid foundation for future optimizations.
+While there are certainly other factors worth measuring, only runtime performance is considered for this project.
+% FIXME: check if this is still true
+
+%There are different approaches to assessing how well a compiler backend translates IR to machine code targeting a specific architecture and platform.
+
+%Relevant metrics by which to compare these variations are naturally the performance of the code generated at runtime, but also the time efficiency at compile-time.
 While the code written isn't expected to outperform LLVM, due to the time complexity of the linear scan and simplification algorithms implemented their respective runtime performances are expected to outperform the builtin graph coloring algorithm.
@@
 
-Another factor worth considering would be memory usage, caches misses, garbage collection (although not relevant to this as it doesn't use GC), vectorizatiion (loop optimizations by SIMD) etc.
+%Another factor worth considering would be memory usage, caches misses, garbage collection (although not relevant to this as it doesn't use GC), vectorizatiion (loop optimizations by SIMD) etc.
 
@@
 
-There are different approaches to assessing how well a compiler backend translates IR to machine code targeting a specific architecture and platform.
-One of the primary means is to simply measure the time it takes to execute the code generated by it.
-
 Since the target architecture is x86-64, all benchmarks are executed natively on the fastest processor available to be used consistently at the time of writing.
@@
 
-Needless to say, CPU cycles isn't a perfect measurement either, but it works to eliminate the clock speed as well as negate much of the time the scheduler allocates for other processes.
-Scheduling still has a negative impact on execution because of the cache misses caused by context switching, but outside of executing each program in immediate mode which isn't possible on Linux or macOS this is a sensible estimate.
-To further reduce context switching, benchmarks are made on a fresh restart with minimal background processes running.
+CPU cycles isn't a perfect measurement either, but it works to eliminate the clock speed as well as negate much of the time the scheduler allocates for other processes.
+Scheduling still has a negative impact on execution because of the cache misses caused by context switching, but outside of executing each program in immediate mode (which isn't possible on Linux or macOS) this is a sensible estimate.
+To further reduce context switching, benchmarks are made on a fresh restart with minimal processes running (Xorg etc.).
 
-Additionally, all benchmarks are deterministic, so the value of each virtual variable will stay the same at every instruction step of execution when compared to another run with the same starting conditions.
+Additionally, all benchmarks are deterministic, meaning relevant values at all points across all runs are consistent.
+So while not completely equivalent due to address randomization, such noise should never leak to value space, and assuming values allocated are initialized before use, no remnants from other processes should leak to value space either.
 %all the way through the execution compared to another run of the same program with the same input.
@@
 This means that the cycles measured over infinitely many runs will converge towards a 'perfect' run with no cache misses as caused by context switching.
-So across \(n\) runs, the run least affected by context switching will be the one with the least CPU cycles, so is the one most representative of the actual performance without noise.
+Although realistically this will never happen in a lifetime, across \(n\) runs the run least affected by context switching will be the one with the least CPU cycles measured, so is the one most representative of the actual performance without noise.
 
 Measurements are made with the \lstinline!perf!
-utility, which relies on the \textit{hardware counters} (HC) registers of modern microprocessors, which are special purpose registers meant to record performance data live during execution.
-Because the analysis is integrated into the chip on which it is executing, very little overhead is incurred, and is therefore the go-to for estimating native performance.
+utility, which relies on the \textit{hardware counters} (HC) registers of modern microprocessors, which are special purpose registers meant to record performance data during execution.
+Because the analysis is integrated into the chip on which it is executing, very little overhead is incurred, and is therefore the go-to for estimating native performance on Linux.
+The \lstinline!perf!
+subcommand is used, as it starts a process and attaches from the very beginning, in addition to padding a \lstinline!-e!
+flag specifying that only cycles are meant to be recorded and the \lstinline!-x!
+flag to help parse the output.
+Both min, avg and max are tracked and reported but only min are represented in the matrices below.
+
+Each \lstinline!.ll! file present in the \lstinline!benches! directory is listed here in a table denoting the input paramater on the left column and type of allocator used.
 
@@
 % FIXME: husk at nævn hvordan størrelsen af kompileret kode har en indflydelse på performance
-
-\subsubsection{Measurements}
-
-Each \lstinline!.ll! file present in the \lstinline!benches! directory is listed here in a table denoting the input paramater on the left column and type of allocator used.
 
@@
 
+\subsubsection{\lstinline!benches/fib.ll!}
+The 'dumb' fib is a very naive approach to finding the \(n\)th number of the Fibonacci sequence by calling itself recursively twice (decrementing \(n\) before each), which causes an exponential growth in function calls.
+\begin{center}
+\begin{NiceTabular}{>{\columncolor{gray!10}}ccccccccc}
+\rowcolor{gray!10}n & clang & simple 12 & simple 2 & briggs 12 & briggs 2 & linear & greedy 12 & greedy 0 \\
+20 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.806640625,0.9990234375,0.75} 1.23& \cellcolor[rgb]{0.806640625,0.9990234375,0.75} 1.23& \cellcolor[rgb]{0.806640625,0.9990234375,0.75} 1.23& \cellcolor[rgb]{0.8056640625,0.9990234375,0.75} 1.23& \cellcolor[rgb]{0.8056640625,0.9990234375,0.75} 1.22& \cellcolor[rgb]{0.8056640625,0.9990234375,0.75} 1.22& \cellcolor[rgb]{0.9990234375,0.9912109375,0.75} 2.23\\
+21 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.8173828125,0.9990234375,0.75} 1.27& \cellcolor[rgb]{0.81640625,0.9990234375,0.75} 1.27& \cellcolor[rgb]{0.8173828125,0.9990234375,0.75} 1.27& \cellcolor[rgb]{0.8212890625,0.9990234375,0.75} 1.29& \cellcolor[rgb]{0.81640625,0.9990234375,0.75} 1.27& \cellcolor[rgb]{0.81640625,0.9990234375,0.75} 1.27& \cellcolor[rgb]{0.9990234375,0.9853515625,0.75} 2.44\\
+22 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.82421875,0.9990234375,0.75} 1.30& \cellcolor[rgb]{0.8232421875,0.9990234375,0.75} 1.29& \cellcolor[rgb]{0.8232421875,0.9990234375,0.75} 1.30& \cellcolor[rgb]{0.8212890625,0.9990234375,0.75} 1.29& \cellcolor[rgb]{0.8232421875,0.9990234375,0.75} 1.30& \cellcolor[rgb]{0.8232421875,0.9990234375,0.75} 1.29& \cellcolor[rgb]{0.9990234375,0.9794921875,0.75} 2.63\\
+23 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.8291015625,0.9990234375,0.75} 1.32& \cellcolor[rgb]{0.8291015625,0.9990234375,0.75} 1.32& \cellcolor[rgb]{0.8291015625,0.9990234375,0.75} 1.32& \cellcolor[rgb]{0.830078125,0.9990234375,0.75} 1.32& \cellcolor[rgb]{0.8310546875,0.9990234375,0.75} 1.33& \cellcolor[rgb]{0.8291015625,0.9990234375,0.75} 1.32& \cellcolor[rgb]{0.9990234375,0.9755859375,0.75} 2.75\\
+24 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.833984375,0.9990234375,0.75} 1.34& \cellcolor[rgb]{0.833984375,0.9990234375,0.75} 1.34& \cellcolor[rgb]{0.8349609375,0.9990234375,0.75} 1.34& \cellcolor[rgb]{0.8330078125,0.9990234375,0.75} 1.33& \cellcolor[rgb]{0.8330078125,0.9990234375,0.75} 1.34& \cellcolor[rgb]{0.8486328125,0.9990234375,0.75} 1.40& \cellcolor[rgb]{0.9990234375,0.97265625,0.75} 2.83\\
+25 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.8388671875,0.9990234375,0.75} 1.36& \cellcolor[rgb]{0.837890625,0.9990234375,0.75} 1.35& \cellcolor[rgb]{0.8369140625,0.9990234375,0.75} 1.35& \cellcolor[rgb]{0.8369140625,0.9990234375,0.75} 1.35& \cellcolor[rgb]{0.8359375,0.9990234375,0.75} 1.35& \cellcolor[rgb]{0.8359375,0.9990234375,0.75} 1.35& \cellcolor[rgb]{0.9990234375,0.970703125,0.75} 2.89\\
+26 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.837890625,0.9990234375,0.75} 1.36& \cellcolor[rgb]{0.8359375,0.9990234375,0.75} 1.35& \cellcolor[rgb]{0.8369140625,0.9990234375,0.75} 1.35& \cellcolor[rgb]{0.8369140625,0.9990234375,0.75} 1.35& \cellcolor[rgb]{0.837890625,0.9990234375,0.75} 1.36& \cellcolor[rgb]{0.8486328125,0.9990234375,0.75} 1.40& \cellcolor[rgb]{0.9990234375,0.9697265625,0.75} 2.92\\
+27 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.8388671875,0.9990234375,0.75} 1.36& \cellcolor[rgb]{0.837890625,0.9990234375,0.75} 1.35& \cellcolor[rgb]{0.83984375,0.9990234375,0.75} 1.36& \cellcolor[rgb]{0.8388671875,0.9990234375,0.75} 1.36& \cellcolor[rgb]{0.837890625,0.9990234375,0.75} 1.36& \cellcolor[rgb]{0.83984375,0.9990234375,0.75} 1.36& \cellcolor[rgb]{0.9990234375,0.96875,0.75} 2.96\\
+28 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.841796875,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.83984375,0.9990234375,0.75} 1.36& \cellcolor[rgb]{0.845703125,0.9990234375,0.75} 1.38& \cellcolor[rgb]{0.8427734375,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.8408203125,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.83984375,0.9990234375,0.75} 1.36& \cellcolor[rgb]{0.9990234375,0.96875,0.75} 2.97\\
+29 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.841796875,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.8408203125,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.841796875,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.83984375,0.9990234375,0.75} 1.36& \cellcolor[rgb]{0.8408203125,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.841796875,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.9990234375,0.9677734375,0.75} 2.98\\
+30 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.8349609375,0.9990234375,0.75} 1.35& \cellcolor[rgb]{0.8349609375,0.9990234375,0.75} 1.34& \cellcolor[rgb]{0.8359375,0.9990234375,0.75} 1.35& \cellcolor[rgb]{0.8330078125,0.9990234375,0.75} 1.34& \cellcolor[rgb]{0.8330078125,0.9990234375,0.75} 1.33& \cellcolor[rgb]{0.8349609375,0.9990234375,0.75} 1.34& \cellcolor[rgb]{0.9990234375,0.9697265625,0.75} 2.91\\
+31 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.83984375,0.9990234375,0.75} 1.36& \cellcolor[rgb]{0.8408203125,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.841796875,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.83984375,0.9990234375,0.75} 1.36& \cellcolor[rgb]{0.8408203125,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.83984375,0.9990234375,0.75} 1.36& \cellcolor[rgb]{0.9990234375,0.9677734375,0.75} 2.98\\
+32 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.8408203125,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.8408203125,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.8427734375,0.9990234375,0.75} 1.38& \cellcolor[rgb]{0.8408203125,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.841796875,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.8427734375,0.9990234375,0.75} 1.38& \cellcolor[rgb]{0.9990234375,0.9677734375,0.75} 2.99\\
+33 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.8427734375,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.841796875,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.8427734375,0.9990234375,0.75} 1.38& \cellcolor[rgb]{0.841796875,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.8408203125,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.841796875,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.9990234375,0.9677734375,0.75} 2.99\\
+34 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.841796875,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.841796875,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.841796875,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.841796875,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.8427734375,0.9990234375,0.75} 1.38& \cellcolor[rgb]{0.8408203125,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.9990234375,0.9677734375,0.75} 2.99\\
+35 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.841796875,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.841796875,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.8408203125,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.8427734375,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.841796875,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.8408203125,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.9990234375,0.9677734375,0.75} 2.99\\
+36 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.841796875,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.841796875,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.8427734375,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.8408203125,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.841796875,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.841796875,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.9990234375,0.9677734375,0.75} 2.99\\
+\end{NiceTabular}
+\end{center}
+The above measures are significantly equivalent to one another, with all allocators (except greedy 0) seeming to converge towards being 37\% slower than Clang.
+This isn't particularly surprising, as the x86 translation is more concerned with preserving callee saved registers than necessary in pushing each and every one despite not being in use.
+Also, instead of pushing with the \lstinline!push!
+instruction it would be faster to simply subtract from the \lstinline!%rsp!
+register directly and using \lstinline!mov!
+directly.
+
+What's more interesting and related to the allocation is the fact that greedy 0 is more than twice as slow.
+This is presumably because the \lstinline!fib!
+function that has 7 variables can assign each of them to a register for the first six allocators listed.
+However, the last one (greedy 0), has no assignable registers, so is forced to start spilling from the very beginning.
+To illusidate this hypothesis, here is the same benchmark but only for the greedy of decreasing \(k\) assignable registers:
+
+\begin{center}
+\begin{NiceTabular}{>{\columncolor{gray!10}}ccccccccc}
+\rowcolor{gray!10}n & clang & greedy 12 & greedy 8 & greedy 6 & greedy 4 & greedy 2 & greedy 1 & greedy 0 \\
+20 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.8115234375,0.9990234375,0.75} 1.25& \cellcolor[rgb]{0.806640625,0.9990234375,0.75} 1.23& \cellcolor[rgb]{0.81640625,0.9990234375,0.75} 1.27& \cellcolor[rgb]{0.837890625,0.9990234375,0.75} 1.36& \cellcolor[rgb]{0.8505859375,0.9990234375,0.75} 1.41& \cellcolor[rgb]{0.9990234375,0.9921875,0.75} 2.20& \cellcolor[rgb]{0.9990234375,0.9912109375,0.75} 2.24\\
+21 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.8173828125,0.9990234375,0.75} 1.27& \cellcolor[rgb]{0.81640625,0.9990234375,0.75} 1.27& \cellcolor[rgb]{0.830078125,0.9990234375,0.75} 1.32& \cellcolor[rgb]{0.853515625,0.9990234375,0.75} 1.42& \cellcolor[rgb]{0.8642578125,0.9990234375,0.75} 1.46& \cellcolor[rgb]{0.9990234375,0.9853515625,0.75} 2.42& \cellcolor[rgb]{0.9990234375,0.984375,0.75} 2.46\\
+22 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.822265625,0.9990234375,0.75} 1.29& \cellcolor[rgb]{0.82421875,0.9990234375,0.75} 1.30& \cellcolor[rgb]{0.837890625,0.9990234375,0.75} 1.35& \cellcolor[rgb]{0.86328125,0.9990234375,0.75} 1.46& \cellcolor[rgb]{0.8759765625,0.9990234375,0.75} 1.51& \cellcolor[rgb]{0.9990234375,0.98046875,0.75} 2.57& \cellcolor[rgb]{0.9990234375,0.9794921875,0.75} 2.62\\
+23 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.8291015625,0.9990234375,0.75} 1.32& \cellcolor[rgb]{0.8291015625,0.9990234375,0.75} 1.32& \cellcolor[rgb]{0.84375,0.9990234375,0.75} 1.38& \cellcolor[rgb]{0.8740234375,0.9990234375,0.75} 1.50& \cellcolor[rgb]{0.88671875,0.9990234375,0.75} 1.55& \cellcolor[rgb]{0.9990234375,0.9775390625,0.75} 2.69& \cellcolor[rgb]{0.9990234375,0.9755859375,0.75} 2.74\\
+24 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.8359375,0.9990234375,0.75} 1.35& \cellcolor[rgb]{0.8349609375,0.9990234375,0.75} 1.34& \cellcolor[rgb]{0.8642578125,0.9990234375,0.75} 1.46& \cellcolor[rgb]{0.8837890625,0.9990234375,0.75} 1.54& \cellcolor[rgb]{0.8935546875,0.9990234375,0.75} 1.58& \cellcolor[rgb]{0.9990234375,0.974609375,0.75} 2.78& \cellcolor[rgb]{0.9990234375,0.97265625,0.75} 2.83\\
+25 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.8359375,0.9990234375,0.75} 1.35& \cellcolor[rgb]{0.837890625,0.9990234375,0.75} 1.35& \cellcolor[rgb]{0.8525390625,0.9990234375,0.75} 1.41& \cellcolor[rgb]{0.8837890625,0.9990234375,0.75} 1.54& \cellcolor[rgb]{0.896484375,0.9990234375,0.75} 1.59& \cellcolor[rgb]{0.9990234375,0.97265625,0.75} 2.83& \cellcolor[rgb]{0.9990234375,0.970703125,0.75} 2.89\\
+26 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.837890625,0.9990234375,0.75} 1.36& \cellcolor[rgb]{0.8388671875,0.9990234375,0.75} 1.36& \cellcolor[rgb]{0.853515625,0.9990234375,0.75} 1.42& \cellcolor[rgb]{0.88671875,0.9990234375,0.75} 1.55& \cellcolor[rgb]{0.8994140625,0.9990234375,0.75} 1.60& \cellcolor[rgb]{0.9990234375,0.9716796875,0.75} 2.87& \cellcolor[rgb]{0.9990234375,0.9697265625,0.75} 2.93\\
+27 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.8388671875,0.9990234375,0.75} 1.36& \cellcolor[rgb]{0.8388671875,0.9990234375,0.75} 1.36& \cellcolor[rgb]{0.857421875,0.9990234375,0.75} 1.43& \cellcolor[rgb]{0.8876953125,0.9990234375,0.75} 1.56& \cellcolor[rgb]{0.90234375,0.9990234375,0.75} 1.61& \cellcolor[rgb]{0.9990234375,0.970703125,0.75} 2.89& \cellcolor[rgb]{0.9990234375,0.96875,0.75} 2.95\\
+28 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.83984375,0.9990234375,0.75} 1.36& \cellcolor[rgb]{0.8408203125,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.8564453125,0.9990234375,0.75} 1.43& \cellcolor[rgb]{0.8916015625,0.9990234375,0.75} 1.57& \cellcolor[rgb]{0.904296875,0.9990234375,0.75} 1.62& \cellcolor[rgb]{0.9990234375,0.9697265625,0.75} 2.92& \cellcolor[rgb]{0.9990234375,0.96875,0.75} 2.97\\
+29 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.83984375,0.9990234375,0.75} 1.36& \cellcolor[rgb]{0.8408203125,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.8583984375,0.9990234375,0.75} 1.44& \cellcolor[rgb]{0.890625,0.9990234375,0.75} 1.57& \cellcolor[rgb]{0.9052734375,0.9990234375,0.75} 1.62& \cellcolor[rgb]{0.9990234375,0.9697265625,0.75} 2.92& \cellcolor[rgb]{0.9990234375,0.9677734375,0.75} 2.98\\
+30 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.83984375,0.9990234375,0.75} 1.36& \cellcolor[rgb]{0.8408203125,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.8564453125,0.9990234375,0.75} 1.43& \cellcolor[rgb]{0.890625,0.9990234375,0.75} 1.57& \cellcolor[rgb]{0.90625,0.9990234375,0.75} 1.63& \cellcolor[rgb]{0.9990234375,0.9697265625,0.75} 2.93& \cellcolor[rgb]{0.9990234375,0.9677734375,0.75} 2.98\\
+31 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.8408203125,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.8388671875,0.9990234375,0.75} 1.36& \cellcolor[rgb]{0.857421875,0.9990234375,0.75} 1.43& \cellcolor[rgb]{0.8916015625,0.9990234375,0.75} 1.57& \cellcolor[rgb]{0.9052734375,0.9990234375,0.75} 1.62& \cellcolor[rgb]{0.9990234375,0.9697265625,0.75} 2.93& \cellcolor[rgb]{0.9990234375,0.9677734375,0.75} 2.99\\
+32 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.8408203125,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.83984375,0.9990234375,0.75} 1.36& \cellcolor[rgb]{0.8564453125,0.9990234375,0.75} 1.43& \cellcolor[rgb]{0.890625,0.9990234375,0.75} 1.57& \cellcolor[rgb]{0.90625,0.9990234375,0.75} 1.63& \cellcolor[rgb]{0.9990234375,0.9697265625,0.75} 2.92& \cellcolor[rgb]{0.9990234375,0.9677734375,0.75} 2.98\\
+33 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.8408203125,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.83984375,0.9990234375,0.75} 1.36& \cellcolor[rgb]{0.857421875,0.9990234375,0.75} 1.43& \cellcolor[rgb]{0.8916015625,0.9990234375,0.75} 1.57& \cellcolor[rgb]{0.90625,0.9990234375,0.75} 1.63& \cellcolor[rgb]{0.9990234375,0.9697265625,0.75} 2.94& \cellcolor[rgb]{0.9990234375,0.9677734375,0.75} 2.99\\
+34 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.841796875,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.8408203125,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.857421875,0.9990234375,0.75} 1.43& \cellcolor[rgb]{0.8916015625,0.9990234375,0.75} 1.57& \cellcolor[rgb]{0.9052734375,0.9990234375,0.75} 1.63& \cellcolor[rgb]{0.9990234375,0.9697265625,0.75} 2.94& \cellcolor[rgb]{0.9990234375,0.9677734375,0.75} 2.99\\
+35 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.8408203125,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.8408203125,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.857421875,0.9990234375,0.75} 1.43& \cellcolor[rgb]{0.8916015625,0.9990234375,0.75} 1.57& \cellcolor[rgb]{0.90625,0.9990234375,0.75} 1.63& \cellcolor[rgb]{0.9990234375,0.9697265625,0.75} 2.93& \cellcolor[rgb]{0.9990234375,0.9677734375,0.75} 2.99\\
+36 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.841796875,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.83984375,0.9990234375,0.75} 1.36& \cellcolor[rgb]{0.857421875,0.9990234375,0.75} 1.43& \cellcolor[rgb]{0.8916015625,0.9990234375,0.75} 1.57& \cellcolor[rgb]{0.90625,0.9990234375,0.75} 1.63& \cellcolor[rgb]{0.9990234375,0.9697265625,0.75} 2.94& \cellcolor[rgb]{0.9990234375,0.9677734375,0.75} 2.99\\
+\end{NiceTabular}
+\end{center}
+
+\subsubsection{\lstinline!benches/ackermann.ll!}
+
+Another benchmark that derives highly from the recursive structure of \lstinline!fib.ll! above is the Ackermann function.
+
+%TODO: bedøm om det giver mening
+
+%\begin{center}
+%\begin{tabular}{|c||c|c|c|c|c|c|c|c|c|c|c|} 
+ %\hline
+%& clang & simple 12 & simple 2 & briggs 12 & briggs 2 & linear & greedy 12 & greedy 0 \\
+ %\hline
+%34 & \cellcolor{1.0, 1.0 , 1.0}1.00& 1.37& 1.37& 1.37& 1.37& 1.36& 1.41& 2.99\\
+%35 & 1.00& 1.39& 1.37& 1.41& 1.37& 1.37& 1.38& 3.00\\
+%36 & 1.00& 1.37& 1.37& 1.37& 1.37& 1.37& 1.38& 2.99\\
+%37 & 1.00& 1.38& 1.37& 1.37& 1.37& 1.39& 1.37& 2.99\\
+%38 & 1.00& 1.37& 1.37& 1.38& 1.37& 1.37& 1.37& 2.99\\
+%39 & 1.00& 1.37& 1.38& 1.37& 1.39& 1.37& 1.37& 2.99\\
+%40 & 1.00& 1.37& 1.37& 1.39& 1.37& 1.37& 1.37& 2.99\\
+%41 & 1.00& 1.37& 1.38& 1.37& 1.37& 1.37& 1.37& 2.99\\
+%42 & 1.00& 1.38& 1.38& 1.37& 1.37& 1.37& 1.37& 3.04\\
+ %\hline
+%\end{tabular}
+%\end{center}
+
+%\begin{center}
+%\begin{tabular}{|c||c|c|c|c|c|c|c|c|c|c|c|} 
+ %\hline
+%&clang& greedy 12 & greedy 8 & greedy 4 & greedy 2 & greedy 1 & greedy 0 \\
+ %\hline
+%34 & 1.00& 1.41& 1.37& 1.57& 1.63& 2.93& 3.04\\
+%35 & 1.00& 1.37& 1.38& 1.58& 1.66& 2.95& 3.01\\
+%36 & 1.00& 1.36& 1.40& 1.56& 1.66& 2.93& 2.98\\
+%37 & 1.00& 1.38& 1.38& 1.58& 1.64& 2.95& 3.01\\
+%38 & 1.00& 1.37& 1.37& 1.57& 1.64& 2.93& 3.04\\
+%39 & 1.00& 1.36& 1.37& 1.57& 1.63& 2.94& 3.04\\
+%40 & 1.00& 1.37& 1.37& 1.57& 1.63& 2.93& 3.04\\
+%41 & 1.00& 1.38& 1.41& 1.58& 1.67& 2.97& 3.03\\
+%42 & 1.00& 1.36& 1.37& 1.56& 1.62& 2.92& 2.98\\
+% \hline
+%\end{tabular}
+%\end{center}
+
+\subsubsection{\lstinline!benches/factori32.ll!}
+Prime factorization of \lstinline!i32!.
+\begin{center}
+\begin{tabular}{|c||c|c|c|c|} 
+ \hline
+ fib & clang & greedy & simple & linear \\ [0.5ex] 
+ \hline
+ 40 & 1.000000 & 1.438693 & 1.441270 & 0.73 \\ [1ex] 
+ 41 & 1.000000 & 1.441035 & 1.440626 & 0.73 \\ [1ex] 
+ 42 & 	1.000000 & 1.447140 & 1.446094 & 0\\
+ \hline
+\end{tabular}
+\end{center}
+
 %TODO:
@@
 %\end{center}
-
-\paragraph{fib.ll}
-The 'dumb' fib is a very naive approach to finding the \(n\)th number of the Fibonacci sequence by calling itself recursively twice (decrementing \(n\) before each), which causes an exponential growth in function calls.
-\begin{center}
-\begin{tabular}{|c||c|c|c|c|} 
- \hline
- fib & clang & greedy & simple & linear \\ [0.5ex] 
- \hline
- 40 & 1.000000 & 1.438693 & 1.441270 & 0.73 \\ [1ex] 
- 41 & 1.000000 & 1.441035 & 1.440626 & 0.73 \\ [1ex] 
- 42 & 	1.000000 & 1.447140 & 1.446094 & 0\\
- \hline
-\end{tabular}
-\end{center}
 
@@
 \end{enumerate}
-
-\section{Benchmarks}
-
-\subsection{\lstinline!benches/fib.ll!}
-\begin{lstlisting}[language=LLVM]
-declare i32 @atoi(i8*)
-define i32 @fib(i32 %n0) {
-  %cn =  icmp sle i32 %n0, 2
-  br i1 %cn, label %base, label %rec
-base:
-  ret i32 1
-rec:
-  %n1 = sub i32 %n0, 1
-  %v0 = call i32 @fib(i32 %n1)
-  %n2 = sub i32 %n0, 2
-  %v1 = call i32 @fib(i32 %n2)
-  %v2 = add i32 %v1, %v2
-  ret i32 %v2
-}
-define i32 @main(i32 %argc, i8** %argv) {
-  %arg1ptr = getelementptr i8*, i8** %argv, i64 1
-  %arg1 = load i8*, i8** %arg1ptr
-  %n = call i32 @atoi(i8* %arg1)
-  call i32 @fib(i32 %n)
-  ret i32 0
-}
-
-\end{lstlisting}
-
-\subsubsection{\lstinline!make fib-clang!}
-\begin{lstlisting}[language=none]
-$ make fib-clang
-clang -O0 -target x86_64-unknown-darwin benches/fib.ll -o fib-clang
-\end{lstlisting}
-\paragraph{\lstinline!make bench fib-clang 42!}
-\begin{lstlisting}[language=none]
-$ make bench fib-clang 42
-/usr/bin/time -al ./fib-clang 42
-        1.58 real         1.56 user         0.00 sys
-             2678784  maximum resident set size
-                   0  average shared memory size
-                   0  average unshared data size
-                   0  average unshared stack size
-                 765  page reclaims
-                   0  page faults
-                   0  swaps
-                   0  block input operations
-                   0  block output operations
-                   0  messages sent
-                   0  messages received
-                   0  signals received
-                   0  voluntary context switches
-                  29  involuntary context switches
-         11000239251  instructions retired
-          4962802652  cycles elapsed
-             1708928  peak memory footprint
-\end{lstlisting}
-\paragraph{\lstinline!make bench fib-clang 43!}
-\begin{lstlisting}[language=none]
-make bench fib-clang 43
-$ /usr/bin/time -al ./fib-clang 43
-        2.71 real         2.49 user         0.00 sys
-             2678784  maximum resident set size
-                   0  average shared memory size
-                   0  average unshared data size
-                   0  average unshared stack size
-                 700  page reclaims
-                  66  page faults
-                   0  swaps
-                   0  block input operations
-                   0  block output operations
-                   0  messages sent
-                   0  messages received
-                   0  signals received
-                  13  voluntary context switches
-                  32  involuntary context switches
-         17797124508  instructions retired
-          8035601598  cycles elapsed
-             1708928  peak memory footprint
-\end{lstlisting}
-
-\paragraph{\lstinline!make bench fib-clang 44!}
-\begin{lstlisting}[language=none]
-$ make bench fib-clang 44
-/usr/bin/time -al ./fib-clang 44
-        4.06 real         4.04 user         0.00 sys
-             2678784  maximum resident set size
-                   0  average shared memory size
-                   0  average unshared data size
-                   0  average unshared stack size
-                 700  page reclaims
-                  66  page faults
-                   0  swaps
-                   0  block input operations
-                   0  block output operations
-                   0  messages sent
-                   0  messages received
-                   0  signals received
-                   5  voluntary context switches
-                  29  involuntary context switches
-         28781328564  instructions retired
-         12988681489  cycles elapsed
-             1708928  peak memory footprint
-\end{lstlisting}
-
-\paragraph{\lstinline!make bench fib-clang 45!}
-\begin{lstlisting}[language=none]
-$ make bench fib-clang 45
-/usr/bin/time -al ./fib-clang 45
-        6.54 real         6.51 user         0.00 sys
-             2678784  maximum resident set size
-                   0  average shared memory size
-                   0  average unshared data size
-                   0  average unshared stack size
-                 700  page reclaims
-                  66  page faults
-                   0  swaps
-                   0  block input operations
-                   0  block output operations
-                   0  messages sent
-                   0  messages received
-                   0  signals received
-                   6  voluntary context switches
-                  39  involuntary context switches
-         46555259189  instructions retired
-         20968319385  cycles elapsed
-             1708928  peak memory footprint
-\end{lstlisting}
-
-\paragraph{\lstinline!make bench fib-clang 46!}
-\begin{lstlisting}[language=none]
-$ make bench fib-clang 46
-/usr/bin/time -al ./fib-clang 46
-       10.58 real        10.54 user         0.00 sys
-             2678784  maximum resident set size
-                   0  average shared memory size
-                   0  average unshared data size
-                   0  average unshared stack size
-                 700  page reclaims
-                  66  page faults
-                   0  swaps
-                   0  block input operations
-                   0  block output operations
-                   0  messages sent
-                   0  messages received
-                   0  signals received
-                   5  voluntary context switches
-                 107  involuntary context switches
-         75314385847  instructions retired
-         33939361073  cycles elapsed
-             1725376  peak memory footprint
-\end{lstlisting}
-
-\paragraph{\lstinline!make bench fib-clang 47!}
-\begin{lstlisting}[language=none]
-$ make bench fib-clang 47
-/usr/bin/time -al ./fib-clang 47
-       17.09 real        17.06 user         0.00 sys
-             2678784  maximum resident set size
-                   0  average shared memory size
-                   0  average unshared data size
-                   0  average unshared stack size
-                 766  page reclaims
-                   0  page faults
-                   0  swaps
-                   0  block input operations
-                   0  block output operations
-                   0  messages sent
-                   0  messages received
-                   0  signals received
-                   0  voluntary context switches
-                 100  involuntary context switches
-        121838577947  instructions retired
-         54901966523  cycles elapsed
-             1708928  peak memory footprint
-\end{lstlisting}
-
-\paragraph{\lstinline!make bench fib-clang 48!}
-\begin{lstlisting}[language=none]
-$ make bench fib-clang 48
-/usr/bin/time -al ./fib-clang 48
-       27.65 real        27.59 user         0.00 sys
-             2678784  maximum resident set size
-                   0  average shared memory size
-                   0  average unshared data size
-                   0  average unshared stack size
-                 766  page reclaims
-                   0  page faults
-                   0  swaps
-                   0  block input operations
-                   0  block output operations
-                   0  messages sent
-                   0  messages received
-                   0  signals received
-                   0  voluntary context switches
-                 161  involuntary context switches
-        197130073055  instructions retired
-         88835375488  cycles elapsed
-             1708928  peak memory footprint
-\end{lstlisting}
-
 \end{document}

### 2023-12-27 article.tex
Pre breaking ref

@@
 \usepackage{varwidth}
+\usepackage{booktabs}
 
@@
 This can largely be attributed to performance, as register machines routinely outperform \textit{stack machines} \parencite{ShiYunhe2008VmsS} that are often used in VMs.
-Although register machines generally also have a stack available, as being limited to mere bytes of storage is simply unfeasible for most large scale applications, using it compared to GPRs is several orders of magnitude slower.
+Although register machines generally also have a stack available, as being limited to mere bytes of storage is simply infeasible for most large scale applications, using it compared to GPRs is several orders of magnitudes slower.
 Because of this, a crucial part of the backend stage for an optimizing compiler is assigning each variable  of the source program to a GPR in such a way that maximizes performance without sacrificing correctness.
@@
 The primary sources will be \textit{Modern Compiler Implementation in ML} \parencite{tiger} and \textit{Compilers: Principles, techniques, and tools} \parencite{dragon}, in addition to publications concerning the linear scan approach.
+
+The implementation is written in OCaml for the most part, although there is some marginal use of Python to interact with the LLDB debugger.
 
@@
 
-\newpage
+\section{Intermediate Representation}
+
+Although compilers can feasibly translate from the source language to machine code of the target platform directly, and doing so could even be more efficient, doing so hinders the portability as machine code targeting architecture \(A\) isn't necessarily useful for targeting architecture \(B\) further down the line.
+
+Instead of translating directly from a source language to target machine code, most compiler frontends emit IR which is intended as an abstraction over particular CPU implementations.
+% FIXME: er det korrekt?
+lol
+It is by no means executable on either \(A\) or \(B\) but is more often than not much closer to the abstract level of a CPU than the source language originally fed to the frontend.
+This helps portability immensely, as a frontend emitting an IR no longer needs to specialize to a particular architecture, instead it can target as many as the backend supports.
+Likewise, a compiler backend will support any frontend provided that they emit correct IR.
+
+Because of this, compilation of high-level programming languages will often go no further than emitting IR, then leave the rest for a subsequent backend toolchain.
+This naturally leads to the LLVM toolchain, which is by far the most used compiler backend.
+Most languages, if targeting native execution on a CPU, will have an LLVM implementation.
+This is true for C and C++ (clang and clang++ respectively), D (with the ldc compiler), Swift and Rust, to name a few.
+
+Conversely, managed languages such as Java and C# typically don't depend on the LLVM toolchain directly, research projects that attempt to compile associated VM bytecode to a native platform using LLVM.
+
+The modularization of the frontend and backend stages also help immensely.
+
+\begin{figure}[!ht]
+   \begin{minipage}{0.48\textwidth}
+     \centering
+     \begin{minted}[bgcolor=LightGray]{c}
+int add(int a, int b) {
+  int sum = a + b;
+  return sum;
+}
+     \end{minted}
+     \caption{Arithmetic function implemented in C}\label{fig:addc}
+   \end{minipage}\hfill
+   \begin{minipage}{0.48\textwidth}
+     \centering
+     \begin{minted}[bgcolor=LightGray]{llvm}
+define i32 @add(i32 %a, i32 %b) {
+  %sum = add i32 %a, %b
+  ret i32 %sum
+}
+     \end{minted}
+     \caption{Corresponding represented in LLVM}\label{fig:addll}
+   \end{minipage}
+   \begin{center}
+   \begin{minipage}{0.48\textwidth}
+     \centering
+     \begin{minted}[bgcolor=LightGray]{gas}
+  .globl _add
+_add:
+  movl	%edi, %eax
+  addl	%esi, %eax
+  retq
+     \end{minted}
+     \caption{Prettified result of \mintinline{bash}{clang -O0}}\label{fig:addx86}
+   \end{minipage}
+   \end{center}
+\end{figure}
 
@@
 
-\subsection{Building a graph}
+\subsection{Building a CFG}
 
@@
 
-With an input stream of instructions
+Building a graph is relatively straight forward, with the input for every function declared parsed into a tuple type \mintinline{ocaml}{cfg}  of the form:
+\begin{center}
+\begin{minted}{ocaml}
+type cfg = (lbl option * block) * (lbl * block) list
+\end{minted}
+\end{center}
+which consists of an optionally named entry block in the first part and a list of trailing blocks that are always named in the second.
 
@@
 
-Another approach, which is a much more precise approximation, is a specific variant of the dataflow analysis as described in \textit{Modern Compiler Implementation in ML} (\cite{tiger}) and \textit{Compilers: Principles, Techniques, and Tools} (\cite{dragon}).
+Another approach, which is a more precise approximation, is a specific variant of the dataflow analysis as described in \textit{Modern Compiler Implementation in ML} (\cite{tiger}) and \textit{Compilers: Principles, Techniques, and Tools} (\cite{dragon}).
 In general, dataflow analysis is the process of finding the possible paths in which data may propagate through various branches of execution.
@@
   let op o s = match o with Ll.Id i -> S.SS.add i s | _ -> s in
-  let po s o = op o s in
+  let po = Fun.flip op in
   match insn with
-  | Insn (_, AllocaN (_, (_, o)))
-  | Insn (_, Bitcast (_, o, _))
-  | Insn (_, Load (_, o))
-  | Insn (_, Ptrtoint (_, o, _))
-  | Insn (_, Sext (_, o, _))
-  | Insn (_, Trunc (_, o, _))
-  | Insn (_, Zext (_, o, _)) ->
+  | Insn (_, AllocaN (_, (_, o))) (* | Bitcast _ | ... | Zext _ *) ->
       op o s
-  | Insn (_, Binop (_, _, l, r))
-  | Insn (_, Icmp (_, _, l, r))
-  | Insn (_, Store (_, l, r)) ->
+  | Insn (_, Binop (_, _, l, r)) (* | Icmp _ | Store _ *) ->
       op l s |> op r
@@
 \[
-  \forall b_j \in (\mathit{out}[i] \setminus B), \mathit{add\_edge}(a, b_j) \text{ where } B = \{b_1, ..., b_n\}
+  \forall b_j \in \mathit{out}[i] \setminus \{b_1, ..., b_n\}, \text{add\_edge}(a, b_j)
 \]
@@
 \[
-  \forall b_j \in B, \mathit{add\_edge}(a, b_j).
+  \forall b_j \in \{b_1, ..., b_n\}, \text{add\_edge}(a, b_j).
 \]
@@
 \section{Linear Scan}
+Another approach that does not depend on liveness analysis i the linear scan algorithm.
+It is a simple and efficient approach that performs a linear pass over IR to produce a register assignment significantly faster than all of the steps involved with graph coloring.
 
+The primary text is \textit{Linear Scan Register Allocation} (\cite{linear}), the execution of which also relies on prior liveness analysis.
+
+% TODO: find ud af hvad binpacking er
+%
 %Write about Appel book.
@@
 The 'dumb' fib is a very naive approach to finding the \(n\)th number of the Fibonacci sequence by calling itself recursively twice (decrementing \(n\) before each), which causes an exponential growth in function calls.
+\begin{table}[!h]
 \begin{center}
-\begin{NiceTabular}{>{\columncolor{gray!10}}ccccccccc}
-\rowcolor{gray!10}n & clang & simple 12 & simple 2 & briggs 12 & briggs 2 & linear & greedy 12 & greedy 0 \\
-20 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.806640625,0.9990234375,0.75} 1.23& \cellcolor[rgb]{0.806640625,0.9990234375,0.75} 1.23& \cellcolor[rgb]{0.806640625,0.9990234375,0.75} 1.23& \cellcolor[rgb]{0.8056640625,0.9990234375,0.75} 1.23& \cellcolor[rgb]{0.8056640625,0.9990234375,0.75} 1.22& \cellcolor[rgb]{0.8056640625,0.9990234375,0.75} 1.22& \cellcolor[rgb]{0.9990234375,0.9912109375,0.75} 2.23\\
-21 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.8173828125,0.9990234375,0.75} 1.27& \cellcolor[rgb]{0.81640625,0.9990234375,0.75} 1.27& \cellcolor[rgb]{0.8173828125,0.9990234375,0.75} 1.27& \cellcolor[rgb]{0.8212890625,0.9990234375,0.75} 1.29& \cellcolor[rgb]{0.81640625,0.9990234375,0.75} 1.27& \cellcolor[rgb]{0.81640625,0.9990234375,0.75} 1.27& \cellcolor[rgb]{0.9990234375,0.9853515625,0.75} 2.44\\
-22 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.82421875,0.9990234375,0.75} 1.30& \cellcolor[rgb]{0.8232421875,0.9990234375,0.75} 1.29& \cellcolor[rgb]{0.8232421875,0.9990234375,0.75} 1.30& \cellcolor[rgb]{0.8212890625,0.9990234375,0.75} 1.29& \cellcolor[rgb]{0.8232421875,0.9990234375,0.75} 1.30& \cellcolor[rgb]{0.8232421875,0.9990234375,0.75} 1.29& \cellcolor[rgb]{0.9990234375,0.9794921875,0.75} 2.63\\
-23 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.8291015625,0.9990234375,0.75} 1.32& \cellcolor[rgb]{0.8291015625,0.9990234375,0.75} 1.32& \cellcolor[rgb]{0.8291015625,0.9990234375,0.75} 1.32& \cellcolor[rgb]{0.830078125,0.9990234375,0.75} 1.32& \cellcolor[rgb]{0.8310546875,0.9990234375,0.75} 1.33& \cellcolor[rgb]{0.8291015625,0.9990234375,0.75} 1.32& \cellcolor[rgb]{0.9990234375,0.9755859375,0.75} 2.75\\
-24 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.833984375,0.9990234375,0.75} 1.34& \cellcolor[rgb]{0.833984375,0.9990234375,0.75} 1.34& \cellcolor[rgb]{0.8349609375,0.9990234375,0.75} 1.34& \cellcolor[rgb]{0.8330078125,0.9990234375,0.75} 1.33& \cellcolor[rgb]{0.8330078125,0.9990234375,0.75} 1.34& \cellcolor[rgb]{0.8486328125,0.9990234375,0.75} 1.40& \cellcolor[rgb]{0.9990234375,0.97265625,0.75} 2.83\\
-25 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.8388671875,0.9990234375,0.75} 1.36& \cellcolor[rgb]{0.837890625,0.9990234375,0.75} 1.35& \cellcolor[rgb]{0.8369140625,0.9990234375,0.75} 1.35& \cellcolor[rgb]{0.8369140625,0.9990234375,0.75} 1.35& \cellcolor[rgb]{0.8359375,0.9990234375,0.75} 1.35& \cellcolor[rgb]{0.8359375,0.9990234375,0.75} 1.35& \cellcolor[rgb]{0.9990234375,0.970703125,0.75} 2.89\\
-26 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.837890625,0.9990234375,0.75} 1.36& \cellcolor[rgb]{0.8359375,0.9990234375,0.75} 1.35& \cellcolor[rgb]{0.8369140625,0.9990234375,0.75} 1.35& \cellcolor[rgb]{0.8369140625,0.9990234375,0.75} 1.35& \cellcolor[rgb]{0.837890625,0.9990234375,0.75} 1.36& \cellcolor[rgb]{0.8486328125,0.9990234375,0.75} 1.40& \cellcolor[rgb]{0.9990234375,0.9697265625,0.75} 2.92\\
-27 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.8388671875,0.9990234375,0.75} 1.36& \cellcolor[rgb]{0.837890625,0.9990234375,0.75} 1.35& \cellcolor[rgb]{0.83984375,0.9990234375,0.75} 1.36& \cellcolor[rgb]{0.8388671875,0.9990234375,0.75} 1.36& \cellcolor[rgb]{0.837890625,0.9990234375,0.75} 1.36& \cellcolor[rgb]{0.83984375,0.9990234375,0.75} 1.36& \cellcolor[rgb]{0.9990234375,0.96875,0.75} 2.96\\
-28 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.841796875,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.83984375,0.9990234375,0.75} 1.36& \cellcolor[rgb]{0.845703125,0.9990234375,0.75} 1.38& \cellcolor[rgb]{0.8427734375,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.8408203125,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.83984375,0.9990234375,0.75} 1.36& \cellcolor[rgb]{0.9990234375,0.96875,0.75} 2.97\\
-29 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.841796875,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.8408203125,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.841796875,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.83984375,0.9990234375,0.75} 1.36& \cellcolor[rgb]{0.8408203125,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.841796875,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.9990234375,0.9677734375,0.75} 2.98\\
-30 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.8349609375,0.9990234375,0.75} 1.35& \cellcolor[rgb]{0.8349609375,0.9990234375,0.75} 1.34& \cellcolor[rgb]{0.8359375,0.9990234375,0.75} 1.35& \cellcolor[rgb]{0.8330078125,0.9990234375,0.75} 1.34& \cellcolor[rgb]{0.8330078125,0.9990234375,0.75} 1.33& \cellcolor[rgb]{0.8349609375,0.9990234375,0.75} 1.34& \cellcolor[rgb]{0.9990234375,0.9697265625,0.75} 2.91\\
-31 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.83984375,0.9990234375,0.75} 1.36& \cellcolor[rgb]{0.8408203125,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.841796875,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.83984375,0.9990234375,0.75} 1.36& \cellcolor[rgb]{0.8408203125,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.83984375,0.9990234375,0.75} 1.36& \cellcolor[rgb]{0.9990234375,0.9677734375,0.75} 2.98\\
-32 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.8408203125,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.8408203125,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.8427734375,0.9990234375,0.75} 1.38& \cellcolor[rgb]{0.8408203125,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.841796875,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.8427734375,0.9990234375,0.75} 1.38& \cellcolor[rgb]{0.9990234375,0.9677734375,0.75} 2.99\\
-33 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.8427734375,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.841796875,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.8427734375,0.9990234375,0.75} 1.38& \cellcolor[rgb]{0.841796875,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.8408203125,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.841796875,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.9990234375,0.9677734375,0.75} 2.99\\
-34 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.841796875,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.841796875,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.841796875,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.841796875,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.8427734375,0.9990234375,0.75} 1.38& \cellcolor[rgb]{0.8408203125,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.9990234375,0.9677734375,0.75} 2.99\\
-35 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.841796875,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.841796875,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.8408203125,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.8427734375,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.841796875,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.8408203125,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.9990234375,0.9677734375,0.75} 2.99\\
-36 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.841796875,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.841796875,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.8427734375,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.8408203125,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.841796875,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.841796875,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.9990234375,0.9677734375,0.75} 2.99\\
+\begin{NiceTabular}{|c|cccccccc|}
+\hline
+n & clang & simple 12 & simple 2 & briggs 12 & briggs 2 & linear & greedy 12 & greedy 0 \\
+\hline
+8 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.7529296875,0.9990234375,0.75} 1.01248x& \cellcolor[rgb]{0.75390625,0.9990234375,0.75} 1.01766x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00146x& \cellcolor[rgb]{0.7548828125,0.9990234375,0.75} 1.02261x& \cellcolor[rgb]{0.7509765625,0.9990234375,0.75} 1.00606x& \cellcolor[rgb]{0.7529296875,0.9990234375,0.75} 1.01436x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99702x\\
+10 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99586x& \cellcolor[rgb]{0.7529296875,0.9990234375,0.75} 1.01369x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99508x& \cellcolor[rgb]{0.751953125,0.9990234375,0.75} 1.00861x& \cellcolor[rgb]{0.7548828125,0.9990234375,0.75} 1.01981x& \cellcolor[rgb]{0.751953125,0.9990234375,0.75} 1.00861x& \cellcolor[rgb]{0.755859375,0.9990234375,0.75} 1.02598x\\
+12 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.7548828125,0.9990234375,0.75} 1.02349x& \cellcolor[rgb]{0.76171875,0.9990234375,0.75} 1.04789x& \cellcolor[rgb]{0.759765625,0.9990234375,0.75} 1.04270x& \cellcolor[rgb]{0.7587890625,0.9990234375,0.75} 1.03716x& \cellcolor[rgb]{0.765625,0.9990234375,0.75} 1.06517x& \cellcolor[rgb]{0.763671875,0.9990234375,0.75} 1.05770x& \cellcolor[rgb]{0.77734375,0.9990234375,0.75} 1.11132x\\
+14 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.759765625,0.9990234375,0.75} 1.04054x& \cellcolor[rgb]{0.763671875,0.9990234375,0.75} 1.05851x& \cellcolor[rgb]{0.759765625,0.9990234375,0.75} 1.04050x& \cellcolor[rgb]{0.7626953125,0.9990234375,0.75} 1.05301x& \cellcolor[rgb]{0.7568359375,0.9990234375,0.75} 1.02863x& \cellcolor[rgb]{0.7607421875,0.9990234375,0.75} 1.04599x& \cellcolor[rgb]{0.7900390625,0.9990234375,0.75} 1.16166x\\
+16 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.771484375,0.9990234375,0.75} 1.08630x& \cellcolor[rgb]{0.7724609375,0.9990234375,0.75} 1.09112x& \cellcolor[rgb]{0.7666015625,0.9990234375,0.75} 1.06935x& \cellcolor[rgb]{0.7724609375,0.9990234375,0.75} 1.09077x& \cellcolor[rgb]{0.771484375,0.9990234375,0.75} 1.08669x& \cellcolor[rgb]{0.765625,0.9990234375,0.75} 1.06631x& \cellcolor[rgb]{0.8525390625,0.9990234375,0.75} 1.41519x\\
+18 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.7919921875,0.9990234375,0.75} 1.17172x& \cellcolor[rgb]{0.791015625,0.9990234375,0.75} 1.16684x& \cellcolor[rgb]{0.7880859375,0.9990234375,0.75} 1.15447x& \cellcolor[rgb]{0.7900390625,0.9990234375,0.75} 1.16224x& \cellcolor[rgb]{0.7919921875,0.9990234375,0.75} 1.17110x& \cellcolor[rgb]{0.7900390625,0.9990234375,0.75} 1.16144x& \cellcolor[rgb]{0.9560546875,0.9990234375,0.75} 1.83133x\\
+20 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.806640625,0.9990234375,0.75} 1.23037x& \cellcolor[rgb]{0.80859375,0.9990234375,0.75} 1.23533x& \cellcolor[rgb]{0.8076171875,0.9990234375,0.75} 1.23433x& \cellcolor[rgb]{0.8095703125,0.9990234375,0.75} 1.23986x& \cellcolor[rgb]{0.806640625,0.9990234375,0.75} 1.23120x& \cellcolor[rgb]{0.8046875,0.9990234375,0.75} 1.22283x& \cellcolor[rgb]{0.9990234375,0.990234375,0.75} 2.26430x\\
+22 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.826171875,0.9990234375,0.75} 1.30843x& \cellcolor[rgb]{0.8251953125,0.9990234375,0.75} 1.30498x& \cellcolor[rgb]{0.826171875,0.9990234375,0.75} 1.30814x& \cellcolor[rgb]{0.826171875,0.9990234375,0.75} 1.30594x& \cellcolor[rgb]{0.8251953125,0.9990234375,0.75} 1.30564x& \cellcolor[rgb]{0.82421875,0.9990234375,0.75} 1.29918x& \cellcolor[rgb]{0.9990234375,0.978515625,0.75} 2.64427x\\
+24 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.8408203125,0.9990234375,0.75} 1.36606x& \cellcolor[rgb]{0.83984375,0.9990234375,0.75} 1.36299x& \cellcolor[rgb]{0.8408203125,0.9990234375,0.75} 1.36538x& \cellcolor[rgb]{0.83984375,0.9990234375,0.75} 1.36151x& \cellcolor[rgb]{0.83984375,0.9990234375,0.75} 1.36449x& \cellcolor[rgb]{0.83984375,0.9990234375,0.75} 1.36204x& \cellcolor[rgb]{0.9990234375,0.970703125,0.75} 2.89705x\\
+26 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.8369140625,0.9990234375,0.75} 1.35207x& \cellcolor[rgb]{0.8369140625,0.9990234375,0.75} 1.35059x& \cellcolor[rgb]{0.837890625,0.9990234375,0.75} 1.35397x& \cellcolor[rgb]{0.8369140625,0.9990234375,0.75} 1.34954x& \cellcolor[rgb]{0.8369140625,0.9990234375,0.75} 1.35255x& \cellcolor[rgb]{0.8369140625,0.9990234375,0.75} 1.34932x& \cellcolor[rgb]{0.9990234375,0.9697265625,0.75} 2.92968x\\
+28 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.8388671875,0.9990234375,0.75} 1.35918x& \cellcolor[rgb]{0.837890625,0.9990234375,0.75} 1.35626x& \cellcolor[rgb]{0.8388671875,0.9990234375,0.75} 1.36006x& \cellcolor[rgb]{0.837890625,0.9990234375,0.75} 1.35674x& \cellcolor[rgb]{0.8388671875,0.9990234375,0.75} 1.35812x& \cellcolor[rgb]{0.837890625,0.9990234375,0.75} 1.35462x& \cellcolor[rgb]{0.9990234375,0.96875,0.75} 2.96506x\\
+30 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.8515625,0.9990234375,0.75} 1.41137x& \cellcolor[rgb]{0.8505859375,0.9990234375,0.75} 1.40745x& \cellcolor[rgb]{0.8515625,0.9990234375,0.75} 1.41125x& \cellcolor[rgb]{0.8505859375,0.9990234375,0.75} 1.40768x& \cellcolor[rgb]{0.8515625,0.9990234375,0.75} 1.40935x& \cellcolor[rgb]{0.8505859375,0.9990234375,0.75} 1.40589x& \cellcolor[rgb]{0.9990234375,0.96484375,0.75} 3.08506x\\
+32 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.853515625,0.9990234375,0.75} 1.41590x& \cellcolor[rgb]{0.8525390625,0.9990234375,0.75} 1.41255x& \cellcolor[rgb]{0.853515625,0.9990234375,0.75} 1.41652x& \cellcolor[rgb]{0.8525390625,0.9990234375,0.75} 1.41300x& \cellcolor[rgb]{0.8525390625,0.9990234375,0.75} 1.41399x& \cellcolor[rgb]{0.8515625,0.9990234375,0.75} 1.41115x& \cellcolor[rgb]{0.9990234375,0.9638671875,0.75} 3.09890x\\
+\hline
 \end{NiceTabular}
+\caption{\label{demo-table}Benchmark of \texttt{benches/fib.ll} output by \mintinline{bash}{dune exec bench -- -f fib -n 1000}}
 \end{center}
+\end{table}
 The above measures are significantly equivalent to one another, with all allocators (except greedy 0) seeming to converge towards being 37\% slower than Clang.
@@
 To illusidate this hypothesis, here is the same benchmark but only for the greedy of decreasing \(k\) assignable registers:
-
+\begin{table}[!h]
 \begin{center}
-\begin{NiceTabular}{>{\columncolor{gray!10}}ccccccccc}
-\rowcolor{gray!10}n & clang & greedy 12 & greedy 8 & greedy 6 & greedy 4 & greedy 2 & greedy 1 & greedy 0 \\
-20 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.8115234375,0.9990234375,0.75} 1.25& \cellcolor[rgb]{0.806640625,0.9990234375,0.75} 1.23& \cellcolor[rgb]{0.81640625,0.9990234375,0.75} 1.27& \cellcolor[rgb]{0.837890625,0.9990234375,0.75} 1.36& \cellcolor[rgb]{0.8505859375,0.9990234375,0.75} 1.41& \cellcolor[rgb]{0.9990234375,0.9921875,0.75} 2.20& \cellcolor[rgb]{0.9990234375,0.9912109375,0.75} 2.24\\
-21 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.8173828125,0.9990234375,0.75} 1.27& \cellcolor[rgb]{0.81640625,0.9990234375,0.75} 1.27& \cellcolor[rgb]{0.830078125,0.9990234375,0.75} 1.32& \cellcolor[rgb]{0.853515625,0.9990234375,0.75} 1.42& \cellcolor[rgb]{0.8642578125,0.9990234375,0.75} 1.46& \cellcolor[rgb]{0.9990234375,0.9853515625,0.75} 2.42& \cellcolor[rgb]{0.9990234375,0.984375,0.75} 2.46\\
-22 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.822265625,0.9990234375,0.75} 1.29& \cellcolor[rgb]{0.82421875,0.9990234375,0.75} 1.30& \cellcolor[rgb]{0.837890625,0.9990234375,0.75} 1.35& \cellcolor[rgb]{0.86328125,0.9990234375,0.75} 1.46& \cellcolor[rgb]{0.8759765625,0.9990234375,0.75} 1.51& \cellcolor[rgb]{0.9990234375,0.98046875,0.75} 2.57& \cellcolor[rgb]{0.9990234375,0.9794921875,0.75} 2.62\\
-23 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.8291015625,0.9990234375,0.75} 1.32& \cellcolor[rgb]{0.8291015625,0.9990234375,0.75} 1.32& \cellcolor[rgb]{0.84375,0.9990234375,0.75} 1.38& \cellcolor[rgb]{0.8740234375,0.9990234375,0.75} 1.50& \cellcolor[rgb]{0.88671875,0.9990234375,0.75} 1.55& \cellcolor[rgb]{0.9990234375,0.9775390625,0.75} 2.69& \cellcolor[rgb]{0.9990234375,0.9755859375,0.75} 2.74\\
-24 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.8359375,0.9990234375,0.75} 1.35& \cellcolor[rgb]{0.8349609375,0.9990234375,0.75} 1.34& \cellcolor[rgb]{0.8642578125,0.9990234375,0.75} 1.46& \cellcolor[rgb]{0.8837890625,0.9990234375,0.75} 1.54& \cellcolor[rgb]{0.8935546875,0.9990234375,0.75} 1.58& \cellcolor[rgb]{0.9990234375,0.974609375,0.75} 2.78& \cellcolor[rgb]{0.9990234375,0.97265625,0.75} 2.83\\
-25 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.8359375,0.9990234375,0.75} 1.35& \cellcolor[rgb]{0.837890625,0.9990234375,0.75} 1.35& \cellcolor[rgb]{0.8525390625,0.9990234375,0.75} 1.41& \cellcolor[rgb]{0.8837890625,0.9990234375,0.75} 1.54& \cellcolor[rgb]{0.896484375,0.9990234375,0.75} 1.59& \cellcolor[rgb]{0.9990234375,0.97265625,0.75} 2.83& \cellcolor[rgb]{0.9990234375,0.970703125,0.75} 2.89\\
-26 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.837890625,0.9990234375,0.75} 1.36& \cellcolor[rgb]{0.8388671875,0.9990234375,0.75} 1.36& \cellcolor[rgb]{0.853515625,0.9990234375,0.75} 1.42& \cellcolor[rgb]{0.88671875,0.9990234375,0.75} 1.55& \cellcolor[rgb]{0.8994140625,0.9990234375,0.75} 1.60& \cellcolor[rgb]{0.9990234375,0.9716796875,0.75} 2.87& \cellcolor[rgb]{0.9990234375,0.9697265625,0.75} 2.93\\
-27 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.8388671875,0.9990234375,0.75} 1.36& \cellcolor[rgb]{0.8388671875,0.9990234375,0.75} 1.36& \cellcolor[rgb]{0.857421875,0.9990234375,0.75} 1.43& \cellcolor[rgb]{0.8876953125,0.9990234375,0.75} 1.56& \cellcolor[rgb]{0.90234375,0.9990234375,0.75} 1.61& \cellcolor[rgb]{0.9990234375,0.970703125,0.75} 2.89& \cellcolor[rgb]{0.9990234375,0.96875,0.75} 2.95\\
-28 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.83984375,0.9990234375,0.75} 1.36& \cellcolor[rgb]{0.8408203125,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.8564453125,0.9990234375,0.75} 1.43& \cellcolor[rgb]{0.8916015625,0.9990234375,0.75} 1.57& \cellcolor[rgb]{0.904296875,0.9990234375,0.75} 1.62& \cellcolor[rgb]{0.9990234375,0.9697265625,0.75} 2.92& \cellcolor[rgb]{0.9990234375,0.96875,0.75} 2.97\\
-29 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.83984375,0.9990234375,0.75} 1.36& \cellcolor[rgb]{0.8408203125,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.8583984375,0.9990234375,0.75} 1.44& \cellcolor[rgb]{0.890625,0.9990234375,0.75} 1.57& \cellcolor[rgb]{0.9052734375,0.9990234375,0.75} 1.62& \cellcolor[rgb]{0.9990234375,0.9697265625,0.75} 2.92& \cellcolor[rgb]{0.9990234375,0.9677734375,0.75} 2.98\\
-30 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.83984375,0.9990234375,0.75} 1.36& \cellcolor[rgb]{0.8408203125,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.8564453125,0.9990234375,0.75} 1.43& \cellcolor[rgb]{0.890625,0.9990234375,0.75} 1.57& \cellcolor[rgb]{0.90625,0.9990234375,0.75} 1.63& \cellcolor[rgb]{0.9990234375,0.9697265625,0.75} 2.93& \cellcolor[rgb]{0.9990234375,0.9677734375,0.75} 2.98\\
-31 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.8408203125,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.8388671875,0.9990234375,0.75} 1.36& \cellcolor[rgb]{0.857421875,0.9990234375,0.75} 1.43& \cellcolor[rgb]{0.8916015625,0.9990234375,0.75} 1.57& \cellcolor[rgb]{0.9052734375,0.9990234375,0.75} 1.62& \cellcolor[rgb]{0.9990234375,0.9697265625,0.75} 2.93& \cellcolor[rgb]{0.9990234375,0.9677734375,0.75} 2.99\\
-32 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.8408203125,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.83984375,0.9990234375,0.75} 1.36& \cellcolor[rgb]{0.8564453125,0.9990234375,0.75} 1.43& \cellcolor[rgb]{0.890625,0.9990234375,0.75} 1.57& \cellcolor[rgb]{0.90625,0.9990234375,0.75} 1.63& \cellcolor[rgb]{0.9990234375,0.9697265625,0.75} 2.92& \cellcolor[rgb]{0.9990234375,0.9677734375,0.75} 2.98\\
-33 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.8408203125,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.83984375,0.9990234375,0.75} 1.36& \cellcolor[rgb]{0.857421875,0.9990234375,0.75} 1.43& \cellcolor[rgb]{0.8916015625,0.9990234375,0.75} 1.57& \cellcolor[rgb]{0.90625,0.9990234375,0.75} 1.63& \cellcolor[rgb]{0.9990234375,0.9697265625,0.75} 2.94& \cellcolor[rgb]{0.9990234375,0.9677734375,0.75} 2.99\\
-34 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.841796875,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.8408203125,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.857421875,0.9990234375,0.75} 1.43& \cellcolor[rgb]{0.8916015625,0.9990234375,0.75} 1.57& \cellcolor[rgb]{0.9052734375,0.9990234375,0.75} 1.63& \cellcolor[rgb]{0.9990234375,0.9697265625,0.75} 2.94& \cellcolor[rgb]{0.9990234375,0.9677734375,0.75} 2.99\\
-35 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.8408203125,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.8408203125,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.857421875,0.9990234375,0.75} 1.43& \cellcolor[rgb]{0.8916015625,0.9990234375,0.75} 1.57& \cellcolor[rgb]{0.90625,0.9990234375,0.75} 1.63& \cellcolor[rgb]{0.9990234375,0.9697265625,0.75} 2.93& \cellcolor[rgb]{0.9990234375,0.9677734375,0.75} 2.99\\
-36 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00& \cellcolor[rgb]{0.841796875,0.9990234375,0.75} 1.37& \cellcolor[rgb]{0.83984375,0.9990234375,0.75} 1.36& \cellcolor[rgb]{0.857421875,0.9990234375,0.75} 1.43& \cellcolor[rgb]{0.8916015625,0.9990234375,0.75} 1.57& \cellcolor[rgb]{0.90625,0.9990234375,0.75} 1.63& \cellcolor[rgb]{0.9990234375,0.9697265625,0.75} 2.94& \cellcolor[rgb]{0.9990234375,0.9677734375,0.75} 2.99\\
+\begin{NiceTabular}{|c|ccccccccc|}
+\hline
+n & clang & greedy 12 & greedy 8 & greedy 6 & greedy 4 & greedy 2 & greedy 1 & greedy 0 & tiger \\
+\hline
+8 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99429x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.98513x& \cellcolor[rgb]{0.751953125,0.9990234375,0.75} 1.00816x& \cellcolor[rgb]{0.7529296875,0.9990234375,0.75} 1.01180x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99806x& \cellcolor[rgb]{0.751953125,0.9990234375,0.75} 1.00869x& \cellcolor[rgb]{0.751953125,0.9990234375,0.75} 1.00883x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99979x\\
+10 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99057x& \cellcolor[rgb]{0.7578125,0.9990234375,0.75} 1.03464x& \cellcolor[rgb]{0.7578125,0.9990234375,0.75} 1.03147x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99206x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00151x& \cellcolor[rgb]{0.7607421875,0.9990234375,0.75} 1.04382x& \cellcolor[rgb]{0.76171875,0.9990234375,0.75} 1.04785x& \cellcolor[rgb]{0.763671875,0.9990234375,0.75} 1.05639x\\
+12 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00006x& \cellcolor[rgb]{0.7509765625,0.9990234375,0.75} 1.00578x& \cellcolor[rgb]{0.7529296875,0.9990234375,0.75} 1.01540x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99370x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.97925x& \cellcolor[rgb]{0.7666015625,0.9990234375,0.75} 1.06915x& \cellcolor[rgb]{0.76953125,0.9990234375,0.75} 1.07918x& \cellcolor[rgb]{0.7578125,0.9990234375,0.75} 1.03213x\\
+14 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.755859375,0.9990234375,0.75} 1.02737x& \cellcolor[rgb]{0.7568359375,0.9990234375,0.75} 1.02834x& \cellcolor[rgb]{0.76171875,0.9990234375,0.75} 1.04708x& \cellcolor[rgb]{0.7568359375,0.9990234375,0.75} 1.02997x& \cellcolor[rgb]{0.7646484375,0.9990234375,0.75} 1.06066x& \cellcolor[rgb]{0.79296875,0.9990234375,0.75} 1.17635x& \cellcolor[rgb]{0.7841796875,0.9990234375,0.75} 1.14104x& \cellcolor[rgb]{0.78125,0.9990234375,0.75} 1.12727x\\
+16 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.771484375,0.9990234375,0.75} 1.08717x& \cellcolor[rgb]{0.7734375,0.9990234375,0.75} 1.09426x& \cellcolor[rgb]{0.77734375,0.9990234375,0.75} 1.11143x& \cellcolor[rgb]{0.7841796875,0.9990234375,0.75} 1.13760x& \cellcolor[rgb]{0.78125,0.9990234375,0.75} 1.12774x& \cellcolor[rgb]{0.8525390625,0.9990234375,0.75} 1.41194x& \cellcolor[rgb]{0.849609375,0.9990234375,0.75} 1.40329x& \cellcolor[rgb]{0.8212890625,0.9990234375,0.75} 1.28770x\\
+18 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.78125,0.9990234375,0.75} 1.12836x& \cellcolor[rgb]{0.7861328125,0.9990234375,0.75} 1.14704x& \cellcolor[rgb]{0.7900390625,0.9990234375,0.75} 1.16301x& \cellcolor[rgb]{0.806640625,0.9990234375,0.75} 1.23098x& \cellcolor[rgb]{0.806640625,0.9990234375,0.75} 1.22815x& \cellcolor[rgb]{0.94140625,0.9990234375,0.75} 1.77085x& \cellcolor[rgb]{0.947265625,0.9990234375,0.75} 1.79503x& \cellcolor[rgb]{0.8720703125,0.9990234375,0.75} 1.49336x\\
+20 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.806640625,0.9990234375,0.75} 1.22913x& \cellcolor[rgb]{0.8056640625,0.9990234375,0.75} 1.22392x& \cellcolor[rgb]{0.8173828125,0.9990234375,0.75} 1.27241x& \cellcolor[rgb]{0.8369140625,0.9990234375,0.75} 1.35182x& \cellcolor[rgb]{0.845703125,0.9990234375,0.75} 1.38569x& \cellcolor[rgb]{0.9990234375,0.9912109375,0.75} 2.22689x& \cellcolor[rgb]{0.9990234375,0.990234375,0.75} 2.26200x& \cellcolor[rgb]{0.9501953125,0.9990234375,0.75} 1.80663x\\
+22 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.822265625,0.9990234375,0.75} 1.29405x& \cellcolor[rgb]{0.8232421875,0.9990234375,0.75} 1.29616x& \cellcolor[rgb]{0.8349609375,0.9990234375,0.75} 1.34432x& \cellcolor[rgb]{0.865234375,0.9990234375,0.75} 1.46289x& \cellcolor[rgb]{0.8759765625,0.9990234375,0.75} 1.50765x& \cellcolor[rgb]{0.9990234375,0.98046875,0.75} 2.57480x& \cellcolor[rgb]{0.9990234375,0.9794921875,0.75} 2.62632x& \cellcolor[rgb]{0.9990234375,0.9970703125,0.75} 2.05105x\\
+24 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.83203125,0.9990234375,0.75} 1.33240x& \cellcolor[rgb]{0.83203125,0.9990234375,0.75} 1.33275x& \cellcolor[rgb]{0.8486328125,0.9990234375,0.75} 1.39726x& \cellcolor[rgb]{0.87890625,0.9990234375,0.75} 1.52070x& \cellcolor[rgb]{0.892578125,0.9990234375,0.75} 1.57452x& \cellcolor[rgb]{0.9990234375,0.974609375,0.75} 2.77989x& \cellcolor[rgb]{0.9990234375,0.97265625,0.75} 2.83443x& \cellcolor[rgb]{0.9990234375,0.9931640625,0.75} 2.18442x\\
+26 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.8359375,0.9990234375,0.75} 1.34858x& \cellcolor[rgb]{0.8359375,0.9990234375,0.75} 1.34784x& \cellcolor[rgb]{0.853515625,0.9990234375,0.75} 1.41620x& \cellcolor[rgb]{0.8828125,0.9990234375,0.75} 1.53692x& \cellcolor[rgb]{0.8994140625,0.9990234375,0.75} 1.60380x& \cellcolor[rgb]{0.9990234375,0.9716796875,0.75} 2.87114x& \cellcolor[rgb]{0.9990234375,0.9697265625,0.75} 2.92874x& \cellcolor[rgb]{0.9990234375,0.9912109375,0.75} 2.24149x\\
+28 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.837890625,0.9990234375,0.75} 1.35448x& \cellcolor[rgb]{0.837890625,0.9990234375,0.75} 1.35457x& \cellcolor[rgb]{0.8544921875,0.9990234375,0.75} 1.42324x& \cellcolor[rgb]{0.8857421875,0.9990234375,0.75} 1.54897x& \cellcolor[rgb]{0.90234375,0.9990234375,0.75} 1.61429x& \cellcolor[rgb]{0.9990234375,0.970703125,0.75} 2.90730x& \cellcolor[rgb]{0.9990234375,0.96875,0.75} 2.96219x& \cellcolor[rgb]{0.9990234375,0.990234375,0.75} 2.26833x\\
+30 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.8388671875,0.9990234375,0.75} 1.35807x& \cellcolor[rgb]{0.8388671875,0.9990234375,0.75} 1.35784x& \cellcolor[rgb]{0.85546875,0.9990234375,0.75} 1.42732x& \cellcolor[rgb]{0.888671875,0.9990234375,0.75} 1.55842x& \cellcolor[rgb]{0.904296875,0.9990234375,0.75} 1.62109x& \cellcolor[rgb]{0.9990234375,0.9697265625,0.75} 2.92283x& \cellcolor[rgb]{0.9990234375,0.9677734375,0.75} 2.98079x& \cellcolor[rgb]{0.9990234375,0.990234375,0.75} 2.28148x\\
+32 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.841796875,0.9990234375,0.75} 1.36987x& \cellcolor[rgb]{0.841796875,0.9990234375,0.75} 1.36911x& \cellcolor[rgb]{0.8583984375,0.9990234375,0.75} 1.43879x& \cellcolor[rgb]{0.892578125,0.9990234375,0.75} 1.57453x& \cellcolor[rgb]{0.9072265625,0.9990234375,0.75} 1.63423x& \cellcolor[rgb]{0.9990234375,0.96875,0.75} 2.94896x& \cellcolor[rgb]{0.9990234375,0.966796875,0.75} 3.00741x& \cellcolor[rgb]{0.9990234375,0.9892578125,0.75} 2.29882x\\
+\hline
 \end{NiceTabular}
+\caption{\label{demo-table}Benchmark of \texttt{benches/fib.ll} output by \mintinline{bash}{dune exec bench -- -f fib -n 1000}}
 \end{center}
+\end{table}
 
@@
 
-%\begin{center}
-%\begin{tabular}{|c||c|c|c|c|c|c|c|c|c|c|c|} 
- %\hline
-%& clang & simple 12 & simple 2 & briggs 12 & briggs 2 & linear & greedy 12 & greedy 0 \\
- %\hline
-%34 & \cellcolor{1.0, 1.0 , 1.0}1.00& 1.37& 1.37& 1.37& 1.37& 1.36& 1.41& 2.99\\
-%35 & 1.00& 1.39& 1.37& 1.41& 1.37& 1.37& 1.38& 3.00\\
-%36 & 1.00& 1.37& 1.37& 1.37& 1.37& 1.37& 1.38& 2.99\\
-%37 & 1.00& 1.38& 1.37& 1.37& 1.37& 1.39& 1.37& 2.99\\
-%38 & 1.00& 1.37& 1.37& 1.38& 1.37& 1.37& 1.37& 2.99\\
-%39 & 1.00& 1.37& 1.38& 1.37& 1.39& 1.37& 1.37& 2.99\\
-%40 & 1.00& 1.37& 1.37& 1.39& 1.37& 1.37& 1.37& 2.99\\
-%41 & 1.00& 1.37& 1.38& 1.37& 1.37& 1.37& 1.37& 2.99\\
-%42 & 1.00& 1.38& 1.38& 1.37& 1.37& 1.37& 1.37& 3.04\\
- %\hline
-%\end{tabular}
-%\end{center}
+\subsubsection{\lstinline!benches/factori32.ll!}
 
-%\begin{center}
-%\begin{tabular}{|c||c|c|c|c|c|c|c|c|c|c|c|} 
- %\hline
-%&clang& greedy 12 & greedy 8 & greedy 4 & greedy 2 & greedy 1 & greedy 0 \\
- %\hline
-%34 & 1.00& 1.41& 1.37& 1.57& 1.63& 2.93& 3.04\\
-%35 & 1.00& 1.37& 1.38& 1.58& 1.66& 2.95& 3.01\\
-%36 & 1.00& 1.36& 1.40& 1.56& 1.66& 2.93& 2.98\\
-%37 & 1.00& 1.38& 1.38& 1.58& 1.64& 2.95& 3.01\\
-%38 & 1.00& 1.37& 1.37& 1.57& 1.64& 2.93& 3.04\\
-%39 & 1.00& 1.36& 1.37& 1.57& 1.63& 2.94& 3.04\\
-%40 & 1.00& 1.37& 1.37& 1.57& 1.63& 2.93& 3.04\\
-%41 & 1.00& 1.38& 1.41& 1.58& 1.67& 2.97& 3.03\\
-%42 & 1.00& 1.36& 1.37& 1.56& 1.62& 2.92& 2.98\\
-% \hline
-%\end{tabular}
-%\end{center}
+Prime factorization of \lstinline!i32!.
+\begin{table}[!h]
+\begin{center}
+\begin{NiceTabular}{|c|cccccccc|}
+\hline
+n & clang & simple 12 & simple 2 & briggs 12 & briggs 2 & linear & greedy 12 & greedy 0 \\
+\hline
+268435399 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.7783203125,0.9990234375,0.75} 1.11499x& \cellcolor[rgb]{0.77734375,0.9990234375,0.75} 1.11025x& \cellcolor[rgb]{0.7783203125,0.9990234375,0.75} 1.11595x& \cellcolor[rgb]{0.779296875,0.9990234375,0.75} 1.11943x& \cellcolor[rgb]{0.916015625,0.9990234375,0.75} 1.66718x& \cellcolor[rgb]{0.9990234375,0.9833984375,0.75} 2.47709x& \cellcolor[rgb]{0.9990234375,0.984375,0.75} 2.44535x\\
+536870909 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.64555x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.64645x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.63974x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.64095x& \cellcolor[rgb]{0.7607421875,0.9990234375,0.75} 1.04505x& \cellcolor[rgb]{0.9013671875,0.9990234375,0.75} 1.60979x& \cellcolor[rgb]{0.904296875,0.9990234375,0.75} 1.61997x\\
+1073741789 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.7958984375,0.9990234375,0.75} 1.18438x& \cellcolor[rgb]{0.7978515625,0.9990234375,0.75} 1.19584x& \cellcolor[rgb]{0.796875,0.9990234375,0.75} 1.19047x& \cellcolor[rgb]{0.7958984375,0.9990234375,0.75} 1.18472x& \cellcolor[rgb]{0.9990234375,0.9951171875,0.75} 2.09863x& \cellcolor[rgb]{0.9990234375,0.9560546875,0.75} 3.36121x& \cellcolor[rgb]{0.9990234375,0.9560546875,0.75} 3.37087x\\
+2147483647 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.8017578125,0.9990234375,0.75} 1.20872x& \cellcolor[rgb]{0.80078125,0.9990234375,0.75} 1.20665x& \cellcolor[rgb]{0.80078125,0.9990234375,0.75} 1.20405x& \cellcolor[rgb]{0.8017578125,0.9990234375,0.75} 1.20826x& \cellcolor[rgb]{0.9990234375,0.98828125,0.75} 2.32725x& \cellcolor[rgb]{0.9990234375,0.9453125,0.75} 3.71253x& \cellcolor[rgb]{0.9990234375,0.9453125,0.75} 3.71693x\\
+4294967291 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.79296875,0.9990234375,0.75} 1.17597x& \cellcolor[rgb]{0.794921875,0.9990234375,0.75} 1.18070x& \cellcolor[rgb]{0.794921875,0.9990234375,0.75} 1.18067x& \cellcolor[rgb]{0.794921875,0.9990234375,0.75} 1.18159x& \cellcolor[rgb]{0.9990234375,0.986328125,0.75} 2.37710x& \cellcolor[rgb]{0.9990234375,0.9345703125,0.75} 4.04637x& \cellcolor[rgb]{0.9990234375,0.9345703125,0.75} 4.06492x\\
+8589934583 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.8203125,0.9990234375,0.75} 1.28534x& \cellcolor[rgb]{0.822265625,0.9990234375,0.75} 1.29317x& \cellcolor[rgb]{0.822265625,0.9990234375,0.75} 1.29145x& \cellcolor[rgb]{0.8212890625,0.9990234375,0.75} 1.28941x& \cellcolor[rgb]{0.9990234375,0.9755859375,0.75} 2.73679x& \cellcolor[rgb]{0.9990234375,0.9130859375,0.75} 4.74271x& \cellcolor[rgb]{0.9990234375,0.912109375,0.75} 4.76719x\\
+17179869143 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.8291015625,0.9990234375,0.75} 1.31790x& \cellcolor[rgb]{0.830078125,0.9990234375,0.75} 1.32297x& \cellcolor[rgb]{0.8310546875,0.9990234375,0.75} 1.32585x& \cellcolor[rgb]{0.830078125,0.9990234375,0.75} 1.32452x& \cellcolor[rgb]{0.9990234375,0.96875,0.75} 2.96961x& \cellcolor[rgb]{0.9990234375,0.900390625,0.75} 5.15949x& \cellcolor[rgb]{0.9990234375,0.9033203125,0.75} 5.07055x\\
+34359738337 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.8369140625,0.9990234375,0.75} 1.34905x& \cellcolor[rgb]{0.8369140625,0.9990234375,0.75} 1.34969x& \cellcolor[rgb]{0.8359375,0.9990234375,0.75} 1.34780x& \cellcolor[rgb]{0.845703125,0.9990234375,0.75} 1.38507x& \cellcolor[rgb]{0.9990234375,0.96484375,0.75} 3.08190x& \cellcolor[rgb]{0.9990234375,0.8935546875,0.75} 5.37886x& \cellcolor[rgb]{0.9990234375,0.892578125,0.75} 5.39434x\\
+\hline
+\end{NiceTabular}
+\caption{\label{demo-table}Benchmark of \texttt{benches/fib.ll} output by \mintinline{bash}{dune exec bench -- -f fib -n 1000}}
+\end{center}
+\end{table}
 
-\subsubsection{\lstinline!benches/factori32.ll!}
-Prime factorization of \lstinline!i32!.
+\subsubsection{\lstinline!benches/factori64.ll!}
+\begin{table}[!h]
 \begin{center}
-\begin{tabular}{|c||c|c|c|c|} 
+\begin{NiceTabular}{|c|cccccccc|}
 \hline
- fib & clang & greedy & simple & linear \\ [0.5ex] 
+n & clang & simple 12 & simple 2 & briggs 12 & briggs 2 & linear & greedy 12 & greedy 0 \\
 \hline
- 40 & 1.000000 & 1.438693 & 1.441270 & 0.73 \\ [1ex] 
- 41 & 1.000000 & 1.441035 & 1.440626 & 0.73 \\ [1ex] 
- 42 & 	1.000000 & 1.447140 & 1.446094 & 0\\
+268435399 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.7783203125,0.9990234375,0.75} 1.11499x& \cellcolor[rgb]{0.77734375,0.9990234375,0.75} 1.11025x& \cellcolor[rgb]{0.7783203125,0.9990234375,0.75} 1.11595x& \cellcolor[rgb]{0.779296875,0.9990234375,0.75} 1.11943x& \cellcolor[rgb]{0.916015625,0.9990234375,0.75} 1.66718x& \cellcolor[rgb]{0.9990234375,0.9833984375,0.75} 2.47709x& \cellcolor[rgb]{0.9990234375,0.984375,0.75} 2.44535x\\
+536870909 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.64555x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.64645x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.63974x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.64095x& \cellcolor[rgb]{0.7607421875,0.9990234375,0.75} 1.04505x& \cellcolor[rgb]{0.9013671875,0.9990234375,0.75} 1.60979x& \cellcolor[rgb]{0.904296875,0.9990234375,0.75} 1.61997x\\
+1073741789 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.7958984375,0.9990234375,0.75} 1.18438x& \cellcolor[rgb]{0.7978515625,0.9990234375,0.75} 1.19584x& \cellcolor[rgb]{0.796875,0.9990234375,0.75} 1.19047x& \cellcolor[rgb]{0.7958984375,0.9990234375,0.75} 1.18472x& \cellcolor[rgb]{0.9990234375,0.9951171875,0.75} 2.09863x& \cellcolor[rgb]{0.9990234375,0.9560546875,0.75} 3.36121x& \cellcolor[rgb]{0.9990234375,0.9560546875,0.75} 3.37087x\\
+2147483647 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.8017578125,0.9990234375,0.75} 1.20872x& \cellcolor[rgb]{0.80078125,0.9990234375,0.75} 1.20665x& \cellcolor[rgb]{0.80078125,0.9990234375,0.75} 1.20405x& \cellcolor[rgb]{0.8017578125,0.9990234375,0.75} 1.20826x& \cellcolor[rgb]{0.9990234375,0.98828125,0.75} 2.32725x& \cellcolor[rgb]{0.9990234375,0.9453125,0.75} 3.71253x& \cellcolor[rgb]{0.9990234375,0.9453125,0.75} 3.71693x\\
+4294967291 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.79296875,0.9990234375,0.75} 1.17597x& \cellcolor[rgb]{0.794921875,0.9990234375,0.75} 1.18070x& \cellcolor[rgb]{0.794921875,0.9990234375,0.75} 1.18067x& \cellcolor[rgb]{0.794921875,0.9990234375,0.75} 1.18159x& \cellcolor[rgb]{0.9990234375,0.986328125,0.75} 2.37710x& \cellcolor[rgb]{0.9990234375,0.9345703125,0.75} 4.04637x& \cellcolor[rgb]{0.9990234375,0.9345703125,0.75} 4.06492x\\
+8589934583 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.8203125,0.9990234375,0.75} 1.28534x& \cellcolor[rgb]{0.822265625,0.9990234375,0.75} 1.29317x& \cellcolor[rgb]{0.822265625,0.9990234375,0.75} 1.29145x& \cellcolor[rgb]{0.8212890625,0.9990234375,0.75} 1.28941x& \cellcolor[rgb]{0.9990234375,0.9755859375,0.75} 2.73679x& \cellcolor[rgb]{0.9990234375,0.9130859375,0.75} 4.74271x& \cellcolor[rgb]{0.9990234375,0.912109375,0.75} 4.76719x\\
+17179869143 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.8291015625,0.9990234375,0.75} 1.31790x& \cellcolor[rgb]{0.830078125,0.9990234375,0.75} 1.32297x& \cellcolor[rgb]{0.8310546875,0.9990234375,0.75} 1.32585x& \cellcolor[rgb]{0.830078125,0.9990234375,0.75} 1.32452x& \cellcolor[rgb]{0.9990234375,0.96875,0.75} 2.96961x& \cellcolor[rgb]{0.9990234375,0.900390625,0.75} 5.15949x& \cellcolor[rgb]{0.9990234375,0.9033203125,0.75} 5.07055x\\
+34359738337 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.8369140625,0.9990234375,0.75} 1.34905x& \cellcolor[rgb]{0.8369140625,0.9990234375,0.75} 1.34969x& \cellcolor[rgb]{0.8359375,0.9990234375,0.75} 1.34780x& \cellcolor[rgb]{0.845703125,0.9990234375,0.75} 1.38507x& \cellcolor[rgb]{0.9990234375,0.96484375,0.75} 3.08190x& \cellcolor[rgb]{0.9990234375,0.8935546875,0.75} 5.37886x& \cellcolor[rgb]{0.9990234375,0.892578125,0.75} 5.39434x\\
 \hline
-\end{tabular}
+\end{NiceTabular}
+\caption{\label{demo-table}Benchmark of \texttt{benches/fib.ll} output by \mintinline{bash}{dune exec bench -- -f fib -n 1000}}
 \end{center}
+\end{table}
 
-%TODO:
-%
-%* should I use shell time, or /usr/bin/time?
-%* how do I measure cpu cycles?
-%\begin{center}
-%\begin{tabular}{ c c c }
- %cell1 & cell2 & cell3 \\ 
- %cell4 & cell5 & cell6 \\  
- %cell7 & cell8 & cell9    
-%\end{tabular}
-%\end{center}
+\subsubsection{\lstinline!benches/sieven.ll!}
+\begin{table}[!h]
+\begin{center}
+\begin{NiceTabular}{|c|cccccccc|}
+\hline
+n & clang & simple 12 & simple 2 & briggs 12 & briggs 2 & linear & greedy 12 & greedy 0 \\
+\hline
+128 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.98451x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.98378x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.98295x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00304x& \cellcolor[rgb]{0.8154296875,0.9990234375,0.75} 1.26346x& \cellcolor[rgb]{0.8125,0.9990234375,0.75} 1.25285x& \cellcolor[rgb]{0.8115234375,0.9990234375,0.75} 1.24850x\\
+256 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.7529296875,0.9990234375,0.75} 1.01514x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00375x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.98807x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00262x& \cellcolor[rgb]{0.869140625,0.9990234375,0.75} 1.47906x& \cellcolor[rgb]{0.8876953125,0.9990234375,0.75} 1.55358x& \cellcolor[rgb]{0.8857421875,0.9990234375,0.75} 1.54756x\\
+512 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.751953125,0.9990234375,0.75} 1.00899x& \cellcolor[rgb]{0.7578125,0.9990234375,0.75} 1.03156x& \cellcolor[rgb]{0.7578125,0.9990234375,0.75} 1.03499x& \cellcolor[rgb]{0.7529296875,0.9990234375,0.75} 1.01201x& \cellcolor[rgb]{0.9990234375,0.998046875,0.75} 2.00339x& \cellcolor[rgb]{0.9990234375,0.9951171875,0.75} 2.09435x& \cellcolor[rgb]{0.9990234375,0.994140625,0.75} 2.14397x\\
+1024 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.7568359375,0.9990234375,0.75} 1.03016x& \cellcolor[rgb]{0.7568359375,0.9990234375,0.75} 1.02839x& \cellcolor[rgb]{0.7529296875,0.9990234375,0.75} 1.01247x& \cellcolor[rgb]{0.7568359375,0.9990234375,0.75} 1.03097x& \cellcolor[rgb]{0.9990234375,0.97265625,0.75} 2.84545x& \cellcolor[rgb]{0.9990234375,0.96484375,0.75} 3.07906x& \cellcolor[rgb]{0.9990234375,0.96484375,0.75} 3.07390x\\
+2048 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.7587890625,0.9990234375,0.75} 1.03732x& \cellcolor[rgb]{0.7607421875,0.9990234375,0.75} 1.04387x& \cellcolor[rgb]{0.7802734375,0.9990234375,0.75} 1.12328x& \cellcolor[rgb]{0.76171875,0.9990234375,0.75} 1.04998x& \cellcolor[rgb]{0.9990234375,0.9287109375,0.75} 4.24679x& \cellcolor[rgb]{0.9990234375,0.919921875,0.75} 4.52175x& \cellcolor[rgb]{0.9990234375,0.91796875,0.75} 4.60346x\\
+4096 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.7646484375,0.9990234375,0.75} 1.05883x& \cellcolor[rgb]{0.78125,0.9990234375,0.75} 1.12849x& \cellcolor[rgb]{0.7626953125,0.9990234375,0.75} 1.05254x& \cellcolor[rgb]{0.775390625,0.9990234375,0.75} 1.10378x& \cellcolor[rgb]{0.9990234375,0.8720703125,0.75} 6.05831x& \cellcolor[rgb]{0.9990234375,0.8583984375,0.75} 6.50294x& \cellcolor[rgb]{0.9990234375,0.8544921875,0.75} 6.61920x\\
+8192 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.7724609375,0.9990234375,0.75} 1.09357x& \cellcolor[rgb]{0.8095703125,0.9990234375,0.75} 1.24057x& \cellcolor[rgb]{0.7744140625,0.9990234375,0.75} 1.10146x& \cellcolor[rgb]{0.7783203125,0.9990234375,0.75} 1.11378x& \cellcolor[rgb]{0.9990234375,0.80859375,0.75} 8.09302x& \cellcolor[rgb]{0.9990234375,0.7900390625,0.75} 8.69270x& \cellcolor[rgb]{0.9990234375,0.783203125,0.75} 8.92530x\\
+16384 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.7763671875,0.9990234375,0.75} 1.10882x& \cellcolor[rgb]{0.7822265625,0.9990234375,0.75} 1.13274x& \cellcolor[rgb]{0.7763671875,0.9990234375,0.75} 1.10937x& \cellcolor[rgb]{0.78515625,0.9990234375,0.75} 1.14416x& \cellcolor[rgb]{0.9990234375,0.7548828125,0.75} 9.84245x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 10.69728x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 10.80864x\\
+\hline
+\end{NiceTabular}
+\caption{\label{demo-table}Benchmark of \texttt{benches/fib.ll} output by \mintinline{bash}{dune exec bench -- -f fib -n 1000}}
+\end{center}
+\end{table}
 
-%\begin{center}
-%\begin{tabular}{|c||c|c|c|c|} 
- %\hline
- %sha256 & clang & greedy & briggs & linear \\ [0.5ex] 
- %\hline
- %1000000 & 1 & 0.73 & 0.73 & 0.73 \\ [1ex] 
- %\hline
-%\end{tabular}
-%\end{center}
+\subsubsection{\lstinline!benches/subset.ll!}
+\begin{NiceTabular}{|c|cccccccc|}
+\hline
+n & clang & simple 12 & simple 2 & briggs 12 & briggs 2 & linear & greedy 12 & greedy 0 \\
+\hline
+14 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.771484375,0.9990234375,0.75} 1.08770x& \cellcolor[rgb]{0.787109375,0.9990234375,0.75} 1.15188x& \cellcolor[rgb]{0.771484375,0.9990234375,0.75} 1.08832x& \cellcolor[rgb]{0.7890625,0.9990234375,0.75} 1.15813x& \cellcolor[rgb]{0.98046875,0.9990234375,0.75} 1.92564x& \cellcolor[rgb]{0.9990234375,0.98046875,0.75} 2.59504x& \cellcolor[rgb]{0.9990234375,0.96875,0.75} 2.95798x\\
+15 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.7705078125,0.9990234375,0.75} 1.08304x& \cellcolor[rgb]{0.78515625,0.9990234375,0.75} 1.14485x& \cellcolor[rgb]{0.7705078125,0.9990234375,0.75} 1.08536x& \cellcolor[rgb]{0.787109375,0.9990234375,0.75} 1.14973x& \cellcolor[rgb]{0.9833984375,0.9990234375,0.75} 1.93885x& \cellcolor[rgb]{0.9990234375,0.9794921875,0.75} 2.60300x& \cellcolor[rgb]{0.9990234375,0.96875,0.75} 2.96001x\\
+16 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.7705078125,0.9990234375,0.75} 1.08622x& \cellcolor[rgb]{0.78515625,0.9990234375,0.75} 1.14309x& \cellcolor[rgb]{0.7705078125,0.9990234375,0.75} 1.08345x& \cellcolor[rgb]{0.78515625,0.9990234375,0.75} 1.14388x& \cellcolor[rgb]{0.9853515625,0.9990234375,0.75} 1.94770x& \cellcolor[rgb]{0.9990234375,0.978515625,0.75} 2.62809x& \cellcolor[rgb]{0.9990234375,0.9677734375,0.75} 2.99463x\\
+17 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.7705078125,0.9990234375,0.75} 1.08533x& \cellcolor[rgb]{0.78515625,0.9990234375,0.75} 1.14335x& \cellcolor[rgb]{0.771484375,0.9990234375,0.75} 1.08652x& \cellcolor[rgb]{0.7861328125,0.9990234375,0.75} 1.14605x& \cellcolor[rgb]{0.986328125,0.9990234375,0.75} 1.94922x& \cellcolor[rgb]{0.9990234375,0.978515625,0.75} 2.64844x& \cellcolor[rgb]{0.9990234375,0.966796875,0.75} 3.01363x\\
+\hline
+\end{NiceTabular}
+
+\begin{NiceTabular}{|c|cccccccc|}
+\hline
+n & clang & simple 12 & simple 2 & briggs 12 & briggs 2 & linear & greedy 12 & greedy 0 \\
+\hline
+100 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.9990234375,0.9619140625,0.75} 3.17253x& \cellcolor[rgb]{0.9990234375,0.9560546875,0.75} 3.35873x& \cellcolor[rgb]{0.9990234375,0.9619140625,0.75} 3.17637x& \cellcolor[rgb]{0.9990234375,0.9560546875,0.75} 3.37014x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 19.50758x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 19.44391x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 19.61921x\\
+1000 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.9990234375,0.9189453125,0.75} 4.54197x& \cellcolor[rgb]{0.9990234375,0.91015625,0.75} 4.83710x& \cellcolor[rgb]{0.9990234375,0.9189453125,0.75} 4.54593x& \cellcolor[rgb]{0.9990234375,0.91015625,0.75} 4.83359x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 31.35583x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 31.26454x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 31.56475x\\
+10000 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.9990234375,0.9111328125,0.75} 4.80821x& \cellcolor[rgb]{0.9990234375,0.9013671875,0.75} 5.11478x& \cellcolor[rgb]{0.9990234375,0.9111328125,0.75} 4.80648x& \cellcolor[rgb]{0.9990234375,0.9013671875,0.75} 5.12019x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 33.50659x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 33.41428x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 33.72138x\\
+\hline
+\end{NiceTabular}
 

### 2023-12-27 article.tex
More text

@@
 
-Conversely, managed languages such as Java and C# typically don't depend on the LLVM toolchain directly, research projects that attempt to compile associated VM bytecode to a native platform using LLVM.
+Conversely, managed languages such as Java and C\# typically don't depend on the LLVM toolchain directly, research projects that attempt to compile associated VM bytecode to a native platform using LLVM (\cite{dotnet}).
+Meanwhile interpreted languages like JavaScript eagerly attempt use LLVM for \textit{just-in-time} (JIT) compilation % FIXME: this is not true at all lol
 
@@
 \end{figure}
+
+\subsection{Static Single-Assignment form}
+In compiler backends, \textit{Static Single-Assignment} (SSA) form is a property that applies to some IR, including LLVM.
+It is a way of representing a program in such a way that each variable is assigned only once throughout
+the scope of the function. Each use of the variable after definition then refers to the value  of that initial assignment.
+%In other words, every variable is assigned a unique version number for each assignment, and these versions are used consistently throughout the program.
 
@@
 The 'dumb' fib is a very naive approach to finding the \(n\)th number of the Fibonacci sequence by calling itself recursively twice (decrementing \(n\) before each), which causes an exponential growth in function calls.
-\begin{table}[!h]
+\begin{table}[!ht]
 \begin{center}
@@
 \end{NiceTabular}
-\caption{\label{demo-table}Benchmark of \texttt{benches/fib.ll} output by \mintinline{bash}{dune exec bench -- -f fib -n 1000}}
+\caption{\label{tab:dumb-fib}Benchmark of \texttt{benches/fib.ll} output by \mintinline{bash}{dune exec bench -- -f fib -n 1000}}
 \end{center}
@@
 To illusidate this hypothesis, here is the same benchmark but only for the greedy of decreasing \(k\) assignable registers:
-\begin{table}[!h]
+\begin{table}[!ht]
 \begin{center}
@@
 \end{NiceTabular}
-\caption{\label{demo-table}Benchmark of \texttt{benches/fib.ll} output by \mintinline{bash}{dune exec bench -- -f fib -n 1000}}
+\caption{\label{tab:dumb-fib-greedy}Benchmark of \texttt{benches/fib.ll} output by \mintinline{bash}{dune exec bench -- -f fib -n 1000}}
 \end{center}
@@
 Prime factorization of \lstinline!i32!.
-\begin{table}[!h]
+\begin{table}[!ht]
 \begin{center}
@@
 \end{NiceTabular}
-\caption{\label{demo-table}Benchmark of \texttt{benches/fib.ll} output by \mintinline{bash}{dune exec bench -- -f fib -n 1000}}
+\caption{\label{tab:factori32}Benchmark of \texttt{benches/fib.ll} output by \mintinline{bash}{dune exec bench -- -f fib -n 1000}}
 \end{center}
@@
 \subsubsection{\lstinline!benches/factori64.ll!}
-\begin{table}[!h]
+\begin{table}[!ht]
 \begin{center}
@@
 \end{NiceTabular}
-\caption{\label{demo-table}Benchmark of \texttt{benches/fib.ll} output by \mintinline{bash}{dune exec bench -- -f fib -n 1000}}
+\caption{\label{tab:factori64-table}Benchmark of \texttt{benches/fib.ll} output by \mintinline{bash}{dune exec bench -- -f fib -n 1000}}
 \end{center}
@@
 \subsubsection{\lstinline!benches/sieven.ll!}
-\begin{table}[!h]
+\begin{table}[!ht]
 \begin{center}
@@
 \hline
-n & clang & simple 12 & simple 2 & briggs 12 & briggs 2 & linear & greedy 12 & greedy 0 \\
-\hline
-128 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.98451x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.98378x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.98295x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00304x& \cellcolor[rgb]{0.8154296875,0.9990234375,0.75} 1.26346x& \cellcolor[rgb]{0.8125,0.9990234375,0.75} 1.25285x& \cellcolor[rgb]{0.8115234375,0.9990234375,0.75} 1.24850x\\
-256 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.7529296875,0.9990234375,0.75} 1.01514x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00375x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.98807x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00262x& \cellcolor[rgb]{0.869140625,0.9990234375,0.75} 1.47906x& \cellcolor[rgb]{0.8876953125,0.9990234375,0.75} 1.55358x& \cellcolor[rgb]{0.8857421875,0.9990234375,0.75} 1.54756x\\
-512 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.751953125,0.9990234375,0.75} 1.00899x& \cellcolor[rgb]{0.7578125,0.9990234375,0.75} 1.03156x& \cellcolor[rgb]{0.7578125,0.9990234375,0.75} 1.03499x& \cellcolor[rgb]{0.7529296875,0.9990234375,0.75} 1.01201x& \cellcolor[rgb]{0.9990234375,0.998046875,0.75} 2.00339x& \cellcolor[rgb]{0.9990234375,0.9951171875,0.75} 2.09435x& \cellcolor[rgb]{0.9990234375,0.994140625,0.75} 2.14397x\\
-1024 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.7568359375,0.9990234375,0.75} 1.03016x& \cellcolor[rgb]{0.7568359375,0.9990234375,0.75} 1.02839x& \cellcolor[rgb]{0.7529296875,0.9990234375,0.75} 1.01247x& \cellcolor[rgb]{0.7568359375,0.9990234375,0.75} 1.03097x& \cellcolor[rgb]{0.9990234375,0.97265625,0.75} 2.84545x& \cellcolor[rgb]{0.9990234375,0.96484375,0.75} 3.07906x& \cellcolor[rgb]{0.9990234375,0.96484375,0.75} 3.07390x\\
-2048 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.7587890625,0.9990234375,0.75} 1.03732x& \cellcolor[rgb]{0.7607421875,0.9990234375,0.75} 1.04387x& \cellcolor[rgb]{0.7802734375,0.9990234375,0.75} 1.12328x& \cellcolor[rgb]{0.76171875,0.9990234375,0.75} 1.04998x& \cellcolor[rgb]{0.9990234375,0.9287109375,0.75} 4.24679x& \cellcolor[rgb]{0.9990234375,0.919921875,0.75} 4.52175x& \cellcolor[rgb]{0.9990234375,0.91796875,0.75} 4.60346x\\
-4096 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.7646484375,0.9990234375,0.75} 1.05883x& \cellcolor[rgb]{0.78125,0.9990234375,0.75} 1.12849x& \cellcolor[rgb]{0.7626953125,0.9990234375,0.75} 1.05254x& \cellcolor[rgb]{0.775390625,0.9990234375,0.75} 1.10378x& \cellcolor[rgb]{0.9990234375,0.8720703125,0.75} 6.05831x& \cellcolor[rgb]{0.9990234375,0.8583984375,0.75} 6.50294x& \cellcolor[rgb]{0.9990234375,0.8544921875,0.75} 6.61920x\\
-8192 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.7724609375,0.9990234375,0.75} 1.09357x& \cellcolor[rgb]{0.8095703125,0.9990234375,0.75} 1.24057x& \cellcolor[rgb]{0.7744140625,0.9990234375,0.75} 1.10146x& \cellcolor[rgb]{0.7783203125,0.9990234375,0.75} 1.11378x& \cellcolor[rgb]{0.9990234375,0.80859375,0.75} 8.09302x& \cellcolor[rgb]{0.9990234375,0.7900390625,0.75} 8.69270x& \cellcolor[rgb]{0.9990234375,0.783203125,0.75} 8.92530x\\
-16384 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.7763671875,0.9990234375,0.75} 1.10882x& \cellcolor[rgb]{0.7822265625,0.9990234375,0.75} 1.13274x& \cellcolor[rgb]{0.7763671875,0.9990234375,0.75} 1.10937x& \cellcolor[rgb]{0.78515625,0.9990234375,0.75} 1.14416x& \cellcolor[rgb]{0.9990234375,0.7548828125,0.75} 9.84245x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 10.69728x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 10.80864x\\
+arg(s) & clang & simple 12 & simple 2 & briggs 12 & briggs 2 & linear & greedy 12 & greedy 0 \\
+\hline
+128 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 159130& \cellcolor[rgb]{0.7509765625,0.9990234375,0.75} 1.00777x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00293x& \cellcolor[rgb]{0.7509765625,0.9990234375,0.75} 1.00644x& \cellcolor[rgb]{0.7529296875,0.9990234375,0.75} 1.01562x& \cellcolor[rgb]{0.8173828125,0.9990234375,0.75} 1.27104x& \cellcolor[rgb]{0.82421875,0.9990234375,0.75} 1.29856x& \cellcolor[rgb]{0.82421875,0.9990234375,0.75} 1.29894x\\
+256 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 163800& \cellcolor[rgb]{0.7509765625,0.9990234375,0.75} 1.00726x& \cellcolor[rgb]{0.7529296875,0.9990234375,0.75} 1.01200x& \cellcolor[rgb]{0.751953125,0.9990234375,0.75} 1.00852x& \cellcolor[rgb]{0.7529296875,0.9990234375,0.75} 1.01440x& \cellcolor[rgb]{0.8818359375,0.9990234375,0.75} 1.53225x& \cellcolor[rgb]{0.89453125,0.9990234375,0.75} 1.58048x& \cellcolor[rgb]{0.8974609375,0.9990234375,0.75} 1.59431x\\
+512 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 172096& \cellcolor[rgb]{0.75390625,0.9990234375,0.75} 1.01902x& \cellcolor[rgb]{0.755859375,0.9990234375,0.75} 1.02488x& \cellcolor[rgb]{0.75390625,0.9990234375,0.75} 1.01789x& \cellcolor[rgb]{0.75390625,0.9990234375,0.75} 1.01876x& \cellcolor[rgb]{0.9990234375,0.9970703125,0.75} 2.03762x& \cellcolor[rgb]{0.9990234375,0.994140625,0.75} 2.13731x& \cellcolor[rgb]{0.9990234375,0.9931640625,0.75} 2.15792x\\
+1024 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 188591& \cellcolor[rgb]{0.7587890625,0.9990234375,0.75} 1.03687x& \cellcolor[rgb]{0.7587890625,0.9990234375,0.75} 1.03706x& \cellcolor[rgb]{0.7568359375,0.9990234375,0.75} 1.02964x& \cellcolor[rgb]{0.7587890625,0.9990234375,0.75} 1.03689x& \cellcolor[rgb]{0.9990234375,0.9697265625,0.75} 2.93413x& \cellcolor[rgb]{0.9990234375,0.9638671875,0.75} 3.09867x& \cellcolor[rgb]{0.9990234375,0.962890625,0.75} 3.15326x\\
+2048 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 221875& \cellcolor[rgb]{0.76171875,0.9990234375,0.75} 1.04971x& \cellcolor[rgb]{0.765625,0.9990234375,0.75} 1.06586x& \cellcolor[rgb]{0.76171875,0.9990234375,0.75} 1.04806x& \cellcolor[rgb]{0.765625,0.9990234375,0.75} 1.06306x& \cellcolor[rgb]{0.9990234375,0.92578125,0.75} 4.34620x& \cellcolor[rgb]{0.9990234375,0.916015625,0.75} 4.64145x& \cellcolor[rgb]{0.9990234375,0.9140625,0.75} 4.72466x\\
+4096 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 288132& \cellcolor[rgb]{0.771484375,0.9990234375,0.75} 1.08649x& \cellcolor[rgb]{0.7744140625,0.9990234375,0.75} 1.09864x& \cellcolor[rgb]{0.76953125,0.9990234375,0.75} 1.08082x& \cellcolor[rgb]{0.7744140625,0.9990234375,0.75} 1.10033x& \cellcolor[rgb]{0.9990234375,0.8662109375,0.75} 6.25886x& \cellcolor[rgb]{0.9990234375,0.8525390625,0.75} 6.69614x& \cellcolor[rgb]{0.9990234375,0.84765625,0.75} 6.84597x\\
+8192 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 424179& \cellcolor[rgb]{0.7763671875,0.9990234375,0.75} 1.10601x& \cellcolor[rgb]{0.78125,0.9990234375,0.75} 1.12721x& \cellcolor[rgb]{0.7763671875,0.9990234375,0.75} 1.10683x& \cellcolor[rgb]{0.78125,0.9990234375,0.75} 1.12700x& \cellcolor[rgb]{0.9990234375,0.8037109375,0.75} 8.26470x& \cellcolor[rgb]{0.9990234375,0.78515625,0.75} 8.85856x& \cellcolor[rgb]{0.9990234375,0.7783203125,0.75} 9.07318x\\
+16384 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 697378& \cellcolor[rgb]{0.78125,0.9990234375,0.75} 1.12603x& \cellcolor[rgb]{0.787109375,0.9990234375,0.75} 1.15191x& \cellcolor[rgb]{0.78125,0.9990234375,0.75} 1.12581x& \cellcolor[rgb]{0.787109375,0.9990234375,0.75} 1.15269x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 10.00521x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 10.73828x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 10.98249x\\
 \hline
 \end{NiceTabular}
-\caption{\label{demo-table}Benchmark of \texttt{benches/fib.ll} output by \mintinline{bash}{dune exec bench -- -f fib -n 1000}}
+\caption{\label{demo-table}Benchmark of \texttt{benches/sieven.ll} output by \mintinline{bash}{dune exec bench -- -f fib -n 1000}}
 \end{center}
@@
 
+\subsubsection{\lstinline!benches/sha256.ll!}
 \begin{NiceTabular}{|c|cccccccc|}

### 2023-12-28 article.tex
Pre-jane

@@
 In compiler backends, \textit{Static Single-Assignment} (SSA) form is a property that applies to some IR, including LLVM.
-It is a way of representing a program in such a way that each variable is assigned only once throughout
-the scope of the function. Each use of the variable after definition then refers to the value  of that initial assignment.
+It is a way of representing a program such that each variable is assigned only once throughout
+the scope of the function.
+Each use of the variable after definition then refers to the value  of that  single assignment, and any operations are applied by assigning the result of said operation to a freshly defined variable.
+
+%Obviously, that 
 %In other words, every variable is assigned a unique version number for each assignment, and these versions are used consistently throughout the program.

### 2023-12-28 article.tex
Before send

@@
 which consists of an optionally named entry block in the first part and a list of trailing blocks that are always named in the second.
+The reason for the first block to be optionally named is the fact that \textit{some} LLVM programs transfer control back to the point of entry, whereas every subsequent block needs to be named because branching can only target it by its name.
 
@@
 
-Each function defined in an LLVM program is constructed with the help of
+%Each function defined in an LLVM program is constructed with the help of
 
@@
 
-The intermediate representation emitted by the frontend of a compiler serves as a stepping stone independent of the target architecture.
+%The intermediate representation emitted by the frontend of a compiler serves as a stepping stone independent of the target architecture.
 The \lstinline!LLVM!
@@
 
-Because of this, these instructions are referred to as \textit{terminators}, as their purpose is to disrupt what had otherwise been a linear flow from the beginning of this continus sequence of instructions, henceforth referred to as a \textit{basic block}.
+%Because of this, these instructions are referred to as \textit{terminators}, as their purpose is to disrupt what had otherwise been a linear flow from the beginning of this continus sequence of instructions, henceforth referred to as a \textit{basic block}.
+
+\label{appendix:llvmis}
+\begin{enumerate}
+    \item binary operations \lstinline!add!, \lstinline!and!, \lstinline!ashr!, \lstinline!lshr!, \lstinline!mul!, \lstinline!or!, \lstinline!sdiv!,\\\lstinline!srem!, \lstinline!shl!, \lstinline!sub!
+and \lstinline!xor!
+    \item integer comparison \lstinline!icmp! (conditions being \lstinline!eq!, \lstinline!ne!, \lstinline!sge!, \lstinline!sgt!, \lstinline!sle! and \lstinline!slt!)
+    \item memory/address operations \lstinline!alloca!, \lstinline!gep!, \lstinline!load! and \lstinline!store!
+    \item  \lstinline!mov! operations \lstinline!gep!, \lstinline!phi!, \lstinline!ptrtoint!, \lstinline!trunc!, and \lstinline!zext!
+    \item control flow operations \lstinline!br!, \lstinline!call! and \lstinline!ret!
+    \item a  nullary block terminator \lstinline!unreachable!
+\end{enumerate}
 
@@
 \end{center}
+
+\subsubsection{Implementing \mintinline{llvm}{call} operation}
+
+The \mintinline{llvm}{call} operation is by far the most complicated one. Mostly because of calling conventions,
+The \mintinline{gas}{%rax} register is zeroed before
 
@@
 
+\subsubsection{\lstinline!benches/fannkuch-redux.ll!}
+\begin{table}[!ht]
+\begin{center}
+\begin{NiceTabular}{|c|cccccccc|}
+\hline
+arg(s) & clang & simple 12 & simple 2 & briggs 12 & briggs 2 & linear & greedy 12 & greedy 0 \\
+\hline
+4 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 172157& \cellcolor[rgb]{0.759765625,0.9990234375,0.75} 1.04312x& \cellcolor[rgb]{0.7568359375,0.9990234375,0.75} 1.02806x& \cellcolor[rgb]{0.751953125,0.9990234375,0.75} 1.00936x& \cellcolor[rgb]{0.767578125,0.9990234375,0.75} 1.07378x& \cellcolor[rgb]{0.806640625,0.9990234375,0.75} 1.23096x& \cellcolor[rgb]{0.7998046875,0.9990234375,0.75} 1.20263x& \cellcolor[rgb]{0.810546875,0.9990234375,0.75} 1.24619x\\
+5 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 277849& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.71875x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.74478x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.72175x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.74462x& \cellcolor[rgb]{0.7978515625,0.9990234375,0.75} 1.19407x& \cellcolor[rgb]{0.794921875,0.9990234375,0.75} 1.18339x& \cellcolor[rgb]{0.8017578125,0.9990234375,0.75} 1.20806x\\
+6 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 307924& \cellcolor[rgb]{0.8125,0.9990234375,0.75} 1.25112x& \cellcolor[rgb]{0.82421875,0.9990234375,0.75} 1.30051x& \cellcolor[rgb]{0.810546875,0.9990234375,0.75} 1.24357x& \cellcolor[rgb]{0.830078125,0.9990234375,0.75} 1.32219x& \cellcolor[rgb]{0.9990234375,0.9423828125,0.75} 3.80349x& \cellcolor[rgb]{0.9990234375,0.9404296875,0.75} 3.86493x& \cellcolor[rgb]{0.9990234375,0.94140625,0.75} 3.84259x\\
+7 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1312806& \cellcolor[rgb]{0.87890625,0.9990234375,0.75} 1.52113x& \cellcolor[rgb]{0.9052734375,0.9990234375,0.75} 1.62393x& \cellcolor[rgb]{0.8779296875,0.9990234375,0.75} 1.51710x& \cellcolor[rgb]{0.9052734375,0.9990234375,0.75} 1.62624x& \cellcolor[rgb]{0.9990234375,0.86328125,0.75} 6.35235x& \cellcolor[rgb]{0.9990234375,0.8515625,0.75} 6.70982x& \cellcolor[rgb]{0.9990234375,0.8642578125,0.75} 6.31152x\\
+8 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 11168259& \cellcolor[rgb]{0.900390625,0.9990234375,0.75} 1.60627x& \cellcolor[rgb]{0.9189453125,0.9990234375,0.75} 1.67856x& \cellcolor[rgb]{0.947265625,0.9990234375,0.75} 1.79351x& \cellcolor[rgb]{0.9140625,0.9990234375,0.75} 1.66070x& \cellcolor[rgb]{0.9990234375,0.8349609375,0.75} 7.25991x& \cellcolor[rgb]{0.9990234375,0.833984375,0.75} 7.29830x& \cellcolor[rgb]{0.9990234375,0.83984375,0.75} 7.09878x\\
+9 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 115754944& \cellcolor[rgb]{0.900390625,0.9990234375,0.75} 1.60476x& \cellcolor[rgb]{0.9296875,0.9990234375,0.75} 1.72352x& \cellcolor[rgb]{0.9052734375,0.9990234375,0.75} 1.62536x& \cellcolor[rgb]{0.9208984375,0.9990234375,0.75} 1.68769x& \cellcolor[rgb]{0.9990234375,0.828125,0.75} 7.46779x& \cellcolor[rgb]{0.9990234375,0.8173828125,0.75} 7.81576x& \cellcolor[rgb]{0.9990234375,0.83203125,0.75} 7.36417x\\
+10 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1352747472& \cellcolor[rgb]{0.8857421875,0.9990234375,0.75} 1.54700x& \cellcolor[rgb]{0.935546875,0.9990234375,0.75} 1.74875x& \cellcolor[rgb]{0.89453125,0.9990234375,0.75} 1.58381x& \cellcolor[rgb]{0.9169921875,0.9990234375,0.75} 1.67440x& \cellcolor[rgb]{0.9990234375,0.818359375,0.75} 7.77269x& \cellcolor[rgb]{0.9990234375,0.8193359375,0.75} 7.74449x& \cellcolor[rgb]{0.9990234375,0.8173828125,0.75} 7.82170x\\
+11 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 16942071879& \cellcolor[rgb]{0.88671875,0.9990234375,0.75} 1.55031x& \cellcolor[rgb]{0.9267578125,0.9990234375,0.75} 1.71033x& \cellcolor[rgb]{0.8876953125,0.9990234375,0.75} 1.55567x& \cellcolor[rgb]{0.9228515625,0.9990234375,0.75} 1.69492x& \cellcolor[rgb]{0.9990234375,0.814453125,0.75} 7.92913x& \cellcolor[rgb]{0.9990234375,0.8017578125,0.75} 8.32639x& \cellcolor[rgb]{0.9990234375,0.8212890625,0.75} 7.70198x\\
+12 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 235117543385& \cellcolor[rgb]{0.8974609375,0.9990234375,0.75} 1.59287x& \cellcolor[rgb]{0.923828125,0.9990234375,0.75} 1.70029x& \cellcolor[rgb]{0.8798828125,0.9990234375,0.75} 1.52228x& \cellcolor[rgb]{0.927734375,0.9990234375,0.75} 1.71456x& \cellcolor[rgb]{0.9990234375,0.8095703125,0.75} 8.05498x& \cellcolor[rgb]{0.9990234375,0.8095703125,0.75} 8.08426x& \cellcolor[rgb]{0.9990234375,0.806640625,0.75} 8.17452x\\
+\hline
+\end{NiceTabular}
+\caption{\label{tab:fannkuch}Benchmark of \texttt{benches/sieven.ll} output by \mintinline{bash}{dune exec bench -- -f fib -n 1000}}
+\end{center}
+\end{table}
+
 \subsection{Comparison to other work and ideas for future work}
@@
 %%% NOTE: Don't include subsections of tthe benchmarks section
-\addtocontents{toc}{\protect\setcounter{tocdepth}{1}}
-
-\section{LLVM-- instruction set}
-\label{appendix:llvmis}
-\begin{enumerate}
-    \item binary operations \lstinline!add!, \lstinline!and!, \lstinline!ashr!, \lstinline!lshr!, \lstinline!mul!, \lstinline!or!, \lstinline!sdiv!,\\\lstinline!srem!, \lstinline!shl!, \lstinline!sub!
-and \lstinline!xor!
-    \item integer comparison \lstinline!icmp! (conditions being \lstinline!eq!, \lstinline!ne!, \lstinline!sge!, \lstinline!sgt!, \lstinline!sle! and \lstinline!slt!)
-    \item memory/address operations \lstinline!alloca!, \lstinline!gep!, \lstinline!load! and \lstinline!store!
-    \item  \lstinline!mov! operations \lstinline!gep!, \lstinline!phi!, \lstinline!ptrtoint!, \lstinline!trunc!, and \lstinline!zext!
-    \item control flow operations \lstinline!br!, \lstinline!call! and \lstinline!ret!
-    \item a  nullary block terminator \lstinline!unreachable!
-\end{enumerate}
+%\addtocontents{toc}{\protect\setcounter{tocdepth}{1}}
+
 \end{document}

### 2023-12-30 article.tex
Jane approved

--- 2023-12-28 article.tex
@@
 Compilation refers to the process of translating from one language to another, most often from a high-level programming language intended for humans to work with, to machine- or bytecode intended to be executed on a target architecture.
-This process can be divided into several distinct phases, which are grouped into one of two stages colloquially referred to as the \textit{frontend} and \textit{backend}, the former translating a high-level programming language to an \textit{intermediate representation} (IR) and the latter translating IR to executable machine code of a target architecture or bytecode of a target \textit{virtual machine} (VM).
+This process can be divided into several distinct phases, which are grouped into one of two stages colloquially referred to as the \textit{frontend} and \textit{backend}, The former is translating a high-level programming language to an \textit{intermediate representation} (IR) and the latter is translating IR to executable machine code of a target architecture or bytecode of a target \textit{virtual machine} (VM).
 
@@
 
-Although compilers can feasibly translate from the source language to machine code of the target platform directly, and doing so could even be more efficient, doing so hinders the portability as machine code targeting architecture \(A\) isn't necessarily useful for targeting architecture \(B\) further down the line.
+Although compilers can feasibly translate from the source language to machine code of the target architecture directly, which could even be more efficient, doing so hinders the portability as machine code targeting architecture \(A\) isn't necessarily useful for targeting architecture \(B\) further down the line.
 
@@
 lol
-It is by no means executable on either \(A\) or \(B\) but is more often than not much closer to the abstract level of a CPU than the source language originally fed to the frontend.
+It is by no means executable on either \(A\) or \(B\), but it is much closer to the instruction set executed on a CPU than the source language originally fed to the frontend.
 This helps portability immensely, as a frontend emitting an IR no longer needs to specialize to a particular architecture, instead it can target as many as the backend supports.
@@
 
-Conversely, managed languages such as Java and C\# typically don't depend on the LLVM toolchain directly, research projects that attempt to compile associated VM bytecode to a native platform using LLVM (\cite{dotnet}).
-Meanwhile interpreted languages like JavaScript eagerly attempt use LLVM for \textit{just-in-time} (JIT) compilation % FIXME: this is not true at all lol
-
-The modularization of the frontend and backend stages also help immensely.
+Conversely, managed languages such as Java and C\# typically don't depend on the LLVM toolchain directly, some research projects attempt to compile associated VM bytecode to a native platform using LLVM (\cite{dotnet}).
+%Meanwhile interpreted languages like JavaScript eagerly attempt use LLVM for \textit{just-in-time} (JIT) compilation % FIXME: this is not true at all lol
+
+The modularization of the frontend and backend stages also helps immensely.
 
@@
 
+\subsubsection{Implementing \mintinline{llvm}{call} operation}
+
+The \mintinline{llvm}{call} operation is by far the most complicated one. Mostly because of calling conventions,
+The \mintinline{gas}{%rax} register is zeroed before
+
 %* safest possible x86 translation
@@
 \subsubsection{Debugger Harness}
-A more elaborate way of testing correctness of translation is to assert that there are no effects outside of those intended.
+A more elaborate way of testing correctness of translation is to assert that there are no effects except for those intended.
 One way of achieving this is to check if a group of x86 instructions representing a single LLVM-- operation change any values except those expected.
@@
 type before the assembly emitted of that specific operation.
-In this a unique identifier as well as a list changed expected to be made is encoded.
-This includes a bitmask of GPRs that can possibly be changed (aside from the scrath registers of couse) as well as which parts of the stack may be modified. % TODO: implement the stack checker
+A unique identifier as well as a list of changes expected to be made is encoded.
+This includes a bitmask of GPRs that can possibly be changed (aside from the scratch registers of course) as well as which parts of the stack may be modified. % TODO: implement the stack checker
 In addition to this, each of the labels are output with a \lstinline!.globl! directive in the file header. A file is built with these extra symbols using the \lstinline!-d! flag.
@@
 \end{table}
-The above measures are significantly equivalent to one another, with all allocators (except greedy 0) seeming to converge towards being 37\% slower than Clang.
+
+The measures in \ref{tab:dumb-fib}  are significantly equivalent to one another, with all allocators (except greedy 0) seeming to converge towards being 37\% slower than Clang.
 This isn't particularly surprising, as the x86 translation is more concerned with preserving callee saved registers than necessary in pushing each and every one despite not being in use.
@@
 function that has 7 variables can assign each of them to a register for the first six allocators listed.
-However, the last one (greedy 0), has no assignable registers, so is forced to start spilling from the very beginning.
-To illusidate this hypothesis, here is the same benchmark but only for the greedy of decreasing \(k\) assignable registers:
+However, the last one (greedy 0), has no assignable registers, so it must necessarily start spilling from the very beginning.
+To elucidate this hypothesis, here is the same benchmark but only for the greedy of decreasing \(k\) assignable registers:
 \begin{table}[!ht]
@@
 
+\subsubsection{\lstinline!benches/fannkuch-redux.ll!}
+\begin{table}[!ht]
+\begin{center}
+\begin{NiceTabular}{|c|cccccccc|}
+\hline
+arg & clang & simple 12 & simple 2 & briggs 12 & briggs 2 & linear & greedy 12 & greedy 0 \\
+\hline
+4 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 197975& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.93202x& \cellcolor[rgb]{0.771484375,0.9990234375,0.75} 1.08772x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.94359x& \cellcolor[rgb]{0.84375,0.9990234375,0.75} 1.37648x& \cellcolor[rgb]{0.7900390625,0.9990234375,0.75} 1.16456x& \cellcolor[rgb]{0.771484375,0.9990234375,0.75} 1.09012x& \cellcolor[rgb]{0.7861328125,0.9990234375,0.75} 1.14835x\\
+5 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 193800& \cellcolor[rgb]{0.7626953125,0.9990234375,0.75} 1.05122x& \cellcolor[rgb]{0.7744140625,0.9990234375,0.75} 1.10079x& \cellcolor[rgb]{0.7685546875,0.9990234375,0.75} 1.07640x& \cellcolor[rgb]{0.77734375,0.9990234375,0.75} 1.11231x& \cellcolor[rgb]{0.9296875,0.9990234375,0.75} 1.72450x& \cellcolor[rgb]{0.9453125,0.9990234375,0.75} 1.78685x& \cellcolor[rgb]{0.931640625,0.9990234375,0.75} 1.73329x\\
+6 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 307184& \cellcolor[rgb]{0.8203125,0.9990234375,0.75} 1.28426x& \cellcolor[rgb]{0.8291015625,0.9990234375,0.75} 1.32022x& \cellcolor[rgb]{0.814453125,0.9990234375,0.75} 1.25973x& \cellcolor[rgb]{0.83203125,0.9990234375,0.75} 1.33286x& \cellcolor[rgb]{0.9990234375,0.9423828125,0.75} 3.81882x& \cellcolor[rgb]{0.9990234375,0.9404296875,0.75} 3.85798x& \cellcolor[rgb]{0.9990234375,0.939453125,0.75} 3.90989x\\
+7 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1312573& \cellcolor[rgb]{0.875,0.9990234375,0.75} 1.50533x& \cellcolor[rgb]{0.8896484375,0.9990234375,0.75} 1.56355x& \cellcolor[rgb]{0.8779296875,0.9990234375,0.75} 1.51619x& \cellcolor[rgb]{0.8828125,0.9990234375,0.75} 1.53672x& \cellcolor[rgb]{0.9990234375,0.861328125,0.75} 6.40226x& \cellcolor[rgb]{0.9990234375,0.8603515625,0.75} 6.44050x& \cellcolor[rgb]{0.9990234375,0.8642578125,0.75} 6.32867x\\
+8 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 11252155& \cellcolor[rgb]{0.900390625,0.9990234375,0.75} 1.60713x& \cellcolor[rgb]{0.908203125,0.9990234375,0.75} 1.63817x& \cellcolor[rgb]{0.8818359375,0.9990234375,0.75} 1.52961x& \cellcolor[rgb]{0.90234375,0.9990234375,0.75} 1.61191x& \cellcolor[rgb]{0.9990234375,0.8349609375,0.75} 7.25974x& \cellcolor[rgb]{0.9990234375,0.8349609375,0.75} 7.25458x& \cellcolor[rgb]{0.9990234375,0.833984375,0.75} 7.28693x\\
+9 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 115513112& \cellcolor[rgb]{0.8935546875,0.9990234375,0.75} 1.57771x& \cellcolor[rgb]{0.904296875,0.9990234375,0.75} 1.62155x& \cellcolor[rgb]{0.888671875,0.9990234375,0.75} 1.55882x& \cellcolor[rgb]{0.9990234375,0.98046875,0.75} 2.57223x& \cellcolor[rgb]{0.9990234375,0.82421875,0.75} 7.60550x& \cellcolor[rgb]{0.9990234375,0.8154296875,0.75} 7.87522x& \cellcolor[rgb]{0.9990234375,0.8212890625,0.75} 7.69289x\\
+10 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1352747472& \cellcolor[rgb]{0.8857421875,0.9990234375,0.75} 1.54700x& \cellcolor[rgb]{0.935546875,0.9990234375,0.75} 1.74875x& \cellcolor[rgb]{0.89453125,0.9990234375,0.75} 1.58381x& \cellcolor[rgb]{0.9169921875,0.9990234375,0.75} 1.67440x& \cellcolor[rgb]{0.9990234375,0.818359375,0.75} 7.77269x& \cellcolor[rgb]{0.9990234375,0.8193359375,0.75} 7.74449x& \cellcolor[rgb]{0.9990234375,0.8173828125,0.75} 7.82170x\\
+\hline
+\end{NiceTabular}
+\caption{\label{tab:fannkuch}Benchmark of \texttt{benches/sieven.ll} output by \mintinline{bash}{dune exec bench -- -f fib -n 1000}}
+\end{center}
+\end{table}
+
 \subsection{Comparison to other work and ideas for future work}

### 2024-01-01 article.tex
More sources

@@
 \documentclass{article}
-\usepackage[style=apa]{biblatex}
+\usepackage[style=ieee]{biblatex}
 
@@
 Because of this, compilation of high-level programming languages will often go no further than emitting IR, then leave the rest for a subsequent backend toolchain.
-This naturally leads to the LLVM toolchain, which is by far the most used compiler backend.
+This naturally leads to the LLVM toolchain, which is by far the most used compiler backend in practice.
 Most languages, if targeting native execution on a CPU, will have an LLVM implementation.
-This is true for C and C++ (clang and clang++ respectively), D (with the ldc compiler), Swift and Rust, to name a few.
-
-Conversely, managed languages such as Java and C\# typically don't depend on the LLVM toolchain directly, some research projects attempt to compile associated VM bytecode to a native platform using LLVM (\cite{dotnet}).
+This is true for C/C++ (Clang compiler \cite{clang}), D (LDC compiler \cite{dlang}), Swift (official Swift compiler \cite{swift}) and Rust (officlal \texttt{rustc} compiler \cite{rustc}).
+The use of LLVM allows for them to be compiled to completely unforeseen architectures, like web browsers with WebAssembly or even graphics cards with compute kernels \cite{backend}.
+
+Conversely, managed languages such as Java and C\# target their own respective VMs, so typically don't have much to gain from the LLVM toolchain.
+That said, some projects have attempted to implement LLVM in their translation of bytecode to native architectures,  such as LLILC which promised both \textit{just-in-time} (JIT) as well as \textit{ahead-of-time} (AOT) compilation \cite{dotnet}.
+Similarly, Kotlin (the spiritual successor to Java) that ordinarily targets the JVM has a native backend that compiles directly to machine code using LLVM, circumventing the need for a VM entirely \cite{kotlin}.
 %Meanwhile interpreted languages like JavaScript eagerly attempt use LLVM for \textit{just-in-time} (JIT) compilation % FIXME: this is not true at all lol
@@
      \begin{minted}[bgcolor=LightGray]{llvm}
-define i32 @add(i32 %a, i32 %b) {
-  %sum = add i32 %a, %b
-  ret i32 %sum
+define i32 @add (i32 %0, i32 %1) {
+ %3 = add i32 %1, %0
+ ret i32 %3
 }
      \end{minted}
-     \caption{Corresponding represented in LLVM}\label{fig:addll}
+     \caption{Stripped \texttt{clang -O1 -S -emit-llvm}}\label{fig:addll}
    \end{minipage}
    \begin{center}
-   \begin{minipage}{0.48\textwidth}
+   \begin{minipage}{0.83\textwidth}
      \centering
      \begin{minted}[bgcolor=LightGray]{gas}
-  .globl _add
-_add:
-  movl	%edi, %eax
-  addl	%esi, %eax
+	.section	__TEXT,__text,regular,pure_instructions
+	.build_version macos, 14, 0
+	.globl	_add              ## -- Begin function add
+	.p2align	4, 0x90
+_add:                                   ## @add
+	.cfi_startproc
+## %bb.0:
+	movl	%esi, %eax
+	addl	%edi, %eax
 	retq
+	.cfi_endproc
+                                        ## -- End function
+.subsections_via_symbols
      \end{minted}
@@
 Each use of the variable after definition then refers to the value  of that  single assignment, and any operations are applied by assigning the result of said operation to a freshly defined variable.
+
+This approach seeks to benefit
 
@@
 
+\begin{figure}[h]
 \begin{minted}[linenos,bgcolor=LightGray]{llvm}
@@
   %x1 = add i32 %x0, 1
+  call void @printInt(i32 %x1)
   call ptr @subproc()
@@
 \end{minted}
-Because of the halting problem, static analysis cannot determine if the call to \lstinline!@subproc!
-will return for every possible implementation.
-%which may or may not return control to the calling function:
-So when assigning \lstinline!%x1!
-a register it is undecidable whether the variable must be live in the last return instruction, i.e.
-needs to live across the call to another.
+
+\caption{\label{fig:undecidable}Increment, print and possibly return argument}
+\end{figure}
+\noindent Because of the halting problem, static analysis cannot necessarily determine if the call to \lstinline!@subproc! will return.  %which may or may not return control to the calling function:
+So when assigning \lstinline!%x1! a register it is undecidable whether the variable must be live in the return instruction on line 5 in Figure \ref{fig:undecidable}.
+A register must be assigned to \texttt{\%x1}, as it is used on line 3, although whether it must live across function calls can reduce the possible registers depending on the target platform and its calling conventions.
+% TODO: kan udvides
 Because of this, any sound approach to liveness analysis will be an approximation.
@@
 %can be thought of as a \textit{transfer function} in terms of liveness, denoting which variables may be live before and after execution.
-Initially, two sets are associated with each instruction: the \textit{live-in} and \textit{live-out} sets, which are the sets of variables that are live respectively before and after execution.
+Initially, two empty sets are associated with each instruction: the \textit{live-in} and \textit{live-out} sets, which are the sets of variables that are live respectively before and after execution.
 Then the two equations above are applied iteratively until a fixed point is reached, i.e.
@@
 \end{minted}
-With the actual fixed-point iteration performed as follows:
+With the actual dataflow analysis performed recursively as follows:
 \begin{minted}[linenos,bgcolor=LightGray]{ocaml}
@@
 \end{minted}
+To
 

### 2024-01-01 article.tex
More livenss analysis

@@
 Compilation refers to the process of translating from one language to another, most often from a high-level programming language intended for humans to work with, to machine- or bytecode intended to be executed on a target architecture.
-This process can be divided into several distinct phases, which are grouped into one of two stages colloquially referred to as the \textit{frontend} and \textit{backend}, The former is translating a high-level programming language to an \textit{intermediate representation} (IR) and the latter is translating IR to executable machine code of a target architecture or bytecode of a target \textit{virtual machine} (VM).
+This process can be divided into several distinct phases, which are grouped into one of two stages colloquially referred to as the \textit{frontend} and \textit{backend} (see Figure \ref{fig:phases}).
+The former is translating a high-level programming language to an \textit{intermediate representation} (IR) and the latter is translating IR to executable machine code of a target architecture or bytecode of a target \textit{virtual machine} (VM).
 
@@
 \end{tikzpicture}}
-\caption{Compiler phases, backend highlighted}
+\caption{Compiler phases, backend highlighted}\label{fig:phases}
 \end{figure}
-Most operations of a general-purpose programming language are translated to a set of control, logic, and arithmetic instructions to be executed sequentially on a computer processor: a single circuit/chip, referred to as the \textit{central processing unit} (CPU), the design of which has varied and evolved over time.
+\noindent Most operations of a general-purpose programming language are translated to a set of control, logic, and arithmetic instructions to be executed sequentially on a computer processor: a single circuit/chip, referred to as the \textit{central processing unit} (CPU), the design of which has varied and evolved over time.
 
@@
 This can largely be attributed to performance, as register machines routinely outperform \textit{stack machines} \parencite{ShiYunhe2008VmsS} that are often used in VMs.
-Although register machines generally also have a stack available, as being limited to mere bytes of storage is simply infeasible for most large scale applications, using it compared to GPRs is several orders of magnitudes slower.
+Although register machines generally also have a stack available, as being limited to mere bytes of storage is simply infeasible for most large scale applications, using it compared to GPRs is orders of magnitudes slower \cite{latency}.
+% FIXME: find source
 Because of this, a crucial part of the backend stage for an optimizing compiler is assigning each variable  of the source program to a GPR in such a way that maximizes performance without sacrificing correctness.
@@
 Most languages, if targeting native execution on a CPU, will have an LLVM implementation.
-This is true for C/C++ (Clang compiler \cite{clang}), D (LDC compiler \cite{dlang}), Swift (official Swift compiler \cite{swift}) and Rust (officlal \texttt{rustc} compiler \cite{rustc}).
+This is true for C/C++ (Clang compiler \cite{clang}), D (LDC compiler \cite{dlang}), Swift (official Swift compiler \cite{swift}) and Rust (officilal \texttt{rustc} compiler \cite{rustc}).
 The use of LLVM allows for them to be compiled to completely unforeseen architectures, like web browsers with WebAssembly or even graphics cards with compute kernels \cite{backend}.
@@
 
-The modularization of the frontend and backend stages also helps immensely.
+%The modularization of the frontend and backend stages also helps immensely.
+For example, the add function implemented in C as seen in Figure \ref{fig:addc} is compiled to the LLVM IR as seen in Figure \ref{fig:addll} after stripping optimization attributes with the \texttt{strip} utility.
+At this level of complexity they are largely equivalent, with the primary difference appearing to be syntactic.
+LLVM IR keeps the typed binary operations as well as the function construct with a return statement, although the variable names are discarded as IR is generally not for human interaction, except the function name, as this is used for linking with other object files.
 
@@
      \begin{minted}[bgcolor=LightGray]{gas}
-	.section	__TEXT,__text,regular,pure_instructions
-	.build_version macos, 14, 0
-	.globl	_add              ## -- Begin function add
-	.p2align	4, 0x90
-_add:                                   ## @add
-	.cfi_startproc
-## %bb.0:
+	.text
+	.globl	add               # -- Begin function add
+add:                                    # @add
 	movl	%esi, %eax
@@
 	retq
-	.cfi_endproc
-                                        ## -- End function
-.subsections_via_symbols
-     \end{minted}
-     \caption{Prettified result of \mintinline{bash}{clang -O0}}\label{fig:addx86}
+     \end{minted}
+     \caption{Output of \mintinline{bash}{clang -S} without debug markers}\label{fig:addx86}
    \end{minipage}
@@
 \end{figure}
+When the IR is translated to x86 as seen in Figure \ref{fig:addx86}, the virtual variables \texttt{\%0}, \texttt{\%1} and \texttt{\%2} are assigned the registers \texttt{\%esi}, \texttt{\%edi} and \texttt{\%eax} respectively.
+This is presumably because of several optimizations, as \texttt{\%rdi} and \texttt{\%rsi} are used to pass the first and second parameters and the \texttt{\%rax} as the return value according to the System V AMD64 ABI \cite{sysv}, so by assigning the variables to those, several \texttt{mov}s are saved.
+Either way, the assignment is sound and the instruction selection is reasonably close to the  source code.
+Obviously there are other changes as well, as marking the target section with \texttt{.text} and the \texttt{.globl} annotation to export \texttt{add} in the symbol table which permits linking with other object files, but that isn't stricly related to register allocation.
 
@@
 In compiler backends, \textit{Static Single-Assignment} (SSA) form is a property that applies to some IR, including LLVM.
-It is a way of representing a program such that each variable is assigned only once throughout
-the scope of the function.
+It is a way of representing a program such that each variable is assigned only once in
+the scope of a function.
 Each use of the variable after definition then refers to the value  of that  single assignment, and any operations are applied by assigning the result of said operation to a freshly defined variable.
 
-This approach seeks to benefit
+This property eliminates redefinitions entirely, seeking to simplify analysis and in turn also speed up analysis.
+Although, this property also 
+\begin{figure}[!ht]
+   \begin{minipage}{0.48\textwidth}
+     \centering
+     \begin{minted}[bgcolor=LightGray]{c}
+extern int add(int a, int b);
+
+int mul(int a, int b) {
+  int product = 0;
+  for (int i = 0; i < a; i++)
+    product = add(product, b);
+  return product;
+}
+     \end{minted}
+     \caption{Multiplication function implemented in C}\label{fig:mulc}
+   \end{minipage}\hfill
+   \begin{minipage}{0.48\textwidth}
+     \centering
+     \begin{minted}[bgcolor=LightGray]{llvm}
+declare i32 @add(i32, i32)
+
+define i32 @mul (i32 %0, i32 %1) {
+ %3 = icmp sgt i32 %0, 0
+ br i1 %3, label %6, label %4
+4:
+ %5 = phi i32 [0, %2], [%9, %6]
+ ret i32 %5
+6:
+ %7 = phi i32 [%10, %6], [0, %2]
+ %8 = phi i32 [%9, %6], [0, %2]
+ %9 = call i32 @add (i32 %8, i32 %1)
+ %10 = add i32 %7, 1
+ %11 = icmp eq i32 %10, %0
+ br i1 %11, label %4, label %6
+}
+     \end{minted}
+     \caption{Stripped \texttt{clang -O1 -S -emit-llvm}}\label{fig:mulll}
+   \end{minipage}
+   \begin{center}
+   \begin{minipage}{0.83\textwidth}
+     \centering
+     \begin{minted}[bgcolor=LightGray]{gas}
+	.text
+	.globl	add               # -- Begin function add
+add:                                    # @add
+	movl	%esi, %eax
+	addl	%edi, %eax
+	retq
+     \end{minted}
+     \caption{Output of \mintinline{bash}{clang -S} without debug markers}\label{fig:addx86}
+   \end{minipage}
+   \end{center}
+\end{figure}
 

### 2024-01-02 article.tex
ok

@@
 
-This property eliminates redefinitions entirely, seeking to simplify analysis and in turn also speed up analysis.
-Although, this property also 
+This property eliminates redefinitions entirely, making it easier to reason about data flow and in turn also speed up analysis.
+It's important to note that SSA form itself doesn't inherently mandate immutability, despite only allowing for single assignment.
+Rather, it provides two approaches to mutable variables: the phi nodes and load/store operations.
+
+As an extension of the earlier example, consider the implementation of multiplication in Figure \ref{fig:mulc}.
+Translating to LLVM IR with no optimization  can be seen in Figure \ref{fig:mulll0}, which achieves mutability by allocating a variable on the stack (lines 2-5), loading the value on use and storing on reassignment.
+This means the new value is loaded and assigned upon reentering a block, so any changes made at runtime by storing new values will take effect.
+
+%Although, this property also However, it's important to note that SSA form itself doesn't inherently introduce mutability;
 \begin{figure}[!ht]
@@
   int product = 0;
-  for (int i = 0; i < a; i++)
+
+  for (int i = 0; i < a; i++) {
     product = add(product, b);
+  }
+
   return product;
@@
      \caption{Multiplication function implemented in C}\label{fig:mulc}
-   \end{minipage}\hfill
-   \begin{minipage}{0.48\textwidth}
-     \centering
      \begin{minted}[bgcolor=LightGray]{llvm}
@@
      \end{minted}
-     \caption{Stripped \texttt{clang -O1 -S -emit-llvm}}\label{fig:mulll}
+     \caption{Stripped \texttt{clang -O1 -S -emit-llvm}}\label{fig:mulll1}
+   \end{minipage}\hfill
+   \begin{minipage}{0.48\textwidth}
+     \centering
+     \begin{minted}[bgcolor=LightGray]{llvm}
+define i32 @mul (i32 %0, i32 %1) {
+ %3 = alloca i32
+ %4 = alloca i32
+ %5 = alloca i32
+ %6 = alloca i32
+ store i32 %0, i32* %3
+ store i32 %1, i32* %4
+ store i32 0, i32* %5
+ store i32 0, i32* %6
+ br label %7
+7:
+ %8 = load i32, i32* %6
+ %9 = load i32, i32* %3
+ %10 = icmp slt i32 %8, %9
+ br i1 %10, label %11, label %18
+11:
+ %12 = load i32, i32* %5
+ %13 = load i32, i32* %4
+ %14 = call i32 @add (i32 %12, i32 %13)
+ store i32 %14, i32* %5
+ br label %15
+15:
+ %16 = load i32, i32* %6
+ %17 = add i32 %16, 1
+ store i32 %17, i32* %6
+ br label %7
+18:
+ %19 = load i32, i32* %5
+ ret i32 %19
+}
+     \end{minted}
+     \caption{Stripped \texttt{clang -O0 -S -emit-llvm}}\label{fig:mulll0}
    \end{minipage}
-   \begin{center}
-   \begin{minipage}{0.83\textwidth}
-     \centering
-     \begin{minted}[bgcolor=LightGray]{gas}
-	.text
-	.globl	add               # -- Begin function add
-add:                                    # @add
-	movl	%esi, %eax
-	addl	%edi, %eax
-	retq
-     \end{minted}
-     \caption{Output of \mintinline{bash}{clang -S} without debug markers}\label{fig:addx86}
-   \end{minipage}
-   \end{center}
 \end{figure}

### 2024-01-02 article.tex
describe phi nodes

@@
 
-As an extension of the earlier example, consider the implementation of multiplication in Figure \ref{fig:mulc}.
-Translating to LLVM IR with no optimization  can be seen in Figure \ref{fig:mulll0}, which achieves mutability by allocating a variable on the stack (lines 2-5), loading the value on use and storing on reassignment.
-This means the new value is loaded and assigned upon reentering a block, so any changes made at runtime by storing new values will take effect.
+Building on the earlier example, consider multiplication as  implemented in Figure \ref{fig:mulc}.
+Translated to LLVM IR with no optimization  can be seen in Figure \ref{fig:mulll0}, which achieves mutability by allocating a variable on the stack (lines 2-5), loading the value on use and storing on reassignment.
+This means new values are loaded and assigned upon reentering a block, without reassigning the value of the source variable, which points to whatever slot was available at the time of definition.
+%so any changes made at runtime by storing new values will take effect, without reassigning the source variable as it remains the address of whatever slot may be allocated on the stack.
 
@@
   int product = 0;
-
   for (int i = 0; i < a; i++) {
@@
   }
-
   return product;
@@
      \caption{Multiplication function implemented in C}\label{fig:mulc}
+     \vspace{2.1em}
      \begin{minted}[bgcolor=LightGray]{llvm}
@@
 \end{figure}
+Another approach is that of Figure \ref{fig:mulll1}, which uses so-called 'phi nodes' instead.
+These are a type of pseudo instruction that evaluate to a specified operand depending on which predecessor block was executed immediately before.
+This circumvents the need for the same level of interior mutability as is achieved by reading/writing to memory, without violating the properties of SSA.
+By copying a value from the end of a predecessor to the beginning of the current block.
+A Phi node represents a point in the program where the control flow merges, and it selects the appropriate version of a variable based on the path taken.
+This ensures that the data flow is well-defined and allows for easy analysis across different control flow paths.
 

### 2024-01-02 article.tex
Center tables on page

@@
 \usepackage{booktabs}
+\usepackage{float}
 
@@
 It is a directed graph whose edges represent transfer of control.
-The type of node varies over the source material, with CFGs of the Appel text constructed over individual instructions (\cite{tiger}) and the Aho text over basic blocks (\cite{dragon}).
-Either way, leaders have a set of predecessors and terminators a set of successors associated with them.
+The type of node varies over the source material, with CFGs of the Appel text \cite{tiger} constructed over individual instructions  and the Aho text \cite{dragon} over basic blocks.
+Either way, leaders will have a set of predecessors and terminators a set of successors associated with them.
 
 Terminators have one or two successors in the case of branching or none at all in the case of function exit. % These can either be associated basic blocks or individual instructions that make it up.
-Unconditional branching always transfers control to the block labelled, meaning only one successor follows, whereas conditional branching could transfer to either of the two, but because control flow analysis is not concerned with data, both are considered as possible successors.
+Unconditional branching always transfers control to the block labelled, meaning only one successor follow, whereas conditional branching could transfer to either of the two, but because control flow analysis is not concerned with data, both are considered as possible successors.
 % TODO: rets
@@
 The 'dumb' fib is a very naive approach to finding the \(n\)th number of the Fibonacci sequence by calling itself recursively twice (decrementing \(n\) before each), which causes an exponential growth in function calls.
-\begin{table}[!ht]
-\begin{center}
+\begin{table}[H]
+\makebox[\textwidth][c]{
 \begin{NiceTabular}{|c|cccccccc|}
@@
 \hline
-\end{NiceTabular}
+\end{NiceTabular}}
 \caption{\label{tab:dumb-fib}Benchmark of \texttt{benches/fib.ll} output by \mintinline{bash}{dune exec bench -- -f fib -n 1000}}
-\end{center}
 \end{table}
@@
 To elucidate this hypothesis, here is the same benchmark but only for the greedy of decreasing \(k\) assignable registers:
-\begin{table}[!ht]
-\begin{center}
+\begin{table}[H]
+\makebox[\textwidth][c]{
 \begin{NiceTabular}{|c|ccccccccc|}
@@
 \hline
-\end{NiceTabular}
+\end{NiceTabular}}
 \caption{\label{tab:dumb-fib-greedy}Benchmark of \texttt{benches/fib.ll} output by \mintinline{bash}{dune exec bench -- -f fib -n 1000}}
-\end{center}
+
 \end{table}
@@
 Prime factorization of \lstinline!i32!.
-\begin{table}[!ht]
-\begin{center}
+\begin{table}[H]
+\makebox[\textwidth][c]{
 \begin{NiceTabular}{|c|cccccccc|}
@@
 \hline
-\end{NiceTabular}
+\end{NiceTabular}}
 \caption{\label{tab:factori32}Benchmark of \texttt{benches/fib.ll} output by \mintinline{bash}{dune exec bench -- -f fib -n 1000}}
-\end{center}
 \end{table}
@@
 \subsubsection{\lstinline!benches/factori64.ll!}
-\begin{table}[!ht]
-\begin{center}
+\begin{table}[H]
+\makebox[\textwidth][c]{
 \begin{NiceTabular}{|c|cccccccc|}
@@
 \hline
-\end{NiceTabular}
+\end{NiceTabular}}
 \caption{\label{tab:factori64-table}Benchmark of \texttt{benches/fib.ll} output by \mintinline{bash}{dune exec bench -- -f fib -n 1000}}
-\end{center}
 \end{table}
@@
 \subsubsection{\lstinline!benches/sieven.ll!}
-\begin{table}[!ht]
-\begin{center}
+\begin{table}[H]
+\makebox[\textwidth][c]{
 \begin{NiceTabular}{|c|cccccccc|}
@@
 \hline
-\end{NiceTabular}
+\end{NiceTabular}}
 \caption{\label{demo-table}Benchmark of \texttt{benches/sieven.ll} output by \mintinline{bash}{dune exec bench -- -f fib -n 1000}}
-\end{center}
 \end{table}
@@
 \subsubsection{\lstinline!benches/subset.ll!}
+\begin{table}[H]
+\makebox[\textwidth][c]{
 \begin{NiceTabular}{|c|cccccccc|}
@@
 \hline
-\end{NiceTabular}
+\end{NiceTabular}}
+\caption{\label{demo-table}Benchmark of \texttt{benches/sieven.ll} output by \mintinline{bash}{dune exec bench -- -f fib -n 1000}}
+\end{table}
 
 \subsubsection{\lstinline!benches/sha256.ll!}
+\begin{table}[H]
+\makebox[\textwidth][c]{
 \begin{NiceTabular}{|c|cccccccc|}
@@
 \hline
-\end{NiceTabular}
+\end{NiceTabular}}
+\caption{\label{tab:fannkuch}Benchmark of \texttt{benches/sha256.ll} output by \mintinline{bash}{dune exec bench -- -f fib -n 1000}}
+\end{table}
 
 \subsubsection{\lstinline!benches/fannkuch-redux.ll!}
-\begin{table}[!ht]
-\begin{center}
+\begin{table}[H]
+\makebox[\textwidth][c]{
 \begin{NiceTabular}{|c|cccccccc|}
@@
 \hline
-\end{NiceTabular}
+\end{NiceTabular}}
 \caption{\label{tab:fannkuch}Benchmark of \texttt{benches/sieven.ll} output by \mintinline{bash}{dune exec bench -- -f fib -n 1000}}
-\end{center}
 \end{table}

### 2024-01-03 article.tex
Begin work on phi nodes

@@
 The control flow of a given program refers to the order in which instructions are executed.
-While the flow of most instructions is linear, in the sense that the next instruction executed is located immediately after, some transfer the flow of execution elsewhere or terminate it.
+While the flow of most instructions is linear, in the sense that the next instruction executed is located immediately after, some transfer the flow of execution elsewhere or outright terminate it.
 %% FIXME: why is this important?
@@
 The type of node varies over the source material, with CFGs of the Appel text \cite{tiger} constructed over individual instructions  and the Aho text \cite{dragon} over basic blocks.
+% TODO: comment on precision
 Either way, leaders will have a set of predecessors and terminators a set of successors associated with them.
@@
 
-Building a graph is relatively straight forward, with the input for every function declared parsed into a tuple type \mintinline{ocaml}{cfg}  of the form:
+Building a graph is relatively straight forward, with the input for every function declared parsed as a tuple type named \mintinline{ocaml}{cfg}  of the form:
 \begin{center}
@@
 \end{center}
-which consists of an optionally named entry block in the first part and a list of trailing blocks that are always named in the second.
+Which consists of an optionally named entry block in the first part and a list of trailing blocks that are always named in the second.
 The reason for the first block to be optionally named is the fact that \textit{some} LLVM programs transfer control back to the point of entry, whereas every subsequent block needs to be named because branching can only target it by its name.
+
+From this, a CFG is constructed  using the \texttt{OCamlgraph} library \cite{ocamlgraph}, which provides several approaches with  various  structures, but the one used in this case is \texttt{Imperative.Digraph.Abstract} where the index is parameterized over \texttt{int} which corresponds to the index of the instruction starting at 0 for the first instruction to be executed.
+At first, a graph is constructed with the input flattened such that
 
@@
 Because of this, for any phi node of the form 
-\[
-  a = \Phi (b_1, ..., b_n)
-\]
+\begin{equation}
+  a = \Phi (b_1, ..., b_n) \label{phi}
+\end{equation}
 add edges to all \textit{live-out} variables not in \(B\):
@@
 
-\subsubsection{Implementing \mintinline{llvm}{call} operation}
+\subsubsection{Implementing \mintinline{llvm}{call} instructions}
 
@@
 The \mintinline{gas}{%rax} register is zeroed before
+
+\subsubsection{Implementing \mintinline{llvm}{phi} instructions}
+Another irregular instruction is the \mintinline{llvm}{phi} instruction representing the \(\Phi\)-functions. Their  intended purpose is to copy data from whichever predecessor node was executed.
+%, so intuitively, additional branching must be incurred to select the correct value to merge, but this isn't the case. Rather, 
+As opposed to the other instructions that are only concerned with transferring the current state to another, this operation needs to concern itself with what was executed immediately prior.
+Fortunately there is no need to further complicate the scenario with additional state for the past executed block, as simply moving from \(\mathit{assign}[b_i]\) to \(\mathit{assign}[a]\) immediately prior to branching into the block whose header contains a phi node of the form \eqref{phi} will suffice.
+
+Notice for instance Figure \ref{fig:square0ll} which has three phi nodes: two in  block 3 and one in block 9, i.e.
+\texttt{\%0} needs to be copied to \texttt{\%4} when control is transferred from block 1 and \texttt{\%7} when transferred from block 3.
+Likewise \texttt{\$0} must be copied to \texttt{\%5} when transferred from block 1 and \texttt{\%6} when transferred from block 3.
+
+\begin{figure}[H]
+   \begin{minipage}{0.48\textwidth}
+     \centering
+     \begin{minted}[bgcolor=LightGray]{llvm}
+define i32 @main (i32 %0) {
+1:
+ %2 = icmp sgt i32 %0, 0
+ br i1 %2, label %3, label %9
+3:
+ %4 = phi i32 [%0, %1], [%7, %3]
+ %5 = phi i32 [0, %1], [%6, %3]
+ %6 = add i32 %5, %0
+ %7 = sub i32 %4, 1
+ %8 = icmp ugt i32 %7, 0
+ br i1 %8, label %3, label %9
+9:
+ %10 = phi i32 [0, %1], [%6, %3]
+ ret i32 %10
+}
+     \end{minted}
+     \caption{\texttt{tests/square0.ll}}\label{fig:square0ll}
+   \end{minipage}\hfill
+   \begin{minipage}{0.48\textwidth}
+     \centering
+     \begin{minted}[bgcolor=LightGray]{llvm}
+define i32 @add (i32 %0, i32 %1) {
+ %3 = add i32 %1, %0
+ ret i32 %3
+}
+     \end{minted}
+     \caption{Stripped \texttt{clang -O1 -S -emit-llvm}}\label{fig:square0c}
+   \end{minipage}
+\end{figure}
+
+%and concerning with which node was executed immediately before, this can be achieved by approaching the problem differently.
 

### 2024-01-03 article.tex
More elaborate phi implementation

@@
 
-Notice for instance Figure \ref{fig:square0ll} which has three phi nodes: two in  block 3 and one in block 9, i.e.
-\texttt{\%0} needs to be copied to \texttt{\%4} when control is transferred from block 1 and \texttt{\%7} when transferred from block 3.
-Likewise \texttt{\$0} must be copied to \texttt{\%5} when transferred from block 1 and \texttt{\%6} when transferred from block 3.
-
 \begin{figure}[H]
@@
      \centering
-     \begin{minted}[bgcolor=LightGray]{llvm}
+     \begin{minted}[linenos,bgcolor=LightGray]{llvm}
 define i32 @main (i32 %0) {
@@
      \end{minted}
-     \caption{\texttt{tests/square0.ll}}\label{fig:square0ll}
+     \caption{\texttt{cat tests/square0.ll}}\label{fig:square0ll}
    \end{minipage}\hfill
@@
      \centering
-     \begin{minted}[bgcolor=LightGray]{llvm}
-define i32 @add (i32 %0, i32 %1) {
- %3 = add i32 %1, %0
- ret i32 %3
-}
-     \end{minted}
-     \caption{Stripped \texttt{clang -O1 -S -emit-llvm}}\label{fig:square0c}
+     \begin{minted}[linenos,bgcolor=LightGray]{gas}
+_main$3:
+  # ...
+  # br i1 %8, label %3, label %9
+  movq %rdx, %rax
+  cmpq $0, %rax
+  movq %rbx, %rax
+  movq %rax, %rdx
+  je _main$9
+  movq %rsi, %rax
+  movq %rax, %rsi
+  movq %rbx, %rax
+  movq %rax, %rdx
+  jmp _main$3
+_main$9:
+  # ...
+     \end{minted}
+     \caption{\texttt{dune exec build -{}- -t x86}}\label{fig:square0s}
    \end{minipage}
 \end{figure}
+
+Notice for instance Figure \ref{fig:square0ll} which has three phi nodes: two in  block 3 and one in block 9, i.e.
+\texttt{\%0} needs to be copied to \texttt{\%4} when control is transferred from block 1 and \texttt{\%7} when transferred from block 3.
+Likewise the immediate value \texttt{\$0} must be copied to \texttt{\%5} when transferred from block 1 and \texttt{\%6} when transferred from block 3.
+All of the relevant translation can be seen in the fragment on Figure \ref{fig:square0s}.
+After the comparison operation on lines 4-6 that check if the boolean condition of \texttt{\%2} is nonzero and setting \texttt{FLAGS} register accordingly, \texttt{\$0} is moved into
+\(\mathit{assign}[\texttt{\%1}]\) in preparation
+of transferring to \texttt{\_main\$9} if the zero flag indeed is set.
+If not, execution is continued in which case both \(\mathit{assign}[\texttt{\%6}]\) and \(\mathit{assign}[\texttt{\%7}]\) are moved into \(\mathit{assign}[\texttt{\%4}]\) and \(\mathit{assign}[\texttt{\%5}]\) respectively, in preperation for continuing execution at \texttt{\_main\$3}.
+
+As \texttt{mov} operations do not change the \texttt{FLAGS} register, having \texttt{mov} instructions between the branching instructions of \texttt{je} and \texttt{jmp} do not affect the branching.
+Likewise, because of the \(\mathit{use}[n]\) semantics of \(\Phi\)-functions, all \(b_i\) are in interference in one another, so may not be overwritten by the \texttt{mov}s inserted.
+
+% FIXME: check if above is sound
 

### 2024-01-04 article.tex
Broken but jane approved

@@
 
-Conversely, managed languages such as Java and C\# target their own respective VMs, so typically don't have much to gain from the LLVM toolchain.
+Conversely, managed languages such as Java and C\# target their own respective VMs, so they typically don't have much to gain from the LLVM toolchain.
 That said, some projects have attempted to implement LLVM in their translation of bytecode to native architectures,  such as LLILC which promised both \textit{just-in-time} (JIT) as well as \textit{ahead-of-time} (AOT) compilation \cite{dotnet}.
-Similarly, Kotlin (the spiritual successor to Java) that ordinarily targets the JVM has a native backend that compiles directly to machine code using LLVM, circumventing the need for a VM entirely \cite{kotlin}.
+Similarly, Kotlin (the spiritual successor of Java) that ordinarily targets the JVM has a native backend that compiles directly to machine code using LLVM, circumventing the need for a VM entirely \cite{kotlin}.
 %Meanwhile interpreted languages like JavaScript eagerly attempt use LLVM for \textit{just-in-time} (JIT) compilation % FIXME: this is not true at all lol
@@
 
-\begin{figure}[!ht]
+\begin{figure}[H]
    \begin{minipage}{0.48\textwidth}
@@
 Either way, the assignment is sound and the instruction selection is reasonably close to the  source code.
-Obviously there are other changes as well, as marking the target section with \texttt{.text} and the \texttt{.globl} annotation to export \texttt{add} in the symbol table which permits linking with other object files, but that isn't stricly related to register allocation.
+There are other changes as well, as marking the target section with \texttt{.text} and the \texttt{.globl} annotation to export \texttt{add} in the symbol table which permits linking with other object files, but that isn't stricly related to register allocation.
 
@@
 Another approach is that of Figure \ref{fig:mulll1}, which uses so-called 'phi nodes' instead.
-These are a type of pseudo instruction that evaluate to a specified operand depending on which predecessor block was executed immediately before.
+These are a type of instruction that evaluates to a specified operand depending on which predecessor block was executed immediately before.
 This circumvents the need for the same level of interior mutability as is achieved by reading/writing to memory, without violating the properties of SSA.
@@
 Successors of node \(n\) are denoted \( \mathit{succ}[n], \)
-and while each node also has predecessors associated, which consist of an  unbounded number of nodes from which control may be transferred, it isn't particularly useful in the following analysis.
+and while each node also has a set of predecessors associated, which consists of an  unbounded number of nodes from which control may be transferred, it isn't particularly useful in the following analysis.
 
@@
 \end{center}
-Which consists of an optionally named entry block in the first part and a list of trailing blocks that are always named in the second.
+which consists of an optionally named entry block in the first part and a list of trailing blocks that are always named in the second.
 The reason for the first block to be optionally named is the fact that \textit{some} LLVM programs transfer control back to the point of entry, whereas every subsequent block needs to be named because branching can only target it by its name.
@@
 \end{equation}
-add edges to all \textit{live-out} variables not in \(B\):
+add edges to all \textit{live-out} variables not in \(\{b_1, ..., b_n\}\):
 \[
@@
 The instructions included are
-\lstinline!trunc! has only been added in order to help cover more generated LLVM.
+\lstinline!trunc! has only been added in order to help cover more generated LLVM. % FIXME
 %with only four distinct block terminators 
@@
 
-As \texttt{mov} operations do not change the \texttt{FLAGS} register, having \texttt{mov} instructions between the branching instructions of \texttt{je} and \texttt{jmp} do not affect the branching.
+As \texttt{mov} operations do not change the \texttt{FLAGS} register \cite{x86asm}, having \texttt{mov} instructions between the branching instructions of \texttt{je} and \texttt{jmp} do not affect the branching.
 Likewise, because of the \(\mathit{use}[n]\) semantics of \(\Phi\)-functions, all \(b_i\) are in interference in one another, so may not be overwritten by the \texttt{mov}s inserted.
@@
 The result of applying a binary operation,  for instance, should only affect the assigned register.
+
+% TODO: kommenter på stripper
 

### 2024-01-05 article.tex
Insert flatten code

@@
 Building a graph is relatively straight forward, with the input for every function declared parsed as a tuple type named \mintinline{ocaml}{cfg}  of the form:
-\begin{center}
 \begin{minted}{ocaml}
@@
 \end{minted}
-\end{center}
 which consists of an optionally named entry block in the first part and a list of trailing blocks that are always named in the second.
@@
 From this, a CFG is constructed  using the \texttt{OCamlgraph} library \cite{ocamlgraph}, which provides several approaches with  various  structures, but the one used in this case is \texttt{Imperative.Digraph.Abstract} where the index is parameterized over \texttt{int} which corresponds to the index of the instruction starting at 0 for the first instruction to be executed.
-At first, a graph is constructed with the input flattened such that
+At first, a graph is constructed with the input flattened such that labels, instructions and terminators are represented as a contiguous sequence as opposed to the basic blocks that is parsed from the input.
+
+\begin{minted}[bgcolor=LightGray]{ocaml}
+let flatten ((head, tail) : Ll.cfg) : insn list =
+  let label l = Label l and insn i = Insn i in
+  let block (b : Ll.block) = List.map insn b.insns @ [ Term b.terminator ] in
+  let named_opt (n, b) = (Option.map label n |> Option.to_list) @ block b in
+  let named (n, b) = Label n :: block b in
+  named_opt head @ (List.map named tail |> List.flatten)
+\end{minted}
 
@@
 \hline
-n & clang & simple 12 & simple 2 & briggs 12 & briggs 2 & linear & greedy 12 & greedy 0 \\
-\hline
-8 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.7529296875,0.9990234375,0.75} 1.01248x& \cellcolor[rgb]{0.75390625,0.9990234375,0.75} 1.01766x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00146x& \cellcolor[rgb]{0.7548828125,0.9990234375,0.75} 1.02261x& \cellcolor[rgb]{0.7509765625,0.9990234375,0.75} 1.00606x& \cellcolor[rgb]{0.7529296875,0.9990234375,0.75} 1.01436x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99702x\\
-10 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99586x& \cellcolor[rgb]{0.7529296875,0.9990234375,0.75} 1.01369x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99508x& \cellcolor[rgb]{0.751953125,0.9990234375,0.75} 1.00861x& \cellcolor[rgb]{0.7548828125,0.9990234375,0.75} 1.01981x& \cellcolor[rgb]{0.751953125,0.9990234375,0.75} 1.00861x& \cellcolor[rgb]{0.755859375,0.9990234375,0.75} 1.02598x\\
-12 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.7548828125,0.9990234375,0.75} 1.02349x& \cellcolor[rgb]{0.76171875,0.9990234375,0.75} 1.04789x& \cellcolor[rgb]{0.759765625,0.9990234375,0.75} 1.04270x& \cellcolor[rgb]{0.7587890625,0.9990234375,0.75} 1.03716x& \cellcolor[rgb]{0.765625,0.9990234375,0.75} 1.06517x& \cellcolor[rgb]{0.763671875,0.9990234375,0.75} 1.05770x& \cellcolor[rgb]{0.77734375,0.9990234375,0.75} 1.11132x\\
-14 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.759765625,0.9990234375,0.75} 1.04054x& \cellcolor[rgb]{0.763671875,0.9990234375,0.75} 1.05851x& \cellcolor[rgb]{0.759765625,0.9990234375,0.75} 1.04050x& \cellcolor[rgb]{0.7626953125,0.9990234375,0.75} 1.05301x& \cellcolor[rgb]{0.7568359375,0.9990234375,0.75} 1.02863x& \cellcolor[rgb]{0.7607421875,0.9990234375,0.75} 1.04599x& \cellcolor[rgb]{0.7900390625,0.9990234375,0.75} 1.16166x\\
-16 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.771484375,0.9990234375,0.75} 1.08630x& \cellcolor[rgb]{0.7724609375,0.9990234375,0.75} 1.09112x& \cellcolor[rgb]{0.7666015625,0.9990234375,0.75} 1.06935x& \cellcolor[rgb]{0.7724609375,0.9990234375,0.75} 1.09077x& \cellcolor[rgb]{0.771484375,0.9990234375,0.75} 1.08669x& \cellcolor[rgb]{0.765625,0.9990234375,0.75} 1.06631x& \cellcolor[rgb]{0.8525390625,0.9990234375,0.75} 1.41519x\\
-18 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.7919921875,0.9990234375,0.75} 1.17172x& \cellcolor[rgb]{0.791015625,0.9990234375,0.75} 1.16684x& \cellcolor[rgb]{0.7880859375,0.9990234375,0.75} 1.15447x& \cellcolor[rgb]{0.7900390625,0.9990234375,0.75} 1.16224x& \cellcolor[rgb]{0.7919921875,0.9990234375,0.75} 1.17110x& \cellcolor[rgb]{0.7900390625,0.9990234375,0.75} 1.16144x& \cellcolor[rgb]{0.9560546875,0.9990234375,0.75} 1.83133x\\
-20 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.806640625,0.9990234375,0.75} 1.23037x& \cellcolor[rgb]{0.80859375,0.9990234375,0.75} 1.23533x& \cellcolor[rgb]{0.8076171875,0.9990234375,0.75} 1.23433x& \cellcolor[rgb]{0.8095703125,0.9990234375,0.75} 1.23986x& \cellcolor[rgb]{0.806640625,0.9990234375,0.75} 1.23120x& \cellcolor[rgb]{0.8046875,0.9990234375,0.75} 1.22283x& \cellcolor[rgb]{0.9990234375,0.990234375,0.75} 2.26430x\\
-22 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.826171875,0.9990234375,0.75} 1.30843x& \cellcolor[rgb]{0.8251953125,0.9990234375,0.75} 1.30498x& \cellcolor[rgb]{0.826171875,0.9990234375,0.75} 1.30814x& \cellcolor[rgb]{0.826171875,0.9990234375,0.75} 1.30594x& \cellcolor[rgb]{0.8251953125,0.9990234375,0.75} 1.30564x& \cellcolor[rgb]{0.82421875,0.9990234375,0.75} 1.29918x& \cellcolor[rgb]{0.9990234375,0.978515625,0.75} 2.64427x\\
-24 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.8408203125,0.9990234375,0.75} 1.36606x& \cellcolor[rgb]{0.83984375,0.9990234375,0.75} 1.36299x& \cellcolor[rgb]{0.8408203125,0.9990234375,0.75} 1.36538x& \cellcolor[rgb]{0.83984375,0.9990234375,0.75} 1.36151x& \cellcolor[rgb]{0.83984375,0.9990234375,0.75} 1.36449x& \cellcolor[rgb]{0.83984375,0.9990234375,0.75} 1.36204x& \cellcolor[rgb]{0.9990234375,0.970703125,0.75} 2.89705x\\
-26 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.8369140625,0.9990234375,0.75} 1.35207x& \cellcolor[rgb]{0.8369140625,0.9990234375,0.75} 1.35059x& \cellcolor[rgb]{0.837890625,0.9990234375,0.75} 1.35397x& \cellcolor[rgb]{0.8369140625,0.9990234375,0.75} 1.34954x& \cellcolor[rgb]{0.8369140625,0.9990234375,0.75} 1.35255x& \cellcolor[rgb]{0.8369140625,0.9990234375,0.75} 1.34932x& \cellcolor[rgb]{0.9990234375,0.9697265625,0.75} 2.92968x\\
-28 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.8388671875,0.9990234375,0.75} 1.35918x& \cellcolor[rgb]{0.837890625,0.9990234375,0.75} 1.35626x& \cellcolor[rgb]{0.8388671875,0.9990234375,0.75} 1.36006x& \cellcolor[rgb]{0.837890625,0.9990234375,0.75} 1.35674x& \cellcolor[rgb]{0.8388671875,0.9990234375,0.75} 1.35812x& \cellcolor[rgb]{0.837890625,0.9990234375,0.75} 1.35462x& \cellcolor[rgb]{0.9990234375,0.96875,0.75} 2.96506x\\
-30 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.8515625,0.9990234375,0.75} 1.41137x& \cellcolor[rgb]{0.8505859375,0.9990234375,0.75} 1.40745x& \cellcolor[rgb]{0.8515625,0.9990234375,0.75} 1.41125x& \cellcolor[rgb]{0.8505859375,0.9990234375,0.75} 1.40768x& \cellcolor[rgb]{0.8515625,0.9990234375,0.75} 1.40935x& \cellcolor[rgb]{0.8505859375,0.9990234375,0.75} 1.40589x& \cellcolor[rgb]{0.9990234375,0.96484375,0.75} 3.08506x\\
-32 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.853515625,0.9990234375,0.75} 1.41590x& \cellcolor[rgb]{0.8525390625,0.9990234375,0.75} 1.41255x& \cellcolor[rgb]{0.853515625,0.9990234375,0.75} 1.41652x& \cellcolor[rgb]{0.8525390625,0.9990234375,0.75} 1.41300x& \cellcolor[rgb]{0.8525390625,0.9990234375,0.75} 1.41399x& \cellcolor[rgb]{0.8515625,0.9990234375,0.75} 1.41115x& \cellcolor[rgb]{0.9990234375,0.9638671875,0.75} 3.09890x\\
+arg(s) & clang & simple 12 & simple 2 & briggs 12 & briggs 2 & linear & greedy 12 & greedy 0 \\
+\hline
+8 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 139964& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00041x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.98897x& \cellcolor[rgb]{0.7509765625,0.9990234375,0.75} 1.00518x& \cellcolor[rgb]{0.7509765625,0.9990234375,0.75} 1.00397x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99405x& \cellcolor[rgb]{0.7509765625,0.9990234375,0.75} 1.00683x& \cellcolor[rgb]{0.7548828125,0.9990234375,0.75} 1.02014x\\
+10 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 141471& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00025x& \cellcolor[rgb]{0.7509765625,0.9990234375,0.75} 1.00662x& \cellcolor[rgb]{0.7509765625,0.9990234375,0.75} 1.00701x& \cellcolor[rgb]{0.751953125,0.9990234375,0.75} 1.01013x& \cellcolor[rgb]{0.751953125,0.9990234375,0.75} 1.01107x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99899x& \cellcolor[rgb]{0.7578125,0.9990234375,0.75} 1.03276x\\
+12 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 143729& \cellcolor[rgb]{0.7529296875,0.9990234375,0.75} 1.01286x& \cellcolor[rgb]{0.751953125,0.9990234375,0.75} 1.00840x& \cellcolor[rgb]{0.755859375,0.9990234375,0.75} 1.02403x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.97714x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00222x& \cellcolor[rgb]{0.755859375,0.9990234375,0.75} 1.02376x& \cellcolor[rgb]{0.7578125,0.9990234375,0.75} 1.03188x\\
+14 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 152717& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00032x& \cellcolor[rgb]{0.759765625,0.9990234375,0.75} 1.03967x& \cellcolor[rgb]{0.7578125,0.9990234375,0.75} 1.03472x& \cellcolor[rgb]{0.7587890625,0.9990234375,0.75} 1.03637x& \cellcolor[rgb]{0.7568359375,0.9990234375,0.75} 1.02969x& \cellcolor[rgb]{0.7578125,0.9990234375,0.75} 1.03224x& \cellcolor[rgb]{0.79296875,0.9990234375,0.75} 1.17572x\\
+16 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 173395& \cellcolor[rgb]{0.76953125,0.9990234375,0.75} 1.08069x& \cellcolor[rgb]{0.7626953125,0.9990234375,0.75} 1.05265x& \cellcolor[rgb]{0.76953125,0.9990234375,0.75} 1.08045x& \cellcolor[rgb]{0.7705078125,0.9990234375,0.75} 1.08271x& \cellcolor[rgb]{0.7666015625,0.9990234375,0.75} 1.06793x& \cellcolor[rgb]{0.7666015625,0.9990234375,0.75} 1.06925x& \cellcolor[rgb]{0.8515625,0.9990234375,0.75} 1.40948x\\
+18 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 227988& \cellcolor[rgb]{0.7890625,0.9990234375,0.75} 1.15964x& \cellcolor[rgb]{0.7880859375,0.9990234375,0.75} 1.15565x& \cellcolor[rgb]{0.791015625,0.9990234375,0.75} 1.16758x& \cellcolor[rgb]{0.791015625,0.9990234375,0.75} 1.16818x& \cellcolor[rgb]{0.7900390625,0.9990234375,0.75} 1.16349x& \cellcolor[rgb]{0.791015625,0.9990234375,0.75} 1.16611x& \cellcolor[rgb]{0.951171875,0.9990234375,0.75} 1.80801x\\
+20 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 375309& \cellcolor[rgb]{0.8095703125,0.9990234375,0.75} 1.24200x& \cellcolor[rgb]{0.8095703125,0.9990234375,0.75} 1.23989x& \cellcolor[rgb]{0.810546875,0.9990234375,0.75} 1.24662x& \cellcolor[rgb]{0.810546875,0.9990234375,0.75} 1.24412x& \cellcolor[rgb]{0.8095703125,0.9990234375,0.75} 1.24000x& \cellcolor[rgb]{0.810546875,0.9990234375,0.75} 1.24497x& \cellcolor[rgb]{0.9990234375,0.990234375,0.75} 2.26177x\\
+22 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 760328& \cellcolor[rgb]{0.826171875,0.9990234375,0.75} 1.30708x& \cellcolor[rgb]{0.8271484375,0.9990234375,0.75} 1.31289x& \cellcolor[rgb]{0.8271484375,0.9990234375,0.75} 1.31045x& \cellcolor[rgb]{0.8271484375,0.9990234375,0.75} 1.31230x& \cellcolor[rgb]{0.826171875,0.9990234375,0.75} 1.30795x& \cellcolor[rgb]{0.826171875,0.9990234375,0.75} 1.30885x& \cellcolor[rgb]{0.9990234375,0.978515625,0.75} 2.63144x\\
+24 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1763965& \cellcolor[rgb]{0.8369140625,0.9990234375,0.75} 1.35080x& \cellcolor[rgb]{0.8369140625,0.9990234375,0.75} 1.35029x& \cellcolor[rgb]{0.8369140625,0.9990234375,0.75} 1.35039x& \cellcolor[rgb]{0.8369140625,0.9990234375,0.75} 1.35129x& \cellcolor[rgb]{0.8369140625,0.9990234375,0.75} 1.35274x& \cellcolor[rgb]{0.8369140625,0.9990234375,0.75} 1.35185x& \cellcolor[rgb]{0.9990234375,0.97265625,0.75} 2.83868x\\
+26 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 4321515& \cellcolor[rgb]{0.8466796875,0.9990234375,0.75} 1.38984x& \cellcolor[rgb]{0.8466796875,0.9990234375,0.75} 1.39041x& \cellcolor[rgb]{0.8466796875,0.9990234375,0.75} 1.38939x& \cellcolor[rgb]{0.8466796875,0.9990234375,0.75} 1.39050x& \cellcolor[rgb]{0.8466796875,0.9990234375,0.75} 1.39175x& \cellcolor[rgb]{0.8466796875,0.9990234375,0.75} 1.39096x& \cellcolor[rgb]{0.9990234375,0.9677734375,0.75} 2.98115x\\
+28 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 11291147& \cellcolor[rgb]{0.841796875,0.9990234375,0.75} 1.37185x& \cellcolor[rgb]{0.8427734375,0.9990234375,0.75} 1.37282x& \cellcolor[rgb]{0.8427734375,0.9990234375,0.75} 1.37280x& \cellcolor[rgb]{0.8427734375,0.9990234375,0.75} 1.37369x& \cellcolor[rgb]{0.8427734375,0.9990234375,0.75} 1.37371x& \cellcolor[rgb]{0.8427734375,0.9990234375,0.75} 1.37320x& \cellcolor[rgb]{0.9990234375,0.96875,0.75} 2.96675x\\
+30 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 28225827& \cellcolor[rgb]{0.8564453125,0.9990234375,0.75} 1.42877x& \cellcolor[rgb]{0.8564453125,0.9990234375,0.75} 1.42993x& \cellcolor[rgb]{0.8564453125,0.9990234375,0.75} 1.42896x& \cellcolor[rgb]{0.8564453125,0.9990234375,0.75} 1.43085x& \cellcolor[rgb]{0.857421875,0.9990234375,0.75} 1.43172x& \cellcolor[rgb]{0.8564453125,0.9990234375,0.75} 1.43069x& \cellcolor[rgb]{0.9990234375,0.96484375,0.75} 3.09771x\\
+32 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 73749658& \cellcolor[rgb]{0.8564453125,0.9990234375,0.75} 1.43033x& \cellcolor[rgb]{0.8564453125,0.9990234375,0.75} 1.42973x& \cellcolor[rgb]{0.8564453125,0.9990234375,0.75} 1.42936x& \cellcolor[rgb]{0.8564453125,0.9990234375,0.75} 1.42954x& \cellcolor[rgb]{0.8564453125,0.9990234375,0.75} 1.43088x& \cellcolor[rgb]{0.8564453125,0.9990234375,0.75} 1.42974x& \cellcolor[rgb]{0.9990234375,0.9638671875,0.75} 3.10158x\\
 \hline
@@
 \hline
-n & clang & greedy 12 & greedy 8 & greedy 6 & greedy 4 & greedy 2 & greedy 1 & greedy 0 & tiger \\
-\hline
-8 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99429x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.98513x& \cellcolor[rgb]{0.751953125,0.9990234375,0.75} 1.00816x& \cellcolor[rgb]{0.7529296875,0.9990234375,0.75} 1.01180x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99806x& \cellcolor[rgb]{0.751953125,0.9990234375,0.75} 1.00869x& \cellcolor[rgb]{0.751953125,0.9990234375,0.75} 1.00883x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99979x\\
-10 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99057x& \cellcolor[rgb]{0.7578125,0.9990234375,0.75} 1.03464x& \cellcolor[rgb]{0.7578125,0.9990234375,0.75} 1.03147x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99206x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00151x& \cellcolor[rgb]{0.7607421875,0.9990234375,0.75} 1.04382x& \cellcolor[rgb]{0.76171875,0.9990234375,0.75} 1.04785x& \cellcolor[rgb]{0.763671875,0.9990234375,0.75} 1.05639x\\
-12 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00006x& \cellcolor[rgb]{0.7509765625,0.9990234375,0.75} 1.00578x& \cellcolor[rgb]{0.7529296875,0.9990234375,0.75} 1.01540x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99370x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.97925x& \cellcolor[rgb]{0.7666015625,0.9990234375,0.75} 1.06915x& \cellcolor[rgb]{0.76953125,0.9990234375,0.75} 1.07918x& \cellcolor[rgb]{0.7578125,0.9990234375,0.75} 1.03213x\\
-14 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.755859375,0.9990234375,0.75} 1.02737x& \cellcolor[rgb]{0.7568359375,0.9990234375,0.75} 1.02834x& \cellcolor[rgb]{0.76171875,0.9990234375,0.75} 1.04708x& \cellcolor[rgb]{0.7568359375,0.9990234375,0.75} 1.02997x& \cellcolor[rgb]{0.7646484375,0.9990234375,0.75} 1.06066x& \cellcolor[rgb]{0.79296875,0.9990234375,0.75} 1.17635x& \cellcolor[rgb]{0.7841796875,0.9990234375,0.75} 1.14104x& \cellcolor[rgb]{0.78125,0.9990234375,0.75} 1.12727x\\
-16 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.771484375,0.9990234375,0.75} 1.08717x& \cellcolor[rgb]{0.7734375,0.9990234375,0.75} 1.09426x& \cellcolor[rgb]{0.77734375,0.9990234375,0.75} 1.11143x& \cellcolor[rgb]{0.7841796875,0.9990234375,0.75} 1.13760x& \cellcolor[rgb]{0.78125,0.9990234375,0.75} 1.12774x& \cellcolor[rgb]{0.8525390625,0.9990234375,0.75} 1.41194x& \cellcolor[rgb]{0.849609375,0.9990234375,0.75} 1.40329x& \cellcolor[rgb]{0.8212890625,0.9990234375,0.75} 1.28770x\\
-18 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.78125,0.9990234375,0.75} 1.12836x& \cellcolor[rgb]{0.7861328125,0.9990234375,0.75} 1.14704x& \cellcolor[rgb]{0.7900390625,0.9990234375,0.75} 1.16301x& \cellcolor[rgb]{0.806640625,0.9990234375,0.75} 1.23098x& \cellcolor[rgb]{0.806640625,0.9990234375,0.75} 1.22815x& \cellcolor[rgb]{0.94140625,0.9990234375,0.75} 1.77085x& \cellcolor[rgb]{0.947265625,0.9990234375,0.75} 1.79503x& \cellcolor[rgb]{0.8720703125,0.9990234375,0.75} 1.49336x\\
-20 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.806640625,0.9990234375,0.75} 1.22913x& \cellcolor[rgb]{0.8056640625,0.9990234375,0.75} 1.22392x& \cellcolor[rgb]{0.8173828125,0.9990234375,0.75} 1.27241x& \cellcolor[rgb]{0.8369140625,0.9990234375,0.75} 1.35182x& \cellcolor[rgb]{0.845703125,0.9990234375,0.75} 1.38569x& \cellcolor[rgb]{0.9990234375,0.9912109375,0.75} 2.22689x& \cellcolor[rgb]{0.9990234375,0.990234375,0.75} 2.26200x& \cellcolor[rgb]{0.9501953125,0.9990234375,0.75} 1.80663x\\
-22 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.822265625,0.9990234375,0.75} 1.29405x& \cellcolor[rgb]{0.8232421875,0.9990234375,0.75} 1.29616x& \cellcolor[rgb]{0.8349609375,0.9990234375,0.75} 1.34432x& \cellcolor[rgb]{0.865234375,0.9990234375,0.75} 1.46289x& \cellcolor[rgb]{0.8759765625,0.9990234375,0.75} 1.50765x& \cellcolor[rgb]{0.9990234375,0.98046875,0.75} 2.57480x& \cellcolor[rgb]{0.9990234375,0.9794921875,0.75} 2.62632x& \cellcolor[rgb]{0.9990234375,0.9970703125,0.75} 2.05105x\\
-24 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.83203125,0.9990234375,0.75} 1.33240x& \cellcolor[rgb]{0.83203125,0.9990234375,0.75} 1.33275x& \cellcolor[rgb]{0.8486328125,0.9990234375,0.75} 1.39726x& \cellcolor[rgb]{0.87890625,0.9990234375,0.75} 1.52070x& \cellcolor[rgb]{0.892578125,0.9990234375,0.75} 1.57452x& \cellcolor[rgb]{0.9990234375,0.974609375,0.75} 2.77989x& \cellcolor[rgb]{0.9990234375,0.97265625,0.75} 2.83443x& \cellcolor[rgb]{0.9990234375,0.9931640625,0.75} 2.18442x\\
-26 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.8359375,0.9990234375,0.75} 1.34858x& \cellcolor[rgb]{0.8359375,0.9990234375,0.75} 1.34784x& \cellcolor[rgb]{0.853515625,0.9990234375,0.75} 1.41620x& \cellcolor[rgb]{0.8828125,0.9990234375,0.75} 1.53692x& \cellcolor[rgb]{0.8994140625,0.9990234375,0.75} 1.60380x& \cellcolor[rgb]{0.9990234375,0.9716796875,0.75} 2.87114x& \cellcolor[rgb]{0.9990234375,0.9697265625,0.75} 2.92874x& \cellcolor[rgb]{0.9990234375,0.9912109375,0.75} 2.24149x\\
-28 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.837890625,0.9990234375,0.75} 1.35448x& \cellcolor[rgb]{0.837890625,0.9990234375,0.75} 1.35457x& \cellcolor[rgb]{0.8544921875,0.9990234375,0.75} 1.42324x& \cellcolor[rgb]{0.8857421875,0.9990234375,0.75} 1.54897x& \cellcolor[rgb]{0.90234375,0.9990234375,0.75} 1.61429x& \cellcolor[rgb]{0.9990234375,0.970703125,0.75} 2.90730x& \cellcolor[rgb]{0.9990234375,0.96875,0.75} 2.96219x& \cellcolor[rgb]{0.9990234375,0.990234375,0.75} 2.26833x\\
-30 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.8388671875,0.9990234375,0.75} 1.35807x& \cellcolor[rgb]{0.8388671875,0.9990234375,0.75} 1.35784x& \cellcolor[rgb]{0.85546875,0.9990234375,0.75} 1.42732x& \cellcolor[rgb]{0.888671875,0.9990234375,0.75} 1.55842x& \cellcolor[rgb]{0.904296875,0.9990234375,0.75} 1.62109x& \cellcolor[rgb]{0.9990234375,0.9697265625,0.75} 2.92283x& \cellcolor[rgb]{0.9990234375,0.9677734375,0.75} 2.98079x& \cellcolor[rgb]{0.9990234375,0.990234375,0.75} 2.28148x\\
-32 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.841796875,0.9990234375,0.75} 1.36987x& \cellcolor[rgb]{0.841796875,0.9990234375,0.75} 1.36911x& \cellcolor[rgb]{0.8583984375,0.9990234375,0.75} 1.43879x& \cellcolor[rgb]{0.892578125,0.9990234375,0.75} 1.57453x& \cellcolor[rgb]{0.9072265625,0.9990234375,0.75} 1.63423x& \cellcolor[rgb]{0.9990234375,0.96875,0.75} 2.94896x& \cellcolor[rgb]{0.9990234375,0.966796875,0.75} 3.00741x& \cellcolor[rgb]{0.9990234375,0.9892578125,0.75} 2.29882x\\
+arg(s) & clang & greedy 12 & greedy 8 & greedy 6 & greedy 4 & greedy 2 & greedy 1 & greedy 0 & tiger \\
+\hline
+8 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 139813& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99868x& \cellcolor[rgb]{0.751953125,0.9990234375,0.75} 1.00816x& \cellcolor[rgb]{0.7509765625,0.9990234375,0.75} 1.00434x& \cellcolor[rgb]{0.7509765625,0.9990234375,0.75} 1.00410x& \cellcolor[rgb]{0.7529296875,0.9990234375,0.75} 1.01222x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00127x& \cellcolor[rgb]{0.7529296875,0.9990234375,0.75} 1.01345x& \cellcolor[rgb]{0.7529296875,0.9990234375,0.75} 1.01526x\\
+10 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 141482& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99391x& \cellcolor[rgb]{0.751953125,0.9990234375,0.75} 1.01001x& \cellcolor[rgb]{0.7529296875,0.9990234375,0.75} 1.01231x& \cellcolor[rgb]{0.7529296875,0.9990234375,0.75} 1.01469x& \cellcolor[rgb]{0.7509765625,0.9990234375,0.75} 1.00529x& \cellcolor[rgb]{0.7548828125,0.9990234375,0.75} 1.02271x& \cellcolor[rgb]{0.7578125,0.9990234375,0.75} 1.03397x& \cellcolor[rgb]{0.7548828125,0.9990234375,0.75} 1.02248x\\
+12 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 145091& \cellcolor[rgb]{0.7509765625,0.9990234375,0.75} 1.00774x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99627x& \cellcolor[rgb]{0.75390625,0.9990234375,0.75} 1.01584x& \cellcolor[rgb]{0.751953125,0.9990234375,0.75} 1.01047x& \cellcolor[rgb]{0.75390625,0.9990234375,0.75} 1.01694x& \cellcolor[rgb]{0.7666015625,0.9990234375,0.75} 1.07056x& \cellcolor[rgb]{0.765625,0.9990234375,0.75} 1.06622x& \cellcolor[rgb]{0.7607421875,0.9990234375,0.75} 1.04396x\\
+14 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 149598& \cellcolor[rgb]{0.7646484375,0.9990234375,0.75} 1.06186x& \cellcolor[rgb]{0.765625,0.9990234375,0.75} 1.06462x& \cellcolor[rgb]{0.7646484375,0.9990234375,0.75} 1.06184x& \cellcolor[rgb]{0.765625,0.9990234375,0.75} 1.06665x& \cellcolor[rgb]{0.7705078125,0.9990234375,0.75} 1.08504x& \cellcolor[rgb]{0.7998046875,0.9990234375,0.75} 1.20101x& \cellcolor[rgb]{0.80078125,0.9990234375,0.75} 1.20679x& \cellcolor[rgb]{0.78515625,0.9990234375,0.75} 1.14215x\\
+16 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 170028& \cellcolor[rgb]{0.7724609375,0.9990234375,0.75} 1.09372x& \cellcolor[rgb]{0.7734375,0.9990234375,0.75} 1.09738x& \cellcolor[rgb]{0.7802734375,0.9990234375,0.75} 1.12226x& \cellcolor[rgb]{0.783203125,0.9990234375,0.75} 1.13603x& \cellcolor[rgb]{0.7890625,0.9990234375,0.75} 1.15900x& \cellcolor[rgb]{0.85546875,0.9990234375,0.75} 1.42535x& \cellcolor[rgb]{0.857421875,0.9990234375,0.75} 1.43162x& \cellcolor[rgb]{0.8232421875,0.9990234375,0.75} 1.29614x\\
+18 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 230073& \cellcolor[rgb]{0.7890625,0.9990234375,0.75} 1.15757x& \cellcolor[rgb]{0.7880859375,0.9990234375,0.75} 1.15498x& \cellcolor[rgb]{0.79296875,0.9990234375,0.75} 1.17517x& \cellcolor[rgb]{0.802734375,0.9990234375,0.75} 1.21433x& \cellcolor[rgb]{0.8134765625,0.9990234375,0.75} 1.25770x& \cellcolor[rgb]{0.93359375,0.9990234375,0.75} 1.73753x& \cellcolor[rgb]{0.947265625,0.9990234375,0.75} 1.79427x& \cellcolor[rgb]{0.876953125,0.9990234375,0.75} 1.51243x\\
+20 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 376308& \cellcolor[rgb]{0.8095703125,0.9990234375,0.75} 1.24275x& \cellcolor[rgb]{0.810546875,0.9990234375,0.75} 1.24539x& \cellcolor[rgb]{0.8193359375,0.9990234375,0.75} 1.28023x& \cellcolor[rgb]{0.837890625,0.9990234375,0.75} 1.35349x& \cellcolor[rgb]{0.8515625,0.9990234375,0.75} 1.41074x& \cellcolor[rgb]{0.9990234375,0.9912109375,0.75} 2.22527x& \cellcolor[rgb]{0.9990234375,0.990234375,0.75} 2.26082x& \cellcolor[rgb]{0.9521484375,0.9990234375,0.75} 1.81290x\\
+22 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 762044& \cellcolor[rgb]{0.826171875,0.9990234375,0.75} 1.30899x& \cellcolor[rgb]{0.826171875,0.9990234375,0.75} 1.30974x& \cellcolor[rgb]{0.8388671875,0.9990234375,0.75} 1.35873x& \cellcolor[rgb]{0.8623046875,0.9990234375,0.75} 1.45410x& \cellcolor[rgb]{0.8798828125,0.9990234375,0.75} 1.52442x& \cellcolor[rgb]{0.9990234375,0.98046875,0.75} 2.57640x& \cellcolor[rgb]{0.9990234375,0.9794921875,0.75} 2.62467x& \cellcolor[rgb]{0.9990234375,0.9970703125,0.75} 2.05193x\\
+24 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1765499& \cellcolor[rgb]{0.8369140625,0.9990234375,0.75} 1.34932x& \cellcolor[rgb]{0.8369140625,0.9990234375,0.75} 1.34929x& \cellcolor[rgb]{0.8505859375,0.9990234375,0.75} 1.40559x& \cellcolor[rgb]{0.876953125,0.9990234375,0.75} 1.51369x& \cellcolor[rgb]{0.896484375,0.9990234375,0.75} 1.59160x& \cellcolor[rgb]{0.9990234375,0.974609375,0.75} 2.78143x& \cellcolor[rgb]{0.9990234375,0.97265625,0.75} 2.83690x& \cellcolor[rgb]{0.9990234375,0.9931640625,0.75} 2.18519x\\
+26 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 4399288& \cellcolor[rgb]{0.8408203125,0.9990234375,0.75} 1.36595x& \cellcolor[rgb]{0.8408203125,0.9990234375,0.75} 1.36628x& \cellcolor[rgb]{0.85546875,0.9990234375,0.75} 1.42456x& \cellcolor[rgb]{0.8837890625,0.9990234375,0.75} 1.53853x& \cellcolor[rgb]{0.904296875,0.9990234375,0.75} 1.62035x& \cellcolor[rgb]{0.9990234375,0.9716796875,0.75} 2.87046x& \cellcolor[rgb]{0.9990234375,0.9697265625,0.75} 2.92749x& \cellcolor[rgb]{0.9990234375,0.9912109375,0.75} 2.24179x\\
+28 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 11290936& \cellcolor[rgb]{0.8427734375,0.9990234375,0.75} 1.37293x& \cellcolor[rgb]{0.8427734375,0.9990234375,0.75} 1.37379x& \cellcolor[rgb]{0.857421875,0.9990234375,0.75} 1.43301x& \cellcolor[rgb]{0.8857421875,0.9990234375,0.75} 1.54853x& \cellcolor[rgb]{0.9072265625,0.9990234375,0.75} 1.63198x& \cellcolor[rgb]{0.9990234375,0.970703125,0.75} 2.90889x& \cellcolor[rgb]{0.9990234375,0.96875,0.75} 2.96589x& \cellcolor[rgb]{0.9990234375,0.990234375,0.75} 2.27035x\\
+30 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 28716847& \cellcolor[rgb]{0.8505859375,0.9990234375,0.75} 1.40466x& \cellcolor[rgb]{0.8505859375,0.9990234375,0.75} 1.40503x& \cellcolor[rgb]{0.865234375,0.9990234375,0.75} 1.46656x& \cellcolor[rgb]{0.8955078125,0.9990234375,0.75} 1.58655x& \cellcolor[rgb]{0.9169921875,0.9990234375,0.75} 1.67357x& \cellcolor[rgb]{0.9990234375,0.9677734375,0.75} 2.98560x& \cellcolor[rgb]{0.9990234375,0.9658203125,0.75} 3.04538x& \cellcolor[rgb]{0.9990234375,0.98828125,0.75} 2.32981x\\
+32 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 76582019& \cellcolor[rgb]{0.84375,0.9990234375,0.75} 1.37696x& \cellcolor[rgb]{0.8427734375,0.9990234375,0.75} 1.37647x& \cellcolor[rgb]{0.8583984375,0.9990234375,0.75} 1.43704x& \cellcolor[rgb]{0.8876953125,0.9990234375,0.75} 1.55462x& \cellcolor[rgb]{0.9091796875,0.9990234375,0.75} 1.64074x& \cellcolor[rgb]{0.9990234375,0.9697265625,0.75} 2.92847x& \cellcolor[rgb]{0.9990234375,0.9677734375,0.75} 2.98745x& \cellcolor[rgb]{0.9990234375,0.9892578125,0.75} 2.28358x\\
 \hline

### 2024-01-05 article.tex
Resolve duplicate labels

@@
 \end{NiceTabular}}
-\caption{\label{demo-table}Benchmark of \texttt{benches/sieven.ll} output by \mintinline{bash}{dune exec bench -- -f fib -n 1000}}
+\caption{\label{tab:sieven}Benchmark of \texttt{benches/sieven.ll} output by \mintinline{bash}{dune exec bench -- -f fib -n 1000}}
 \end{table}
@@
 \end{NiceTabular}}
-\caption{\label{demo-table}Benchmark of \texttt{benches/sieven.ll} output by \mintinline{bash}{dune exec bench -- -f fib -n 1000}}
+\caption{\label{tab:subset}Benchmark of \texttt{benches/sieven.ll} output by \mintinline{bash}{dune exec bench -- -f fib -n 1000}}
 \end{table}
@@
 \end{NiceTabular}}
-\caption{\label{tab:fannkuch}Benchmark of \texttt{benches/sha256.ll} output by \mintinline{bash}{dune exec bench -- -f fib -n 1000}}
+\caption{\label{tab:sha256}Benchmark of \texttt{benches/sha256.ll} output by \mintinline{bash}{dune exec bench -- -f fib -n 1000}}
 \end{table}

### 2024-01-05 article.tex
Remove parens around ieee citations

@@
 
-Another approach, which is a more precise approximation, is a specific variant of the dataflow analysis as described in \textit{Modern Compiler Implementation in ML} (\cite{tiger}) and \textit{Compilers: Principles, Techniques, and Tools} (\cite{dragon}).
+Another approach, which is a more precise approximation, is a specific variant of the dataflow analysis as described in \textit{Modern Compiler Implementation in ML} \cite{tiger} and \textit{Compilers: Principles, Techniques, and Tools} \cite{dragon}.
 In general, dataflow analysis is the process of finding the possible paths in which data may propagate through various branches of execution.
@@
 \]
-Although the Appel text (\cite{tiger}) describes interference of variables with concrete registers as well as overlapping variables, this isn't considered in this implementation.
+Although the Appel text \cite{tiger} describes interference of variables with concrete registers as well as overlapping variables, this isn't considered in this implementation.
 
@@
 
-The primary text is \textit{Linear Scan Register Allocation} (\cite{linear}), the execution of which also relies on prior liveness analysis.
+The primary text is \textit{Linear Scan Register Allocation} \cite{linear}, the execution of which also relies on prior liveness analysis.
 

### 2024-01-05 article.tex
Add frontpage picture
Add placeholder

@@
 \usepackage{float}
+\usepackage{svg}
 
@@
 
-\title{Implementation and comparison of register allocation to translate \lstinline!LLVM--! to x86}
-\author{William Welle Tange}
-\date{2023}
+\font\myfont=cmr12 at 24pt
+\title{\myfont Implementation and evaluation of register allocation used to translate \texttt{LLVM-{}-} to x86-64}
+\author{William Welle Tange, 201809080}
+
+\pagenumbering{roman}
 
@@
 \maketitle
+\begin{center}
+\includesvg[width=0.5\textwidth]{yawning-egg}
+\end{center}
+\newpage
+
+\section*{Abstract}
+Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
+Tellus at urna condimentum mattis.
+Enim facilisis gravida neque convallis.
+Enim facilisis gravida neque convallis a cras semper auctor.
+Vitae proin sagittis nisl rhoncus mattis rhoncus urna neque.
+Justo nec ultrices dui sapien eget mi proin.
+Id aliquet risus feugiat in ante.
+A cras semper auctor neque vitae tempus quam pellentesque nec.
+Nulla porttitor massa id neque aliquam vestibulum morbi.
+Facilisi cras fermentum odio eu.
+Odio pellentesque diam volutpat commodo sed egestas egestas fringilla.
+Amet consectetur adipiscing elit duis tristique sollicitudin nibh sit amet.
+Adipiscing at in tellus integer feugiat.
+A pellentesque sit amet porttitor.
+Sed elementum tempus egestas sed sed risus pretium quam.
+Quam viverra orci sagittis eu volutpat odio facilisis mauris sit.
+Arcu dictum varius duis at.
+Sociis natoque penatibus et magnis dis parturient montes.
+Varius sit amet mattis vulputate enim nulla aliquet.
+Ac ut consequat semper viverra.
+Scelerisque eleifend donec pretium vulputate sapien nec.
+Orci ac auctor augue mauris augue.
+Pulvinar mattis nunc sed blandit.
+Semper auctor neque vitae tempus quam pellentesque nec nam.
+Risus pretium quam vulputate dignissim.
+Dui faucibus in ornare quam viverra orci sagittis eu volutpat.
+Ac tortor vitae purus faucibus.
+Eu mi bibendum neque egestas congue quisque egestas diam in.
+Curabitur vitae nunc sed velit dignissim sodales ut eu sem.
+Arcu odio ut sem nulla.
+\newpage
 
@@
     %* jo mindre man skal overskue af gangen jo lettere er det at komme i gang med at skrive
+
+\setcounter{page}{1}
+\pagenumbering{arabic}
 
@@
 Another approach is that of Figure \ref{fig:mulll1}, which uses so-called 'phi nodes' instead.
-These are a type of instruction that evaluates to a specified operand depending on which predecessor block was executed immediately before.
+These are a type of instruction that evaluates to a specified operand depending on which predecessor block was executed immediately prior.
 This circumvents the need for the same level of interior mutability as is achieved by reading/writing to memory, without violating the properties of SSA.
@@
 which consists of an optionally named entry block in the first part and a list of trailing blocks that are always named in the second.
-The reason for the first block to be optionally named is the fact that \textit{some} LLVM programs transfer control back to the point of entry, whereas every subsequent block needs to be named because branching can only target it by its name.
+The reason for the first block to be optionally named is the fact that \textit{some} LLVM programs transfer control back to the point of entry, whereas every subsequent block needs to be named because branching can only target it by referring to its name.
 
 From this, a CFG is constructed  using the \texttt{OCamlgraph} library \cite{ocamlgraph}, which provides several approaches with  various  structures, but the one used in this case is \texttt{Imperative.Digraph.Abstract} where the index is parameterized over \texttt{int} which corresponds to the index of the instruction starting at 0 for the first instruction to be executed.
-At first, a graph is constructed with the input flattened such that labels, instructions and terminators are represented as a contiguous sequence as opposed to the basic blocks that is parsed from the input.
+At first, a graph is constructed with the input flattened such that labels, instructions and terminators are represented as a continuous sequence as opposed to the basic blocks that is parsed from the input.
 
@@
 \end{minted}
+
+This is done to more easily look up which graph vertex corresponds to the \(n\)th instruction/terminator.
 

### 2024-01-05 article.tex
Looks okay

@@
 Such a heuristic is greedy in the sense that it picks an assignment known to be safe for the least amount of preprocessing possible. However, this can be a very inefficient assignment at runtime. 
-When the number of variables is greater than the number of assignable registers, the variable is spilled.
-A number of \(n\)  variables greater than \(k\) working registers causes \(n-k\) variables to be spilled to memory, % TODO: skal der stå stack?
+When the number of variables is greater than the number of assignable registers, a variable must be spilled.
+A number of \(n\)  variables greater than \(k\) working registers causes \(n-k\) variables to be spilled, % TODO: skal der stå stack?
 which can greatly reduce performance, but  allows for translation in linear time. %FIXME: is this even true?
@@
 with the \(\mathit{use}[n]\) set being defined as any variables that may be used and the \(\mathit{def}[n]\) as the set of variables defined in node \(n\).
-Because the chosen IR is in SSA form the \(\mathit{def}[n]\) either consists of one variable or equal to \(\emptyset\).
+The \(\mathit{def}[n]\) either consists of one variable or equal to \(\emptyset\).
 The  \(\mathit{use}[n]\) set is effectively unbounded as some instructions take any \(m\) variables as parameters.
@@
 \end{minted}
+
+\noindent The \texttt{flowin} function corresponds to the \textit{live-in} equation \eqref{flowin} and is implemented as follows:
+\begin{minted}[linenos,bgcolor=LightGray]{ocaml}
+let flowin (i, insn) =
+  let newin = S.SS.union (use insn) (S.SS.diff out.(i) (def insn)) in
+  let changed = not (S.SS.equal newin in_.(i)) in
+  if changed then in_.(i) <- newin;
+  changed
+\end{minted}
+And the \texttt{flowout} function which corresponds to the \textit{live-out} equation \eqref{flowout} implemented as follows:
+\begin{minted}[linenos,bgcolor=LightGray]{ocaml}
+let flowout (i, _) =
+  let newout =
+    let succ = Cfg.G.succ g ids.(i) in
+    List.fold_left
+      (fun s v -> S.SS.union s in_.(Cfg.G.V.label v))
+      S.SS.empty succ
+  in
+  let changed = not (S.SS.equal newout out.(i)) in
+  if changed then out.(i) <- newout;
+  changed
+\end{minted}
+
 With the actual dataflow analysis performed recursively as follows:
+\begin{figure}[H]
 \begin{minted}[linenos,bgcolor=LightGray]{ocaml}
@@
 \end{minted}
-The \texttt{flowin} function correponds to the \textit{live-in} equation \eqref{flowin} and is implemented as follows:
-\begin{minted}[linenos,bgcolor=LightGray]{ocaml}
-let flowin (i, insn) =
-  let newin = S.SS.union (use insn) (S.SS.diff out.(i) (def insn)) in
-  let changed = not (S.SS.equal newin in_.(i)) in
-  if changed then in_.(i) <- newin;
-  changed
-\end{minted}
-And the \texttt{flowout} function which correponds to the \textit{live-out} equation \eqref{flowout} implemented as follows:
-\begin{minted}[linenos,bgcolor=LightGray]{ocaml}
-let flowout (i, _) =
-  let newout =
-    let succ = Cfg.G.succ g ids.(i) in
-    List.fold_left
-      (fun s v -> S.SS.union s in_.(Cfg.G.V.label v))
-      S.SS.empty succ
-  in
-  let changed = not (S.SS.equal newout out.(i)) in
-  if changed then out.(i) <- newout;
-  changed
-\end{minted}
-To
+\caption{Overall implementation of \texttt{dataflow}\label{fig:dataflow}}
+\end{figure}
+\noindent One thing that is particularly noteworthy is reversing of the flattened list of instructions on line 2 of Figure \ref{fig:dataflow}.
+After enumerating each entry so that the corresponding vertex is an array lookup away, it is reversed.
+The motivation for this is the fact that the \(\mathit{out}[n]\) of node \(n\) depends on \(\mathit{in}[s]\) for all successors \(s\).
+Since most nodes have one successor immediately after, populating this before hands is significantly more efficient.
+
+%\begin{NiceTabular}{|W{c}{0.5cm}|W{c}{0.5cm} W{c}{0.5cm}|W{c}{0.5cm} W{c}{0.5cm}|W{c}{0.5cm} W{c}{0.5cm}|W{c}{0.5cm} W{c}{0.5cm}|W{c}{0.5cm} W{c}{0.5cm}|W{c}{0.5cm} W{c}{0.5cm}|W{c}{0.5cm} W{c}{0.5cm}|W{c}{0.5cm} W{c}{0.5cm}|}
+
+\begin{table}[H]
+\makebox[\textwidth][c]{
+\begin{NiceTabular}{|c|c c|c c|c c|c c|c c|c c|c c|}
+  \hline
+  i & use & def & in & out & in & out & in & out & in & out & in & out & in & out\\
+  \hline
+  0  & 0  & 3 & 0  & 3 & 0  & 0,3,8 & 0,8  & 0,3,8 & 0,8  & 0,3,8 & 0,8  & 0,3,8 & 0,8  & 0,3,8\\
+  1  & 3  &  & 3  & 0,8 & 0,3,8  & 0,8 & 0,3,8  & 0,8 & 0,3,8  & 0,8 & 0,3,8  & 0,8 & 0,3,8  & 0,8\\
+  2  &   &  &   &  &   &  &   &  &   &  &   &  &   & \\
+  3  &   &  &   &  &   &  &   &  &   &  &   &  &   & \\
+  4  &   &  &   & 0,8 & 0,8  & 0,8 & 0,8  & 0,8 & 0,8  & 0,8 & 0,8  & 0,8 & 0,8  & 0,8\\
+  5  & 0,8  & 6 & 0,8  & 6 & 0,8  & 6 & 0,8  & 6 & 0,8  & 6 & 0,8  & 0,6 & 0,8  & 0,6\\
+  6  & 6  & 7 & 6  & 6 & 6  & 6 & 6  & 6 & 6  & 0,6 & 0,6  & 0,6 & 0,6  & 0,6\\
+  7  & 6  & 8 & 6  &  & 6  &  & 6  & 0,8 & 0,6  & 0,8 & 0,6  & 0,8 & 0,6  & 0,8\\
+  8  &   &  &   &  &   & 0,8 & 0,8  & 0,8 & 0,8  & 0,8 & 0,8  & 0,8 & 0,8  & 0,8\\
+  \hline
+\end{NiceTabular}
+}
+\caption{Output of \texttt{dune exec build -{}- -t lva -v tests/count1.ll}}
+\label{tab:countll-rev-steps}
+\end{table}
+
+\begin{table}[H]
+\makebox[\textwidth][c]{
+\begin{NiceTabular}{|c|c c|c c|c c|c c|c c|c c|c c|c c|c c|}
+  \hline
+  i & use & def & in & out & in & out & in & out & in & out & in & out & in & out & in & out & in & out\\
+  \hline
+  0  & 0  & 3 & 0  &  & 0  & 3 & 0  & 3 & 0  & 0,3,8 & 0,8  & 0,3,8 & 0,8  & 0,3,8 & 0,8  & 0,3,8 & 0,8  & 0,3,8\\
+  1  & 3  &  & 3  &  & 3  & 0,8 & 0,3,8  & 0,8 & 0,3,8  & 0,8 & 0,3,8  & 0,8 & 0,3,8  & 0,8 & 0,3,8  & 0,8 & 0,3,8  & 0,8\\
+  2  &   &  &   &  &   &  &   &  &   &  &   &  &   &  &   &  &   & \\
+  3  &   &  &   &  &   &  &   &  &   &  &   &  &   &  &   &  &   & \\
+  4  &   &  &   &  &   & 0,8 & 0,8  & 0,8 & 0,8  & 0,8 & 0,8  & 0,8 & 0,8  & 0,8 & 0,8  & 0,8 & 0,8  & 0,8\\
+  5  & 0,8  & 6 & 0,8  &  & 0,8  & 6 & 0,8  & 6 & 0,8  & 6 & 0,8  & 6 & 0,8  & 6 & 0,8  & 0,6 & 0,8  & 0,6\\
+  6  & 6  & 7 & 6  &  & 6  & 6 & 6  & 6 & 6  & 6 & 6  & 0,6 & 0,6  & 0,6 & 0,6  & 0,6 & 0,6  & 0,6\\
+  7  & 6  & 8 & 6  &  & 6  &  & 6  & 0,8 & 0,6  & 0,8 & 0,6  & 0,8 & 0,6  & 0,8 & 0,6  & 0,8 & 0,6  & 0,8\\
+  8  &   &  &   & 0,8 & 0,8  & 0,8 & 0,8  & 0,8 & 0,8  & 0,8 & 0,8  & 0,8 & 0,8  & 0,8 & 0,8  & 0,8 & 0,8  & 0,8\\
+  \hline
+\end{NiceTabular}
+}
+\caption{Output of \texttt{dune exec build -{}- -t lva -v -r tests/count1.ll}}
+\label{tab:countll-steps}
+\end{table}
 

### 2024-01-06 article.tex
Finalize dataflow example

@@
 \end{figure}
-\noindent One thing that is particularly noteworthy is reversing of the flattened list of instructions on line 2 of Figure \ref{fig:dataflow}.
-After enumerating each entry so that the corresponding vertex is an array lookup away, it is reversed.
+\noindent One particularly noteworthy thing is reversing of the flattened list of instructions on line 2 of Figure \ref{fig:dataflow}.
+After enumerating each entry so that the corresponding vertex can be looked up in the \texttt{ids} array, it is reversed.
 The motivation for this is the fact that the \(\mathit{out}[n]\) of node \(n\) depends on \(\mathit{in}[s]\) for all successors \(s\).
@@
 
+Consider a simple program like Figure \ref{fig:countc} that only prints integers incrementally up until \texttt{argc}, in LLVM-{}-  that would consist of one loop with one phi node as seen in Figure \ref{fig:count1ll}.
+The number of iterations is nearly halfed when reverseing the list of instructions prior to dataflow as can be seen on \ref{tab:countll-rev-steps} when compared to \ref{tab:countll-steps}.
+
+\begin{figure}[H]
+  \begin{minipage}[b]{0.48\textwidth}
+     \centering
+     \begin{minted}[linenos,bgcolor=LightGray]{c}
+#include <stdio.h>
+
+int main(int argc, char **argv) {
+
+  for (int i = 1; i <= argc; i++) {
+    printf("%d!\n", i);
+  }
+
+  return 0;
+}
+     \end{minted}
+     \caption{\texttt{cat tests/count.ll}}\label{fig:countc}
+   \end{minipage}\hfill
+   \begin{minipage}[b]{0.48\textwidth}
+     \centering
+     \begin{minted}[linenos,bgcolor=LightGray]{llvm}
+define i32 @main (i32 %0, i8* %1) {
+ %3 = icmp slt i32 %0, 1
+ br i1 %3, label %4, label %5
+4:
+ ret i32 0
+5:
+ %6 = phi i32 [%8, %5], [1, %2]
+ %7 = call i32 (i8*, ...) @printf (i8* @.str, i32 %6)
+ %8 = add i32 %6, 1
+ %9 = icmp eq i32 %6, %0
+ br i1 %9, label %4, label %5
+}
+     \end{minted}
+     \caption{Stripped \texttt{clang -O1 -S -emit-llvm}}\label{fig:count1ll}
+   \end{minipage}
+\end{figure}
+
 %\begin{NiceTabular}{|W{c}{0.5cm}|W{c}{0.5cm} W{c}{0.5cm}|W{c}{0.5cm} W{c}{0.5cm}|W{c}{0.5cm} W{c}{0.5cm}|W{c}{0.5cm} W{c}{0.5cm}|W{c}{0.5cm} W{c}{0.5cm}|W{c}{0.5cm} W{c}{0.5cm}|W{c}{0.5cm} W{c}{0.5cm}|W{c}{0.5cm} W{c}{0.5cm}|}
@@
 \makebox[\textwidth][c]{
-\begin{NiceTabular}{|c|c c|c c|c c|c c|c c|c c|c c|}
-  \hline
-  i & use & def & in & out & in & out & in & out & in & out & in & out & in & out\\
-  \hline
-  0  & 0  & 3 & 0  & 3 & 0  & 0,3,8 & 0,8  & 0,3,8 & 0,8  & 0,3,8 & 0,8  & 0,3,8 & 0,8  & 0,3,8\\
-  1  & 3  &  & 3  & 0,8 & 0,3,8  & 0,8 & 0,3,8  & 0,8 & 0,3,8  & 0,8 & 0,3,8  & 0,8 & 0,3,8  & 0,8\\
-  2  &   &  &   &  &   &  &   &  &   &  &   &  &   & \\
-  3  &   &  &   &  &   &  &   &  &   &  &   &  &   & \\
-  4  &   &  &   & 0,8 & 0,8  & 0,8 & 0,8  & 0,8 & 0,8  & 0,8 & 0,8  & 0,8 & 0,8  & 0,8\\
-  5  & 0,8  & 6 & 0,8  & 6 & 0,8  & 6 & 0,8  & 6 & 0,8  & 6 & 0,8  & 0,6 & 0,8  & 0,6\\
-  6  & 6  & 7 & 6  & 6 & 6  & 6 & 6  & 6 & 6  & 0,6 & 0,6  & 0,6 & 0,6  & 0,6\\
-  7  & 6  & 8 & 6  &  & 6  &  & 6  & 0,8 & 0,6  & 0,8 & 0,6  & 0,8 & 0,6  & 0,8\\
-  8  &   &  &   &  &   & 0,8 & 0,8  & 0,8 & 0,8  & 0,8 & 0,8  & 0,8 & 0,8  & 0,8\\
+\begin{NiceTabular}{|c|c c|c c|c c|c c|c c|c c|c c|c c|}
+  \hline
+  i & use & def & in & out & in & out & in & out & in & out & in & out & in & out & in & out\\
+  \hline
+  0  & 0  & 3 & 0  & 3 & 0  & 3,8 & 0,8  & 3,8 & 0,8  & 3,8 & 0,8  & 0,3,8 & 0,8  & 0,3,8 & 0,8  & 0,3,8\\
+  1  & 3  &  & 3  & 8 & 3,8  & 8 & 3,8  & 8 & 3,8  & 0,8 & 0,3,8  & 0,8 & 0,3,8  & 0,8 & 0,3,8  & 0,8\\
+  2  &   &  &   &  &   &  &   &  &   &  &   &  &   &  &   & \\
+  3  &   &  &   &  &   &  &   &  &   &  &   &  &   &  &   & \\
+  4  &   &  &   & 8 & 8  & 8 & 8  & 8 & 8  & 0,8 & 0,8  & 0,8 & 0,8  & 0,8 & 0,8  & 0,8\\
+  5  & 8  & 6 & 8  & 6 & 8  & 6 & 8  & 0,6 & 0,8  & 0,6 & 0,8  & 0,6 & 0,8  & 0,6 & 0,8  & 0,6\\
+  6  & 6  & 7 & 6  & 6 & 6  & 0,6 & 0,6  & 0,6 & 0,6  & 0,6 & 0,6  & 0,6 & 0,6  & 0,6 & 0,6  & 0,6\\
+  7  & 6  & 8 & 6  & 0,6 & 0,6  & 0,6 & 0,6  & 0,6 & 0,6  & 0,6,8 & 0,6  & 0,6,8 & 0,6  & 0,6,8 & 0,6  & 0,6,8\\
+  8  & 0,6  & 9 & 0,6  & 9 & 0,6  & 9 & 0,6  & 8,9 & 0,6,8  & 8,9 & 0,6,8  & 8,9 & 0,6,8  & 0,8,9 & 0,6,8  & 0,8,9\\
+  9  & 9  &  & 9  &  & 9  & 8 & 8,9  & 8 & 8,9  & 8 & 8,9  & 0,8 & 0,8,9  & 0,8 & 0,8,9  & 0,8\\
   \hline
@@
 \makebox[\textwidth][c]{
-\begin{NiceTabular}{|c|c c|c c|c c|c c|c c|c c|c c|c c|c c|}
-  \hline
-  i & use & def & in & out & in & out & in & out & in & out & in & out & in & out & in & out & in & out\\
-  \hline
-  0  & 0  & 3 & 0  &  & 0  & 3 & 0  & 3 & 0  & 0,3,8 & 0,8  & 0,3,8 & 0,8  & 0,3,8 & 0,8  & 0,3,8 & 0,8  & 0,3,8\\
-  1  & 3  &  & 3  &  & 3  & 0,8 & 0,3,8  & 0,8 & 0,3,8  & 0,8 & 0,3,8  & 0,8 & 0,3,8  & 0,8 & 0,3,8  & 0,8 & 0,3,8  & 0,8\\
-  2  &   &  &   &  &   &  &   &  &   &  &   &  &   &  &   &  &   & \\
-  3  &   &  &   &  &   &  &   &  &   &  &   &  &   &  &   &  &   & \\
-  4  &   &  &   &  &   & 0,8 & 0,8  & 0,8 & 0,8  & 0,8 & 0,8  & 0,8 & 0,8  & 0,8 & 0,8  & 0,8 & 0,8  & 0,8\\
-  5  & 0,8  & 6 & 0,8  &  & 0,8  & 6 & 0,8  & 6 & 0,8  & 6 & 0,8  & 6 & 0,8  & 6 & 0,8  & 0,6 & 0,8  & 0,6\\
-  6  & 6  & 7 & 6  &  & 6  & 6 & 6  & 6 & 6  & 6 & 6  & 0,6 & 0,6  & 0,6 & 0,6  & 0,6 & 0,6  & 0,6\\
-  7  & 6  & 8 & 6  &  & 6  &  & 6  & 0,8 & 0,6  & 0,8 & 0,6  & 0,8 & 0,6  & 0,8 & 0,6  & 0,8 & 0,6  & 0,8\\
-  8  &   &  &   & 0,8 & 0,8  & 0,8 & 0,8  & 0,8 & 0,8  & 0,8 & 0,8  & 0,8 & 0,8  & 0,8 & 0,8  & 0,8 & 0,8  & 0,8\\
+\setlength{\tabcolsep}{1pt}
+\begin{NiceTabular}{|c|c c|c c|c c|c c|c c|c c|c c|c c|c c|c c|c c|c c|}
+  \hline
+  i & use & def & in & out & in & out & in & out & in & out & in & out & in & out & in & out & in & out & in & out & in & out & in & out\\
+  \hline
+  0  & 0  & 3 & 0  &  & 0  & 3 & 0  & 3 & 0  & 3,8 & 0,8  & 3,8 & 0,8  & 3,8 & 0,8  & 3,8 & 0,8  & 3,8 & 0,8  & 3,8 & 0,8  & 0,3,8 & 0,8  & 0,3,8\\
+  1  & 3  &  & 3  &  & 3  & 8 & 3,8  & 8 & 3,8  & 8 & 3,8  & 8 & 3,8  & 8 & 3,8  & 8 & 3,8  & 0,8 & 0,3,8  & 0,8 & 0,3,8  & 0,8 & 0,3,8  & 0,8\\
+  2  &   &  &   &  &   &  &   &  &   &  &   &  &   &  &   &  &   &  &   &  &   &  &   & \\
+  3  &   &  &   &  &   &  &   &  &   &  &   &  &   &  &   &  &   &  &   &  &   &  &   & \\
+  4  &   &  &   &  &   & 8 & 8  & 8 & 8  & 8 & 8  & 8 & 8  & 8 & 8  & 8 & 8  & 0,8 & 0,8  & 0,8 & 0,8  & 0,8 & 0,8  & 0,8\\
+  5  & 8  & 6 & 8  &  & 8  & 6 & 8  & 6 & 8  & 6 & 8  & 6 & 8  & 0,6 & 0,8  & 0,6 & 0,8  & 0,6 & 0,8  & 0,6 & 0,8  & 0,6 & 0,8  & 0,6\\
+  6  & 6  & 7 & 6  &  & 6  & 6 & 6  & 6 & 6  & 0,6 & 0,6  & 0,6 & 0,6  & 0,6 & 0,6  & 0,6 & 0,6  & 0,6 & 0,6  & 0,6 & 0,6  & 0,6 & 0,6  & 0,6\\
+  7  & 6  & 8 & 6  &  & 6  & 0,6 & 0,6  & 0,6 & 0,6  & 0,6 & 0,6  & 0,6,8 & 0,6  & 0,6,8 & 0,6  & 0,6,8 & 0,6  & 0,6,8 & 0,6  & 0,6,8 & 0,6  & 0,6,8 & 0,6  & 0,6,8\\
+  8  & 0,6  & 9 & 0,6  &  & 0,6  & 9 & 0,6  & 8,9 & 0,6,8  & 8,9 & 0,6,8  & 8,9 & 0,6,8  & 8,9 & 0,6,8  & 8,9 & 0,6,8  & 8,9 & 0,6,8  & 0,8,9 & 0,6,8  & 0,8,9 & 0,6,8  & 0,8,9\\
+  9  & 9  &  & 9  & 8 & 8,9  & 8 & 8,9  & 8 & 8,9  & 8 & 8,9  & 8 & 8,9  & 8 & 8,9  & 0,8 & 0,8,9  & 0,8 & 0,8,9  & 0,8 & 0,8,9  & 0,8 & 0,8,9  & 0,8\\
   \hline
@@
 
+This is unnoticeable for snippets of marginal size like most examples included, but when applied to functions with a significant amount of def-use paths, it can be quite consequential.
+For \texttt{tests/sha256.ll} for instance, the running times are effectively doubled on x86-64:
+
+\begin{minted}{bash}
+$ time dune exec build -- tests/sha256.ll  -t lva -r
+[..]
+dune exec build [..] -t lva -r  33,69s user 0,04s system 99% cpu 33,765 total
+$ time dune exec build -- tests/sha256.ll  -t lva
+[..]
+dune exec build [..] -t lva  17,26s user 0,05s system 99% cpu 17,335 total
+\end{minted}
+
 \subsection{Interference Graph}
@@
 This is done by building an interference graph, which is an undirected graph, whose nodes represent variables and edges that signify interference between them, i.e.
-variables \(a\) and \(b\) live at overlapping program points is represented with an edge \((a,b)\).
+variables \(a\) and \(b\) live at intersecting program points is represented with an edge \((a,b)\).
 

### 2024-01-06 article.tex
Expand CG section

@@
 \font\myfont=cmr12 at 24pt
-\title{\myfont Implementation and evaluation of register allocation used to translate \texttt{LLVM-{}-} to x86-64}
+\title{\myfont Implementation and evaluation of register allocation used to translate LLVM-{}- to x86-64}
 \author{William Welle Tange, 201809080}
@@
 
-This is done to more easily look up which graph vertex corresponds to the \(n\)th instruction/terminator.
+This is done to simplify lookup of which graph vertex corresponds to the \(n\)th instruction/terminator.
 
@@
 \]
-Although the Appel text \cite{tiger} describes interference of variables with concrete registers as well as overlapping variables, this isn't considered in this implementation.
+Although the Appel text \cite{tiger} describes interference of variables with concrete registers as well as overlapping variables, this isn't considered in this implementation. % FIXME: why??
 
@@
 Once an interference graph is constructed, the actual assignments can be found using graph coloring.
-Although this has long been known to be NP-complete, the heuristic as introduced in both the Appel and Aho et al.
-texts is a linear time approximation to this problem.
-It is based on an iterative approach wherein nodes known to be colorable are removed until either an empty graph remains, in which case the original graph \(G\) is \(k\)-colorable  or nodes with more than \(k\) neighbours remain.
-In this case a node is chosen to be spilled to the stack and removed.
-This is then repeated.
-
-Let \(k\) be the number of working registers available on the target architecture.
+Graph coloring is a problem in graph theory where the goal is to assign colors to the vertices of a graph in such a way that no adjacent vertices are assigned the same color.
+A graph that can be colored using \(k\) different colors is referred to as \(k\)-colorable and the minimum number of colors needed to be assigned is also called its' chromatic number.
+Although this problem is known to be NP-complete  \cites[229]{tiger}[510]{dragon}, the heuristic as introduced in
+both the Appel \cite[229]{tiger}, and Aho et al. \cites[557]{dragon} texts,
+is a linear time approximation to this problem is described below.
+
+\subsection{Coloring by simplification}
+
+The \textit{simplification} algorithm, also known as Chaitin's algorithm \cite{chaitin}, 
+ is an iterative approach wherein nodes known to be colorable are removed and pushed to a stack until either an empty graph remains, in which case the original graph is \(k\)-colorable,  or nodes with more than \(k\) neighbours remain.
+In this case a node is chosen to be spilled after which the process is started over.
+This is repeated until every node has been removed successfully, after which the vertices are assigned by popping them from the stack.
+Each of these are known to be \(k\)-colorable because of the earlier criteria of having less than \(k\) neighbors.
+
+%Let \(k\) be the number of working registers available on the target architecture.
 After building the interference graph \(G\), a node \(n\) with fewer than \(k\) neighbors is chosen.
@@
 %It must hold that \(G-\{n\}\) is \(k\)-colorable if \(G\) is \(k\)-colorable
-
-\subsection{Coloring by simplification}
 

### 2024-01-06 article.tex
Insert some \noindents, change all LLVM-- to LLVM-{}-

@@
 
-The implementation is written in OCaml for the most part, although there is some marginal use of Python to interact with the LLDB debugger.
+The implementation is written in OCaml for the most part, although there is some marginal use of Python to interact with the LLVM toolchain's debugger 'LLDB'.
 
@@
 \end{figure}
-Another approach is that of Figure \ref{fig:mulll1}, which uses so-called 'phi nodes' instead.
+\noindent Another approach is that of Figure \ref{fig:mulll1}, which uses so-called 'phi nodes' instead.
 These are a type of instruction that evaluates to a specified operand depending on which predecessor block was executed immediately prior.
@@
 
-This is done to simplify lookup of which graph vertex corresponds to the \(n\)th instruction/terminator.
+\noindent This is done to simplify lookup of which graph vertex corresponds to the \(n\)th instruction/terminator.
+Labels, while present in the flattened list, aren't considered as they have no inherent function other than labelling the leader instruction to branch to.
 
@@
 Variables that are in use at a given program point are considered \textit{live}, and although variables can be assigned the same register, %if they are not in interference
-variables that are live at the same time (i.e. at intersecting program points) cannot, in which case they are also said to be in interference with one another.
+variables that are live at the same time (i.e. at intersecting program points) cannot, %FIXME: why not?
+in which case they are also said to be in interference with one another.
  Variables that are not in interference can be assigned the same register, and finding the precise points at which any variable is live is trivial for linear sequences of instructions.
@@
 
-This is unnoticeable for snippets of marginal size like most examples included, but when applied to functions with a significant amount of def-use paths, it can be quite consequential.
-For \texttt{tests/sha256.ll} for instance, the running times are effectively doubled on x86-64:
+\noindent This is unnoticeable for snippets of marginal size like most examples included, but when applied to functions with a significant amount of def-use paths, it can be quite consequential.
+For \texttt{tests/sha256.ll} for instance, the running time of this analysis is effectively doubled:
 
@@
 $ time dune exec build -- tests/sha256.ll  -t lva -r
-[..]
-dune exec build [..] -t lva -r  33,69s user 0,04s system 99% cpu 33,765 total
+33,69s user 0,04s system 99% cpu 33,765 total
 $ time dune exec build -- tests/sha256.ll  -t lva
-[..]
-dune exec build [..] -t lva  17,26s user 0,05s system 99% cpu 17,335 total
+17,26s user 0,05s system 99% cpu 17,335 total
 \end{minted}
@@
 
-The purpose for conducting dataflow analysis as above is finding variables that may be assigned the same register.
+The purpose of conducting dataflow analysis as above is finding variables that may be assigned the same register.
 This is done by building an interference graph, which is an undirected graph, whose nodes represent variables and edges that signify interference between them, i.e.
@@
 \end{equation}
-add edges to all \textit{live-out} variables not in \(\{b_1, ..., b_n\}\):
+add edges to all \textit{live-out} variables not in \(\{b_1, ..., b_n\}\)
 \[
-  \forall b_j \in \mathit{out}[i] \setminus \{b_1, ..., b_n\}, \text{add\_edge}(a, b_j)
+  \forall b_j \in \mathit{out}[i] \setminus \{b_1, ..., b_n\}, \text{add\_edge}(a, b_j),
 \]
-For any other instruction that defines a variable \(a\)
+and for any other instruction that defines a variable \(a\)
 \[
@@
 \]
-Although the Appel text \cite{tiger} describes interference of variables with concrete registers as well as overlapping variables, this isn't considered in this implementation. % FIXME: why??
+Although the Appel text \cite{tiger} describes interference of variables with concrete registers as well as overlapping variables, this isn't considered in this implementation % FIXME: why??
+for simplicity's sake.
+
+Again the \texttt{OCamlgraph} library is used, but this time \texttt{Imperative.Graph.Abstract} is parameterized over the aforementioned set type \texttt{S.SS.t}. % FIXME: whyyy????
+
+%module G = Imperative.Graph.Abstract (struct
+  %type t = S.SS.t
+%end)
 
@@
 
-\subsection{\lstinline!LLVM--! instruction set}
+\subsection{LLVM-{}- instruction set}
 
@@
 %%%
-\subsection{Translating \lstinline!LLVM--! to x86}
-
-Translating each IR instruction to x86 correctly is a matter of eliminating unintended side-effects. Each LLVM-- instruction is defined to have only one purpose, %% FIXME is purpose the right word?
+\subsection{Translating LLVM-{}- to x86}
+
+Translating each IR instruction to x86 correctly is a matter of eliminating unintended side-effects. Each LLVM-{}- instruction is defined to have only one purpose, %% FIXME is purpose the right word?
 as concepts such as calling conventions, stack frames, or a \lstinline!FLAGS! register are completely abstracted over in order to remain platform independent.
@@
 
-Notice for instance Figure \ref{fig:square0ll} which has three phi nodes: two in  block 3 and one in block 9, i.e.
+\noindent Notice for instance Figure \ref{fig:square0ll} which has three phi nodes: two in  block 3 and one in block 9, i.e.
 \texttt{\%0} needs to be copied to \texttt{\%4} when control is transferred from block 1 and \texttt{\%7} when transferred from block 3.
@@
 A more elaborate way of testing correctness of translation is to assert that there are no effects except for those intended.
-One way of achieving this is to check if a group of x86 instructions representing a single LLVM-- operation change any values except those expected.
-An approach to this is to attach a debugger, insert a breakpoint before every LLVM-- instruction, and ensure that only the intended registers are altered.
+One way of achieving this is to check if a group of x86 instructions representing a single LLVM-{}- operation change any values except those expected.
+An approach to this is to attach a debugger, insert a breakpoint before every LLVM-{}- instruction, and ensure that only the intended registers are altered.
 
@@
 While there are many approaches to assessing the quality of translation, one of the primary means is to simply measure the time taken to execute the code generated. 
-Assuming the translation is sound given the previous steps taken to validate correctness in translating \lstinline!LLVM--!
-to x86, it is a solid foundation for future optimizations.
+Assuming the translation is sound given the previous steps taken to validate correctness in translating LLVM-{}- to x86, it is a solid foundation for future optimizations.
 While there are certainly other factors worth measuring, only runtime performance is considered for this project.

### 2024-01-06 article.tex
Add some of interf  func

@@
 
+\begin{figure}[H]
+\begin{minted}[linenos,bgcolor=LightGray]{ocaml}
+let interf (params : Ll.uid list) (insns : Cfg.insn list) _ (out : S.SS.t array)
+    : G.V.t S.table * G.t =
+  let g = G.create () in
+  let vert e t = (* .. *) in
+  let vert2 t e = vert e t |> snd and vert3 e t = vert e t |> snd in
+  (* add params *)
+  let t = List.fold_left vert2 S.ST.empty params in
+  (* connect all params *)
+  let edge v1 v2 t = (* .. *) in
+  let param t p1 =
+    let param t p2 = if p1 <> p2 then edge p1 p2 t else t in
+    List.fold_left param t params
+  in
+  let t = List.fold_left param t params in
+  (* add all defs *)
+  let setverts t s = S.SS.fold vert3 s t in
+  let t = List.map (def S.SS.empty) insns |> List.fold_left setverts t in
+  let defoutedges t (i, n) =
+    let defs = def S.SS.empty n in
+    let outedge e1 t = S.SS.fold (edge e1) out.(i) t in
+    S.SS.fold outedge defs t
+  in
+  let t = List.mapi (fun i n -> (i, n)) insns |> List.fold_left defoutedges t in
+  (t, g)
+\end{minted}
+\caption{Overall implementation of \texttt{dataflow}\label{fig:dataflow}}
+\end{figure}
+
 %module G = Imperative.Graph.Abstract (struct

### 2024-01-07 article.tex
Begin work on graph coloring

@@
 \makebox[\textwidth][c]{
-\setlength{\tabcolsep}{1pt}
+\setlength{\tabcolsep}{0.9pt}
 \begin{NiceTabular}{|c|c c|c c|c c|c c|c c|c c|c c|c c|c c|c c|c c|c c|}
@@
 \end{minted}
-\caption{Overall implementation of \texttt{dataflow}\label{fig:dataflow}}
+\caption{Overall implementation of \texttt{interf}\label{fig:interf}}
+\end{figure}
+
+\noindent For instance, the LLVM-{}- program of Figure \ref{fig:countdown.ll}, a simple handwritten program that counts down from \texttt{\%argc}, would translate the interference graph of Figure \ref{fig:countdown.dot}.
+% TODO: explain step wise?
+
+\begin{figure}[H]
+     \centering
+     \begin{minipage}[b]{0.57\textwidth}
+     \begin{minted}[bgcolor=LightGray]{llvm}
+define i32 @main (i32 %argc, i8** %argv) {
+entry:
+ br label %loop
+loop:
+ %c1 = phi i32 [%argc, %entry], [%c2, %loop]
+ call void @printf (i8* @fmt, i32 %c1)
+ %c2 = sub i32 %c1, 1
+ %cn = icmp sgt i32 %c2, 0
+ br i1 %cn, label %loop, label %exit
+exit:
+ ret i32 0
+}
+     \end{minted}
+     \caption{\texttt{cat tests/countdown.ll}}\label{fig:countdown.ll}
+   \end{minipage}
+   \begin{minipage}[b]{0.42\textwidth}
+     \centering
+     \includesvg[height=15em]{countdown.simple}
+     \caption{\texttt{dune exec build -- -t dot}}\label{fig:countdown.dot}
+   \end{minipage}
 \end{figure}
@@
 It is pushed to a stack to preserve the order in which they are removed so that \(G\) can be rebuilt and the corresponding registers can be assigned correctly once a \(k\)-colorable assignment is found.
+
+To perform the simplication algorithm on the previous example of Figure \ref{fig:countdown.ll}, its interference graph seen on Figure \ref{fig:countdown.dot} only has one node of degree 4 with the ones remaining all being \(\leq 2\).
+This wouldn't be a problem on modern day processors most of which have \textasciitilde 16 or so GPRs, but for the sake of clearness one can imagine a register machine of \(k=2\) working registers.
 

### 2024-01-07 article.tex
Nice progress on grapph erxamples

@@
 
+If at first a single node of insignificant degree (i.e.
+\(< 2\)) is selected, it is then removed and pushed to a stack.
+In Figure \ref{fig:countdown.dot} only \texttt{\%argv} and \texttt{\%c1} are of insignificant degree, so either of those is chosen to be removed (represented with dashed borders and edges).
+
+\begin{figure}[H]
+\centering
+  \begin{minipage}[b]{0.48\textwidth}
+    \begin{subfigure}{0.3\textwidth}
+      \includesvg[height=12em]{countdown.simple2}
+      \caption{Graph}\label{fig:countdown.dot2}
+    \end{subfigure}
+\hfill
+    \begin{subfigure}{0.18\textwidth}
+      \texttt{\%c1}
+      \caption{Stack}
+      \label{fig:second}
+    \end{subfigure}
+   \end{minipage}
+  \begin{minipage}[b]{0.48\textwidth}
+    \begin{subfigure}{0.3\textwidth}
+      \includesvg[height=12em]{countdown.simple2}
+      \caption{Graph}\label{fig:countdown.dot2}
+    \end{subfigure}
+\hfill
+    \begin{subfigure}{0.18\textwidth}
+      \texttt{\%c1}
+      \caption{Stack}
+      \label{fig:second}
+    \end{subfigure}
+   \end{minipage}
+\end{figure}
+
 %  (i.e. \(\mathit{deg}(n) < k\))

### 2024-01-07 article.tex
Even more examples
Finalize graphs with colors and shit
Elaborate all the things

--- 2024-01-07 article.tex
@@
 
+If at first a single node of insignificant degree (i.e.
+\(< 2\)) is selected, it is then removed and pushed to a stack.
+In Figure \ref{fig:countdown.dot} only \texttt{\%argv} and \texttt{\%c1} are of insignificant degree, so either of those is chosen to be removed (represented with dashed borders and edges) as can be seen on Figure \ref{fig:countdown.dot2} and \ref{fig:countdown.dot3} and pushed to the stack of variables to be colored as can be seen on Figure \ref{fig:countdown.stack2} and \ref{fig:countdown.stack3}.
+
+% Side by side figures 
+\begin{figure}[H]
+  \centering
+  \begin{minipage}[b]{0.32\linewidth}
+    \begin{subfigure}[b]{0.70\textwidth}
+      \includesvg[height=10em]{countdown.simple2}
+      \caption{Graph}\label{fig:countdown.dot2}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.27\textwidth}
+      \centering
+        \texttt{\%c1}
+        \vspace{0.5em}
+      \caption{Stack}\label{fig:countdown.stack2}
+    \end{subfigure}
+    \caption{Push \texttt{\%c1}}
+  \end{minipage}%
+  \begin{minipage}[b]{0.32\linewidth}
+    \begin{subfigure}[b]{0.70\textwidth}
+      \includesvg[height=10em]{countdown.simple3}
+      \caption{Graph}\label{fig:countdown.dot3}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.27\textwidth}
+      \centering
+        \texttt{\%argv}\\
+        \texttt{\%c1}
+        \vspace{0.5em}
+      \caption{Stack}\label{fig:countdown.stack3}
+    \end{subfigure}
+    \caption{Push \texttt{\%argv}}
+  \end{minipage}%
+  \begin{minipage}[b]{0.32\linewidth}
+    \begin{subfigure}[b]{0.70\textwidth}
+      \includesvg[height=6em]{countdown.simple4}
+      \caption{Graph}\label{fig:countdown.dot4}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.27\textwidth}
+      \caption{Stack}\label{fig:countdown.stack4}
+    \end{subfigure}
+    \caption{Spill \texttt{\%c2}}
+  \end{minipage}%
+\end{figure}
+\noindent From this point onwards  nothing can be done to further simplify the graph as all of the remaining nodes are of significant degree (\(\geq 2\)).
+Therefore, one of the remaining nodes is spilled to memory and entirely removed from the graph as can be seen on Figure \ref{fig:countdown.dot4} and the stack reset as can be seen on Figure \ref{fig:countdown.stack4}.
+
+\begin{figure}[H]
+  \centering
+  \begin{minipage}[c]{0.32\linewidth}
+    \begin{subfigure}[b]{0.70\textwidth}
+      \includesvg[height=6em]{countdown.simple5} 
+      \caption{Graph}\label{fig:countdown.dot5}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.27\textwidth}
+      \centering
+        \texttt{\%cn}
+        \vspace{0.5em}
+      \caption{Stack}\label{fig:countdown.stack5}
+    \end{subfigure}
+    \caption{Push \texttt{\%cn}}
+  \end{minipage}%
+  \begin{minipage}[c]{0.32\linewidth}
+    \begin{subfigure}[b]{0.70\textwidth}
+      \includesvg[height=6em]{countdown.simple6} 
+      \caption{Graph}\label{fig:countdown.dot6}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.27\textwidth}
+      \centering
+        \texttt{\%argv}\\
+        \texttt{\%cn}
+        \vspace{0.5em}
+      \caption{Stack}\label{fig:countdown.stack6}
+    \end{subfigure}
+    \caption{Push \texttt{\%argv}}
+  \end{minipage}%
+  \begin{minipage}[c]{0.32\linewidth}
+    \begin{subfigure}[b]{0.70\textwidth}
+      \includesvg[height=6em]{countdown.simple7} 
+      \caption{Graph}\label{fig:countdown.dot7}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.27\textwidth}
+      \centering
+      \texttt{\%argc}\\
+      \texttt{\%argv}\\
+      \texttt{\%cn}
+      \vspace{0.5em}
+      \caption{Stack}\label{fig:countdown.stack7}
+    \end{subfigure}
+    \caption{Push \texttt{\%argc}}
+  \end{minipage}%
+\end{figure}
+
+\noindent After spilling \texttt{\%c2}, the process is repeated.
+First by pushing \texttt{\%cn} (Figure \ref{fig:countdown.dot5} and \ref{fig:countdown.stack5}), then \texttt{\%argv} (Figure \ref{fig:countdown.dot6} and \ref{fig:countdown.stack6}) and finally \texttt{\%argc} (Figure \ref{fig:countdown.dot7} and \ref{fig:countdown.stack7}).
+One thing of note is the fact that the order in which this is happening is irrelevant, so long as variables pushed are of insignificant degree.
+For instance, \texttt{\%argc} was of degree 3 in Figure \ref{fig:countdown.dot4}, but as more and more of its neighbors are pushed it's degree decreases to a point where only \texttt{\%argc} and \texttt{\%c1} are yet to be pushed to the stack.
+
+\begin{figure}[H]
+  \centering
+  \begin{minipage}[c]{0.32\linewidth}
+    \begin{subfigure}[b]{0.70\textwidth}
+      \includesvg[height=6em]{countdown.simple8} 
+      \caption{Graph}\label{fig:countdown.dot8}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.27\textwidth}
+      \centering
+      \texttt{\%c1}\\
+      \texttt{\%argc}\\
+      \texttt{\%argv}\\
+      \texttt{\%cn}
+      \vspace{0.5em}
+      \caption{Stack}\label{fig:countdown.stack8}
+    \end{subfigure}
+    \caption{Push \texttt{\%c1}}
+  \end{minipage}%
+  \begin{minipage}[c]{0.32\linewidth}
+    \begin{subfigure}[b]{0.70\textwidth}
+      \includesvg[height=6em]{countdown.simple9} 
+      \caption{Graph}\label{fig:countdown.dot9}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.27\textwidth}
+      \centering
+      \texttt{\%argc}\\
+      \texttt{\%argv}\\
+      \texttt{\%cn}
+      \vspace{0.5em}
+      \caption{Stack}\label{fig:countdown.stack9}
+    \end{subfigure}
+    \caption{Pop \texttt{\%c1}}
+  \end{minipage}%
+  \begin{minipage}[c]{0.32\linewidth}
+    \begin{subfigure}[b]{0.70\textwidth}
+      \includesvg[height=6em]{countdown.simple10} 
+      \caption{Graph}\label{fig:countdown.dot10}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.27\textwidth}
+      \centering
+      \texttt{\%argv}\\
+      \texttt{\%cn}
+      \vspace{0.5em}
+      \caption{Stack}\label{fig:countdown.stack10}
+    \end{subfigure}
+    \caption{Pop \texttt{\%argc}}
+  \end{minipage}%
+\end{figure}
+
+Dolor sit amet badungadong and so on
+
+\begin{figure}[H]
+  \centering
+  \begin{minipage}[c]{0.32\linewidth}
+    \begin{subfigure}[b]{0.70\textwidth}
+      \includesvg[height=6em]{countdown.simple11} 
+      \caption{Graph}\label{fig:countdown.dot11}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.27\textwidth}
+      \centering
+      \texttt{\%cn}
+      \vspace{0.5em}
+      \caption{Stack}\label{fig:countdown.stack11}
+    \end{subfigure}
+    \caption{Pop \texttt{\%argv}} 
+  \end{minipage}%
+  \begin{minipage}[c]{0.32\linewidth}
+    \begin{subfigure}[b]{0.70\textwidth}
+      \includesvg[height=6em]{countdown.simple12} 
+      \caption{Graph}\label{fig:countdown.dot12}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.27\textwidth}
+      \caption{Stack}\label{fig:countdown.stack12} 
+    \end{subfigure}
+    \caption{Pop \texttt{\%c1}}
+  \end{minipage}%
+\end{figure}
+
 %  (i.e. \(\mathit{deg}(n) < k\))

### 2024-01-07 article.tex
Explain the last graphs

@@
     \end{subfigure}
-    \caption{Push \texttt{\%c1}}
+    \caption{Push \texttt{\%c1}}\label{fig:countdown8}
   \end{minipage}%
@@
     \end{subfigure}
-    \caption{Pop \texttt{\%c1}}
+    \caption{Pop \texttt{\%c1}}\label{fig:countdown9}
   \end{minipage}%
@@
 
-Dolor sit amet badungadong and so on
+\noindent As the last variable \texttt{\%c1} is pushed in Figure \ref{fig:countdown8} and an empty graph remains, the process of reconstructing the graph can begin.
+When popping \texttt{\%c1} in Figure \ref{fig:countdown9}, it can be assigned any of the two colors as it has no neigbors adjacent, so either is chosen.
+On the other hand, when assigning \texttt{\%argc}, it's only adjacent naighbor \texttt{\%c1} is already assigned red so only blue remains.
+Likewise, for \texttt{\%argv} and \texttt{\%cn} their only immediately adjacent neighbor of \texttt{\%argc} causes them to be assigned red as can be seen on Figure \ref{fig:countdown11} and \ref{fig:countdown12}.
 
@@
     \end{subfigure}
-    \caption{Pop \texttt{\%argv}} 
+    \caption{Pop \texttt{\%argv}} \label{fig:countdown11}
   \end{minipage}%
@@
     \end{subfigure}
-    \caption{Pop \texttt{\%c1}}
+    \caption{Pop \texttt{\%c1}}\label{fig:countdown12} 
   \end{minipage}%

### 2024-01-07 article.tex
Begin coalescing

@@
 Because of this, only strategies that produce a \(k\)-colorable graph are worth considering, as additional spills negate the purpose entirely.
+One of the  strategies introduced in the Appel text is the \textit{Briggs} strategy  \cite[232]{tiger}, in which nodes \(a\) and \(n\) can be coalesced if and only if the resulting node \(ab\) is of insignificant degree  (i.e.
+\(\leq K\)).
+
+The strategy Since the only move operations in use in LLVM-{}- are phi nodes,
 

### 2024-01-07 article.tex
Attempt to readjust width of graphs
Display resulting assignments
Talk a bit more about coalescing

--- 2024-01-07 article.tex
@@
   \centering
-  \begin{minipage}[b]{0.32\linewidth}
+  \begin{minipage}[b]{0.31\linewidth}
     \begin{subfigure}[b]{0.70\textwidth}
@@
   \end{minipage}%
-  \begin{minipage}[b]{0.32\linewidth}
+  \begin{minipage}[b]{0.31\linewidth}
     \begin{subfigure}[b]{0.70\textwidth}
@@
 \end{figure}
+\noindent The final coloring of Figure \ref{fig:countdown12} leads to the assignment 
+\begin{align*}
+  \mathit{assign[\texttt{\%argv}]} = \mathit{assign[\texttt{\%c1}]} = \mathit{assign[\texttt{\%cn}]} &= \texttt{\%red}\\
+  \mathit{assign[\texttt{\%argc}]} &= \texttt{\%blue}
+\end{align*}
+and the spilled variable \texttt{\%c2}
+\[
+  \mathit{assign[\texttt{\%c2}]} = \texttt{stackslot(0)}
+\]
 
@@
 Because of this, only strategies that produce a \(k\)-colorable graph are worth considering, as additional spills negate the purpose entirely.
+One of the  strategies introduced in the Appel text is the \textit{Briggs} strategy  \cite[232]{tiger}, in which nodes \(a\) and \(b\) can be coalesced if and only if the resulting node \(ab\) is of insignificant degree  (i.e.
+\(\leq K\)).
+
+Since the only move operations in use in LLVM-{}- are phi nodes, only variables that are subject to this sort of operation are eligible to be coalesced in this form. 
+Variables that may be coalesced can be thought of as having a different sort of edge, quite the opposite of interference edges, in fact they can be thought of as a sort of 'preference edge'.
+Strictly speaking, one could optimize many more mov instructions away if
 

### 2024-01-07 article.tex
More coalescing

@@
      \includesvg[height=15em]{countdown.simple}
-     \caption{\texttt{dune exec build -- -t dot}}\label{fig:countdown.dot}
+     \caption{\texttt{dune exec build -{}- -t dot}}\label{fig:countdown.dot}
    \end{minipage}
@@
 Variables that may be coalesced can be thought of as having a different sort of edge, quite the opposite of interference edges, in fact they can be thought of as a sort of 'preference edge'.
-Strictly speaking, one could optimize many more mov instructions away if
+
+When applied to the previous example (Figure \ref{fig:countdown.ll}), only \texttt{\%argc} and \texttt{\%c2} may be coalesced with \texttt{\%c1} as this is the only mov instruction.
+
+\begin{figure}[H]
+     \centering
+   \begin{minipage}[b]{0.42\textwidth}
+     \centering
+     \includesvg[height=10em]{countdown.briggs}
+     \caption{\texttt{dune exec build -{}- -t dot}}\label{fig:countdown.dot}
+   \end{minipage}
+\end{figure}
+
+%Strictly speaking, one could optimize many more mov instructions away if
 

### 2024-01-08 article.tex
Sven approved

@@
 \font\myfont=cmr12 at 24pt
-\title{\myfont Implementation and evaluation of register allocation used to translate LLVM-{}- to x86-64}
+\title{\myfont Implementation and evaluation of register allocation for translating LLVM-{}- to x86-64}
 \author{William Welle Tange, 201809080}
@@
 The motivation for this is the fact that the \(\mathit{out}[n]\) of node \(n\) depends on \(\mathit{in}[s]\) for all successors \(s\).
-Since most nodes have one successor immediately after, populating this before hands is significantly more efficient.
+Since most nodes have one successor immediately after, populating this beforehands is significantly more efficient.
 
@@
 %\begin{NiceTabular}{|W{c}{0.5cm}|W{c}{0.5cm} W{c}{0.5cm}|W{c}{0.5cm} W{c}{0.5cm}|W{c}{0.5cm} W{c}{0.5cm}|W{c}{0.5cm} W{c}{0.5cm}|W{c}{0.5cm} W{c}{0.5cm}|W{c}{0.5cm} W{c}{0.5cm}|W{c}{0.5cm} W{c}{0.5cm}|W{c}{0.5cm} W{c}{0.5cm}|}
+
+% FIXME: forklarer tabellerne nedenunder
 
@@
 One thing of note is the fact that the order in which this is happening is irrelevant, so long as variables pushed are of insignificant degree.
-For instance, \texttt{\%argc} was of degree 3 in Figure \ref{fig:countdown.dot4}, but as more and more of its neighbors are pushed it's degree decreases to a point where only \texttt{\%argc} and \texttt{\%c1} are yet to be pushed to the stack.
+For instance, \texttt{\%argc} was of degree 3 in Figure \ref{fig:countdown.dot4}, but as more and more of its neighbors are pushed its degree decreases to a point where only \texttt{\%argc} and \texttt{\%c1} are yet to be pushed to the stack.
 
@@
 When popping \texttt{\%c1} in Figure \ref{fig:countdown9}, it can be assigned any of the two colors as it has no neigbors adjacent, so either is chosen.
-On the other hand, when assigning \texttt{\%argc}, it's only adjacent naighbor \texttt{\%c1} is already assigned red so only blue remains.
+On the other hand, when assigning \texttt{\%argc}, it's only adjacent neighbor \texttt{\%c1} is already assigned red so only blue remains.
 Likewise, for \texttt{\%argv} and \texttt{\%cn} their only immediately adjacent neighbor of \texttt{\%argc} causes them to be assigned red as can be seen on Figure \ref{fig:countdown11} and \ref{fig:countdown12}.
@@
 
+Consider the above example but with k geq 2, the \%c1 and
+
 \begin{figure}[H]
@@
      \includesvg[height=10em]{countdown.briggs}
-     \caption{\texttt{dune exec build -{}- -t dot}}\label{fig:countdown.dot}
+     \caption{\texttt{dune exec build -{}- -t dot}}\label{fig:countdown.briggs.dot}
    \end{minipage}

### 2024-01-08 article.tex
Fix sizing of graphs

@@
     \begin{subfigure}[b]{0.70\textwidth}
-      \includesvg[height=10em]{countdown.simple2}
+      \includesvg[height=9em]{countdown.simple2}
       \caption{Graph}\label{fig:countdown.dot2}
@@
     \begin{subfigure}[b]{0.70\textwidth}
-      \includesvg[height=10em]{countdown.simple3}
+      \includesvg[height=9em]{countdown.simple3}
       \caption{Graph}\label{fig:countdown.dot3}
@@
     \begin{subfigure}[b]{0.70\textwidth}
-      \includesvg[height=6em]{countdown.simple4}
+      \includesvg[height=5.5em]{countdown.simple4}
       \caption{Graph}\label{fig:countdown.dot4}
@@
     \begin{subfigure}[b]{0.70\textwidth}
-      \includesvg[height=6em]{countdown.simple5} 
+      \includesvg[height=5.5em]{countdown.simple5} 
       \caption{Graph}\label{fig:countdown.dot5}
@@
     \begin{subfigure}[b]{0.70\textwidth}
-      \includesvg[height=6em]{countdown.simple6} 
+      \includesvg[height=5.5em]{countdown.simple6} 
       \caption{Graph}\label{fig:countdown.dot6}
@@
     \begin{subfigure}[b]{0.70\textwidth}
-      \includesvg[height=6em]{countdown.simple7} 
+      \includesvg[height=5.5em]{countdown.simple7} 
       \caption{Graph}\label{fig:countdown.dot7}
@@
     \begin{subfigure}[b]{0.70\textwidth}
-      \includesvg[height=6em]{countdown.simple8} 
+      \includesvg[height=5.5em]{countdown.simple8} 
       \caption{Graph}\label{fig:countdown.dot8}
@@
     \begin{subfigure}[b]{0.70\textwidth}
-      \includesvg[height=6em]{countdown.simple9} 
+      \includesvg[height=5.5em]{countdown.simple9} 
       \caption{Graph}\label{fig:countdown.dot9}
@@
     \begin{subfigure}[b]{0.70\textwidth}
-      \includesvg[height=6em]{countdown.simple10} 
+      \includesvg[height=5.5em]{countdown.simple10} 
       \caption{Graph}\label{fig:countdown.dot10}
@@
     \begin{subfigure}[b]{0.70\textwidth}
-      \includesvg[height=6em]{countdown.simple11} 
+      \includesvg[height=5.5em]{countdown.simple11} 
       \caption{Graph}\label{fig:countdown.dot11}
@@
     \begin{subfigure}[b]{0.70\textwidth}
-      \includesvg[height=6em]{countdown.simple12} 
+      \includesvg[height=5.5em]{countdown.simple12} 
       \caption{Graph}\label{fig:countdown.dot12}
@@
 
-Consider the above example but with k geq 2, the \%c1 and
+Consider the above example but with some \(k \geq 2\). the \%c1 and
 
@@
      \centering
-   \begin{minipage}[b]{0.42\textwidth}
+   \begin{minipage}[b]{0.5\textwidth}
      \centering
@@
 \section{Linear Scan}
-Another approach that does not depend on liveness analysis i the linear scan algorithm.
+Another approach that also depends on liveness analysis is the linear scan algorithm.
 It is a simple and efficient approach that performs a linear pass over IR to produce a register assignment significantly faster than all of the steps involved with graph coloring.
+That said, it may be generate more imprecise assignment and therefore also be less performant at runtime, although the fact that assignment is fast is a benefit in some cases.
 

### 2024-01-08 article.tex
Some linear

@@
 Most CPUs are \textit{register machines}, in that they use a limited set of \textit{general-purpose registers} (GPRs) to store working values in combination with \textit{random access memory} (RAM) for mid-term, and other I/O peripherals for long-term storage.
-This can largely be attributed to performance, as register machines routinely outperform \textit{stack machines} \parencite{ShiYunhe2008VmsS} that are often used in VMs.
+This can largely be attributed to performance, as register machines routinely outperform \textit{stack machines} \parencite{ShiYunhe2008VmsS} that are often used in \textit{virtual machines} (VMs).
 Although register machines generally also have a stack available, as being limited to mere bytes of storage is simply infeasible for most large scale applications, using it compared to GPRs is orders of magnitudes slower \cite{latency}.
@@
 %% TODO: compare with time-complexity/compile-time duration?
-The primary sources will be \textit{Modern Compiler Implementation in ML} \parencite{tiger} and \textit{Compilers: Principles, techniques, and tools} \parencite{dragon}, in addition to publications concerning the linear scan approach.
+The primary sources are \textit{Modern Compiler Implementation in ML} \parencite{tiger} and \textit{Compilers: Principles, techniques, and tools} \parencite{dragon} with regards to static analysis and graph coloring, in addition to \textit{Linear Scan Register Allocation} \cite{linear} concerning linear scan.
 
@@
 Most languages, if targeting native execution on a CPU, will have an LLVM implementation.
-This is true for C/C++ (Clang compiler \cite{clang}), D (LDC compiler \cite{dlang}), Swift (official Swift compiler \cite{swift}) and Rust (officilal \texttt{rustc} compiler \cite{rustc}).
+This is true for C/C++ (Clang compiler \cite{clang}), D (LDC compiler \cite{dlang}), Swift (official Swift compiler \cite{swift}) and Rust (official \texttt{rustc} compiler \cite{rustc}).
 The use of LLVM allows for them to be compiled to completely unforeseen architectures, like web browsers with WebAssembly or even graphics cards with compute kernels \cite{backend}.
@@
      \centering
-   \begin{minipage}[b]{0.5\textwidth}
+   \begin{minipage}[b]{0.42\textwidth}
      \centering
@@
 It is a simple and efficient approach that performs a linear pass over IR to produce a register assignment significantly faster than all of the steps involved with graph coloring.
-That said, it may be generate more imprecise assignment and therefore also be less performant at runtime, although the fact that assignment is fast is a benefit in some cases.
+That said, it may be generate more imprecise assignments and therefore also be less performant at runtime, so the utility of such an approach depends on context.
+As JIT compilation 
+%Whether or not to  a cost depends on the context, as JIT compilation will often 
+%reduced compile-time in  the fact that assignment is fast is a benefit in some cases.
 

### 2024-01-09 article.tex
Some notes

@@
 
+The purpose is only to assign the same registers to different variables. You don't need to note down information about blocks etc.
+
 %Strictly speaking, one could optimize many more mov instructions away if
@@
 It is a simple and efficient approach that performs a linear pass over IR to produce a register assignment significantly faster than all of the steps involved with graph coloring.
-That said, it may be generate more imprecise assignments and therefore also be less performant at runtime, so the utility of such an approach depends on context.
-As JIT compilation 
+That said, it may generate more imprecise assignments and therefore also be less performant at runtime, so the utility of such an approach depends on context.
+Some compilers typically compile code into machine code ahead of time (AOT) while others depend on just-in-time (JIT) compilation which defers the compilation of code until runtime.
+Therefore linear scan is an approach to register allocation that may be used to achieve reasonably performant runtime compilation.
 %Whether or not to  a cost depends on the context, as JIT compilation will often 
@@
 
-The primary text is \textit{Linear Scan Register Allocation} \cite{linear}, the execution of which also relies on prior liveness analysis.
+The primary text is \textit{Linear Scan Register Allocation} \cite{linear}, the application of which also relies on prior liveness analysis.
+Given liveranges, assignments are easily derived from only one pass through IR.
+Similar to graph coloring, given \(K\) available GPRs, it must assign  as many variables as possible to these, and spill the remaining.
 

### 2024-01-09 article.tex
Add both briggs examples

@@
 This is because variables \(a\) and \(b\) may still be assigned different registers or even spilled if,  for instance, either of them are of significant degree.
+
 Coalescing joins nodes \(a\) and \(b\) to node \(ab\) preserving the edges of both to maintain soundness.
-
 Since all edges are preserved, the resulting node \(ab\) may be of a much higher degree.
 Because of this, only strategies that produce a \(k\)-colorable graph are worth considering, as additional spills negate the purpose entirely.
-One of the  strategies introduced in the Appel text is the \textit{Briggs} strategy  \cite[232]{tiger}, in which nodes \(a\) and \(b\) can be coalesced if and only if the resulting node \(ab\) is of insignificant degree  (i.e.
-\(\leq K\)).
+One of the  strategies introduced in the Appel text is the \textit{Briggs} strategy  \cite[232]{tiger}, in which nodes \(a\) and \(b\) can be coalesced if and only if the resulting node \(ab\) has \(\leq K\) nieghbors of significant degree  (i.e.
+\(\geq K\)).
 
 Since the only move operations in use in LLVM-{}- are phi nodes, only variables that are subject to this sort of operation are eligible to be coalesced in this form. 
-Variables that may be coalesced can be thought of as having a different sort of edge, quite the opposite of interference edges, in fact they can be thought of as a sort of 'preference edge'.
-
-When applied to the previous example (Figure \ref{fig:countdown.ll}), only \texttt{\%argc} and \texttt{\%c2} may be coalesced with \texttt{\%c1} as this is the only mov instruction.
-
-Consider the above example but with some \(k \geq 2\). the \%c1 and
+Variables that can be coalesced can be thought of as having a preference edge, and any nodes with such an edge stand to benefit from coalescing, but this may only be done if they are not in interference.
+%different sort of edge, quite the opposite of interference edges, in fact they can be thought of as a sort of 'preference edge'.
+
+%When applied to the previous example , only \texttt{\%argc} and \texttt{\%c2} may be coalesced with \texttt{\%c1} as this is the only mov instruction.
+
+Consider the previous example (Figure \ref{fig:countdown.ll}): when applied with the same \(k=2\) nothing happens regardless of which allocator is used (see Figure \ref{fig:countdown.briggs.dot2} which was generated using the \texttt{briggs} allocator).
+This is because although the \texttt{\%c1} and \texttt{\%c2} nodes are not in interference, both neighbors of \texttt{\%c2} are of significant degree (\(\mathit{deg}(\texttt{\%cn}) = 2 \geq 2\) and \(\mathit{deg}(\texttt{\%argc}) = 3 \geq 2\)).
+There fore it doesn't satisfy the Briggs criteria, and are not coalesced, resulting in the same interference graph as the \texttt{simple} allocator in Figure \ref{fig:countdown.dot}.
+But for some \(k \geq 2\), now there are fewer than \(k\) significant neighbors of \texttt{\%c2} (and \texttt{\%c1}), meaning the two can be coalesced, with the resulting node \texttt{\%c1, \%c2} having two neighbors of insignificant degree.
 
@@
      \centering
-   \begin{minipage}[b]{0.42\textwidth}
-     \centering
-     \includesvg[height=10em]{countdown.briggs}
-     \caption{\texttt{dune exec build -{}- -t dot}}\label{fig:countdown.briggs.dot}
+   \begin{minipage}[b]{0.50\linewidth}
+     \centering
+     \includesvg[height=10.9em]{countdown.briggs2}
+     \caption{\texttt{build -{}- -t dot -a briggs -n 2}}\label{fig:countdown.briggs.dot2}
+   \end{minipage}
+   \begin{minipage}[b]{0.32\linewidth}
+     \centering
+     \includesvg[height=9em]{countdown.briggs}
+     \caption{\texttt{-a briggs -n 3}}\label{fig:countdown.briggs.dot}
    \end{minipage}
@@
 
-The purpose is only to assign the same registers to different variables. You don't need to note down information about blocks etc.
+\noindent The purpose is only to assign the same registers to different variables. You don't need to note down information about blocks etc.
 

### 2024-01-09 article.tex
Add table

@@
 
-Translating each IR instruction to x86 correctly is a matter of eliminating unintended side-effects. Each LLVM-{}- instruction is defined to have only one purpose, %% FIXME is purpose the right word?
+Translating each IR instruction to x86 correctly is a matter of eliminating unintended side-effects.
+Each LLVM-{}- instruction generally only serves one specific function, %% FIXME is purpose the right word?
 as concepts such as calling conventions, stack frames, or a \lstinline!FLAGS! register are completely abstracted over in order to remain platform independent.
@@
 In contrast, the x86 \textit{instruction set architecture} (ISA) is targeting a  \textit{complex instruction set computer} (CISC) family of processors, the instructions of which perform a much broader set of operations \cite[p.~190]{tiger}.
-This is is in part due to pipelining, i.e.
+This is in part due to pipelining, i.e.
 an abstraction over the concrete implementation of the actual processor, which in turn is made for the sake of performance.
 
-An example of this would be the division/remainder operation: since integer division is a non-trivial iterative process wherein both the quotient and remainder is needed throughout, the result of both of these is stored in the \lstinline!%rax!
-and \lstinline!%rdx!
-registers respectively.
+But another more concise example of this would be the division/remainder operation: since integer division is a non-trivial iterative process wherein both the quotient and remainder is needed throughout, the result of both of these is stored in the \texttt{\%rax} and \texttt{\%rdx} registers respectively.
 This means two operations are performed simultaneously regardless of which value is used, hence they need to be restored before executing the next instruction, as any variables assigned to \lstinline!%rax!
@@
 
-\begin{center}
-\begin{tabular}{|c||c|c|c|c|} 
- \hline
- add & addx & \{\}\\
- \hline
- mul & imul & \{ \%rax \} \\
- \hline
- sdiv & idivx & \{\%rax, \%rdx\} \\
- \hline
- srem & idivx & \{\%rax, \%rdx \\
- \hline
-   call & callx & \{\%rax, \textit{caller saved registers}\} \\
- \hline
+To generally map each LLVM-{}- instruction to what amounts to the 'core' function in x86, consider Table \ref{tab:trans}.
+
+\begin{table}[H]
+  \centering
+\begin{tabular}{|>{\tt}l|>{\tt}l|l|} 
+ \hline
+ LLVM-{}- & x86 & Registers affected\\
+ \hline
+add      & addx \%src, \%dst & \{\}\\
+alloca   & subq\ \ \_\_\_, \%rsp & \{\texttt{\%rsp}\}\\
+allocan  & subq\ \ \_\_\_, \%dst & \{\texttt{\%rsp}\}\\
+and      & andx \%src, \%dst & \{\}\\
+ashr     & sarx \%src, \%dst & \{\}\\
+bitcast  & movx \%src, \%dst & \{\}\\
+call     & \textrm{\textit{see \ref{impl-call}}} & \{\texttt{\%rax}, \texttt{\%rcx}, \texttt{\%rdx}, \texttt{\%rsi}, \texttt{\%r8-\%r11}\}\\
+gep      & \textrm{\textit{see \ref{impl-gep}}} & \{\}\\
+icmp     & cmpx \%src, \%dst & \{\texttt{\%rFLAGS}\}\\
+load     & movx (\%src), \%dst & \{\}\\
+lshr     & shrx \%src, \%dst & \{\}\\
+mul      & imulx \%src, \%dst & \{\texttt{\%rax}\}\\
+or       & orx \%src, \%dst & \{\}\\
+phi      & \textrm{\textit{see \ref{impl-phi}}} & \{\}\\
+ptrtoint & movx \%src, \%dst & \{\}\\
+sdiv     & idivx \%src, \%dst & \{\texttt{\%rax}, \texttt{\%rdx}\}\\
+select   &  & \{\}\\
+sext     & movx \%src, \%dst & \{\}\\
+shl      & shl \%src, \%dst & \{\}\\
+srem     & idivx \%src, \%dst & \{\texttt{\%rax}, \texttt{\%rdx}\}\\
+store    & movx \%src, (\%dst) & \{\}\\
+sub      & subx \%src, \%dst & \{\}\\
+trunc    & movx \%src, \%dst & \{\}\\
+udiv     & divx \%src, \%dst & \{\texttt{\%rax}, \texttt{\%rdx}\}\\
+urem     & divx \%src, \%dst & \{\texttt{\%rax}, \texttt{\%rdx}\}\\
+xor      & xorx \%src, \%dst & \{\}\\
+zext     & movx \%src, \%dst & \{\}\\
  \hline
 \end{tabular}
-\end{center}
-
-\subsubsection{Implementing \mintinline{llvm}{call} instructions}
+\caption{LLVM-{}- to x86 translation table}\label{tab:trans}
+\end{table}
+
+ %add & addx & \{\}\\
+ %\hline
+ %mul & imul & \{ \%rax \} \\
+ %\hline
+ %sdiv & idivx & \{\%rax, \%rdx\} \\
+ %\hline
+ %srem & idivx & \{\%rax, \%rdx \\
+ %\hline
+   %call & callx & \{\%rax, \textit{caller saved registers}\} \\
+
+\subsubsection{Implementing \mintinline{llvm}{call} instructions} \label{impl-call}
 
 The \mintinline{llvm}{call} operation is by far the most complicated one. Mostly because of calling conventions,
-The \mintinline{gas}{%rax} register is zeroed before
-
-\subsubsection{Implementing \mintinline{llvm}{phi} instructions}
+The \texttt{\%rax} register is zeroed before
+
+\subsubsection{Implementing \mintinline{llvm}{gep} instructions}\label{impl-gep}
+GEP instructions are used to perform address calculations for accessing elements in  structures or arrays.
+GEP instructions are particularly useful for efficiently addressing elements in complex data structures.
+
+\subsubsection{Implementing \mintinline{llvm}{phi} instructions}\label{impl-phi}
 Another irregular instruction is the \mintinline{llvm}{phi} instruction representing the \(\Phi\)-functions. Their  intended purpose is to copy data from whichever predecessor node was executed.

### 2024-01-09 article.tex
Add gep example

@@
 GEP instructions are used to perform address calculations for accessing elements in  structures or arrays.
-GEP instructions are particularly useful for efficiently addressing elements in complex data structures.
+They are particularly useful for efficiently addressing elements in complex data structures.
+For example, code accessing fields like line 8 in Figure \ref{fig:vector.c}, would  translate to the snippet seen in Figure \ref{fig:vector.ll}.
+
+\begin{figure}[H]
+   \begin{minipage}{0.48\textwidth}
+     \centering
+     \begin{minted}[linenos,bgcolor=LightGray]{c}
+typedef struct {
+  int x;
+  int y;
+} vector;
+
+void dot(vector *a, vector *b) {
+  a->x += b->x;
+  a->y += b->y;     // -------->
+}
+
+     \end{minted}
+     \caption{Subfield access}\label{fig:vector.c}
+   \end{minipage}
+   \begin{minipage}{0.48\textwidth}
+     \centering
+     \begin{minted}[linenos,bgcolor=LightGray]{llvm}
+; ...
+%6 = getelementptr %struct.vector,
+            %i8* %1, i64 0, i32 1
+%7 = load i32, i32* %6
+%8 = getelementptr %struct.vector,
+            %i8* %0, i64 0, i32 1
+%9 = load i32, i32* %8
+%10 = add i32 %9, %7
+store i32 %10, i32* %8
+; ...
+     \end{minted}
+     \caption{\texttt{clang -O1 -S -emit-llvm}}\label{fig:vector.ll}
+   \end{minipage}
+\end{figure}
 

### 2024-01-09 article.tex
Explain between example and tables
Add linearscan as well as brief explanations

@@
 Consider a simple program like Figure \ref{fig:countc} that only prints integers incrementally up until \texttt{argc}, in LLVM-{}-  that would consist of one loop with one phi node as seen in Figure \ref{fig:count1ll}.
-The number of iterations is nearly halfed when reverseing the list of instructions prior to dataflow as can be seen on \ref{tab:countll-rev-steps} when compared to \ref{tab:countll-steps}.
 
@@
 \end{figure}
+
+\noindent If one were to note the \(\mathit{in}[s]\) and \(\mathit{out}[s]\) sets for each instruction \(s\) for each iteration one would arrive at the two tables seen in Table \ref{tab:countll-rev-steps} and \ref{tab:countll-steps}.
+The number of iterations is nearly halfed when reverseing the list of instructions prior to dataflow as can be seen on \ref{tab:countll-rev-steps} when compared to \ref{tab:countll-steps}.
 
@@
    \end{minipage}
-   \begin{minipage}[b]{0.32\linewidth}
-     \centering
-     \includesvg[height=9em]{countdown.briggs}
+   \begin{minipage}[b]{0.33\linewidth}
+     \centering
+     \includesvg[height=8.em]{countdown.briggs}
      \caption{\texttt{-a briggs -n 3}}\label{fig:countdown.briggs.dot}
@@
 Given liveranges, assignments are easily derived from only one pass through IR.
-Similar to graph coloring, given \(K\) available GPRs, it must assign  as many variables as possible to these, and spill the remaining.
+Similar to graph coloring, given \(k\) available GPRs, it assigns  as many variables as possible, and spilling the rest.
+
+It does this by first building the intervals using the \(\mathit{in}[s]\) and \(\mathit{out}[s]\) resulting from prior dataflow analysis.
+This can be done in a single pass with a little creative array indexing as can be seen in Figure \ref{fig:intervals-ocaml}.
+
+\begin{figure}[H]
+     \centering
+     \begin{minted}[linenos,bgcolor=LightGray]{ocaml}
+let intervals insns (in_, out) =
+  let insn (starts, t, active) (i, n) =
+    let init s (added, t, active) = (* .. *) in
+    let end_ s (added, t, active) = (* .. *) in
+    let a, t, c =
+      init in_.(i) (S.SS.empty, t, active) |> init out.(i) |> end_ out.(i)
+    in
+    ((i, n, a) :: starts, t, c)
+  in
+  let starts, _, active =
+    List.fold_left insn
+      ([], S.ST.empty, Array.init (List.length insns) (Fun.const S.SS.empty))
+      insns
+  in
+  List.rev starts |> List.map (fun (i, n, a) -> (i, n, a, active.(i)))
+     \end{minted}
+     \caption{Implementation of \texttt{intervals} in \texttt{linear.ml}}\label{fig:intervals-ocaml}
+\end{figure}
+
+\noindent The array indexing is used in the inner functions of \texttt{init} and \texttt{end\_} to unset the last point at which an interval goes out of scope.
+
+\begin{figure}[H]
+     \centering
+     \begin{minted}[linenos,bgcolor=LightGray]{ocaml}
+let init s (added, t, active) =
+  ( S.SS.filter (fun s -> not (S.ST.mem s t)) s |> S.SS.union added,
+    S.SS.fold
+      (fun s t ->
+        S.ST.update s
+          (function
+            | Some j ->
+                active.(i) <- S.SS.add s active.(i);
+                active.(j) <- S.SS.remove s active.(j);
+                Some i
+            | None ->
+                active.(i) <- S.SS.add s active.(i);
+                Some i)
+          t)
+      s t,
+    active )
+     \end{minted}
+     \caption{Implementation of \texttt{init} in \texttt{linear.ml}}\label{fig:init-ocaml}
+\end{figure}
+
+\begin{figure}[H]
+     \centering
+     \begin{minted}[linenos,bgcolor=LightGray]{ocaml}
+let end_ s (added, t, active) =
+  ( S.SS.filter (fun s -> not (S.ST.mem s t)) s |> S.SS.union added,
+    S.SS.fold
+      (fun s t ->
+        S.ST.update s
+          (function
+            | Some j ->
+                active.(i) <- S.SS.add s active.(i);
+                active.(j) <- S.SS.remove s active.(j);
+                Some i
+            | None -> Some i)
+          t)
+      s t,
+    active )
+     \end{minted}
+     \caption{Implementation of \texttt{end\_} in \texttt{linear.ml}}\label{fig:end_-ocaml}
+\end{figure}
+
+\begin{figure}[H]
+     \centering
+     \begin{minted}[linenos,bgcolor=LightGray]{ocaml}
+let linearscan k insns =
+  let avail = Regs.of_list (List.init k reg_of_int) in
+  let insn (avail, active, assign) (_, _n, s, _e) =
+    (*Printf.printf "%s\n" (Cfg.string_of_insn n);*)
+    let start s (avail, active, assign) =
+      match Regs.choose_opt avail with
+      | Some reg -> (Regs.remove reg avail, S.ST.add s reg active, assign)
+      | None ->
+          (* FIXME: order by remaining or total length *)
+          let z, reg = S.ST.choose active in
+          ( avail,
+            S.ST.remove z active |> S.ST.add s reg,
+            S.ST.add z
+              (Ind3 (Lit (Int64.of_int ((S.ST.cardinal assign * -8) - 8)), Rbp))
+              assign )
+    in
+    S.SS.fold start s (avail, active, assign)
+  in
+  let _, active, assign =
+    List.fold_left insn (avail, S.ST.empty, S.ST.empty) insns
+  in
+  S.ST.fold (fun k v acc -> S.ST.add k (Reg v) acc) active assign
+     \end{minted}
+     \caption{Implementation of \texttt{linearscan} in \texttt{linear.ml}}\label{fig:linearscan-ocaml}
+\end{figure}
 

### 2024-01-10 article.tex
Further explain dumb fib

@@
 
-The measures in \ref{tab:dumb-fib}  are significantly equivalent to one another, with all allocators (except greedy 0) seeming to converge towards being 37\% slower than Clang.
+\noindent The measures in Table \ref{tab:dumb-fib}  are significantly equivalent to one another, with all allocators (except greedy 0) seeming to converge towards being 37\% slower than Clang.
 This isn't particularly surprising, as the x86 translation is more concerned with preserving callee saved registers than necessary in pushing each and every one despite not being in use.
@@
 However, the last one (greedy 0), has no assignable registers, so it must necessarily start spilling from the very beginning.
-To elucidate this hypothesis, here is the same benchmark but only for the greedy of decreasing \(k\) assignable registers:
+To elucidate this hypothesis, the same benchmark was performed but only for the greedy of decreasing \(k\) assignable registers as well as the Tiger compiler which has the same approach to spilling every variable prematurely  and can be seen in Table \ref{tab:dumb-fib-greedy}:
 \begin{table}[H]
@@
 \caption{\label{tab:dumb-fib-greedy}Benchmark of \texttt{benches/fib.ll} output by \mintinline{bash}{dune exec bench -- -f fib -n 1000}}
-
 \end{table}
+
+\noindent While it's not a complete 1:1 in terms of runtime as the Tiger compiler seems to perform slightly better than Greedy 0 (by \textasciitilde 30\% or so), that does make sense as there are marginally more instructions involved with each operation, not to mention the extra care taken to ensure calling conventions are upheld.
 
@@
 
+% TODO: comment that this is basically writing an llvm backend, except redoing all the register allocation stuff.
+
 \subsection{Comparison to other work and ideas for future work}

### 2024-01-11 article.tex
Add newer cfg code

@@
 
-\subsection{Building a CFG}
+\subsection{Building a Control Flow Graph}
 
@@
 Labels, while present in the flattened list, aren't considered as they have no inherent function other than labelling the leader instruction to branch to.
+
+\begin{minted}[linenos,bgcolor=LightGray]{ocaml}
+let graph ((head, tail) : Ll.cfg) : G.V.t array * G.t =
+  let insns = flatten (head, tail) |> List.mapi (fun i n -> (i, n)) in
+  let verts = Array.init (List.length insns) G.V.create in
+  let addbl t = function i, Label l -> S.ST.add l verts.(i + 1) t | _ -> t in
+  let blocks = List.fold_left addbl S.ST.empty insns in
+  let g = G.create () in
+  Array.iter (G.add_vertex g) verts;
+  let addterm i = function
+    | Ll.Ret _ | Unreachable -> ()
+    | Br lbl -> G.add_edge g verts.(i) (S.ST.find lbl blocks)
+    | Cbr (_, l1, l2) ->
+        G.add_edge g verts.(i) (S.ST.find l1 blocks);
+        G.add_edge g verts.(i) (S.ST.find l2 blocks)
+    | Switch (_, _, _, lbls) ->
+        let vert l = S.ST.find l blocks in
+        List.map snd lbls |> List.map vert |> List.iter (G.add_edge g verts.(i))
+  in
+  let addedges = function
+    | _, Label _ -> ()
+    | i, Insn _ -> G.add_edge g verts.(i) verts.(i + 1)
+    | i, Term term -> addterm i term
+  in
+  List.iter addedges insns;
+  (verts, g)
+\end{minted}
 
@@
 
-\subsection{Interference Graph}
+\subsection{Building an Interference Graph}
 

### 2024-01-11 article.tex
Add spilling heuristics example
Elaborate run time of lva

@@
 \noindent This is unnoticeable for snippets of marginal size like most examples included, but when applied to functions with a significant amount of def-use paths, it can be quite consequential.
-For \texttt{tests/sha256.ll} for instance, the running time of this analysis is effectively doubled:
-
-\begin{minted}{bash}
+For \texttt{tests/sha256.ll} for instance, the running time of this analysis is effectively doubled (see Figure \ref{fig:lva-time}).
+
+\begin{figure}[H]
+  \centering
+  \begin{minipage}[b]{0.69\textwidth}
+    \begin{minted}[bgcolor=LightGray]{bash}
 $ time dune exec build -- tests/sha256.ll -t lva -r
-33,69s user 0,04s system 99% cpu 33,765 total
+36.91s user 0,04s system 99% cpu 33,765 total
 $ time dune exec build -- tests/sha256.ll -t lva
-17,26s user 0,05s system 99% cpu 17,335 total
-\end{minted}
+18,61s user 0,03s system 99% cpu 19,055 total
+    \end{minted}
+    \caption{Average runtime over several runs (n=32)}\label{fig:lva-time}
+  \end{minipage}
+\end{figure}
 
@@
 \section{Linear Scan}
-Another approach that also depends on liveness analysis is the linear scan algorithm.
-It is a simple and efficient approach that performs a linear pass over IR to produce a register assignment significantly faster than all of the steps involved with graph coloring.
-That said, it may generate more imprecise assignments and therefore also be less performant at runtime, so the utility of such an approach depends on context.
-Some compilers typically compile code into machine code ahead of time (AOT) while others depend on just-in-time (JIT) compilation which defers the compilation of code until runtime.
-Therefore linear scan is an approach to register allocation that may be used to achieve reasonably performant runtime compilation.
+Another approach that depends on liveness analysis is the linear scan algorithm.
+It is a simple and efficient approach that performs a linear pass over IR to produce a register assignment significantly faster than the steps involved with graph coloring.
+That said, it may generate more imprecise assignments and therefore also be less performant at runtime, so the utility of such an approach depends on the context in which it is used.
+While compilers like clang translate to machine code ahead-of-time (AOT), languages like C\# and Java will often rely on just-in-time (JIT) compilation to translate their respective VM instructions to native machine code,
+
+Such VMs that depend on JIT will often defer compilation to runtime.
+This is done for a myriad of reasons, portability and security among them, but usually at the cost of runtime performance.
+One of the reasons for this can be attributed to the way in which registers are allocated, as a significant downside of JIT is the fact it needs to be compiled in full before execution can begin.
+Therefore, generally a faster class of allocators is employed, linear scan being one of them, as linear scan is a reasonably performant alternative to graph coloring, that may be used to achieve reasonably efficient register assignments at runtime.
 %Whether or not to  a cost depends on the context, as JIT compilation will often 
@@
 Given liveranges, assignments are easily derived from only one pass through IR.
-Similar to graph coloring, given \(k\) available GPRs, it assigns  as many variables as possible, and spilling the rest.
+Similar to graph coloring, given \(k\) available GPRs, it assigns  as many variables as possible, and spills the rest.
 
 It does this by first building the intervals using the \(\mathit{in}[s]\) and \(\mathit{out}[s]\) resulting from prior dataflow analysis.
-This can be done in a single pass with a little creative array indexing as can be seen in Figure \ref{fig:intervals-ocaml}.
+This can be done in a single pass with some creative array indexing, but my approach relies on two passes: one for finding the start points (see Figure \ref{fig:livenessstarts.ml}), the other for finding the end points.
+
+\begin{figure}[H]
+  \centering
+   \begin{minipage}[b]{0.88\textwidth}
+     \centering
+     \begin{minted}[linenos,bgcolor=LightGray]{ocaml}
+let add e = function
+  | Some s -> Some (S.SS.add e s)
+  | None -> Some (S.SS.singleton e)
+
+let intervalstart insns (in_, out) =
+  let insn (ordstarts, starts, active) (i, _n) =
+    let out' = S.SS.diff (S.SS.union in_.(i) out.(i)) active in
+    ( S.SS.fold (fun e a -> IT.update i (add e) a) out' ordstarts,
+      S.SS.fold (fun e a -> S.ST.add e i a) out' starts,
+      S.SS.union active out' )
+  in
+  List.fold_left insn (IT.empty, S.ST.empty, S.SS.empty) insns
+     \end{minted}
+     \caption{Implementation of \texttt{intervalstart} in \texttt{lib/linear.ml}}\label{fig:livenessstarts.ml}
+   \end{minipage}
+\end{figure}
+
+\noindent Of note are the structures over which the folding is done: the \texttt{IT.t} is a \texttt{Map.S} module indexing earlier described symbol sets (i.e.
+\texttt{S.SS.t}).
+The reason for this is to iterate through intervals in increasing and decreasing order of start/length and end points respectively.
+The symbol table (\texttt{S.ST.t}) is to allow for reverse lookup, i.e.
+which program point variable \(s\) is defined, and lastly the symbol set (i.e.
+\texttt{S.SS.t}) which is just there to prevent multiple start points for the same variable.
+
+\begin{figure}[H]
+  \centering
+   \begin{minipage}[b]{0.95\textwidth}
+     \centering
+     \begin{minted}[linenos,bgcolor=LightGray]{ocaml}
+let intervalends insns starts (in_, out) =
+  let insn (i, _n) (ordstarts, lengths, ends, active) =
+    let in' = S.SS.diff (S.SS.union in_.(i) out.(i)) active in
+    ( S.SS.fold (fun e a -> IT.update i (add e) a) in' ordstarts,
+      S.SS.fold
+        (fun e a -> IT.update (-i + S.ST.find e starts) (add e) a)
+        in' lengths,
+      S.SS.fold (fun e a -> S.ST.add e i a) in' ends,
+      S.SS.union active in' )
+  in
+  List.fold_right insn insns (IT.empty, IT.empty, S.ST.empty, S.SS.empty)
+     \end{minted}
+     \caption{Implementation of \texttt{intervalendsintervalstart} in \texttt{lib/linear.ml}}\label{fig:intervalends.ml}
+   \end{minipage}
+\end{figure}
+
+\noindent The opposite is quite similar (see Figure \ref{fig:intervalends.ml}).
+
+as can be seen in Figure \ref{fig:intervals-ocaml}.
+
+\subsection{Heuristics spilling}
+\usetikzlibrary{chains,arrows}
+\begin{figure}[H]
+
+     \centering
+     \begin{minipage}[b]{0.30\textwidth}
+     \centering
+
+     \begin{tikzpicture}[scale=0.442,
+    node distance=0mm, text height=1.5ex, text depth=.25ex,
+    attribute/.style={rectangle,minimum size=8mm,very thin,draw=black!50,font=\ttfamily},
+    link/.style={thin, to path={-- ++(0,8mm) -| (\tikztotarget)}, >= triangle 45, shorten >= 0pt}]
+
+%\node (n1) [attribute] {\small NODE1};
+%\node (n2) [attribute, right=of n1] {\small NODE2};
+
+   \node at (7,0) (11) {};
+   \node at (7,1) (10) {};
+   \node at (7,2) (9) {};
+   \node at (7,3) (8) {};
+   \node at (7,4) (7) {};
+   \node at (7,5) (6) {};
+   \node at (7,6) (5) {};
+   \node at (7,7) (4) {};
+   \node at (7,8) (3) {};
+   \node at (7,9) (2) {};
+
+   \draw[|-|] (2)  node[above,transform canvas={xshift=1.43cm}] {\texttt{\%3}} edge [transform canvas={xshift=1.5cm}] (6);
+   \draw[|-|] (3) node[above,transform canvas={xshift=0.83cm}] {\texttt{\%4}} edge [transform canvas={xshift=0.9cm}] (8);
+   \draw[|-|] (4) node[above] {\texttt{\%5}} edge [ ]  (11);
+
+   %\draw[help lines] (0,0) grid (3,11);
+   %\node at (7,0) (11) {};
+   %\node at (7,1) (10) {};
+   %\node at (7,2) (9) {};
+   %\node at (7,3) (8) {};
+   %\node at (7,4) (7) {};
+   %\node at (7,5) (6) {};
+   %\node at (7,6) (5) {};
+   %\node at (7,7) (4) {};
+   %\node at (7,8) (3) {};
+   %\node at (7,9) (2) {};
+   %\path (2.center) edge node[above] {} (4.center);
+   %\path (5.center) edge node[above] {} (7.center);
+   %\draw[|-|] (8.center)  edge [transform canvas={xshift=3mm]]   (10.center);
+
+\end{tikzpicture}
+\vspace{0.7em}
+     \caption{Live ranges}\label{fig:linear.lra}
+   \end{minipage}
+   \begin{minipage}[b]{0.68\textwidth}
+     \centering
+     \begin{minted}[linenos,bgcolor=LightGray]{llvm}
+define i32 @main (i32 %0, i8* %1) {
+  %3 = getelementptr i8*, i8** %1, i64 1
+  %4 = load i8*, i8** %3
+  %5 = call i32 @atoi (i8* %4)
+  call i32 @printf (i8* @.str, i8** %3, i8* %4, ...
+  %7 = call i32 @isqrt (i32 %5)
+  %8 = call i64 @strtoll (i8* %4, i8* null, i32 10)
+  call i32 @printf (i8* @.str.1, i32 %5, i64 %8)
+  call i32 @printf (i8* @.str.2, i32 %7, i32 %5)
+  ret i32 %5
+}
+     \end{minted}
+     \caption{\texttt{@main} of \texttt{tests/linear.ll}}\label{fig:linear.ll}
+   \end{minipage}
+\end{figure}
 

### 2024-01-11 article.tex
Fix llivespans figure

@@
 \begin{equation}
-  a = \Phi (b_1, ..., b_n) \label{phi}
+i a = \Phi (b_1, ..., b_n) \label{phi}
 \end{equation}
@@
 \begin{figure}[H]
-
-     \centering
-     \begin{minipage}[b]{0.30\textwidth}
-     \centering
-
-     \begin{tikzpicture}[scale=0.442,
-    node distance=0mm, text height=1.5ex, text depth=.25ex,
-    attribute/.style={rectangle,minimum size=8mm,very thin,draw=black!50,font=\ttfamily},
-    link/.style={thin, to path={-- ++(0,8mm) -| (\tikztotarget)}, >= triangle 45, shorten >= 0pt}]
-
-%\node (n1) [attribute] {\small NODE1};
-%\node (n2) [attribute, right=of n1] {\small NODE2};
-
-   \node at (7,0) (11) {};
-   \node at (7,1) (10) {};
-   \node at (7,2) (9) {};
-   \node at (7,3) (8) {};
-   \node at (7,4) (7) {};
-   \node at (7,5) (6) {};
-   \node at (7,6) (5) {};
-   \node at (7,7) (4) {};
-   \node at (7,8) (3) {};
-   \node at (7,9) (2) {};
-
-   \draw[|-|] (2)  node[above,transform canvas={xshift=1.43cm}] {\texttt{\%3}} edge [transform canvas={xshift=1.5cm}] (6);
-   \draw[|-|] (3) node[above,transform canvas={xshift=0.83cm}] {\texttt{\%4}} edge [transform canvas={xshift=0.9cm}] (8);
-   \draw[|-|] (4) node[above] {\texttt{\%5}} edge [ ]  (11);
+    \centering
+    \begin{minipage}[b]{0.31\textwidth}
+      \centering
+      \begin{tikzpicture}[scale=0.4]
+        \node[above] at (0,0) {\texttt{\%3}};
+        \node[above] at (1,-1) {\texttt{\%4}};
+        \node[above] at (2,-2) {\texttt{\%5}};
+        \node[above] at (3,-4) {\texttt{\%7}};
+        \node[above] at (4,-5) {\texttt{\%8}};
+        \draw[|-|] (0,0) edge   (0,-3);
+        \draw[|-|] (1,-1) edge   (1,-5);
+        \draw[|-|] (2,-2) edge   (2,-8);
+        \draw[|-|] (3,-4) edge   (3,-7);
+        \draw[|-|] (4,-5) edge   (4,-6);
 
@@
 \end{tikzpicture}
-\vspace{0.7em}
-     \caption{Live ranges}\label{fig:linear.lra}
+\vspace{2.7em}
+     \caption{Live intervals}\label{fig:linear.lra}
    \end{minipage}

### 2024-01-11 article.tex
Update intervalenvds

@@
      \begin{minted}[linenos,bgcolor=LightGray]{ocaml}
-let intervalends insns starts (in_, out) =
-  let insn (i, _n) (ordstarts, lengths, ends, active) =
+let intervalends insns (in_, out) =
+  let insn (i, _n) (ordends, ends, active) =
     let in' = S.SS.diff (S.SS.union in_.(i) out.(i)) active in
-    ( S.SS.fold (fun e a -> IT.update i (add e) a) in' ordstarts,
-      S.SS.fold
-        (fun e a -> IT.update (-i + S.ST.find e starts) (add e) a)
-        in' lengths,
+    ( S.SS.fold (fun e a -> IT.update i (add e) a) in' ordends,
       S.SS.fold (fun e a -> S.ST.add e i a) in' ends,
@@
   in
-  List.fold_right insn insns (IT.empty, IT.empty, S.ST.empty, S.SS.empty)
+  List.fold_right insn insns (IT.empty, S.ST.empty, S.SS.empty)
      \end{minted}
@@
 
-\noindent The opposite is quite similar (see Figure \ref{fig:intervalends.ml}).
+\noindent The opposite is quite similar (see Figure \ref{fig:intervalends.ml}), with the only real difference being the use of \texttt{List.fold\_right} instead of \texttt{List.fold\_left} to reduce the function body in reverse.
 

### 2024-01-11 article.tex
Jane approved

@@
 These are a type of instruction that evaluates to a specified operand depending on which predecessor block was executed immediately prior.
-This circumvents the need for the same level of interior mutability as is achieved by reading/writing to memory, without violating the properties of SSA.
-By copying a value from the end of a predecessor to the beginning of the current block.
+This circumvents the need for the same level of mutability as is achieved by reading/writing to memory, without violating the properties of SSA, by copying a value from the end of a predecessor to the beginning of the current block.
 A Phi node represents a point in the program where the control flow merges, and it selects the appropriate version of a variable based on the path taken.
@@
 
-Control-flow constraints on the other hand propagate the use of variables to previously executed instructions, expecting these to be defined somewhere further up the CFG.
+Control-flow constraints, on the other hand, propagate the use of variables to previously executed instructions, expecting these to be defined somewhere further up the CFG.
 This is done by propagating the union of the \textit{live-in} set associated with all immediate successor nodes.
@@
 The motivation for this is the fact that the \(\mathit{out}[n]\) of node \(n\) depends on \(\mathit{in}[s]\) for all successors \(s\).
-Since most nodes have one successor immediately after, populating this beforehands is significantly more efficient.
+Since most nodes only have one immediate successor, populating this first is significantly more efficient.
 
@@
 
-Constructing an interference graph only depends on the \textit{live-out} set and type of instruction.
+The construction of an interference graph only depends on the \textit{live-out} set and type of instruction.
 If the instruction defines a variable, said variable is in interference with all variables in the \textit{live-out} set.
@@
 both the Appel \cite[229]{tiger}, and Aho et al. \cites[557]{dragon} texts,
-is a linear time approximation to this problem is described below.
+is a linear time approximation to this problem as described below.
 
@@
 
-To validate that this works as intended,
+%To validate that this works as intended,
 

### 2024-01-12 article.tex
Cite variadic arguments

@@
 
-To generally map each LLVM-{}- instruction to what amounts to the 'core' function in x86, consider Table \ref{tab:trans}.
+To map each LLVM-{}- instruction to what amounts to the 'core' function in x86, consider Table \ref{tab:trans}.
+Instructions for which the size varies, the suffix is denoted with \texttt{x} instead.
+This is then replaced with the appropriate size suffix depending on the size denoted in the LLVM-{}- instruction.
+For example, most binary operations move the left-hand operand into \texttt{\%rax} and the right-hand operand into  \texttt{\%rcx}, after which the associated instruction from the table is performed.
+For x86, most operations are done in place, so for the most part \texttt{\%rax} is modified and moved to the destination after.
 
@@
 add      & addx \%src, \%dst & \{\}\\
-alloca   & subq\ \ \_\_\_, \%rsp & \{\texttt{\%rsp}\}\\
-allocan  & subq\ \ \_\_\_, \%dst & \{\texttt{\%rsp}\}\\
+alloca   & subq \$n, \%rsp & \{\texttt{\%rsp}\}\\
+allocan  & subq \$n, \%rsp & \{\texttt{\%rsp}\}\\
 and      & andx \%src, \%dst & \{\}\\
@@
 \end{tabular}
-\caption{LLVM-{}- to x86 translation table}\label{tab:trans}
+\caption{LLVM-{}- instruction to x86 translation table}\label{tab:trans}
 \end{table}
 
- %add & addx & \{\}\\
- %\hline
- %mul & imul & \{ \%rax \} \\
- %\hline
- %sdiv & idivx & \{\%rax, \%rdx\} \\
- %\hline
- %srem & idivx & \{\%rax, \%rdx \\
- %\hline
-   %call & callx & \{\%rax, \textit{caller saved registers}\} \\
+\begin{table}[H]
+  \centering
+\begin{tabular}{|>{\tt}l|>{\tt}l|l|} 
+ \hline
+ LLVM-{}- & x86 & Registers affected\\
+ \hline
+add      & addx \%src, \%dst & \{\}\\
+alloca   & subq \$n, \%rsp & \{\texttt{\%rsp}\}\\
+allocan  & subq \$n, \%rsp & \{\texttt{\%rsp}\}\\
+and      & andx \%src, \%dst & \{\}\\
+ashr     & sarx \%src, \%dst & \{\}\\
+bitcast  & movx \%src, \%dst & \{\}\\
+call     & \textrm{\textit{see \ref{impl-call}}} & \{\texttt{\%rax}, \texttt{\%rcx}, \texttt{\%rdx}, \texttt{\%rsi}, \texttt{\%r8-\%r11}\}\\
+gep      & \textrm{\textit{see \ref{impl-gep}}} & \{\}\\
+icmp     & cmpx \%src, \%dst & \{\texttt{\%rFLAGS}\}\\
+load     & movx (\%src), \%dst & \{\}\\
+lshr     & shrx \%src, \%dst & \{\}\\
+mul      & imulx \%src, \%dst & \{\texttt{\%rax}\}\\
+or       & orx \%src, \%dst & \{\}\\
+phi      & \textrm{\textit{see \ref{impl-phi}}} & \{\}\\
+ptrtoint & movx \%src, \%dst & \{\}\\
+sdiv     & idivx \%src, \%dst & \{\texttt{\%rax}, \texttt{\%rdx}\}\\
+select   &  & \{\}\\
+sext     & movx \%src, \%dst & \{\}\\
+shl      & shl \%src, \%dst & \{\}\\
+srem     & idivx \%src, \%dst & \{\texttt{\%rax}, \texttt{\%rdx}\}\\
+store    & movx \%src, (\%dst) & \{\}\\
+sub      & subx \%src, \%dst & \{\}\\
+trunc    & movx \%src, \%dst & \{\}\\
+udiv     & divx \%src, \%dst & \{\texttt{\%rax}, \texttt{\%rdx}\}\\
+urem     & divx \%src, \%dst & \{\texttt{\%rax}, \texttt{\%rdx}\}\\
+xor      & xorx \%src, \%dst & \{\}\\
+zext     & movx \%src, \%dst & \{\}\\
+ \hline
+\end{tabular}
+\caption{LLVM-{}- instruction to x86 translation table}\label{tab:trans}
+\end{table}
 
@@
 The \mintinline{llvm}{call} operation is by far the most complicated one. Mostly because of calling conventions,
-The \texttt{\%rax} register is zeroed before
-
-\subsubsection{Implementing \mintinline{llvm}{gep} instructions}\label{impl-gep}
+The \texttt{\%rax} register is zeroed before every call as its lowermost byte register \texttt{\%al} is used to indicate the number of variadic arguments \cite[25]{sysv}.
+
+\subsubsection{Implementing \mintinline{llvm}{getelementptr} instructions}\label{impl-gep}
 GEP instructions are used to perform address calculations for accessing elements in  structures or arrays.

### 2024-01-12 article.tex
Jane approved

@@
 
-\noindent If one were to note the \(\mathit{in}[s]\) and \(\mathit{out}[s]\) sets for each instruction \(s\) for each iteration one would arrive at the two tables seen in Table \ref{tab:countll-rev-steps} and \ref{tab:countll-steps}.
+\noindent If one were to note the \(\mathit{in}[s]\) and \(\mathit{out}[s]\) sets for each instruction \(s\) for each iteration, one would arrive at the two tables seen in Table \ref{tab:countll-rev-steps} and \ref{tab:countll-steps}.
 The number of iterations is nearly halfed when reverseing the list of instructions prior to dataflow as can be seen on \ref{tab:countll-rev-steps} when compared to \ref{tab:countll-steps}.
@@
 This is because although the \texttt{\%c1} and \texttt{\%c2} nodes are not in interference, both neighbors of \texttt{\%c2} are of significant degree (\(\mathit{deg}(\texttt{\%cn}) = 2 \geq 2\) and \(\mathit{deg}(\texttt{\%argc}) = 3 \geq 2\)).
-There fore it doesn't satisfy the Briggs criteria, and are not coalesced, resulting in the same interference graph as the \texttt{simple} allocator in Figure \ref{fig:countdown.dot}.
+Therefore they don't satisfy the Briggs criteria, and are not coalesced, resulting in the same interference graph as the \texttt{simple} allocator in Figure \ref{fig:countdown.dot}.
 But for some \(k \geq 2\), now there are fewer than \(k\) significant neighbors of \texttt{\%c2} (and \texttt{\%c1}), meaning the two can be coalesced, with the resulting node \texttt{\%c1, \%c2} having two neighbors of insignificant degree.
@@
 \texttt{S.SS.t}).
-The reason for this is to iterate through intervals in increasing and decreasing order of start/length and end points respectively.
+The reason for doing this is to iterate through intervals in increasing and decreasing order of start/length and end points respectively.
 The symbol table (\texttt{S.ST.t}) is to allow for reverse lookup, i.e.
@@
 
-But another more concise example of this would be the division/remainder operation: since integer division is a non-trivial iterative process wherein both the quotient and remainder is needed throughout, the result of both of these is stored in the \texttt{\%rax} and \texttt{\%rdx} registers respectively.
+Another more concise example of this would be the division/remainder operation: since integer division is a non-trivial iterative process wherein both the quotient and remainder is needed throughout, the result of both of these is stored in the \texttt{\%rax} and \texttt{\%rdx} registers respectively.
 This means two operations are performed simultaneously regardless of which value is used, hence they need to be restored before executing the next instruction, as any variables assigned to \lstinline!%rax!
@@
 gep      & \textrm{\textit{see \ref{impl-gep}}} & \{\}\\
-icmp     & cmpx \%src, \%dst & \{\texttt{\%rFLAGS}\}\\
+icmp     & \textrm{\textit{see \ref{impl-cond}}} & \{\texttt{\%rFLAGS}\}\\
 load     & movx (\%src), \%dst & \{\}\\
@@
  \hline
-add      & addx \%src, \%dst & \{\}\\
-alloca   & subq \$n, \%rsp & \{\texttt{\%rsp}\}\\
-allocan  & subq \$n, \%rsp & \{\texttt{\%rsp}\}\\
-and      & andx \%src, \%dst & \{\}\\
-ashr     & sarx \%src, \%dst & \{\}\\
-bitcast  & movx \%src, \%dst & \{\}\\
-call     & \textrm{\textit{see \ref{impl-call}}} & \{\texttt{\%rax}, \texttt{\%rcx}, \texttt{\%rdx}, \texttt{\%rsi}, \texttt{\%r8-\%r11}\}\\
-gep      & \textrm{\textit{see \ref{impl-gep}}} & \{\}\\
-icmp     & cmpx \%src, \%dst & \{\texttt{\%rFLAGS}\}\\
-load     & movx (\%src), \%dst & \{\}\\
-lshr     & shrx \%src, \%dst & \{\}\\
-mul      & imulx \%src, \%dst & \{\texttt{\%rax}\}\\
-or       & orx \%src, \%dst & \{\}\\
-phi      & \textrm{\textit{see \ref{impl-phi}}} & \{\}\\
-ptrtoint & movx \%src, \%dst & \{\}\\
-sdiv     & idivx \%src, \%dst & \{\texttt{\%rax}, \texttt{\%rdx}\}\\
-select   &  & \{\}\\
-sext     & movx \%src, \%dst & \{\}\\
-shl      & shl \%src, \%dst & \{\}\\
-srem     & idivx \%src, \%dst & \{\texttt{\%rax}, \texttt{\%rdx}\}\\
-store    & movx \%src, (\%dst) & \{\}\\
-sub      & subx \%src, \%dst & \{\}\\
-trunc    & movx \%src, \%dst & \{\}\\
-udiv     & divx \%src, \%dst & \{\texttt{\%rax}, \texttt{\%rdx}\}\\
-urem     & divx \%src, \%dst & \{\texttt{\%rax}, \texttt{\%rdx}\}\\
-xor      & xorx \%src, \%dst & \{\}\\
-zext     & movx \%src, \%dst & \{\}\\
+ret     & epilogue, retq & \{\}\\
+br label \%l & jmp \%l & \{\}\\
+br i1 \%c, label \%t, label \%f & cmpx \%c & \{\}\\
+ & jz \%f & \\
+ & jmp \%t & \\
  \hline
 \end{tabular}
-\caption{LLVM-{}- instruction to x86 translation table}\label{tab:trans}
+\caption{LLVM-{}- terminators to x86 translation table}\label{tab:trans-terms}
 \end{table}
@@
 \end{figure}
+
+\subsubsection{Implementing \mintinline{llvm}{icmp} instructions} \label{impl-cond}
+The \mintinline{llvm}{icmp} and \mintinline{llvm}{select} instructions both rely on the \texttt{cmpx} x86 instruction for most of the heavy lifting.
 

### 2024-01-12 article.tex
Explain bug

@@
 
-as can be seen in Figure \ref{fig:intervals-ocaml}.
+%as can be seen in Figure \ref{fig:intervals-ocaml}.
 
@@
 
+One particularly nasty oversight that was discovered a couple of days prior to handin stems from phi movs overwriting each other.
+Consider Figure \ref{fig:sqrt1.ll} where two phi instructions either move \texttt{\%6} to \texttt{label \%10} or \texttt{\%7} to \texttt{label \%5} depending on the condition.
+In some cases, even in greedy where each variable is assigned their own discrete register, their respective values are overwritten.
+This is due to \texttt{\%7} being moved to \texttt{\%6} prior to potentially branching back to \texttt{label \%5}.
+If phi moves are conducted unconditionally of branch condition \texttt{\%9}, the value of \texttt{\%6} prior to branching to \texttt{label \%10} is overwritten on the off-chance of returning to \texttt{label \%5}, causing the value assigned to \texttt{\%11} and returned on line 10 to always be equal to \texttt{\%7} instead of the intended \texttt{\%7} instead of the intended  \texttt{\%6}.
+
+\begin{figure}[H]
+   \begin{minipage}{0.58\textwidth}
+     \centering
+     \begin{minted}[linenos,bgcolor=LightGray]{llvm}
+ ; ...
+5:
+ %6 = phi i32 [%7, %5], [1, %3]
+ %7 = add i32 %6, 1
+ %8 = mul i32 %7, %7
+ %9 = icmp sgt i32 %8, %0
+ br i1 %9, label %10, label %5
+10:
+ %11 = phi i32 [%0, %1], [0, %3], [%6, %5]
+ ret i32 %11
+}
+     \end{minted}
+     \caption{\texttt{dune exec build -{}- -t x86}}\label{fig:sqrt1.ll}
+   \end{minipage}
+   \begin{minipage}{0.40\textwidth}
+     \centering
+     \begin{minted}[linenos,bgcolor=LightGray]{gas}
+# ...
+# br i1 %9, label %10, label %5
+movq	%r10, %rax
+cmpq	$0, %rax
+movq	%r8, %rax
+movq	%rax, %rdi
+je	_isqrt$5
+movq	%rdi, %rax
+movq	%rax, %r11
+jmp	_isqrt$10
+# ...
+     \end{minted}
+     \caption{TODO}\label{fig:sqrt1.s}
+   \end{minipage}
+\end{figure}
+This was resolved by relying on conditional movs immediately prior to branching conditionally.
+Specifically, the \texttt{cmoveq} instruction was used, which only moves the value if the equals flag (i.e.
+\texttt{ZF}) is set.
+
 % FIXME: check if above is sound

### 2024-01-12 article.tex
Finish explaining bug

@@
 
-The instructions included are
-\lstinline!trunc! has only been added in order to help cover more generated LLVM. % FIXME
+Some additional instructions are included, like \lstinline!trunc! and \lstinline!switch!, to widen the range of supported frontend languages with the clang compiler.
+%been added in order to help cover more generated LLVM. % FIXME
 %with only four distinct block terminators 
@@
 %\end{enumerate}
-The branching for most of these is trivial: after successful execution the flow of all but \lstinline!br!, \lstinline!ret!
+%The branching for most of these is trivial: after successful execution the flow of all but \lstinline!br!, \lstinline!ret!
 and \lstinline!unreachable!
@@
 
-\label{appendix:llvmis}
-\begin{enumerate}
-    \item binary operations \lstinline!add!, \lstinline!and!, \lstinline!ashr!, \lstinline!lshr!, \lstinline!mul!, \lstinline!or!, \lstinline!sdiv!,\\\lstinline!srem!, \lstinline!shl!, \lstinline!sub!
+%\label{appendix:llvmis}
+%\begin{enumerate}
+    %\item binary operations \lstinline!add!, \lstinline!and!, \lstinline!ashr!, \lstinline!lshr!, \lstinline!mul!, \lstinline!or!, \lstinline!sdiv!,\\\lstinline!srem!, \lstinline!shl!, \lstinline!sub!
 and \lstinline!xor!
-    \item integer comparison \lstinline!icmp! (conditions being \lstinline!eq!, \lstinline!ne!, \lstinline!sge!, \lstinline!sgt!, \lstinline!sle! and \lstinline!slt!)
-    \item memory/address operations \lstinline!alloca!, \lstinline!gep!, \lstinline!load! and \lstinline!store!
-    \item  \lstinline!mov! operations \lstinline!gep!, \lstinline!phi!, \lstinline!ptrtoint!, \lstinline!trunc!, and \lstinline!zext!
-    \item control flow operations \lstinline!br!, \lstinline!call! and \lstinline!ret!
-    \item a  nullary block terminator \lstinline!unreachable!
-\end{enumerate}
+    %\item integer comparison \lstinline!icmp! (conditions being \lstinline!eq!, \lstinline!ne!, \lstinline!sge!, \lstinline!sgt!, \lstinline!sle! and \lstinline!slt!)
+    %\item memory/address operations \lstinline!alloca!, \lstinline!gep!, \lstinline!load! and \lstinline!store!
+    %\item  \lstinline!mov! operations \lstinline!gep!, \lstinline!phi!, \lstinline!ptrtoint!, \lstinline!trunc!, and \lstinline!zext!
+    %\item control flow operations \lstinline!br!, \lstinline!call! and \lstinline!ret!
+    %\item a  nullary block terminator \lstinline!unreachable!
+%\end{enumerate}
 
@@
 \end{figure}
-This was resolved by relying on conditional movs immediately prior to branching conditionally.
-Specifically, the \texttt{cmoveq} instruction was used, which only moves the value if the equals flag (i.e.
+
+\noindent This was resolved by relying on conditional \texttt{mov}s immediately prior to the conditional branch (i.e.
+\texttt{je}).
+Specifically, the \texttt{cmoveq} instruction was used, which only moves from source to destination if the equality flag (i.e.
 \texttt{ZF}) is set.
+Unfortunately the \texttt{cmoveq} instruction doesn't support indirect addressing, so in order to support spilled variables both the values need to be moved into registers as can be seen in Figure \ref{fig:fixed-ind-sqrt1.s}.
+
+\begin{figure}[H]
+   \begin{minipage}{0.48\textwidth}
+     \centering
+     \begin{minted}[linenos,bgcolor=LightGray]{gas}
+# ...
+# br i1 %9, label %10, label %5
+movq	%rdx, %rax
+cmpq	$0, %rax
+movq	%rbx, %rax
+cmoveq	%rax, %rsi
+je	_isqrt$5
+movq	%rsi, %rax
+movq	%rax, %rdx
+jmp	_isqrt$10
+# ...
+     \end{minted}
+     \caption{TODO}\label{fig:fixed-sqrt1.s}
+   \end{minipage}
+   \begin{minipage}{0.48\textwidth}
+     \centering
+     \begin{minted}[linenos,bgcolor=LightGray]{gas}
+# ...
+# br i1 %13, label %14, label %9
+movq	-88(%rbp), %rax
+cmpq	$0, %rax
+movq	-72(%rbp), %rax
+movq	-16(%rbp), %rcx
+cmoveq	%rax, %rcx
+movq	%rcx, -16(%rbp)
+je	_main$9
+movq	-16(%rbp), %rax
+# ...
+     \end{minted}
+     \caption{\texttt{-a greedy -n 0}}\label{fig:fixed-ind-sqrt1.s}
+   \end{minipage}
+\end{figure}
 

### 2024-01-12 article.tex
Fix some placement

@@
 Because of this, only strategies that produce a \(k\)-colorable graph are worth considering, as additional spills negate the purpose entirely.
-One of the  strategies introduced in the Appel text is the \textit{Briggs} strategy  \cite[232]{tiger}, in which nodes \(a\) and \(b\) can be coalesced if and only if the resulting node \(ab\) has \(\leq K\) nieghbors of significant degree  (i.e.
+One of the  strategies introduced in the Appel text is the \textit{Briggs} strategy  \cite[232]{tiger}, in which nodes \(a\) and \(b\) can be coalesced iff the resulting node \(ab\) has \(\leq K\) nieghbors of significant degree  (i.e.
 \(\geq K\)).
@@
 Therefore they don't satisfy the Briggs criteria, and are not coalesced, resulting in the same interference graph as the \texttt{simple} allocator in Figure \ref{fig:countdown.dot}.
-But for some \(k \geq 2\), now there are fewer than \(k\) significant neighbors of \texttt{\%c2} (and \texttt{\%c1}), meaning the two can be coalesced, with the resulting node \texttt{\%c1, \%c2} having two neighbors of insignificant degree.
-
+But for some \(k \geq 2\), now there are fewer than \(k\) significant neighbors of \texttt{\%c2} (and \texttt{\%c1}), meaning the two can be coalesced, with the resulting node \texttt{\%c1, \%c2} having two neighbors of insignificant degree as can be seen in Figure \ref{fig:countdown.briggs.dot}.
 \begin{figure}[H]
@@
 
-\noindent The purpose is only to assign the same registers to different variables. You don't need to note down information about blocks etc.
+\noindent The purpose is only to assign the same registers to different variables, which stands to benefit from less \texttt{mov} instructions being used.
+%. You don't need to note down information about blocks etc.
 

### 2024-01-12 article.tex
Add comment about bug

@@
 \noindent If one were to note the \(\mathit{in}[s]\) and \(\mathit{out}[s]\) sets for each instruction \(s\) for each iteration, one would arrive at the two tables seen in Table \ref{tab:countll-rev-steps} and \ref{tab:countll-steps}.
-The number of iterations is nearly halfed when reverseing the list of instructions prior to dataflow as can be seen on \ref{tab:countll-rev-steps} when compared to \ref{tab:countll-steps}.
+The number of iterations is nearly halfed when reverseing the list of instructions prior to dataflow as can be seen on the amount of recursions of Table \ref{tab:countll-rev-steps} when compared to Table \ref{tab:countll-steps}.
 
@@
 
+\subsection{Test Suite}
+
+In addition to my own cases and other programs from the internet, some cases from the LLVM test suite are also used.
+
+The structure of the test suite repository has a directory of so-called 'single source' C programs totalling to 2113.
+Of these only 1222 compiled successfully with \texttt{clang -S -emit-llvm} and of these only 624 parsed sucessfuly with the strip utility, of which 425 compiled successfully.
+
+% find ./SingleSource -name '*.c' -exec clang --target=x86_64-unknown-darwin -S -emit-llvm {} \;
+
 % TODO: kommenter på stripper

### 2024-01-13 article.tex
stuff about testing framework

@@
 
+\noindent Terminators are translated with branching instructions instead. One significant factor that needs to be accounted for is the phi nodes.
 \begin{table}[H]
@@
  \hline
-ret     & epilogue, retq & \{\}\\
+ret     &  retq & \{\}\\
 br label \%l & jmp \%l & \{\}\\
@@
 The result of applying a binary operation,  for instance, should only affect the assigned register.
-
-\subsection{Test Suite}
-
-In addition to my own cases and other programs from the internet, some cases from the LLVM test suite are also used.
-
+Another is an extensive test suite, although achieving 100\% coverage is unrealistic.
+
+\subsection{Test Suite and  Framework}
+
+The test suite is quite extensive although not anywhere close to complete coverage.
+Some tests have been handwritten during development, mostly meant to assert one specific feature (unit tests), but others are also inteded to assert them working together (integration tests).
+Some of them are hand written (simple ones), others are based on the output of \texttt{clang -S -emit-llvm} stripped with the the strip utility in \texttt{bin/strip.ml}, and lastly some are generated with Python scripts to parameterize certain properties (e.g.
+long move-chains, many operations).
+In addition to this some are copied from other projects
+In addition to my own cases and other programs from the internet (attributed in the file header), some cases from the LLVM test suite are also used.
 The structure of the test suite repository has a directory of so-called 'single source' C programs totalling to 2113.
 Of these only 1222 compiled successfully with \texttt{clang -S -emit-llvm} and of these only 624 parsed sucessfuly with the strip utility, of which 425 compiled successfully.
+In addition to these there are also the test suite used throughout Tiger development, most of which were provided by the compilers lecturers, as well as some LLVM snippets from the Dolphin compiler.
+
+The testing framework is roughly based on the build system that came with the Tiger work environment.
+It is based on a collection of files that can be compiled and executed on their own.
+Each of these are enumerated in the test runner (\texttt{bin/testrunner.ml}) with parameters associated.
+Each of these are then compiled with the \texttt{clang} compiler and executed, capturing all the relevant output data, which is stdout and exit code.
+After this, the same test file is compiled for each of the allocators.
+The executables generated are then also executed, also storing the output data.
+All test cases are deterministic and will halt at some point.
+This is necessary because tests can easily end in an infinite loop because of erroneous translation.
+As each of these halt or time out, their output is compared to what was output from executables generated by the 'correct' compiler of \texttt{clang}.
 

### 2024-01-13 article.tex
Add some subset tables

@@
 
-\subsubsection{\lstinline!benches/ackermann.ll!}
-
-Another benchmark that derives highly from the recursive structure of \lstinline!fib.ll! above is the Ackermann function.
+%\subsubsection{\lstinline!benches/ackermann.ll!}
+
+%Another benchmark that derives highly from the recursive structure of \lstinline!fib.ll! above is the Ackermann function.
 
 %TODO: bedøm om det giver mening
+
+\subsubsection{\lstinline!benches/phis.ll!}
+Coalescing with the Briggs allocator is evidently quite beneficial, as the it does out perform most benchmarks.
+This is obviously because eliminating movs stand to gain.
+A more direct approach to demonstrating this is with parameterized benchmarks.
+While in theory a phi node has the potential to eliminate one \texttt{mov} instruction, in practice this might be more beneficial due to the implementation.
+The \texttt{benches/phis.py <n> <m>} script will generate a \texttt{.ll} program consisting of \(m\) phi nodes over \(n\) iterations.
+%Parameterized over how many phi moves need to be made per iteration.
+Sure most time might be spent in \texttt{@printf}, but this is constant across all runs, and in fact serves a purpose of adding extra load.
+
+\begin{NiceTabular}{|c|ccccccccc|}
+\hline
+arg(s) & clang & simple 12 & simple 2 & briggs 12 & briggs 2 & linear 2 & linear 12 & greedy 12 & greedy 0 \\
+\hline
+ & \cellcolor[rgb]{0.75,0.9990234375,0.75} 33699586& \cellcolor[rgb]{0.9990234375,0.9833984375,0.75} 2.49352x& \cellcolor[rgb]{0.9990234375,0.9833984375,0.75} 2.49358x& \cellcolor[rgb]{0.8115234375,0.9990234375,0.75} 1.24889x& \cellcolor[rgb]{0.8115234375,0.9990234375,0.75} 1.24890x& \cellcolor[rgb]{0.9990234375,0.9833984375,0.75} 2.49352x& \cellcolor[rgb]{0.9990234375,0.9833984375,0.75} 2.49355x& \cellcolor[rgb]{0.9990234375,0.9833984375,0.75} 2.49351x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 11.61953x\\
+ & \cellcolor[rgb]{0.75,0.9990234375,0.75} 33696431& \cellcolor[rgb]{0.9990234375,0.9814453125,0.75} 2.54359x& \cellcolor[rgb]{0.9990234375,0.9814453125,0.75} 2.55986x& \cellcolor[rgb]{0.8740234375,0.9990234375,0.75} 1.49846x& \cellcolor[rgb]{0.873046875,0.9990234375,0.75} 1.49796x& \cellcolor[rgb]{0.9990234375,0.9833984375,0.75} 2.49373x& \cellcolor[rgb]{0.9990234375,0.9833984375,0.75} 2.49378x& \cellcolor[rgb]{0.9990234375,0.9814453125,0.75} 2.54356x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 11.52326x\\
+ & \cellcolor[rgb]{0.75,0.9990234375,0.75} 50475128& \cellcolor[rgb]{0.9453125,0.9990234375,0.75} 1.78694x& \cellcolor[rgb]{0.998046875,0.9990234375,0.75} 1.99722x& \cellcolor[rgb]{0.791015625,0.9990234375,0.75} 1.16621x& \cellcolor[rgb]{0.83203125,0.9990234375,0.75} 1.33250x& \cellcolor[rgb]{0.9150390625,0.9990234375,0.75} 1.66480x& \cellcolor[rgb]{0.9150390625,0.9990234375,0.75} 1.66483x& \cellcolor[rgb]{0.94140625,0.9990234375,0.75} 1.77143x& \cellcolor[rgb]{0.9990234375,0.7939453125,0.75} 8.57567x\\
+ & \cellcolor[rgb]{0.75,0.9990234375,0.75} 67274018& \cellcolor[rgb]{0.935546875,0.9990234375,0.75} 1.74795x& \cellcolor[rgb]{0.966796875,0.9990234375,0.75} 1.87263x& \cellcolor[rgb]{0.76171875,0.9990234375,0.75} 1.04956x& \cellcolor[rgb]{0.7646484375,0.9990234375,0.75} 1.06208x& \cellcolor[rgb]{0.8740234375,0.9990234375,0.75} 1.49853x& \cellcolor[rgb]{0.8115234375,0.9990234375,0.75} 1.24908x& \cellcolor[rgb]{0.87890625,0.9990234375,0.75} 1.52063x& \cellcolor[rgb]{0.9990234375,0.8564453125,0.75} 6.57981x\\
+ & \cellcolor[rgb]{0.75,0.9990234375,0.75} 84029983& \cellcolor[rgb]{0.8486328125,0.9990234375,0.75} 1.39938x& \cellcolor[rgb]{0.9482421875,0.9990234375,0.75} 1.79869x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00002x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00003x& \cellcolor[rgb]{0.82421875,0.9990234375,0.75} 1.29955x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00004x& \cellcolor[rgb]{0.9990234375,0.8681640625,0.75} 6.19846x& \cellcolor[rgb]{0.9990234375,0.875,0.75} 5.98194x\\
+ & \cellcolor[rgb]{0.75,0.9990234375,0.75} 100810530& \cellcolor[rgb]{0.8740234375,0.9990234375,0.75} 1.49925x& \cellcolor[rgb]{0.95703125,0.9990234375,0.75} 1.83217x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99997x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99996x& \cellcolor[rgb]{0.8115234375,0.9990234375,0.75} 1.24967x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99998x& \cellcolor[rgb]{0.9990234375,0.9169921875,0.75} 4.60493x& \cellcolor[rgb]{0.9990234375,0.8798828125,0.75} 5.80149x\\
+ & \cellcolor[rgb]{0.75,0.9990234375,0.75} 117587019& \cellcolor[rgb]{0.837890625,0.9990234375,0.75} 1.35669x& \cellcolor[rgb]{0.9462890625,0.9990234375,0.75} 1.78907x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.88585x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.90488x& \cellcolor[rgb]{0.802734375,0.9990234375,0.75} 1.21403x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.85730x& \cellcolor[rgb]{0.9990234375,0.9326171875,0.75} 4.10513x& \cellcolor[rgb]{0.9990234375,0.888671875,0.75} 5.51438x\\
+ & \cellcolor[rgb]{0.75,0.9990234375,0.75} 134365885& \cellcolor[rgb]{0.8740234375,0.9990234375,0.75} 1.49947x& \cellcolor[rgb]{0.9521484375,0.9990234375,0.75} 1.81229x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.87512x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.87515x& \cellcolor[rgb]{0.7958984375,0.9990234375,0.75} 1.18732x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.87514x& \cellcolor[rgb]{0.9990234375,0.947265625,0.75} 3.63380x& \cellcolor[rgb]{0.9990234375,0.9111328125,0.75} 4.81382x\\
+\hline
+\end{NiceTabular}
+
+1074877208
+% -O0
+\begin{table}[H]
+\makebox[\textwidth][c]{
+\begin{NiceTabular}{|c|ccccccccc|}
+\hline
+phis & clang & simple 12 & simple 2 & briggs 12 & briggs 2 & linear 2 & linear 12 & greedy 12 & greedy 0 \\
+\hline
+ 0 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 150762& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99922x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00386x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00104x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99880x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00325x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99096x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00077x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99764x\\
+ 1 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 263112& \cellcolor[rgb]{0.9453125,0.9990234375,0.75} 1.78765x& \cellcolor[rgb]{0.9580078125,0.9990234375,0.75} 1.83810x& \cellcolor[rgb]{0.8154296875,0.9990234375,0.75} 1.26448x& \cellcolor[rgb]{0.9365234375,0.9990234375,0.75} 1.75223x& \cellcolor[rgb]{0.9384765625,0.9990234375,0.75} 1.75782x& \cellcolor[rgb]{0.939453125,0.9990234375,0.75} 1.76278x& \cellcolor[rgb]{0.9453125,0.9990234375,0.75} 1.78716x& \cellcolor[rgb]{0.9990234375,0.865234375,0.75} 6.27904x\\
+ 2 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 333724& \cellcolor[rgb]{0.8603515625,0.9990234375,0.75} 1.44531x& \cellcolor[rgb]{0.8955078125,0.9990234375,0.75} 1.58670x& \cellcolor[rgb]{0.7734375,0.9990234375,0.75} 1.09621x& \cellcolor[rgb]{0.7978515625,0.9990234375,0.75} 1.19483x& \cellcolor[rgb]{0.8466796875,0.9990234375,0.75} 1.39069x& \cellcolor[rgb]{0.8466796875,0.9990234375,0.75} 1.39096x& \cellcolor[rgb]{0.8642578125,0.9990234375,0.75} 1.45969x& \cellcolor[rgb]{0.9990234375,0.896484375,0.75} 5.28416x\\
+ 3 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 399042& \cellcolor[rgb]{0.8720703125,0.9990234375,0.75} 1.49027x& \cellcolor[rgb]{0.892578125,0.9990234375,0.75} 1.57311x& \cellcolor[rgb]{0.7568359375,0.9990234375,0.75} 1.03136x& \cellcolor[rgb]{0.7587890625,0.9990234375,0.75} 1.03867x& \cellcolor[rgb]{0.8310546875,0.9990234375,0.75} 1.32618x& \cellcolor[rgb]{0.7900390625,0.9990234375,0.75} 1.16354x& \cellcolor[rgb]{0.8310546875,0.9990234375,0.75} 1.32695x& \cellcolor[rgb]{0.9990234375,0.9169921875,0.75} 4.61992x\\
+ 4 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 464640& \cellcolor[rgb]{0.8193359375,0.9990234375,0.75} 1.28048x& \cellcolor[rgb]{0.8896484375,0.9990234375,0.75} 1.56253x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00015x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99836x& \cellcolor[rgb]{0.8017578125,0.9990234375,0.75} 1.21081x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99861x& \cellcolor[rgb]{0.9990234375,0.9375,0.75} 3.96512x& \cellcolor[rgb]{0.9990234375,0.9208984375,0.75} 4.48737x\\
+ 5 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 529535& \cellcolor[rgb]{0.841796875,0.9990234375,0.75} 1.37214x& \cellcolor[rgb]{0.9033203125,0.9990234375,0.75} 1.61569x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99931x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99959x& \cellcolor[rgb]{0.7939453125,0.9990234375,0.75} 1.17952x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99948x& \cellcolor[rgb]{0.9990234375,0.94921875,0.75} 3.58774x& \cellcolor[rgb]{0.9990234375,0.921875,0.75} 4.47259x\\
+ 6 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 594553& \cellcolor[rgb]{0.818359375,0.9990234375,0.75} 1.27726x& \cellcolor[rgb]{0.9013671875,0.9990234375,0.75} 1.61016x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.91261x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.92786x& \cellcolor[rgb]{0.7900390625,0.9990234375,0.75} 1.16424x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.89141x& \cellcolor[rgb]{0.9990234375,0.955078125,0.75} 3.38554x& \cellcolor[rgb]{0.9990234375,0.921875,0.75} 4.46924x\\
+ 7 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 659461& \cellcolor[rgb]{0.8486328125,0.9990234375,0.75} 1.39945x& \cellcolor[rgb]{0.90234375,0.9990234375,0.75} 1.61499x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.90111x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.90219x& \cellcolor[rgb]{0.787109375,0.9990234375,0.75} 1.15128x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.90144x& \cellcolor[rgb]{0.9990234375,0.96484375,0.75} 3.08855x& \cellcolor[rgb]{0.9990234375,0.935546875,0.75} 4.03288x\\
+\end{NiceTabular}}
+\caption{\label{tab:dumb-fib-greedy}Benchmark of \texttt{benches/fib.ll} output by \mintinline{bash}{dune exec bench -- -f fib -n 1000}}
+\end{table}
+
+% -O1
+\begin{table}[H]
+\makebox[\textwidth][c]{
+\begin{NiceTabular}{|c|ccccccccc|}
+\hline
+phis & clang & simple 12 & simple 2 & briggs 12 & briggs 2 & linear 2 & linear 12 & greedy 12 & greedy 0 \\
+\hline
+ & \cellcolor[rgb]{0.75,0.9990234375,0.75} 150756& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00269x& \cellcolor[rgb]{0.7509765625,0.9990234375,0.75} 1.00742x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00356x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00237x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99991x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00269x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99794x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00298x\\
+ & \cellcolor[rgb]{0.75,0.9990234375,0.75} 266509& \cellcolor[rgb]{0.94140625,0.9990234375,0.75} 1.76877x& \cellcolor[rgb]{0.9521484375,0.9990234375,0.75} 1.81332x& \cellcolor[rgb]{0.8115234375,0.9990234375,0.75} 1.24997x& \cellcolor[rgb]{0.93359375,0.9990234375,0.75} 1.74009x& \cellcolor[rgb]{0.93359375,0.9990234375,0.75} 1.73912x& \cellcolor[rgb]{0.9345703125,0.9990234375,0.75} 1.74314x& \cellcolor[rgb]{0.9404296875,0.9990234375,0.75} 1.76571x& \cellcolor[rgb]{0.9990234375,0.8681640625,0.75} 6.19913x\\
+ & \cellcolor[rgb]{0.75,0.9990234375,0.75} 331770& \cellcolor[rgb]{0.8720703125,0.9990234375,0.75} 1.49255x& \cellcolor[rgb]{0.8974609375,0.9990234375,0.75} 1.59420x& \cellcolor[rgb]{0.775390625,0.9990234375,0.75} 1.10326x& \cellcolor[rgb]{0.7998046875,0.9990234375,0.75} 1.20125x& \cellcolor[rgb]{0.8486328125,0.9990234375,0.75} 1.39740x& \cellcolor[rgb]{0.8486328125,0.9990234375,0.75} 1.39905x& \cellcolor[rgb]{0.8701171875,0.9990234375,0.75} 1.48485x& \cellcolor[rgb]{0.9990234375,0.8955078125,0.75} 5.31939x\\
+ & \cellcolor[rgb]{0.75,0.9990234375,0.75} 398255& \cellcolor[rgb]{0.8720703125,0.9990234375,0.75} 1.49334x& \cellcolor[rgb]{0.892578125,0.9990234375,0.75} 1.57569x& \cellcolor[rgb]{0.7578125,0.9990234375,0.75} 1.03280x& \cellcolor[rgb]{0.759765625,0.9990234375,0.75} 1.04240x& \cellcolor[rgb]{0.83203125,0.9990234375,0.75} 1.32999x& \cellcolor[rgb]{0.7900390625,0.9990234375,0.75} 1.16465x& \cellcolor[rgb]{0.83203125,0.9990234375,0.75} 1.33295x& \cellcolor[rgb]{0.9990234375,0.9169921875,0.75} 4.62699x\\
+ & \cellcolor[rgb]{0.75,0.9990234375,0.75} 463887& \cellcolor[rgb]{0.8203125,0.9990234375,0.75} 1.28291x& \cellcolor[rgb]{0.8876953125,0.9990234375,0.75} 1.55488x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00064x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00080x& \cellcolor[rgb]{0.8017578125,0.9990234375,0.75} 1.21103x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99887x& \cellcolor[rgb]{0.9990234375,0.9375,0.75} 3.97028x& \cellcolor[rgb]{0.9990234375,0.9208984375,0.75} 4.49752x\\
+ & \cellcolor[rgb]{0.75,0.9990234375,0.75} 529682& \cellcolor[rgb]{0.841796875,0.9990234375,0.75} 1.37111x& \cellcolor[rgb]{0.9033203125,0.9990234375,0.75} 1.61877x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99918x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00048x& \cellcolor[rgb]{0.7958984375,0.9990234375,0.75} 1.18536x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99971x& \cellcolor[rgb]{0.9990234375,0.94921875,0.75} 3.58938x& \cellcolor[rgb]{0.9990234375,0.921875,0.75} 4.44715x\\
+& \cellcolor[rgb]{0.75,0.9990234375,0.75} 594307& \cellcolor[rgb]{0.8173828125,0.9990234375,0.75} 1.27157x& \cellcolor[rgb]{0.9013671875,0.9990234375,0.75} 1.60988x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.91407x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.92725x& \cellcolor[rgb]{0.791015625,0.9990234375,0.75} 1.16608x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.89112x& \cellcolor[rgb]{0.9990234375,0.955078125,0.75} 3.38938x& \cellcolor[rgb]{0.9990234375,0.921875,0.75} 4.46031x\\
+ & \cellcolor[rgb]{0.75,0.9990234375,0.75} 659352& \cellcolor[rgb]{0.8486328125,0.9990234375,0.75} 1.39923x& \cellcolor[rgb]{0.9033203125,0.9990234375,0.75} 1.61575x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.90244x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.90256x& \cellcolor[rgb]{0.787109375,0.9990234375,0.75} 1.15120x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.90131x& \cellcolor[rgb]{0.9990234375,0.96484375,0.75} 3.08998x& \cellcolor[rgb]{0.9990234375,0.935546875,0.75} 4.03296x\\
+\end{NiceTabular}}
+\caption{\label{tab:dumb-fib-greedy}Benchmark of \texttt{benches/fib.ll} output by \mintinline{bash}{dune exec bench -- -f fib -n 1000}}
+\end{table}
+
+\begin{table}[H]
+\makebox[\textwidth][c]{
+\begin{NiceTabular}{|c|ccccccccc|}
+\hline
+arg(s) & clang & simple 12 & simple 2 & briggs 12 & briggs 2 & linear 2 & linear 12 & greedy 12 & greedy 0 \\
+\hline
+ & \cellcolor[rgb]{0.75,0.9990234375,0.75} 150790& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99873x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00204x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00055x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00202x& \cellcolor[rgb]{0.7509765625,0.9990234375,0.75} 1.00475x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99925x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99983x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00036x\\
+ & \cellcolor[rgb]{0.75,0.9990234375,0.75} 266858& \cellcolor[rgb]{0.939453125,0.9990234375,0.75} 1.76227x& \cellcolor[rgb]{0.951171875,0.9990234375,0.75} 1.81025x& \cellcolor[rgb]{0.8095703125,0.9990234375,0.75} 1.24026x& \cellcolor[rgb]{0.93359375,0.9990234375,0.75} 1.73967x& \cellcolor[rgb]{0.93359375,0.9990234375,0.75} 1.73845x& \cellcolor[rgb]{0.9326171875,0.9990234375,0.75} 1.73599x& \cellcolor[rgb]{0.939453125,0.9990234375,0.75} 1.76279x& \cellcolor[rgb]{0.9990234375,0.8681640625,0.75} 6.18927x\\
+ & \cellcolor[rgb]{0.75,0.9990234375,0.75} 332932& \cellcolor[rgb]{0.8681640625,0.9990234375,0.75} 1.47463x& \cellcolor[rgb]{0.896484375,0.9990234375,0.75} 1.58958x& \cellcolor[rgb]{0.7734375,0.9990234375,0.75} 1.09780x& \cellcolor[rgb]{0.7978515625,0.9990234375,0.75} 1.19567x& \cellcolor[rgb]{0.8466796875,0.9990234375,0.75} 1.39187x& \cellcolor[rgb]{0.84765625,0.9990234375,0.75} 1.39397x& \cellcolor[rgb]{0.86328125,0.9990234375,0.75} 1.45798x& \cellcolor[rgb]{0.9990234375,0.8955078125,0.75} 5.30079x\\
+ & \cellcolor[rgb]{0.75,0.9990234375,0.75} 396964& \cellcolor[rgb]{0.873046875,0.9990234375,0.75} 1.49802x& \cellcolor[rgb]{0.89453125,0.9990234375,0.75} 1.58318x& \cellcolor[rgb]{0.7578125,0.9990234375,0.75} 1.03435x& \cellcolor[rgb]{0.7607421875,0.9990234375,0.75} 1.04439x& \cellcolor[rgb]{0.8310546875,0.9990234375,0.75} 1.32655x& \cellcolor[rgb]{0.7919921875,0.9990234375,0.75} 1.16965x& \cellcolor[rgb]{0.8330078125,0.9990234375,0.75} 1.33411x& \cellcolor[rgb]{0.9990234375,0.916015625,0.75} 4.64216x\\
+ & \cellcolor[rgb]{0.75,0.9990234375,0.75} 463143& \cellcolor[rgb]{0.8193359375,0.9990234375,0.75} 1.27994x& \cellcolor[rgb]{0.8916015625,0.9990234375,0.75} 1.56894x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00193x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00219x& \cellcolor[rgb]{0.802734375,0.9990234375,0.75} 1.21525x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00036x& \cellcolor[rgb]{0.9990234375,0.9375,0.75} 3.97536x& \cellcolor[rgb]{0.9990234375,0.9208984375,0.75} 4.50366x\\
+ & \cellcolor[rgb]{0.75,0.9990234375,0.75} 528926& \cellcolor[rgb]{0.841796875,0.9990234375,0.75} 1.37202x& \cellcolor[rgb]{0.904296875,0.9990234375,0.75} 1.62053x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00265x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00090x& \cellcolor[rgb]{0.7958984375,0.9990234375,0.75} 1.18442x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00260x& \cellcolor[rgb]{0.9990234375,0.94921875,0.75} 3.59263x& \cellcolor[rgb]{0.9990234375,0.921875,0.75} 4.46336x\\
+ & \cellcolor[rgb]{0.75,0.9990234375,0.75} 594806& \cellcolor[rgb]{0.818359375,0.9990234375,0.75} 1.27651x& \cellcolor[rgb]{0.9013671875,0.9990234375,0.75} 1.60931x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.91222x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.92581x& \cellcolor[rgb]{0.791015625,0.9990234375,0.75} 1.16478x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.89116x& \cellcolor[rgb]{0.9990234375,0.955078125,0.75} 3.38589x& \cellcolor[rgb]{0.9990234375,0.921875,0.75} 4.46530x\\
+ & \cellcolor[rgb]{0.75,0.9990234375,0.75} 660398& \cellcolor[rgb]{0.84765625,0.9990234375,0.75} 1.39599x& \cellcolor[rgb]{0.9013671875,0.9990234375,0.75} 1.60987x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.89423x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.90015x& \cellcolor[rgb]{0.7861328125,0.9990234375,0.75} 1.14754x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.90032x& \cellcolor[rgb]{0.9990234375,0.96484375,0.75} 3.08470x& \cellcolor[rgb]{0.9990234375,0.935546875,0.75} 4.02611x\\
+\hline
+\end{NiceTabular}}
+\caption{\label{tab:dumb-fib-greedy}Benchmark of \texttt{benches/fib.ll} output by \mintinline{bash}{dune exec bench -- -f fib -n 1000}}
+\end{table}
+
+\subsubsection{\lstinline!benches/subset.ll!}
+A suitable example for nested loop iteration is \texttt{benches/subset.ll}.
+It's a very intuitive program that simply prints all \(2^n\) of a set of \(n\) elements, given as command line arguments.
+This is done by incrementing \(i\) by 1 from 0 to \(2^n-1\) and then printing the \(j\)th entry of the arguments array iff the \(j\)th bit of \(i\) is set.
+This leads to \(2^n\) repetitions of the outer loop and \(n\) of the inner, leading to \(2^n n\) repetitions.
+Since
+\begin{table}[H]
+\makebox[\textwidth][c]{
+\begin{NiceTabular}{|c|cccccccc|}
+\hline
+n & clang & simple 12 & simple 2 & briggs 12 & briggs 2 & linear & greedy 12 & greedy 0 \\
+\hline
+14 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.771484375,0.9990234375,0.75} 1.08770x& \cellcolor[rgb]{0.787109375,0.9990234375,0.75} 1.15188x& \cellcolor[rgb]{0.771484375,0.9990234375,0.75} 1.08832x& \cellcolor[rgb]{0.7890625,0.9990234375,0.75} 1.15813x& \cellcolor[rgb]{0.98046875,0.9990234375,0.75} 1.92564x& \cellcolor[rgb]{0.9990234375,0.98046875,0.75} 2.59504x& \cellcolor[rgb]{0.9990234375,0.96875,0.75} 2.95798x\\
+15 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.7705078125,0.9990234375,0.75} 1.08304x& \cellcolor[rgb]{0.78515625,0.9990234375,0.75} 1.14485x& \cellcolor[rgb]{0.7705078125,0.9990234375,0.75} 1.08536x& \cellcolor[rgb]{0.787109375,0.9990234375,0.75} 1.14973x& \cellcolor[rgb]{0.9833984375,0.9990234375,0.75} 1.93885x& \cellcolor[rgb]{0.9990234375,0.9794921875,0.75} 2.60300x& \cellcolor[rgb]{0.9990234375,0.96875,0.75} 2.96001x\\
+16 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.7705078125,0.9990234375,0.75} 1.08622x& \cellcolor[rgb]{0.78515625,0.9990234375,0.75} 1.14309x& \cellcolor[rgb]{0.7705078125,0.9990234375,0.75} 1.08345x& \cellcolor[rgb]{0.78515625,0.9990234375,0.75} 1.14388x& \cellcolor[rgb]{0.9853515625,0.9990234375,0.75} 1.94770x& \cellcolor[rgb]{0.9990234375,0.978515625,0.75} 2.62809x& \cellcolor[rgb]{0.9990234375,0.9677734375,0.75} 2.99463x\\
+17 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.7705078125,0.9990234375,0.75} 1.08533x& \cellcolor[rgb]{0.78515625,0.9990234375,0.75} 1.14335x& \cellcolor[rgb]{0.771484375,0.9990234375,0.75} 1.08652x& \cellcolor[rgb]{0.7861328125,0.9990234375,0.75} 1.14605x& \cellcolor[rgb]{0.986328125,0.9990234375,0.75} 1.94922x& \cellcolor[rgb]{0.9990234375,0.978515625,0.75} 2.64844x& \cellcolor[rgb]{0.9990234375,0.966796875,0.75} 3.01363x\\
+\hline
+\end{NiceTabular}}
+\caption{\label{tab:subset}Benchmark of \texttt{benches/sieven.ll} output by \mintinline{bash}{dune exec bench -- -f fib -n 1000}}
+\end{table}
 
@@
 
-\subsubsection{\lstinline!benches/subset.ll!}
-\begin{table}[H]
-\makebox[\textwidth][c]{
-\begin{NiceTabular}{|c|cccccccc|}
-\hline
-n & clang & simple 12 & simple 2 & briggs 12 & briggs 2 & linear & greedy 12 & greedy 0 \\
-\hline
-14 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.771484375,0.9990234375,0.75} 1.08770x& \cellcolor[rgb]{0.787109375,0.9990234375,0.75} 1.15188x& \cellcolor[rgb]{0.771484375,0.9990234375,0.75} 1.08832x& \cellcolor[rgb]{0.7890625,0.9990234375,0.75} 1.15813x& \cellcolor[rgb]{0.98046875,0.9990234375,0.75} 1.92564x& \cellcolor[rgb]{0.9990234375,0.98046875,0.75} 2.59504x& \cellcolor[rgb]{0.9990234375,0.96875,0.75} 2.95798x\\
-15 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.7705078125,0.9990234375,0.75} 1.08304x& \cellcolor[rgb]{0.78515625,0.9990234375,0.75} 1.14485x& \cellcolor[rgb]{0.7705078125,0.9990234375,0.75} 1.08536x& \cellcolor[rgb]{0.787109375,0.9990234375,0.75} 1.14973x& \cellcolor[rgb]{0.9833984375,0.9990234375,0.75} 1.93885x& \cellcolor[rgb]{0.9990234375,0.9794921875,0.75} 2.60300x& \cellcolor[rgb]{0.9990234375,0.96875,0.75} 2.96001x\\
-16 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.7705078125,0.9990234375,0.75} 1.08622x& \cellcolor[rgb]{0.78515625,0.9990234375,0.75} 1.14309x& \cellcolor[rgb]{0.7705078125,0.9990234375,0.75} 1.08345x& \cellcolor[rgb]{0.78515625,0.9990234375,0.75} 1.14388x& \cellcolor[rgb]{0.9853515625,0.9990234375,0.75} 1.94770x& \cellcolor[rgb]{0.9990234375,0.978515625,0.75} 2.62809x& \cellcolor[rgb]{0.9990234375,0.9677734375,0.75} 2.99463x\\
-17 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.7705078125,0.9990234375,0.75} 1.08533x& \cellcolor[rgb]{0.78515625,0.9990234375,0.75} 1.14335x& \cellcolor[rgb]{0.771484375,0.9990234375,0.75} 1.08652x& \cellcolor[rgb]{0.7861328125,0.9990234375,0.75} 1.14605x& \cellcolor[rgb]{0.986328125,0.9990234375,0.75} 1.94922x& \cellcolor[rgb]{0.9990234375,0.978515625,0.75} 2.64844x& \cellcolor[rgb]{0.9990234375,0.966796875,0.75} 3.01363x\\
-\hline
-\end{NiceTabular}}
-\caption{\label{tab:subset}Benchmark of \texttt{benches/sieven.ll} output by \mintinline{bash}{dune exec bench -- -f fib -n 1000}}
-\end{table}
-
 \subsubsection{\lstinline!benches/sha256.ll!}

### 2024-01-13 article.tex
Mostly coalescing code

@@
 \begin{equation}
-i a = \Phi (b_1, ..., b_n) \label{phi}
+ a = \Phi (b_1, ..., b_n) \label{phi}
 \end{equation}
@@
 \end{figure}
-\noindent From this point onwards  nothing can be done to further simplify the graph as all of the remaining nodes are of significant degree (\(\geq 2\)).
+\noindent From this point on  nothing can be done to further simplify the graph as all of the remaining nodes are of significant degree (\(\geq 2\)).
 Therefore, one of the remaining nodes is spilled to memory and entirely removed from the graph as can be seen on Figure \ref{fig:countdown.dot4} and the stack reset as can be seen on Figure \ref{fig:countdown.stack4}.
@@
 \noindent The purpose is only to assign the same registers to different variables, which stands to benefit from less \texttt{mov} instructions being used.
+\subsection{Implementation}
+
+The first step is deriving all preference edges:
+
+\begin{figure}[H]
+  \centering
+  \begin{minted}[linenos,bgcolor=LightGray]{ocaml}
+let prefer (insns : Cfg.insn list) : S.SS.t S.ST.t =
+  let insn t = function
+    | Cfg.Insn (Some d, Ll.PhiNode (_, ops)) ->
+        List.fold_left
+          (fun t o ->
+            match o with
+            | Ll.Id sop, _ ->
+                S.ST.update sop
+                  (function
+                    | Some s -> Some (S.SS.add d s)
+                    | None -> Some (S.SS.singleton d))
+                  t
+            | _ -> t)
+          t ops
+    | _ -> t
+  in
+  List.fold_left insn S.ST.empty insns
+  \end{minted}
+  \caption{Implementation of \texttt{prefer} in \texttt{lib/coalesce.ml}}\label{fig:prefer.ml}
+\end{figure}
+\noindent This is done simply by iterating over all of the instructions, and for every phi node, add a preference edge between \(a\) on the left-hand side of \eqref{phi} and all \(b_i\) on the right.
+Of note is the fact that only operands of the form \(\%i\) (i.e.
+\texttt{Id} variants of the \texttt{operand} type) are considered.
+This is because immediate values may also be specified in an LLVM phi node, whose \texttt{mov} instruction  don't make sense to eliminate.
+% FIXME: is this true?
+
+Briggs coalescing is then implemented as follows:
+\begin{figure}[H]
+  \centering
+  \begin{minted}[linenos,bgcolor=LightGray]{ocaml}
+let coalesce_briggs k (prefs : S.SS.t S.ST.t)
+    ((l, g) : Lva.G.V.t S.ST.t * Lva.G.t) : Lva.G.V.t S.table * Lva.G.t =
+  let try_coalesce sop dop (l, g) =
+    if not (Lva.G.mem_edge g (S.ST.find sop l) (S.ST.find dop l)) then
+      let sneighs = S.ST.find sop l |> Lva.G.succ g |> Lva.VS.of_list
+      and dneighs = S.ST.find dop l |> Lva.G.succ g |> Lva.VS.of_list in
+      let neighs = Lva.VS.union sneighs dneighs in
+      let sign v = Lva.G.succ g v |> List.length >= k in
+      let signeighs = Lva.VS.filter sign neighs in
+      if Lva.VS.cardinal signeighs < k then
+        coalesce (S.ST.find sop l) (S.ST.find dop l) (l, g)
+      else (l, g)
+    else (l, g)
+  in
+  let try_pref sop dops (l, g) = S.SS.fold (try_coalesce sop) dops (l, g) in
+  S.ST.fold try_pref prefs (l, g)
+  \end{minted}
+  \caption{Implementation of \texttt{coalesce\_briggs} in \texttt{lib/coalesce.ml}}\label{fig:coalesce_briggs.ml}
+\end{figure}
+\noindent For every preference edge, attempt to coalesce them into one. The \texttt{try\_coalesce} function is applied to every pair \((a,b)\) for all \(\Phi\text{-nodes}\).
+At first whether \(a\) and \(b\) are in conflict is checked, because only nodes that are not in interference may be coalesced, and if they are there is no point in finding their neighbors.
+The degree of potential node \(ab\) is taking all of the neighbors  of \(a\) and \(b\), and if the union of these is of magnitude \(\geq k\) it doesn't satisfy the Briggs criteria and no further action is taken.
+If it is \(<k\), i.e.
+node \(ab\) is of insignificant degree, the two are combined in the \texttt{coalesce} function.
+
 %. You don't need to note down information about blocks etc.
@@
 %Strictly speaking, one could optimize many more mov instructions away if
+
+The following is a very tedious implementation, stemming from the the \texttt{ocamlgraph} library simply doesn't support contracting a graph.
+Instead one is directed towards constructing a new graph and copying over the vertices and edges in that order.
+This is done by first folding over all existing vertices,  adding the one's that are not in the process of being coalesced, and when the first of the pair being coalesced is met, set the third tuple entry to be some, the associated data being the vertex identifier as well as it's associated symbols.
+Each vertex maps to a symbol set (i.e.
+\texttt{S.SS.t}) of variables already coalesced which is initially a singleton.
+
+\begin{figure}[H]
+  \centering
+  \begin{minted}[linenos,bgcolor=LightGray]{ocaml}
+let coalesce v1 v2 (st, g) =
+  if v1 = v2 then (st, g)
+  else
+    let g' = Lva.G.create () in
+    let st', vt', _ =
+      Lva.G.fold_vertex
+        (fun v (st, vt, jn) ->
+          if v = v1 || v = v2 then (
+            match jn with
+            | None ->
+                let s = Lva.G.V.label v in
+                (st, vt, Some (v, s))
+            | Some (v1, s1) ->
+                let s' = Lva.G.V.label v |> S.SS.union s1 in
+                let v' = Lva.G.V.create s' in
+                Lva.G.add_vertex g' v';
+                ( S.SS.fold (fun e l -> S.ST.add e v' l) s' st,
+                  Lva.VT.add v1 v' vt |> Lva.VT.add v v',
+                  None ))
+          else
+            let s = Lva.G.V.label v in
+            let v' = Lva.G.V.create s in
+            Lva.G.add_vertex g' v';
+            ( S.SS.fold (fun e st -> S.ST.add e v' st) s st,
+              Lva.VT.add v v' vt,
+              jn ))
+        g
+        (S.ST.empty, Lva.VT.empty, None)
+    in
+    Lva.G.iter_edges
+      (fun v1 v2 ->
+        let v1 = Lva.VT.find v1 vt' in
+        let v2 = Lva.VT.find v2 vt' in
+        if v1 <> v2 then Lva.G.add_edge g' v1 v2)
+      g;
+    (st', g')
+  \end{minted}
+  \caption{Implementation of \texttt{coalesce} in \texttt{lib/coalesce.ml}}\label{fig:coalesce.ml}
+\end{figure}
 
@@
 The instruction set used in this paper will be a union of the sets used in the 2022 and 2023 compilers courses in order to work as a drop-in replacement of LLVM for either of the two respective source languages:  Tiger and Dolphin.
-This instruction set is a subset of the one used in practice, as, for instance, neither of the languages implemented support exception handling, floating point operations and so on, and instead only strive to cover the basics of compilers.
+This instruction set is a subset of the one used in practice, as, for instance, neither of the languages implemented support interrupts, floating point operations and so on, and instead only strive to cover the basics of compilers.
 
@@
 
-The testing framework is roughly based on the build system that came with the Tiger work environment.
+The testing framework is roughly based on the build system that came with the Tiger development environment.
 It is based on a collection of files that can be compiled and executed on their own.
 Each of these are enumerated in the test runner (\texttt{bin/testrunner.ml}) with parameters associated.
-Each of these are then compiled with the \texttt{clang} compiler and executed, capturing all the relevant output data, which is stdout and exit code.
+Each of these are then compiled with the \texttt{clang} compiler and executed, capturing all the relevant output, which is stdout and exit code.
 After this, the same test file is compiled for each of the allocators.
 The executables generated are then also executed, also storing the output data.
-All test cases are deterministic and will halt at some point.
+All test cases are deterministic and will be  halted by configurable timeout at some point.
 This is necessary because tests can easily end in an infinite loop because of erroneous translation.

### 2024-01-13 article.tex
Remove harness

@@
 
-\subsubsection{Debugger Harness}
-A more elaborate way of testing correctness of translation is to assert that there are no effects except for those intended.
-One way of achieving this is to check if a group of x86 instructions representing a single LLVM-{}- operation change any values except those expected.
-An approach to this is to attach a debugger, insert a breakpoint before every LLVM-{}- instruction, and ensure that only the intended registers are altered.
-
-To achieve this, only a few changes need to be made to the codebase.
-A debug flag is introduced, which inserts a label before every instruction by adding a \lstinline!Breakpoint!
-variant of the \lstinline!ins!
-type before the assembly emitted of that specific operation.
-A unique identifier as well as a list of changes expected to be made is encoded.
-This includes a bitmask of GPRs that can possibly be changed (aside from the scratch registers of course) as well as which parts of the stack may be modified. % TODO: implement the stack checker
-In addition to this, each of the labels are output with a \lstinline!.globl! directive in the file header. A file is built with these extra symbols using the \lstinline!-d! flag.
-
 %To validate that this works as intended,

### 2024-01-13 article.tex
abc

@@
 The instruction set used in this paper will be a union of the sets used in the 2022 and 2023 compilers courses in order to work as a drop-in replacement of LLVM for either of the two respective source languages:  Tiger and Dolphin.
-This instruction set is a subset of the one used in practice, as, for instance, neither of the languages implemented support interrupts, floating point operations and so on, and instead only strive to cover the basics of compilers.
-
-Some additional instructions are included, like \lstinline!trunc! and \lstinline!switch!, to widen the range of supported frontend languages with the clang compiler.
+This instruction set is a subset of the one used in practice, as, for instance, neither of the languages implemented support interrupts, floating point operations etc., and instead only strive to cover the basics of compilers.
+In additional some other instructions are also included, like \lstinline!trunc!
+and \lstinline!switch!, to widen the range of supported frontend languages with the clang compiler.
 %been added in order to help cover more generated LLVM. % FIXME

### 2024-01-13 article.tex
Cleanup

@@
 
-The implementation is written in OCaml for the most part, although there is some marginal use of Python to interact with the LLVM toolchain's debugger 'LLDB'.
+The implementation is written in OCaml for the most part, although  Python has been put to use to generate further tests paramterized over some value.% to interact with the LLVM toolchain's debugger 'LLDB'.
 
@@
 
-Instead of translating directly from a source language to target machine code, most compiler frontends emit IR which is intended as an abstraction over particular CPU implementations.
-% FIXME: er det korrekt?
-lol
+Instead of translating directly from a source language to target machine code, most compiler frontends emit IR which is intended as an abstraction over CPU architectures. % FIXME: er det korrekt? lol
 It is by no means executable on either \(A\) or \(B\), but it is much closer to the instruction set executed on a CPU than the source language originally fed to the frontend.
@@
 
-Because of this, compilation of high-level programming languages will often go no further than emitting IR, then leave the rest for a subsequent backend toolchain.
+Because of this, compilation of high-level programming languages will often go no further than emitting IR, then leave the rest for a subsequent backend implementation.
 This naturally leads to the LLVM toolchain, which is by far the most used compiler backend in practice.
 Most languages, if targeting native execution on a CPU, will have an LLVM implementation.
-This is true for C/C++ (Clang compiler \cite{clang}), D (LDC compiler \cite{dlang}), Swift (official Swift compiler \cite{swift}) and Rust (official \texttt{rustc} compiler \cite{rustc}).
-The use of LLVM allows for them to be compiled to completely unforeseen architectures, like web browsers with WebAssembly or even graphics cards with compute kernels \cite{backend}.
+This is true for C/C++ (Clang compiler \cite{clang}), D (LDC compiler \cite{dlang}), Swift (official Swift compiler \cite{swift}) and Rust (official \texttt{rustc} compiler \cite{rustc}) to name a few.
+In fact, the use of LLVM allows for them to be compiled to completely unforeseen and unintended targets, like web browsers with WebAssembly or even graphics cards with compute kernels \cite{backend}.
 
@@
 That said, some projects have attempted to implement LLVM in their translation of bytecode to native architectures,  such as LLILC which promised both \textit{just-in-time} (JIT) as well as \textit{ahead-of-time} (AOT) compilation \cite{dotnet}.
-Similarly, Kotlin (the spiritual successor of Java) that ordinarily targets the JVM has a native backend that compiles directly to machine code using LLVM, circumventing the need for a VM entirely \cite{kotlin}.
+Similarly, Kotlin (the spiritual successor to Java) that ordinarily targets the JVM has a native backend that compiles directly to machine code using LLVM, circumventing the need for a VM entirely \cite{kotlin}.
 %Meanwhile interpreted languages like JavaScript eagerly attempt use LLVM for \textit{just-in-time} (JIT) compilation % FIXME: this is not true at all lol
@@
 At this level of complexity they are largely equivalent, with the primary difference appearing to be syntactic.
-LLVM IR keeps the typed binary operations as well as the function construct with a return statement, although the variable names are discarded as IR is generally not for human interaction, except the function name, as this is used for linking with other object files.
+LLVM IR keeps the typed binary operations as well as the function construct with a return statement, although the variable names are discarded as IR is generally not for human interaction, except the function name, as this is used for linking with other object files where the symbol naming is used to look up offsets.
 
@@
 \end{figure}
-When the IR is translated to x86 as seen in Figure \ref{fig:addx86}, the virtual variables \texttt{\%0}, \texttt{\%1} and \texttt{\%2} are assigned the registers \texttt{\%esi}, \texttt{\%edi} and \texttt{\%eax} respectively.
+\noindent When the IR is translated to x86 as seen in Figure \ref{fig:addx86}, the virtual variables \texttt{\%0}, \texttt{\%1} and \texttt{\%2} are assigned the registers \texttt{\%esi}, \texttt{\%edi} and \texttt{\%eax} respectively.
 This is presumably because of several optimizations, as \texttt{\%rdi} and \texttt{\%rsi} are used to pass the first and second parameters and the \texttt{\%rax} as the return value according to the System V AMD64 ABI \cite{sysv}, so by assigning the variables to those, several \texttt{mov}s are saved.
@@
 This circumvents the need for the same level of mutability as is achieved by reading/writing to memory, without violating the properties of SSA, by copying a value from the end of a predecessor to the beginning of the current block.
-A Phi node represents a point in the program where the control flow merges, and it selects the appropriate version of a variable based on the path taken.
+A phi node represents a point in the program where the control flow merges, and it selects the appropriate version of a variable based on the path taken.
 This ensures that the data flow is well-defined and allows for easy analysis across different control flow paths.
@@
 % TODO: comment on precision
+This project has opted to only use the Appel approach.
+One of the benefits is greater precision, in that several variables may be used in the same block and be assigned the same register as long as the two are not in interference within it.
+This wouldn't be possible with the Aho approach which completely discounts which parts of the block it is used in.
 Either way, leaders will have a set of predecessors and terminators a set of successors associated with them.
@@
 % TODO: rets
-
 Successors of node \(n\) are denoted \( \mathit{succ}[n], \)
@@
 The reason for the first block to be optionally named is the fact that \textit{some} LLVM programs transfer control back to the point of entry, whereas every subsequent block needs to be named because branching can only target it by referring to its name.
+Strictly speaking, this isn't how labels are treated in LLVM.
+Instead each block is always named, if not explicitly then by the next unused value from the same counter as temporaries \cite{llvm-functions}, but this was discovered too late in development to account for.
 
@@
 \end{minted}
+
+\noindent As can be seen above, instructions always have one edge to the instruction/terminator immediately following, while terminators can have any number of edges (0 for \texttt{ret} and \texttt{unreachable}, 1-2 for \texttt{br} and an unbounded number for \texttt{switch}).
 
@@
 
-\noindent Terminators are translated with branching instructions instead. One significant factor that needs to be accounted for is the phi nodes.
-\begin{table}[H]
-  \centering
-\begin{tabular}{|>{\tt}l|>{\tt}l|l|} 
- \hline
- LLVM-{}- & x86 & Registers affected\\
- \hline
-ret     &  retq & \{\}\\
-br label \%l & jmp \%l & \{\}\\
-br i1 \%c, label \%t, label \%f & cmpx \%c & \{\}\\
- & jz \%f & \\
- & jmp \%t & \\
- \hline
-\end{tabular}
-\caption{LLVM-{}- terminators to x86 translation table}\label{tab:trans-terms}
-\end{table}
+% TODO explain terminators
 
@@
 
-Prime factorization of \lstinline!i32!.
+Integer factorization is another interesting algorithm. Generally considered to be NP, here are the results of factorizing known primes of inreasing bit length. For \(n=1024\).
 \begin{table}[H]
 \makebox[\textwidth][c]{
-\begin{NiceTabular}{|c|cccccccc|}
-\hline
-n & clang & simple 12 & simple 2 & briggs 12 & briggs 2 & linear & greedy 12 & greedy 0 \\
-\hline
-268435399 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.7783203125,0.9990234375,0.75} 1.11499x& \cellcolor[rgb]{0.77734375,0.9990234375,0.75} 1.11025x& \cellcolor[rgb]{0.7783203125,0.9990234375,0.75} 1.11595x& \cellcolor[rgb]{0.779296875,0.9990234375,0.75} 1.11943x& \cellcolor[rgb]{0.916015625,0.9990234375,0.75} 1.66718x& \cellcolor[rgb]{0.9990234375,0.9833984375,0.75} 2.47709x& \cellcolor[rgb]{0.9990234375,0.984375,0.75} 2.44535x\\
-536870909 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.64555x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.64645x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.63974x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.64095x& \cellcolor[rgb]{0.7607421875,0.9990234375,0.75} 1.04505x& \cellcolor[rgb]{0.9013671875,0.9990234375,0.75} 1.60979x& \cellcolor[rgb]{0.904296875,0.9990234375,0.75} 1.61997x\\
-1073741789 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.7958984375,0.9990234375,0.75} 1.18438x& \cellcolor[rgb]{0.7978515625,0.9990234375,0.75} 1.19584x& \cellcolor[rgb]{0.796875,0.9990234375,0.75} 1.19047x& \cellcolor[rgb]{0.7958984375,0.9990234375,0.75} 1.18472x& \cellcolor[rgb]{0.9990234375,0.9951171875,0.75} 2.09863x& \cellcolor[rgb]{0.9990234375,0.9560546875,0.75} 3.36121x& \cellcolor[rgb]{0.9990234375,0.9560546875,0.75} 3.37087x\\
-2147483647 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.8017578125,0.9990234375,0.75} 1.20872x& \cellcolor[rgb]{0.80078125,0.9990234375,0.75} 1.20665x& \cellcolor[rgb]{0.80078125,0.9990234375,0.75} 1.20405x& \cellcolor[rgb]{0.8017578125,0.9990234375,0.75} 1.20826x& \cellcolor[rgb]{0.9990234375,0.98828125,0.75} 2.32725x& \cellcolor[rgb]{0.9990234375,0.9453125,0.75} 3.71253x& \cellcolor[rgb]{0.9990234375,0.9453125,0.75} 3.71693x\\
-4294967291 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.79296875,0.9990234375,0.75} 1.17597x& \cellcolor[rgb]{0.794921875,0.9990234375,0.75} 1.18070x& \cellcolor[rgb]{0.794921875,0.9990234375,0.75} 1.18067x& \cellcolor[rgb]{0.794921875,0.9990234375,0.75} 1.18159x& \cellcolor[rgb]{0.9990234375,0.986328125,0.75} 2.37710x& \cellcolor[rgb]{0.9990234375,0.9345703125,0.75} 4.04637x& \cellcolor[rgb]{0.9990234375,0.9345703125,0.75} 4.06492x\\
-8589934583 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.8203125,0.9990234375,0.75} 1.28534x& \cellcolor[rgb]{0.822265625,0.9990234375,0.75} 1.29317x& \cellcolor[rgb]{0.822265625,0.9990234375,0.75} 1.29145x& \cellcolor[rgb]{0.8212890625,0.9990234375,0.75} 1.28941x& \cellcolor[rgb]{0.9990234375,0.9755859375,0.75} 2.73679x& \cellcolor[rgb]{0.9990234375,0.9130859375,0.75} 4.74271x& \cellcolor[rgb]{0.9990234375,0.912109375,0.75} 4.76719x\\
-17179869143 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.8291015625,0.9990234375,0.75} 1.31790x& \cellcolor[rgb]{0.830078125,0.9990234375,0.75} 1.32297x& \cellcolor[rgb]{0.8310546875,0.9990234375,0.75} 1.32585x& \cellcolor[rgb]{0.830078125,0.9990234375,0.75} 1.32452x& \cellcolor[rgb]{0.9990234375,0.96875,0.75} 2.96961x& \cellcolor[rgb]{0.9990234375,0.900390625,0.75} 5.15949x& \cellcolor[rgb]{0.9990234375,0.9033203125,0.75} 5.07055x\\
-34359738337 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.8369140625,0.9990234375,0.75} 1.34905x& \cellcolor[rgb]{0.8369140625,0.9990234375,0.75} 1.34969x& \cellcolor[rgb]{0.8359375,0.9990234375,0.75} 1.34780x& \cellcolor[rgb]{0.845703125,0.9990234375,0.75} 1.38507x& \cellcolor[rgb]{0.9990234375,0.96484375,0.75} 3.08190x& \cellcolor[rgb]{0.9990234375,0.8935546875,0.75} 5.37886x& \cellcolor[rgb]{0.9990234375,0.892578125,0.75} 5.39434x\\
+\begin{NiceTabular}{|c|ccccccccc|}
+\hline
+arg(s) & clang & simple 12 & simple 2 & briggs 12 & briggs 2 & linear 12 & linear 2 & greedy 12 & greedy 0 \\
+\hline
+16777213 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 167130& \cellcolor[rgb]{0.759765625,0.9990234375,0.75} 1.04160x& \cellcolor[rgb]{0.7509765625,0.9990234375,0.75} 1.00651x& \cellcolor[rgb]{0.7626953125,0.9990234375,0.75} 1.05411x& \cellcolor[rgb]{0.7626953125,0.9990234375,0.75} 1.05372x& \cellcolor[rgb]{0.7607421875,0.9990234375,0.75} 1.04686x& \cellcolor[rgb]{0.958984375,0.9990234375,0.75} 1.84213x& \cellcolor[rgb]{0.9990234375,0.98828125,0.75} 2.33215x& \cellcolor[rgb]{0.9990234375,0.9873046875,0.75} 2.34511x\\
+33554393 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 175409& \cellcolor[rgb]{0.7646484375,0.9990234375,0.75} 1.06219x& \cellcolor[rgb]{0.7646484375,0.9990234375,0.75} 1.06070x& \cellcolor[rgb]{0.765625,0.9990234375,0.75} 1.06415x& \cellcolor[rgb]{0.7646484375,0.9990234375,0.75} 1.06212x& \cellcolor[rgb]{0.7646484375,0.9990234375,0.75} 1.06039x& \cellcolor[rgb]{0.9990234375,0.9951171875,0.75} 2.12199x& \cellcolor[rgb]{0.9990234375,0.974609375,0.75} 2.78336x& \cellcolor[rgb]{0.9990234375,0.9736328125,0.75} 2.79956x\\
+67108859 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 185779& \cellcolor[rgb]{0.7685546875,0.9990234375,0.75} 1.07588x& \cellcolor[rgb]{0.76953125,0.9990234375,0.75} 1.07923x& \cellcolor[rgb]{0.7705078125,0.9990234375,0.75} 1.08373x& \cellcolor[rgb]{0.76953125,0.9990234375,0.75} 1.08043x& \cellcolor[rgb]{0.76953125,0.9990234375,0.75} 1.07895x& \cellcolor[rgb]{0.9990234375,0.9833984375,0.75} 2.49571x& \cellcolor[rgb]{0.9990234375,0.955078125,0.75} 3.38089x& \cellcolor[rgb]{0.9990234375,0.955078125,0.75} 3.40121x\\
+134217689 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 197302& \cellcolor[rgb]{0.7802734375,0.9990234375,0.75} 1.12173x& \cellcolor[rgb]{0.775390625,0.9990234375,0.75} 1.10225x& \cellcolor[rgb]{0.779296875,0.9990234375,0.75} 1.12120x& \cellcolor[rgb]{0.779296875,0.9990234375,0.75} 1.12080x& \cellcolor[rgb]{0.779296875,0.9990234375,0.75} 1.11794x& \cellcolor[rgb]{0.9990234375,0.9677734375,0.75} 3.00248x& \cellcolor[rgb]{0.9990234375,0.9306640625,0.75} 4.17628x& \cellcolor[rgb]{0.9990234375,0.9296875,0.75} 4.20718x\\
+268435399 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 218313& \cellcolor[rgb]{0.7861328125,0.9990234375,0.75} 1.14571x& \cellcolor[rgb]{0.7861328125,0.9990234375,0.75} 1.14655x& \cellcolor[rgb]{0.7861328125,0.9990234375,0.75} 1.14531x& \cellcolor[rgb]{0.78515625,0.9990234375,0.75} 1.14376x& \cellcolor[rgb]{0.7861328125,0.9990234375,0.75} 1.14572x& \cellcolor[rgb]{0.9990234375,0.9501953125,0.75} 3.55232x& \cellcolor[rgb]{0.9990234375,0.9033203125,0.75} 5.06141x& \cellcolor[rgb]{0.9990234375,0.90234375,0.75} 5.08687x\\
+536870909 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 245682& \cellcolor[rgb]{0.794921875,0.9990234375,0.75} 1.18408x& \cellcolor[rgb]{0.794921875,0.9990234375,0.75} 1.18299x& \cellcolor[rgb]{0.794921875,0.9990234375,0.75} 1.18404x& \cellcolor[rgb]{0.794921875,0.9990234375,0.75} 1.18194x& \cellcolor[rgb]{0.794921875,0.9990234375,0.75} 1.18307x& \cellcolor[rgb]{0.9990234375,0.9296875,0.75} 4.20818x& \cellcolor[rgb]{0.9990234375,0.8720703125,0.75} 6.06747x& \cellcolor[rgb]{0.9990234375,0.8701171875,0.75} 6.13899x\\
+1073741789 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 285913& \cellcolor[rgb]{0.8037109375,0.9990234375,0.75} 1.21667x& \cellcolor[rgb]{0.8037109375,0.9990234375,0.75} 1.21894x& \cellcolor[rgb]{0.8037109375,0.9990234375,0.75} 1.21755x& \cellcolor[rgb]{0.8037109375,0.9990234375,0.75} 1.21668x& \cellcolor[rgb]{0.8037109375,0.9990234375,0.75} 1.21706x& \cellcolor[rgb]{0.9990234375,0.908203125,0.75} 4.89647x& \cellcolor[rgb]{0.9990234375,0.8369140625,0.75} 7.20172x& \cellcolor[rgb]{0.9990234375,0.8359375,0.75} 7.23917x\\
 \hline
@@
 \makebox[\textwidth][c]{
-\begin{NiceTabular}{|c|cccccccc|}
-\hline
-n & clang & simple 12 & simple 2 & briggs 12 & briggs 2 & linear & greedy 12 & greedy 0 \\
-\hline
-268435399 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.7783203125,0.9990234375,0.75} 1.11499x& \cellcolor[rgb]{0.77734375,0.9990234375,0.75} 1.11025x& \cellcolor[rgb]{0.7783203125,0.9990234375,0.75} 1.11595x& \cellcolor[rgb]{0.779296875,0.9990234375,0.75} 1.11943x& \cellcolor[rgb]{0.916015625,0.9990234375,0.75} 1.66718x& \cellcolor[rgb]{0.9990234375,0.9833984375,0.75} 2.47709x& \cellcolor[rgb]{0.9990234375,0.984375,0.75} 2.44535x\\
-536870909 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.64555x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.64645x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.63974x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.64095x& \cellcolor[rgb]{0.7607421875,0.9990234375,0.75} 1.04505x& \cellcolor[rgb]{0.9013671875,0.9990234375,0.75} 1.60979x& \cellcolor[rgb]{0.904296875,0.9990234375,0.75} 1.61997x\\
-1073741789 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.7958984375,0.9990234375,0.75} 1.18438x& \cellcolor[rgb]{0.7978515625,0.9990234375,0.75} 1.19584x& \cellcolor[rgb]{0.796875,0.9990234375,0.75} 1.19047x& \cellcolor[rgb]{0.7958984375,0.9990234375,0.75} 1.18472x& \cellcolor[rgb]{0.9990234375,0.9951171875,0.75} 2.09863x& \cellcolor[rgb]{0.9990234375,0.9560546875,0.75} 3.36121x& \cellcolor[rgb]{0.9990234375,0.9560546875,0.75} 3.37087x\\
-2147483647 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.8017578125,0.9990234375,0.75} 1.20872x& \cellcolor[rgb]{0.80078125,0.9990234375,0.75} 1.20665x& \cellcolor[rgb]{0.80078125,0.9990234375,0.75} 1.20405x& \cellcolor[rgb]{0.8017578125,0.9990234375,0.75} 1.20826x& \cellcolor[rgb]{0.9990234375,0.98828125,0.75} 2.32725x& \cellcolor[rgb]{0.9990234375,0.9453125,0.75} 3.71253x& \cellcolor[rgb]{0.9990234375,0.9453125,0.75} 3.71693x\\
-4294967291 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.79296875,0.9990234375,0.75} 1.17597x& \cellcolor[rgb]{0.794921875,0.9990234375,0.75} 1.18070x& \cellcolor[rgb]{0.794921875,0.9990234375,0.75} 1.18067x& \cellcolor[rgb]{0.794921875,0.9990234375,0.75} 1.18159x& \cellcolor[rgb]{0.9990234375,0.986328125,0.75} 2.37710x& \cellcolor[rgb]{0.9990234375,0.9345703125,0.75} 4.04637x& \cellcolor[rgb]{0.9990234375,0.9345703125,0.75} 4.06492x\\
-8589934583 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.8203125,0.9990234375,0.75} 1.28534x& \cellcolor[rgb]{0.822265625,0.9990234375,0.75} 1.29317x& \cellcolor[rgb]{0.822265625,0.9990234375,0.75} 1.29145x& \cellcolor[rgb]{0.8212890625,0.9990234375,0.75} 1.28941x& \cellcolor[rgb]{0.9990234375,0.9755859375,0.75} 2.73679x& \cellcolor[rgb]{0.9990234375,0.9130859375,0.75} 4.74271x& \cellcolor[rgb]{0.9990234375,0.912109375,0.75} 4.76719x\\
-17179869143 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.8291015625,0.9990234375,0.75} 1.31790x& \cellcolor[rgb]{0.830078125,0.9990234375,0.75} 1.32297x& \cellcolor[rgb]{0.8310546875,0.9990234375,0.75} 1.32585x& \cellcolor[rgb]{0.830078125,0.9990234375,0.75} 1.32452x& \cellcolor[rgb]{0.9990234375,0.96875,0.75} 2.96961x& \cellcolor[rgb]{0.9990234375,0.900390625,0.75} 5.15949x& \cellcolor[rgb]{0.9990234375,0.9033203125,0.75} 5.07055x\\
-34359738337 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.8369140625,0.9990234375,0.75} 1.34905x& \cellcolor[rgb]{0.8369140625,0.9990234375,0.75} 1.34969x& \cellcolor[rgb]{0.8359375,0.9990234375,0.75} 1.34780x& \cellcolor[rgb]{0.845703125,0.9990234375,0.75} 1.38507x& \cellcolor[rgb]{0.9990234375,0.96484375,0.75} 3.08190x& \cellcolor[rgb]{0.9990234375,0.8935546875,0.75} 5.37886x& \cellcolor[rgb]{0.9990234375,0.892578125,0.75} 5.39434x\\
+\begin{NiceTabular}{|c|ccccccccc|}
+\hline
+arg(s) & clang & simple 12 & simple 2 & briggs 12 & briggs 2 & linear 12 & linear 2 & greedy 12 & greedy 0 \\
+\hline
+268435399 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 218996& \cellcolor[rgb]{0.7666015625,0.9990234375,0.75} 1.07020x& \cellcolor[rgb]{0.7763671875,0.9990234375,0.75} 1.10754x& \cellcolor[rgb]{0.767578125,0.9990234375,0.75} 1.07114x& \cellcolor[rgb]{0.7763671875,0.9990234375,0.75} 1.10950x& \cellcolor[rgb]{0.7744140625,0.9990234375,0.75} 1.09908x& \cellcolor[rgb]{0.90234375,0.9990234375,0.75} 1.61452x& \cellcolor[rgb]{0.9990234375,0.986328125,0.75} 2.38305x& \cellcolor[rgb]{0.9990234375,0.986328125,0.75} 2.38509x\\
+536870909 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 247101& \cellcolor[rgb]{0.7705078125,0.9990234375,0.75} 1.08257x& \cellcolor[rgb]{0.7822265625,0.9990234375,0.75} 1.13159x& \cellcolor[rgb]{0.7705078125,0.9990234375,0.75} 1.08546x& \cellcolor[rgb]{0.7822265625,0.9990234375,0.75} 1.13164x& \cellcolor[rgb]{0.783203125,0.9990234375,0.75} 1.13364x& \cellcolor[rgb]{0.947265625,0.9990234375,0.75} 1.79438x& \cellcolor[rgb]{0.9990234375,0.9755859375,0.75} 2.73111x& \cellcolor[rgb]{0.9990234375,0.9755859375,0.75} 2.73643x\\
+1073741789 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 285989& \cellcolor[rgb]{0.775390625,0.9990234375,0.75} 1.10503x& \cellcolor[rgb]{0.7900390625,0.9990234375,0.75} 1.16197x& \cellcolor[rgb]{0.775390625,0.9990234375,0.75} 1.10359x& \cellcolor[rgb]{0.7900390625,0.9990234375,0.75} 1.16233x& \cellcolor[rgb]{0.7900390625,0.9990234375,0.75} 1.16392x& \cellcolor[rgb]{0.9912109375,0.9990234375,0.75} 1.97079x& \cellcolor[rgb]{0.9990234375,0.9638671875,0.75} 3.11293x& \cellcolor[rgb]{0.9990234375,0.9638671875,0.75} 3.12110x\\
+2147483647 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 341802& \cellcolor[rgb]{0.779296875,0.9990234375,0.75} 1.12053x& \cellcolor[rgb]{0.7958984375,0.9990234375,0.75} 1.18807x& \cellcolor[rgb]{0.7802734375,0.9990234375,0.75} 1.12297x& \cellcolor[rgb]{0.7978515625,0.9990234375,0.75} 1.19256x& \cellcolor[rgb]{0.796875,0.9990234375,0.75} 1.19163x& \cellcolor[rgb]{0.9990234375,0.994140625,0.75} 2.14407x& \cellcolor[rgb]{0.9990234375,0.9521484375,0.75} 3.49973x& \cellcolor[rgb]{0.9990234375,0.951171875,0.75} 3.50847x\\
+4294967291 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 419941& \cellcolor[rgb]{0.78515625,0.9990234375,0.75} 1.14339x& \cellcolor[rgb]{0.8046875,0.9990234375,0.75} 1.22187x& \cellcolor[rgb]{0.78515625,0.9990234375,0.75} 1.14366x& \cellcolor[rgb]{0.8046875,0.9990234375,0.75} 1.22078x& \cellcolor[rgb]{0.8046875,0.9990234375,0.75} 1.22093x& \cellcolor[rgb]{0.9990234375,0.98828125,0.75} 2.31858x& \cellcolor[rgb]{0.9990234375,0.9404296875,0.75} 3.87528x& \cellcolor[rgb]{0.9990234375,0.939453125,0.75} 3.88825x\\
+8589934583 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 530002& \cellcolor[rgb]{0.7900390625,0.9990234375,0.75} 1.16154x& \cellcolor[rgb]{0.8115234375,0.9990234375,0.75} 1.24776x& \cellcolor[rgb]{0.7900390625,0.9990234375,0.75} 1.16318x& \cellcolor[rgb]{0.810546875,0.9990234375,0.75} 1.24673x& \cellcolor[rgb]{0.8115234375,0.9990234375,0.75} 1.24866x& \cellcolor[rgb]{0.9990234375,0.9833984375,0.75} 2.47972x& \cellcolor[rgb]{0.9990234375,0.9296875,0.75} 4.22355x& \cellcolor[rgb]{0.9990234375,0.9287109375,0.75} 4.23487x\\
+17179869143 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 685889& \cellcolor[rgb]{0.7939453125,0.9990234375,0.75} 1.17835x& \cellcolor[rgb]{0.8173828125,0.9990234375,0.75} 1.27310x& \cellcolor[rgb]{0.7939453125,0.9990234375,0.75} 1.17878x& \cellcolor[rgb]{0.8173828125,0.9990234375,0.75} 1.27365x& \cellcolor[rgb]{0.8173828125,0.9990234375,0.75} 1.27361x& \cellcolor[rgb]{0.9990234375,0.9794921875,0.75} 2.61753x& \cellcolor[rgb]{0.9990234375,0.919921875,0.75} 4.52244x& \cellcolor[rgb]{0.9990234375,0.919921875,0.75} 4.53682x\\
+34359738337 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 905070& \cellcolor[rgb]{0.796875,0.9990234375,0.75} 1.19204x& \cellcolor[rgb]{0.8232421875,0.9990234375,0.75} 1.29498x& \cellcolor[rgb]{0.7978515625,0.9990234375,0.75} 1.19218x& \cellcolor[rgb]{0.8232421875,0.9990234375,0.75} 1.29535x& \cellcolor[rgb]{0.8232421875,0.9990234375,0.75} 1.29450x& \cellcolor[rgb]{0.9990234375,0.9755859375,0.75} 2.73632x& \cellcolor[rgb]{0.9990234375,0.912109375,0.75} 4.77726x& \cellcolor[rgb]{0.9990234375,0.912109375,0.75} 4.79190x\\
 \hline

### 2024-01-13 article.tex
af

@@
 %, the former of which must satisfy a transfer function a
-The transfer constraint is based on a \textit{transfer function} that describes how liveness is affected across instructions.
+The transfer constraint is based on a \textit{transfer function} \cite[599]{dragon} that describes how liveness is affected across instructions.
 For each instruction, there is a transfer function that describes how liveness changes from one point to the one immediately after.
 For example, as an arithmetic operation needs to be assigned a new temporary variable, the liveness of a new variable is propagated to all  instructions  executed subsequently.
-This is done by applying the transfer function to the current \textit{live-out} set to stop further propagation of variables defined by the currently visited instruction.
-% TODO: In fact, because the IR is in SSA form, the only time
+This is done by applying the transfer function to the current \textit{live-out} set to stop further propagation of variables defined by the currently visited instruction: % TODO: In fact, because the IR is in SSA form, the only time
 \begin{align}\label{flowin}
@@
 This is done by propagating the union of the \textit{live-in} set associated with all immediate successor nodes.
-This is also referred to as the \textit{meet operator}, whose operator depends on the type of dataflow analysis as well, but for liveness analysis a union is performed on the \textit{live-in} sets of  any successive  nodes:
+This is also referred to as the \textit{meet operator} \cite[605]{dragon}, whose operator depends on the type of dataflow analysis as well, but for liveness analysis a union is performed on the \textit{live-in} sets of  any successive  nodes:
 \begin{align}\label{flowout}
@@
 
-The simplest equations to implement are the \lstinline!def!
-and \lstinline!use!
-functions, as all of the values of interest are located immediately within the instruction itself and not hidden behind some layer of indirection:
+The simplest equations to implement are the \texttt{def} and \texttt{use} functions, as all of the values of interest are located immediately within the instruction itself and not hidden behind some layer of indirection:
 \begin{minted}[linenos,bgcolor=LightGray]{ocaml}
@@
 \end{minted}
-Where \lstinline!S.SS! is a \lstinline!Set.S! module built over the \lstinline!symbol! type found in \lstinline!lib/symbol.ml!:
-\begin{minted}[linenos,bgcolor=LightGray]{ocaml}
-type symbol = string * int
-(* ... *)
-module SS = Set.Make (struct
-  type t = symbol
-  let compare (_, n1) (_, n2) = compare n1 n2
-end)
-type set = SS.t
-\end{minted}
-
-\noindent The \texttt{flowin} function corresponds to the \textit{live-in} equation \eqref{flowin} and is implemented as follows:
+Where \texttt{S.SS} is a \texttt{Set.S} module built over the \texttt{Symbol.symbol} type found in \texttt{lib/symbol.ml}.
+The \texttt{flowin} function corresponds to the \textit{live-in} equation \eqref{flowin} and is implemented as follows:
 \begin{minted}[linenos,bgcolor=LightGray]{ocaml}
@@
 \end{minted}
-
-With the actual dataflow analysis performed recursively as follows:
+The set operations are applied as you would with the \texttt{S.SS.t} struct, the only interesting part is noting whether changes were made, as this is needed further down in execution.
+The actual dataflow analysis performed recursively as follows:
 \begin{figure}[H]
@@
     if List.fold_left flow false insns then dataflow () else (in_, out)
-  in
-  dataflow ()
+  in dataflow ()
 \end{minted}
@@
 
-\noindent If one were to note the \(\mathit{in}[s]\) and \(\mathit{out}[s]\) sets for each instruction \(s\) for each iteration, one would arrive at the two tables seen in Table \ref{tab:countll-rev-steps} and \ref{tab:countll-steps}.
-The number of iterations is nearly halfed when reverseing the list of instructions prior to dataflow as can be seen on the amount of recursions of Table \ref{tab:countll-rev-steps} when compared to Table \ref{tab:countll-steps}.
+\noindent If one were to note the \(\mathit{in}[s]\) and \(\mathit{out}[s]\) sets for each instruction \(s\) for each iteration, one would arrive at Tables \ref{tab:countll-rev-steps} and \ref{tab:countll-steps}.
+The number of iterations is nearly halfed when reverseing the list of instructions prior to dataflow analysis as can be seen on
+how many columns are present in each.
+%the amount of recursions of Table \ref{tab:countll-rev-steps} when compared to Table \ref{tab:countll-steps}.
 
@@
   \hline
-  i & use & def & in & out & in & out & in & out & in & out & in & out & in & out & in & out\\
+  i & use & def & in & out & in & out & in & out & in & out & in & out & in & out & in & out\\[-0.5ex]
   \hline
@@
   \hline
-  i & use & def & in & out & in & out & in & out & in & out & in & out & in & out & in & out & in & out & in & out & in & out & in & out\\
+  i & use & def & in & out & in & out & in & out & in & out & in & out & in & out & in & out & in & out & in & out & in & out & in & out\\[-0.5ex]
   \hline
@@
 Again the \texttt{OCamlgraph} library is used, but this time \texttt{Imperative.Graph.Abstract} is parameterized over the aforementioned set type \texttt{S.SS.t}. % FIXME: whyyy????
-
-\begin{figure}[H]
+Such that several variables can be assigned to the same node (more on this in section \ref{coalescing}).
+A helper function is defined, that looks up a variable's symbol in the symbol table over vertices and adds it if it doesn't exist.
+Otherwise it simply returns the existing vertex identifier:
 \begin{minted}[linenos,bgcolor=LightGray]{ocaml}
-let interf (params : Ll.uid list) (insns : Cfg.insn list) _ (out : S.SS.t array)
-    : G.V.t S.table * G.t =
+let vert g e t =
+  match S.ST.find_opt e t with
+  | Some v -> (v, t)
+  | None ->
+      let v = S.SS.add e S.SS.empty |> G.V.create in
+      G.add_vertex g v;
+      (v, S.ST.add e v t)
+\end{minted}
+Then the interference graph is derived in Figure \ref{fig:interf}. First, all parameters are connected, then the instructions are processed according to the book.
+
+\begin{figure}[H]
+\begin{minted}[linenos,bgcolor=LightGray]{ocaml}
+let interf (params : Ll.uid list) (insns : Cfg.insn list) _ (out : S.SS.t array) =
   let g = G.create () in
-  let vert e t = (* .. *) in
-  let vert2 t e = vert e t |> snd and vert3 e t = vert e t |> snd in
-  (* add params *)
+  let vert2 t e = vert g e t |> snd and vert3 e t = vert g e t |> snd in
+  let edge s1 s2 t =
+    let v1, t = vert g s1 t in
+    let v2, t = vert g s2 t in
+    if v1 <> v2 then G.add_edge g v1 v2;
+    t
+  in
   let t = List.fold_left vert2 S.ST.empty params in
-  (* connect all params *)
-  let edge v1 v2 t = (* .. *) in
   let param t p1 =
@@
   let t = List.fold_left param t params in
-  (* add all defs *)
   let setverts t s = S.SS.fold vert3 s t in
@@
 %* optimistic coloring
-\subsection{Coalescing}
+\subsection{Coalescing} \label{coalescing}
 

### 2024-01-13 article.tex
more detail

@@
 \end{minted}
-Then the interference graph is derived in Figure \ref{fig:interf}. First, all parameters are connected, then the instructions are processed according to the book.
+Then the interference graph is derived in Figure \ref{fig:interf}.
+First, all parameters are connected, then the instructions are processed as described in the Appel text \cite[222]{tiger}: phi nodes, as they are the only \texttt{mov} instructions, interference edges are added with all  \(\mathit{out}[n] \setminus \{b_1, ..., b_n\}\) (see \eqref{phi}) while all others are just \(\mathit{out}[n]\).
 
@@
   let t = List.map (def S.SS.empty) insns |> List.fold_left setverts t in
+  let ids = function Ll.Id id -> Some id | _ -> None in
   let defoutedges t (i, n) =
     let defs = def S.SS.empty n in
-    let outedge e1 t = S.SS.fold (edge e1) out.(i) t in
+    let out =
+      match n with
+      | Cfg.Insn (_, Ll.PhiNode (_, ops)) ->
+          List.map fst ops |> List.filter_map ids |> S.SS.of_list
+          |> S.SS.diff out.(i)
+      | _ -> out.(i)
+    in
+    let outedge e1 t = S.SS.fold (edge e1) out t in
     S.SS.fold outedge defs t
@@
 
-The \mintinline{llvm}{call} operation is by far the most complicated one. Mostly because of calling conventions,
-The \texttt{\%rax} register is zeroed before every call as its lowermost byte register \texttt{\%al} is used to indicate the number of variadic arguments \cite[25]{sysv}.
+The \mintinline{llvm}{call} operation is by far the most complicated one.
+Mostly because of calling conventions, caller saved registers listed in Table \ref{tab:trans} need to be saved/restored before/after all.
+This isn't necessary in all cases, only the cases where they are actually used (which ought to be a simple lookup in the assignments table), but is not accounted for in this implementation.
+Instead they are all saved/restored regardless of use this could also be further optimized by simply subtracting from \texttt{\%rsp} and moving them to their respective stack locations manually,  instead of using push/pop.
+Additionally, the \texttt{\%rax} register is zeroed before every call as its lowermost byte register \texttt{\%al} is used to indicate the number of variadic arguments \cite[25]{sysv}.
+This isn't actually accounted for, instead it is just set to zero regardless.
+Furthermore, the way in which arguments are passed is also unnecessarily convoluted, but stems from a corner case in which variables were overwritten by moving to the respective register of the \(n\) function argument.
+This can be remedied by ordering the \texttt{mov}s correctly, and even in the case where there are cycles and \texttt{mov}s cannot be made without overwriting, one of the scratch registers can be used instead.
+But as a workaround for this, the arguments are simply pushed from the source operands and popped to their respective argument registers (in reverse order).
 
@@
 As each of these halt or time out, their output is compared to what was output from executables generated by the 'correct' compiler of \texttt{clang}.
+Though testing the outputs of several stages much like the Tiger compiler from the compilers course, this wasn't realized in time.
+All of the test entries in \texttt{bin/testrunner.ml} were either working at one point and the result of a regression or it was a work in progress that wasn't realized in time either.
+Except the \~ 20\% of the LLVM test suite that were never working.
 
@@
 
-Each \lstinline!.ll! file present in the \lstinline!benches! directory is listed here in a table denoting the input paramater on the left column and type of allocator used.
+Some of the \texttt{.ll} file present in the \texttt{benches} directory are listed here in a table denoting the input paramater on the left column and type of allocator used.
+The data measured is extracted from the \texttt{perf} as mentioned above. This is done in \texttt{bin/bench.ml} which also contains the commands and parameters to execute each benchmark.
 

### 2024-01-13 article.tex
elaborate spilling heuristic

@@
 
-\subsection{Heuristics spilling}
+\subsection{Spilling heuristics}
+The spilling heuristic can vary, but the one mentioned in the paper is chosen from remaining length, as in  \textit{\textquoteleft furthest away from the current point\textquoteright} \cite[900]{linear}.
+The same decision was made for this implementation, as it does stand to reduce the amount of spills necessary.
+See The program of Figure \ref{fig:linear.ll} and its associated live ranges/intervals of Figure \ref{fig:linear.lra}.
+For a target of \(k=2\), upon reaching the start of interval \texttt{\%5},  one variable needs to be spilled.
+If either \texttt{\%3} or \texttt{\%4} are chosen, in the case of \texttt{\%3} being spilled, already upon reaching \texttt{\%7} another needs to be spilled.
+In case of \texttt{\%4}, two variables won't be live simultaneously until \texttt{\%8}.
+Nonetheless, that means two spills instead of spilling the longest interval of \texttt{\%5}.
 \usetikzlibrary{chains,arrows}

### 2024-01-13 article.tex
fix linear order

@@
 That said, it may generate more imprecise assignments and therefore also be less performant at runtime, so the utility of such an approach depends on the context in which it is used.
-While compilers like clang translate to machine code ahead-of-time (AOT), languages like C\# and Java will often rely on just-in-time (JIT) compilation to translate their respective VM instructions to native machine code,
-
+While compilers like clang translate to machine code ahead-of-time (AOT), languages like C\# and Java will often rely on just-in-time (JIT) compilation to translate their respective VM instructions to native machine code.
 Such VMs that depend on JIT will often defer compilation to runtime.
@@
 It does this by first building the intervals using the \(\mathit{in}[s]\) and \(\mathit{out}[s]\) resulting from prior dataflow analysis.
-This can be done in a single pass with some creative array indexing, but my approach relies on two passes: one for finding the start points (see Figure \ref{fig:livenessstarts.ml}), the other for finding the end points.
-
-\begin{figure}[H]
-  \centering
-   \begin{minipage}[b]{0.88\textwidth}
-     \centering
-     \begin{minted}[linenos,bgcolor=LightGray]{ocaml}
-let add e = function
-  | Some s -> Some (S.SS.add e s)
-  | None -> Some (S.SS.singleton e)
-
-let intervalstart insns (in_, out) =
-  let insn (ordstarts, starts, active) (i, _n) =
-    let out' = S.SS.diff (S.SS.union in_.(i) out.(i)) active in
-    ( S.SS.fold (fun e a -> IT.update i (add e) a) out' ordstarts,
-      S.SS.fold (fun e a -> S.ST.add e i a) out' starts,
-      S.SS.union active out' )
-  in
-  List.fold_left insn (IT.empty, S.ST.empty, S.SS.empty) insns
-     \end{minted}
-     \caption{Implementation of \texttt{intervalstart} in \texttt{lib/linear.ml}}\label{fig:livenessstarts.ml}
-   \end{minipage}
-\end{figure}
-
-\noindent Of note are the structures over which the folding is done: the \texttt{IT.t} is a \texttt{Map.S} module indexing earlier described symbol sets (i.e.
-\texttt{S.SS.t}).
-The reason for doing this is to iterate through intervals in increasing and decreasing order of start/length and end points respectively.
-The symbol table (\texttt{S.ST.t}) is to allow for reverse lookup, i.e.
-which program point variable \(s\) is defined, and lastly the symbol set (i.e.
-\texttt{S.SS.t}) which is just there to prevent multiple start points for the same variable.
-
-\begin{figure}[H]
-  \centering
-   \begin{minipage}[b]{0.95\textwidth}
-     \centering
-     \begin{minted}[linenos,bgcolor=LightGray]{ocaml}
-let intervalends insns (in_, out) =
-  let insn (i, _n) (ordends, ends, active) =
-    let in' = S.SS.diff (S.SS.union in_.(i) out.(i)) active in
-    ( S.SS.fold (fun e a -> IT.update i (add e) a) in' ordends,
-      S.SS.fold (fun e a -> S.ST.add e i a) in' ends,
-      S.SS.union active in' )
-  in
-  List.fold_right insn insns (IT.empty, S.ST.empty, S.SS.empty)
-     \end{minted}
-     \caption{Implementation of \texttt{intervalendsintervalstart} in \texttt{lib/linear.ml}}\label{fig:intervalends.ml}
-   \end{minipage}
-\end{figure}
-
-\noindent The opposite is quite similar (see Figure \ref{fig:intervalends.ml}), with the only real difference being the use of \texttt{List.fold\_right} instead of \texttt{List.fold\_left} to reduce the function body in reverse.
+By keeping a list of active variables, on reaching the beginning of a new interval whith \(lk\) variables already actuve, one of them need to be spilled such that the starting interval can be assigned a register.
+It could be chosen at random but that could be costly, instead some spilling heuristic is applied.
 
@@
 \end{figure}
+
+\subsection{Implementation}
+
+At first, intervals are derived from the flattened instruction sequence and associated \(\mathit{in}[n]\) and \(\mathit{out}[n]\) sets. 
+%spilling heuristic can vary, but the one mentioned in the paper is chosen from remaining length, as in  \textit{\textquoteleft furthest away from the current point\textquoteright} \cite[900]{linear}.
+The same decision was made for this implementation, as it does stand to reduce the amount of spills necessary.
+See The program of Figure \ref{fig:linear.ll} and its associated live ranges/intervals of Figure \ref{fig:linear.lra}.
+For a target of \(k=2\), upon reaching the start of interval \texttt{\%5},  one variable needs to be spilled.
+If either \texttt{\%3} or \texttt{\%4} are chosen, in the case of \texttt{\%3} being spilled, already upon reaching \texttt{\%7} another needs to be spilled.
+In case of \texttt{\%4}, two variables won't be live simultaneously until \texttt{\%8}.
+Nonetheless, that means two spills instead of spilling the longest interval of \texttt{\%5}.
+While this can be done in a single pass with some creative array indexing, but my approach relies on two passes: one for finding the start points (see Figure \ref{fig:livenessstarts.ml}), the other for finding the end points.
+
+\begin{figure}[H]
+  \centering
+   \begin{minipage}[b]{0.88\textwidth}
+     \centering
+     \begin{minted}[linenos,bgcolor=LightGray]{ocaml}
+let add e = function
+  | Some s -> Some (S.SS.add e s)
+  | None -> Some (S.SS.singleton e)
+let intervalstart insns (in_, out) =
+  let insn (ordstarts, starts, active) (i, _n) =
+    let out' = S.SS.diff (S.SS.union in_.(i) out.(i)) active in
+    ( S.SS.fold (fun e a -> IT.update i (add e) a) out' ordstarts,
+      S.SS.fold (fun e a -> S.ST.add e i a) out' starts,
+      S.SS.union active out' )
+  in
+  List.fold_left insn (IT.empty, S.ST.empty, S.SS.empty) insns
+     \end{minted}
+     \caption{Implementation of \texttt{intervalstart} in \texttt{lib/linear.ml}}\label{fig:livenessstarts.ml}
+   \end{minipage}
+\end{figure}
+
+\noindent Of note are the structures over which the folding is done: the \texttt{IT.t} is a \texttt{Map.S} module indexing earlier described symbol sets (i.e.
+\texttt{S.SS.t}).
+The reason for doing this is to iterate through intervals in increasing and decreasing order of start/length and end points respectively.
+The symbol table (\texttt{S.ST.t}) is to allow for reverse lookup, i.e.
+which program point variable \(s\) is defined, and lastly the symbol set (i.e.
+\texttt{S.SS.t}) which is just there to prevent multiple start points for the same variable.
+
+\begin{figure}[H]
+  \centering
+   \begin{minipage}[b]{0.95\textwidth}
+     \centering
+     \begin{minted}[linenos,bgcolor=LightGray]{ocaml}
+let intervalends insns (in_, out) =
+  let insn (i, _n) (ordends, ends, active) =
+    let in' = S.SS.diff (S.SS.union in_.(i) out.(i)) active in
+    ( S.SS.fold (fun e a -> IT.update i (add e) a) in' ordends,
+      S.SS.fold (fun e a -> S.ST.add e i a) in' ends,
+      S.SS.union active in' )
+  in
+  List.fold_right insn insns (IT.empty, S.ST.empty, S.SS.empty)
+     \end{minted}
+     \caption{Implementation of \texttt{intervalendsintervalstart} in \texttt{lib/linear.ml}}\label{fig:intervalends.ml}
+   \end{minipage}
+\end{figure}
+
+\noindent The opposite is quite similar (see Figure \ref{fig:intervalends.ml}), with the only real difference being the use of \texttt{List.fold\_right} instead of \texttt{List.fold\_left} to reduce the function body in reverse.
 
@@
 
-\noindent The array indexing is used in the inner functions of \texttt{init} and \texttt{end\_} to unset the last point at which an interval goes out of scope.
+\noindent The array indexing is used in the inner functions of \texttt{init} and \texttt{end\_} to unset the last point at which an interval goes out of scope and are defined in Figure \ref{fig:init-ocaml} and \ref{fig:end_-ocaml} respectively.
 

### 2024-01-13 article.tex
explain sha256

@@
 \subsubsection{\lstinline!benches/sha256.ll!}
+The sha256 benchmark is a massive program when compared to all others.
+It stems from an MIT licensed implementation  \cite{sha256}.
+The core function is \texttt{sha256\_compress} with no loops, but it does perform 64 SHA-2 rounds which are quite involved by design.
+As most binary operations depend on \~5 or so instructions instead of 1-2 as is possible for the majority, that alone ought to reduce performance immensely.
 \begin{table}[H]
 \makebox[\textwidth][c]{
-\begin{NiceTabular}{|c|cccccccc|}
-\hline
-n & clang & simple 12 & simple 2 & briggs 12 & briggs 2 & linear & greedy 12 & greedy 0 \\
-\hline
-100 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.9990234375,0.9619140625,0.75} 3.17253x& \cellcolor[rgb]{0.9990234375,0.9560546875,0.75} 3.35873x& \cellcolor[rgb]{0.9990234375,0.9619140625,0.75} 3.17637x& \cellcolor[rgb]{0.9990234375,0.9560546875,0.75} 3.37014x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 19.50758x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 19.44391x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 19.61921x\\
-1000 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.9990234375,0.9189453125,0.75} 4.54197x& \cellcolor[rgb]{0.9990234375,0.91015625,0.75} 4.83710x& \cellcolor[rgb]{0.9990234375,0.9189453125,0.75} 4.54593x& \cellcolor[rgb]{0.9990234375,0.91015625,0.75} 4.83359x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 31.35583x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 31.26454x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 31.56475x\\
-10000 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.9990234375,0.9111328125,0.75} 4.80821x& \cellcolor[rgb]{0.9990234375,0.9013671875,0.75} 5.11478x& \cellcolor[rgb]{0.9990234375,0.9111328125,0.75} 4.80648x& \cellcolor[rgb]{0.9990234375,0.9013671875,0.75} 5.12019x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 33.50659x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 33.41428x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 33.72138x\\
+\begin{NiceTabular}{|c|ccccccccc|}
+\hline
+arg(s) & clang & simple 12 & simple 2 & briggs 12 & briggs 2 & linear 12 & linear 2 & greedy 12 & greedy 0 \\
+\hline
+4 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 151231& \cellcolor[rgb]{0.810546875,0.9990234375,0.75} 1.24623x& \cellcolor[rgb]{0.8193359375,0.9990234375,0.75} 1.28026x& \cellcolor[rgb]{0.8115234375,0.9990234375,0.75} 1.24897x& \cellcolor[rgb]{0.8193359375,0.9990234375,0.75} 1.28066x& \cellcolor[rgb]{0.9619140625,0.9990234375,0.75} 1.85133x& \cellcolor[rgb]{0.96875,0.9990234375,0.75} 1.88049x& \cellcolor[rgb]{0.9990234375,0.966796875,0.75} 3.00814x& \cellcolor[rgb]{0.9990234375,0.966796875,0.75} 3.02968x\\
+8 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 159289& \cellcolor[rgb]{0.8515625,0.9990234375,0.75} 1.41012x& \cellcolor[rgb]{0.8623046875,0.9990234375,0.75} 1.45279x& \cellcolor[rgb]{0.8525390625,0.9990234375,0.75} 1.41295x& \cellcolor[rgb]{0.861328125,0.9990234375,0.75} 1.44789x& \cellcolor[rgb]{0.9990234375,0.9853515625,0.75} 2.42693x& \cellcolor[rgb]{0.9990234375,0.984375,0.75} 2.46595x& \cellcolor[rgb]{0.9990234375,0.923828125,0.75} 4.38812x& \cellcolor[rgb]{0.9990234375,0.923828125,0.75} 4.40305x\\
+16 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 173067& \cellcolor[rgb]{0.92578125,0.9990234375,0.75} 1.70848x& \cellcolor[rgb]{0.94140625,0.9990234375,0.75} 1.76986x& \cellcolor[rgb]{0.92578125,0.9990234375,0.75} 1.70898x& \cellcolor[rgb]{0.94140625,0.9990234375,0.75} 1.77111x& \cellcolor[rgb]{0.9990234375,0.953125,0.75} 3.45499x& \cellcolor[rgb]{0.9990234375,0.9521484375,0.75} 3.49756x& \cellcolor[rgb]{0.9990234375,0.8466796875,0.75} 6.88402x& \cellcolor[rgb]{0.9990234375,0.8466796875,0.75} 6.88614x\\
+32 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 202474& \cellcolor[rgb]{0.9990234375,0.994140625,0.75} 2.15587x& \cellcolor[rgb]{0.9990234375,0.990234375,0.75} 2.25675x& \cellcolor[rgb]{0.9990234375,0.994140625,0.75} 2.14418x& \cellcolor[rgb]{0.9990234375,0.990234375,0.75} 2.26214x& \cellcolor[rgb]{0.9990234375,0.904296875,0.75} 5.03392x& \cellcolor[rgb]{0.9990234375,0.9013671875,0.75} 5.12076x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 10.69747x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 10.70125x\\
+64 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 256682& \cellcolor[rgb]{0.9990234375,0.9736328125,0.75} 2.78746x& \cellcolor[rgb]{0.9990234375,0.96875,0.75} 2.95124x& \cellcolor[rgb]{0.9990234375,0.9736328125,0.75} 2.79279x& \cellcolor[rgb]{0.9990234375,0.96875,0.75} 2.95499x& \cellcolor[rgb]{0.9990234375,0.8349609375,0.75} 7.25491x& \cellcolor[rgb]{0.9990234375,0.830078125,0.75} 7.39688x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 16.05447x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 16.05908x\\
+128 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 369901& \cellcolor[rgb]{0.9990234375,0.953125,0.75} 3.45784x& \cellcolor[rgb]{0.9990234375,0.9462890625,0.75} 3.67742x& \cellcolor[rgb]{0.9990234375,0.953125,0.75} 3.45438x& \cellcolor[rgb]{0.9990234375,0.9462890625,0.75} 3.67655x& \cellcolor[rgb]{0.9990234375,0.7626953125,0.75} 9.56947x& \cellcolor[rgb]{0.9990234375,0.7568359375,0.75} 9.78032x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 21.68320x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 21.71900x\\
+256 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 598859& \cellcolor[rgb]{0.9990234375,0.935546875,0.75} 4.01075x& \cellcolor[rgb]{0.9990234375,0.927734375,0.75} 4.27823x& \cellcolor[rgb]{0.9990234375,0.935546875,0.75} 4.01290x& \cellcolor[rgb]{0.9990234375,0.927734375,0.75} 4.27501x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 11.53833x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 11.80069x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 26.47159x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 26.47409x\\
+512 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1054540& \cellcolor[rgb]{0.9990234375,0.923828125,0.75} 4.40890x& \cellcolor[rgb]{0.9990234375,0.9140625,0.75} 4.70502x& \cellcolor[rgb]{0.9990234375,0.923828125,0.75} 4.40897x& \cellcolor[rgb]{0.9990234375,0.9140625,0.75} 4.70366x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 12.93893x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 13.22190x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 29.86374x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 29.85531x\\
+1024 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1989472& \cellcolor[rgb]{0.9990234375,0.91796875,0.75} 4.59073x& \cellcolor[rgb]{0.9990234375,0.908203125,0.75} 4.90795x& \cellcolor[rgb]{0.9990234375,0.91796875,0.75} 4.59569x& \cellcolor[rgb]{0.9990234375,0.908203125,0.75} 4.90336x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 13.62511x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 13.93164x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 31.53197x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 31.56367x\\
+2048 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 3831949& \cellcolor[rgb]{0.9990234375,0.9130859375,0.75} 4.73191x& \cellcolor[rgb]{0.9990234375,0.9033203125,0.75} 5.05794x& \cellcolor[rgb]{0.9990234375,0.9130859375,0.75} 4.73420x& \cellcolor[rgb]{0.9990234375,0.9033203125,0.75} 5.05587x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 14.10539x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 14.41266x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 32.69551x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 32.72633x\\
+4096 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 7524589& \cellcolor[rgb]{0.9990234375,0.9111328125,0.75} 4.80394x& \cellcolor[rgb]{0.9990234375,0.9013671875,0.75} 5.13065x& \cellcolor[rgb]{0.9990234375,0.9111328125,0.75} 4.80706x& \cellcolor[rgb]{0.9990234375,0.9013671875,0.75} 5.13594x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 14.33546x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 14.66333x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 33.25560x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 33.30167x\\
+8192 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 14901775& \cellcolor[rgb]{0.9990234375,0.91015625,0.75} 4.84418x& \cellcolor[rgb]{0.9990234375,0.8994140625,0.75} 5.18112x& \cellcolor[rgb]{0.9990234375,0.91015625,0.75} 4.84423x& \cellcolor[rgb]{0.9990234375,0.8994140625,0.75} 5.18259x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 14.46452x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 14.78932x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 33.58380x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 33.61478x\\
 \hline
 \end{NiceTabular}}
-\caption{\label{tab:sha256}Benchmark of \texttt{benches/sha256.ll} output by \mintinline{bash}{dune exec bench -- -f fib -n 1000}}
+\caption{\label{tab:sha256}Benchmark of \texttt{benches/sha256.ll} output by \mintinline{bash}{dune exec bench -- -f sha256 -n 256}}
 \end{table}
@@
 
+Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
+Tellus at urna condimentum mattis.
+Enim facilisis gravida neque convallis.
+Enim facilisis gravida neque convallis a cras semper auctor.
+Vitae proin sagittis nisl rhoncus mattis rhoncus urna neque.
+Justo nec ultrices dui sapien eget mi proin.
+Id aliquet risus feugiat in ante.
+A cras semper auctor neque vitae tempus quam pellentesque nec.
+Nulla porttitor massa id neque aliquam vestibulum morbi.
+Facilisi cras fermentum odio eu.
+Odio pellentesque diam volutpat commodo sed egestas egestas fringilla.
+Amet consectetur adipiscing elit duis tristique sollicitudin nibh sit amet.
+Adipiscing at in tellus integer feugiat.
+A pellentesque sit amet porttitor.
+Sed elementum tempus egestas sed sed risus pretium quam.
+Quam viverra orci sagittis eu volutpat odio facilisis mauris sit.
+Arcu dictum varius duis at.
+Sociis natoque penatibus et magnis dis parturient montes.
+Varius sit amet mattis vulputate enim nulla aliquet.
+Ac ut consequat semper viverra.
+Scelerisque eleifend donec pretium vulputate sapien nec.
+Orci ac auctor augue mauris augue.
+Pulvinar mattis nunc sed blandit.
+Semper auctor neque vitae tempus quam pellentesque nec nam.
+Risus pretium quam vulputate dignissim.
+Dui faucibus in ornare quam viverra orci sagittis eu volutpat.
+Ac tortor vitae purus faucibus.
+Eu mi bibendum neque egestas congue quisque egestas diam in.
+Curabitur vitae nunc sed velit dignissim sodales ut eu sem.
+Arcu odio ut sem nulla.
+
 \section{Conclusion}
+
+Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
+Tellus at urna condimentum mattis.
+Enim facilisis gravida neque convallis.
+Enim facilisis gravida neque convallis a cras semper auctor.
+Vitae proin sagittis nisl rhoncus mattis rhoncus urna neque.
+Justo nec ultrices dui sapien eget mi proin.
+Id aliquet risus feugiat in ante.
+A cras semper auctor neque vitae tempus quam pellentesque nec.
+Nulla porttitor massa id neque aliquam vestibulum morbi.
+Facilisi cras fermentum odio eu.
+Odio pellentesque diam volutpat commodo sed egestas egestas fringilla.
+Amet consectetur adipiscing elit duis tristique sollicitudin nibh sit amet.
+Adipiscing at in tellus integer feugiat.
+A pellentesque sit amet porttitor.
+Sed elementum tempus egestas sed sed risus pretium quam.
+Quam viverra orci sagittis eu volutpat odio facilisis mauris sit.
+Arcu dictum varius duis at.
+Sociis natoque penatibus et magnis dis parturient montes.
+Varius sit amet mattis vulputate enim nulla aliquet.
+Ac ut consequat semper viverra.
+Scelerisque eleifend donec pretium vulputate sapien nec.
+Orci ac auctor augue mauris augue.
+Pulvinar mattis nunc sed blandit.
+Semper auctor neque vitae tempus quam pellentesque nec nam.
+Risus pretium quam vulputate dignissim.
+Dui faucibus in ornare quam viverra orci sagittis eu volutpat.
+Ac tortor vitae purus faucibus.
+Eu mi bibendum neque egestas congue quisque egestas diam in.
+Curabitur vitae nunc sed velit dignissim sodales ut eu sem.
+Arcu odio ut sem nulla.
 

### 2024-01-14 article.tex
more

@@
 Arcu odio ut sem nulla.
+
+\section*{Acknowledgments}
+I would like to express my gratitude and thanks towards all who have facilitated me in making this project a part of reality.
+First of all, I can hardly begin to express my gratitude towards my advisors Aslan and Amin for the weekly meetings where most if not all of my confusion was settled.
+%Firstly, we would like to thank Bas Spitters, our supervisor, for his guidance and expertise on the subject, as
+%well as giving us the idea for the project. Without him this project would not have been possible. Secondly,
+%we thank Franziskus Kiefer for his helpfulness in answering questions about various Hacspec-related quirks,
+%issues, as well as being swift in these responses despite his large workload. Thirdly, we want to thank Las
+%Safin, who helped proofreading and thus make this report more bulletproof. Finally, we thank everyone who
+%helped answer questions over the various stages of the project. Your assistance is greatly appreciated
 \newpage
@@
 Then the interference graph is derived in Figure \ref{fig:interf}.
-First, all parameters are connected, then the instructions are processed as described in the Appel text \cite[222]{tiger}: phi nodes, as they are the only \texttt{mov} instructions, interference edges are added with all  \(\mathit{out}[n] \setminus \{b_1, ..., b_n\}\) (see \eqref{phi}) while all others are just \(\mathit{out}[n]\).
+First, all parameters are connected, then the instructions are processed as described in the Appel text \cite[222]{tiger}.
+As phi nodes are the only \texttt{mov} instructions, their interference edges are   \(\mathit{out}[n] \setminus \{b_1, ..., b_n\}\) (see \eqref{phi}) as opposed to  non-move instructions that are just \(\mathit{out}[n]\).
 
@@
 \subsubsection{\lstinline!benches/fannkuch-redux.ll!}
+
+Fannkuch-redux involves a permutation algorithm that has a non-trivial algorithmic complexity.
+It requires both recursion and iterative processes, making it a good test for a language's ability to handle complex computations efficiently.
+The fannkuch algorithm is
+
 \begin{table}[H]
@@
 10 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1352747472& \cellcolor[rgb]{0.8857421875,0.9990234375,0.75} 1.54700x& \cellcolor[rgb]{0.935546875,0.9990234375,0.75} 1.74875x& \cellcolor[rgb]{0.89453125,0.9990234375,0.75} 1.58381x& \cellcolor[rgb]{0.9169921875,0.9990234375,0.75} 1.67440x& \cellcolor[rgb]{0.9990234375,0.818359375,0.75} 7.77269x& \cellcolor[rgb]{0.9990234375,0.8193359375,0.75} 7.74449x& \cellcolor[rgb]{0.9990234375,0.8173828125,0.75} 7.82170x\\
+\hline
+\end{NiceTabular}}
+\caption{\label{tab:fannkuch}Benchmark of \texttt{benches/sieven.ll} output by \mintinline{bash}{dune exec bench -- -f fib -n 1000}}
+\end{table}
+
+\begin{table}[H]
+\makebox[\textwidth][c]{
+\begin{NiceTabular}{|c|ccccccccc|}
+\hline
+arg(s) & clang & simple 12 & simple 2 & briggs 12 & briggs 2 & linear 12 & linear 2 & greedy 12 & greedy 0 \\
+\hline
+4 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 171300& \cellcolor[rgb]{0.751953125,0.9990234375,0.75} 1.01148x& \cellcolor[rgb]{0.7568359375,0.9990234375,0.75} 1.02792x& \cellcolor[rgb]{0.755859375,0.9990234375,0.75} 1.02449x& \cellcolor[rgb]{0.755859375,0.9990234375,0.75} 1.02731x& \cellcolor[rgb]{0.7880859375,0.9990234375,0.75} 1.15660x& \cellcolor[rgb]{0.7841796875,0.9990234375,0.75} 1.13751x& \cellcolor[rgb]{0.8173828125,0.9990234375,0.75} 1.27163x& \cellcolor[rgb]{0.8037109375,0.9990234375,0.75} 1.21946x\\
+5 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 189030& \cellcolor[rgb]{0.76171875,0.9990234375,0.75} 1.04909x& \cellcolor[rgb]{0.771484375,0.9990234375,0.75} 1.08923x& \cellcolor[rgb]{0.7626953125,0.9990234375,0.75} 1.05266x& \cellcolor[rgb]{0.7705078125,0.9990234375,0.75} 1.08530x& \cellcolor[rgb]{0.8486328125,0.9990234375,0.75} 1.39953x& \cellcolor[rgb]{0.85546875,0.9990234375,0.75} 1.42370x& \cellcolor[rgb]{0.9033203125,0.9990234375,0.75} 1.61577x& \cellcolor[rgb]{0.908203125,0.9990234375,0.75} 1.63543x\\
+6 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 305772& \cellcolor[rgb]{0.8115234375,0.9990234375,0.75} 1.24795x& \cellcolor[rgb]{0.830078125,0.9990234375,0.75} 1.32393x& \cellcolor[rgb]{0.8125,0.9990234375,0.75} 1.25272x& \cellcolor[rgb]{0.83203125,0.9990234375,0.75} 1.33146x& \cellcolor[rgb]{0.9990234375,0.9873046875,0.75} 2.34850x& \cellcolor[rgb]{0.9990234375,0.9873046875,0.75} 2.37422x& \cellcolor[rgb]{0.9990234375,0.9609375,0.75} 3.21562x& \cellcolor[rgb]{0.9990234375,0.9599609375,0.75} 3.23381x\\
+7 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1331652& \cellcolor[rgb]{0.86328125,0.9990234375,0.75} 1.45633x& \cellcolor[rgb]{0.90625,0.9990234375,0.75} 1.62799x& \cellcolor[rgb]{0.8740234375,0.9990234375,0.75} 1.50098x& \cellcolor[rgb]{0.9091796875,0.9990234375,0.75} 1.64250x& \cellcolor[rgb]{0.9990234375,0.9580078125,0.75} 3.31242x& \cellcolor[rgb]{0.9990234375,0.955078125,0.75} 3.39020x& \cellcolor[rgb]{0.9990234375,0.904296875,0.75} 5.03373x& \cellcolor[rgb]{0.9990234375,0.9033203125,0.75} 5.06248x\\
+8 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 11197605& \cellcolor[rgb]{0.904296875,0.9990234375,0.75} 1.62006x& \cellcolor[rgb]{0.9375,0.9990234375,0.75} 1.75454x& \cellcolor[rgb]{0.89453125,0.9990234375,0.75} 1.58268x& \cellcolor[rgb]{0.939453125,0.9990234375,0.75} 1.76405x& \cellcolor[rgb]{0.9990234375,0.9482421875,0.75} 3.61197x& \cellcolor[rgb]{0.9990234375,0.9482421875,0.75} 3.62996x& \cellcolor[rgb]{0.9990234375,0.8876953125,0.75} 5.56708x& \cellcolor[rgb]{0.9990234375,0.888671875,0.75} 5.53880x\\
+9 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 112923594& \cellcolor[rgb]{0.9140625,0.9990234375,0.75} 1.66209x& \cellcolor[rgb]{0.9541015625,0.9990234375,0.75} 1.82157x& \cellcolor[rgb]{0.904296875,0.9990234375,0.75} 1.62219x& \cellcolor[rgb]{0.9453125,0.9990234375,0.75} 1.78657x& \cellcolor[rgb]{0.9990234375,0.947265625,0.75} 3.65915x& \cellcolor[rgb]{0.9990234375,0.94140625,0.75} 3.82090x& \cellcolor[rgb]{0.9990234375,0.880859375,0.75} 5.79421x& \cellcolor[rgb]{0.9990234375,0.8818359375,0.75} 5.75327x\\
+10 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1348654063& \cellcolor[rgb]{0.884765625,0.9990234375,0.75} 1.54471x& \cellcolor[rgb]{0.9306640625,0.9990234375,0.75} 1.72799x& \cellcolor[rgb]{0.896484375,0.9990234375,0.75} 1.59112x& \cellcolor[rgb]{0.943359375,0.9990234375,0.75} 1.77928x& \cellcolor[rgb]{0.9990234375,0.9501953125,0.75} 3.55946x& \cellcolor[rgb]{0.9990234375,0.9462890625,0.75} 3.68201x& \cellcolor[rgb]{0.9990234375,0.8828125,0.75} 5.71345x& \cellcolor[rgb]{0.9990234375,0.8837890625,0.75} 5.67565x\\
 \hline
@@
 
-Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
+The intended goal of this project has been to implement both the graph coloring and linear scan algorithms and evaluate them both in terms of benefits and drawbacks when used to compile LLVM-{}- IR to x86 assembly.
+
+%Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
 Tellus at urna condimentum mattis.

### 2024-01-14 article.tex
Fix naming

@@
 \title{\myfont Implementation and evaluation of register allocation for translating LLVM-{}- to x86-64}
-\author{William Welle Tange, 201809080}
+\author{William Welle Tange -- 201809080\\
+  Advisors: Aslan Askarov \& Amin Timany
+}
 
@@
 \section*{Acknowledgments}
-I would like to express my gratitude and thanks towards all who have facilitated me in making this project a part of reality.
-First of all, I can hardly begin to express my gratitude towards my advisors Aslan and Amin for the weekly meetings where most if not all of my confusion was settled.
+I would like to express my sincere gratitude and thanks to all who have supported me in making this project a part of reality.
+First of all, I can hardly begin to express my gratitude towards my advisors Aslan and Amin for the weekly meetings where most if not all concerns were settled.
 %Firstly, we would like to thank Bas Spitters, our supervisor, for his guidance and expertise on the subject, as
@@
 This is due to \texttt{\%7} being moved to \texttt{\%6} prior to potentially branching back to \texttt{label \%5}.
-If phi moves are conducted unconditionally of branch condition \texttt{\%9}, the value of \texttt{\%6} prior to branching to \texttt{label \%10} is overwritten on the off-chance of returning to \texttt{label \%5}, causing the value assigned to \texttt{\%11} and returned on line 10 to always be equal to \texttt{\%7} instead of the intended \texttt{\%7} instead of the intended  \texttt{\%6}.
+If phi moves are conducted unconditionally of branch condition \texttt{\%9}, the value of \texttt{\%6} prior to branching to \texttt{label \%10} is overwritten on the off-chance of returning to \texttt{label \%5}, causing the value assigned to \texttt{\%11} and returned on line 10 to always be equal to \texttt{\%7}   instead of the intended  \texttt{\%6}.
 

### 2024-01-14 article.tex
Rework linescan

@@
   \hline
-  i & use & def & in & out & in & out & in & out & in & out & in & out & in & out & in & out\\[-0.5ex]
+  i & use & def & in & out & in & out & in & out & in & out & in & out & in & out & in & out\\
   \hline
@@
 \label{tab:countll-rev-steps}
-\end{table}
-
-\begin{table}[H]
+\vspace{0.75em}
 \makebox[\textwidth][c]{
@@
   \hline
-  i & use & def & in & out & in & out & in & out & in & out & in & out & in & out & in & out & in & out & in & out & in & out & in & out\\[-0.5ex]
+  i & use & def & in & out & in & out & in & out & in & out & in & out & in & out & in & out & in & out & in & out & in & out & in & out\\
   \hline
@@
                   None ))
+  \end{minted}
+\end{figure}
+\begin{figure}[H]
+  \centering
+  \begin{minted}[linenos,bgcolor=LightGray,firstnumber=20]{ocaml}
           else
@@
 It is a simple and efficient approach that performs a linear pass over IR to produce a register assignment significantly faster than the steps involved with graph coloring.
-That said, it may generate more imprecise assignments and therefore also be less performant at runtime, so the utility of such an approach depends on the context in which it is used.
-While compilers like clang translate to machine code ahead-of-time (AOT), languages like C\# and Java will often rely on just-in-time (JIT) compilation to translate their respective VM instructions to native machine code.
-Such VMs that depend on JIT will often defer compilation to runtime.
+That said, it may generate more imprecise assignments and therefore also be less performant at runtime, so the utility depends on the context in which it is used.
+
+While compilers like clang generally has an end-goal of machine code, languages like C\# and Java 
+will target a VM instead.
+This is done for a myriad of reasons, portability and security among them, but usually at the cost of runtime performance.
+One of the reasons for this is the fact that VM bytecode needs to be processed beforehands.
+While bytecode is generated ahead-of-time, running it on a target machine by interpreting such instructions in real time is inefficient.
+Therefore, bytecode is generally translated to native machine code immediately prior to being executed, a process known as \textit{just-in-time} (JIT) compilation.
+
+A significant downside of JIT is the fact it needs to be compiled in full before execution can begin, hence processes such as register allocation are deferred to runtime.
+Therefore, generally a faster class of allocators are used, linear scan being one of them, as linear scan is a reasonably performant alternative to graph coloring, that may be used to achieve reasonably efficient register assignments at runtime.
+%rely on just-in-time (JIT) compilation to translate their respective VM instructions to native machine code. Such VMs that depend on JIT will often
+%defer compilation to runtime.
 This is done for a myriad of reasons, portability and security among them, but usually at the cost of runtime performance.
@@
 It does this by first building the intervals using the \(\mathit{in}[s]\) and \(\mathit{out}[s]\) resulting from prior dataflow analysis.
-By keeping a list of active variables, on reaching the beginning of a new interval whith \(lk\) variables already actuve, one of them need to be spilled such that the starting interval can be assigned a register.
+By keeping a list of active variables, on reaching the beginning of a new interval whith \(k\) variables already active, one of them need to be spilled such that the starting interval can be assigned a register.
 It could be chosen at random but that could be costly, instead some spilling heuristic is applied.

### 2024-01-14 article.tex
Finish explaining linear scan

@@
 Nonetheless, that means two spills instead of spilling the longest interval of \texttt{\%5}.
-While this can be done in a single pass with some creative array indexing, but my approach relies on two passes: one for finding the start points (see Figure \ref{fig:livenessstarts.ml}), the other for finding the end points.
+While this can be done in a single pass with some creative array indexing, my approach relies on two passes: one for finding the start points (see Figure \ref{fig:livenessstarts.ml}), the other for finding the end points.
 
@@
 
-\noindent The opposite is quite similar (see Figure \ref{fig:intervalends.ml}), with the only real difference being the use of \texttt{List.fold\_right} instead of \texttt{List.fold\_left} to reduce the function body in reverse.
+\noindent The opposite, i.e.
+finding the interval ends, is quite similar (see Figure \ref{fig:intervalends.ml}), with the only real difference being the use of \texttt{List.fold\_right} instead of \texttt{List.fold\_left} to reduce the function body in reverse.
 
@@
      \begin{minted}[linenos,bgcolor=LightGray]{ocaml}
-let intervals insns (in_, out) =
-  let insn (starts, t, active) (i, n) =
-    let init s (added, t, active) = (* .. *) in
-    let end_ s (added, t, active) = (* .. *) in
-    let a, t, c =
-      init in_.(i) (S.SS.empty, t, active) |> init out.(i) |> end_ out.(i)
+let rec linearscan insns spills ?(v = 0)
+    (avail, active, assign, incstart, incend) =
+  match IT.min_binding_opt incstart with
+  | Some (i, ss) ->
+      let e = S.SS.choose ss in
+      let ss' = S.SS.remove e ss in
+      let incstart' =
+        if S.SS.is_empty ss' then IT.remove i incstart
+        else IT.add i ss' incstart
       in
-    ((i, n, a) :: starts, t, c)
+      let spills, (avail', active', assign', incstart', incend') =
+        match Regs.choose_opt avail with
+        | Some reg ->
+            ( spills,
+              expire i
+                ( Regs.remove reg avail,
+                  S.ST.add e reg active,
+                  S.ST.add e (Reg reg) assign,
+                  incstart',
+                  incend ) )
+        | None ->
+            let _j, spill =
+              IT.filter
+                (fun _k v -> S.SS.exists (fun e -> S.ST.mem e active) v)
+                incend
+              |> IT.max_binding
             in
-  let starts, _, active =
-    List.fold_left insn
-      ([], S.ST.empty, Array.init (List.length insns) (Fun.const S.SS.empty))
-      insns
+            let spill =
+              S.SS.filter (fun e -> S.ST.mem e active) spill |> S.SS.choose
             in
-  List.rev starts |> List.map (fun (i, n, a) -> (i, n, a, active.(i)))
-     \end{minted}
-     \caption{Implementation of \texttt{intervals} in \texttt{linear.ml}}\label{fig:intervals-ocaml}
-\end{figure}
-
-\noindent The array indexing is used in the inner functions of \texttt{init} and \texttt{end\_} to unset the last point at which an interval goes out of scope and are defined in Figure \ref{fig:init-ocaml} and \ref{fig:end_-ocaml} respectively.
+            let reg = S.ST.find spill active in
+            let incend' =
+              IT.filter_map
+                (fun _ ss ->
+                  let ss' = S.SS.remove spill ss in
+                  if S.SS.is_empty ss' then None else Some ss')
+                incend
+            in
+     \end{minted}
+\end{figure}
 
@@
      \begin{minted}[linenos,bgcolor=LightGray]{ocaml}
-let init s (added, t, active) =
-  ( S.SS.filter (fun s -> not (S.ST.mem s t)) s |> S.SS.union added,
-    S.SS.fold
-      (fun s t ->
-        S.ST.update s
-          (function
-            | Some j ->
-                active.(i) <- S.SS.add s active.(i);
-                active.(j) <- S.SS.remove s active.(j);
-                Some i
-            | None ->
-                active.(i) <- S.SS.add s active.(i);
-                Some i)
-          t)
-      s t,
-    active )
-     \end{minted}
-     \caption{Implementation of \texttt{init} in \texttt{linear.ml}}\label{fig:init-ocaml}
-\end{figure}
+            ( spills + 1,
+              expire i
+                ( avail,
+                  S.ST.remove spill active |> S.ST.add e reg,
+                  S.ST.remove spill assign
+                  |> S.ST.add spill (stack spills)
+                  |> S.ST.add e (Reg reg),
+                  incstart',
+                  incend' ) )
+      in
+      linearscan insns spills (avail', active', assign', incstart', incend')
+  | None -> expire (List.length insns) (avail, active, assign, incstart, incend)
+     \end{minted}
+     \caption{Implementation of \texttt{linearscan} in \texttt{linear.ml}}\label{fig:linearscan-ocaml}
+\end{figure}
+\noindent The implementation of the actual linearscan iteration in Figure \ref{fig:linearscan-ocaml} is quite tedious, but the core idea is the outerermost \texttt{match} block.
+The smallest interval is selected using \texttt{min\_binding\_opt}.
+This is done to visit each interval in increasing order of start point.
+If there are none left, the linearscan iteration is essentially over, but calls \texttt{expire} regardless to clean things up.
+The case of intervals remaining is more interesting, as one of the variables starting at \(i\) needs to be assigned.
+First a variable is chosen, the order of this is not important.
+Despite the minimal example used in Figure \ref{fig:linear.ll}, there are cases where several intervals begin at the same point (e.g.
+loops), hence a symbol set is used.
+Lines 7-10 are just concerned with removing the chosen variable from the table of increasing start points.
+Then, an available register is chosen.
+If one is available (line 13), it is 1) removed from the set of available registers, 2) added to the symbol table of active variables, and 3) assigned the same register in the returned assignments table.
+If no registers are available on the other hand (line 21), a variable needs to be spilled.
+It does this by first filtering through and only visiting the end points that are currently active (i.e.
+members of \texttt{active}).
+Then the one of intervals that are currently active with the greatest end points is chosen by first filtering end points for which at least of the variables is active (lines 23-26), and then selecting one of those immediately after (line 29).
+The currently assigned register of that variable is then looked up (line 31) and the end point is removed (lines 33-37).
+
+In the case of spilling, the idea is to 1) remove the previously assigned variable with the greatest remaining length from the active set, 2) assign the previously assigned variable to the nextmost available stack slot, and 3) assign the currently visited start interval to the register just unassigned from the spilled variable.
+
+In either of those cases, the \texttt{expire} function is called right after.
+This is to ensure there are no variables left in the table of increading end points, and that their previously assigned registers are put back in the available set of registers as can be seen in Figure \ref{fig:expire-ocaml}.
 
@@
      \begin{minted}[linenos,bgcolor=LightGray]{ocaml}
-let end_ s (added, t, active) =
-  ( S.SS.filter (fun s -> not (S.ST.mem s t)) s |> S.SS.union added,
-    S.SS.fold
-      (fun s t ->
-        S.ST.update s
-          (function
-            | Some j ->
-                active.(i) <- S.SS.add s active.(i);
-                active.(j) <- S.SS.remove s active.(j);
-                Some i
-            | None -> Some i)
-          t)
-      s t,
-    active )
-     \end{minted}
-     \caption{Implementation of \texttt{end\_} in \texttt{linear.ml}}\label{fig:end_-ocaml}
-\end{figure}
-
+let rec expire i (avail, active, assign, incstart, incend) =
+  match IT.min_binding_opt incend with
+  | Some (j, end_) ->
+      if j >= i then (avail, active, assign, incstart, incend)
+      else
+        let s = S.SS.choose end_ in
+        let reg = S.ST.find s active in
+        let active' = S.ST.remove s active in
+        let avail' = Regs.add reg avail in
+        let incend' = IT.remove j incend in
+        expire i (avail', active', assign, incstart, incend')
+  | None -> (avail, active, assign, incstart, incend)
+     \end{minted}
+     \caption{Implementation of \texttt{expire} in \texttt{lib/linear.ml}}\label{fig:expire-ocaml}
+\end{figure}
+
+\noindent Putting it all together, is the \texttt{alloc} function in \texttt{lib/linear.ml} that builds the set of available registers from the \(k\) passed, builds the increasing interval starts and ends tables, and finally, begins the iteratice process of linear scan.
 \begin{figure}[H]
@@
      \begin{minted}[linenos,bgcolor=LightGray]{ocaml}
-let linearscan k insns =
-  let avail = Regs.of_list (List.init k reg_of_int) in
-  let insn (avail, active, assign) (_, _n, s, _e) =
-    (*Printf.printf "%s\n" (Cfg.string_of_insn n);*)
-    let start s (avail, active, assign) =
-      match Regs.choose_opt avail with
-      | Some reg -> (Regs.remove reg avail, S.ST.add s reg active, assign)
-      | None ->
-          (* FIXME: order by remaining or total length *)
-          let z, reg = S.ST.choose active in
-          ( avail,
-            S.ST.remove z active |> S.ST.add s reg,
-            S.ST.add z
-              (Ind3 (Lit (Int64.of_int ((S.ST.cardinal assign * -8) - 8)), Rbp))
-              assign )
-    in
-    S.SS.fold start s (avail, active, assign)
-  in
-  let _, active, assign =
-    List.fold_left insn (avail, S.ST.empty, S.ST.empty) insns
-  in
-  S.ST.fold (fun k v acc -> S.ST.add k (Reg v) acc) active assign
-     \end{minted}
-     \caption{Implementation of \texttt{linearscan} in \texttt{linear.ml}}\label{fig:linearscan-ocaml}
+let alloc k insns (in_, out) =
+  let avail = List.init k reg_of_int |> Regs.of_list in
+  let insns = List.mapi (fun i n -> (i, n)) insns in
+  let incstart, _, _ = intervalstart insns (in_, out) in
+  let incend, _, _ = intervalends insns (in_, out) in
+  let _, _, asn, _, _ =
+    linearscan insns 0 (avail, S.ST.empty, S.ST.empty, incstart, incend)
+  in asn
+     \end{minted}
+     \caption{Implementation of \texttt{alloc} in \texttt{lib/linear.ml}}\label{fig:intervals-ocaml}
 \end{figure}

### 2024-01-14 article.tex
move factor

@@
 
-The instruction set used in this paper will be a union of the sets used in the 2022 and 2023 compilers courses in order to work as a drop-in replacement of LLVM for either of the two respective source languages:  Tiger and Dolphin.
+The instruction set used in this paper will be a combination of those used in the 2022 and 2023 compilers courses in order to work as a drop-in replacement of LLVM for either of the two respective source languages:  Tiger and Dolphin.
 This instruction set is a subset of the one used in practice, as, for instance, neither of the languages implemented support interrupts, floating point operations etc., and instead only strive to cover the basics of compilers.
@@
 
-\noindent The measures in Table \ref{tab:dumb-fib}  are significantly equivalent to one another, with all allocators (except greedy 0) seeming to converge towards being 37\% slower than Clang.
+\noindent The measures in Table \ref{tab:dumb-fib}  are remarkably similar to one another, with all allocators (except greedy 0) seeming to converge towards being 37\% slower than Clang.
 This isn't particularly surprising, as the x86 translation is more concerned with preserving callee saved registers than necessary in pushing each and every one despite not being in use.
@@
 
+\subsubsection{\lstinline!benches/factori32.ll!}
+
+Integer factorization is another interesting algorithm. Generally considered to be NP, here are the results of factorizing known primes of inreasing bit length. For \(n=1024\).
+\begin{table}[H]
+\makebox[\textwidth][c]{
+\begin{NiceTabular}{|c|ccccccccc|}
+\hline
+arg(s) & clang & simple 12 & simple 2 & briggs 12 & briggs 2 & linear 12 & linear 2 & greedy 12 & greedy 0 \\
+\hline
+16777213 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 167130& \cellcolor[rgb]{0.759765625,0.9990234375,0.75} 1.04160x& \cellcolor[rgb]{0.7509765625,0.9990234375,0.75} 1.00651x& \cellcolor[rgb]{0.7626953125,0.9990234375,0.75} 1.05411x& \cellcolor[rgb]{0.7626953125,0.9990234375,0.75} 1.05372x& \cellcolor[rgb]{0.7607421875,0.9990234375,0.75} 1.04686x& \cellcolor[rgb]{0.958984375,0.9990234375,0.75} 1.84213x& \cellcolor[rgb]{0.9990234375,0.98828125,0.75} 2.33215x& \cellcolor[rgb]{0.9990234375,0.9873046875,0.75} 2.34511x\\
+33554393 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 175409& \cellcolor[rgb]{0.7646484375,0.9990234375,0.75} 1.06219x& \cellcolor[rgb]{0.7646484375,0.9990234375,0.75} 1.06070x& \cellcolor[rgb]{0.765625,0.9990234375,0.75} 1.06415x& \cellcolor[rgb]{0.7646484375,0.9990234375,0.75} 1.06212x& \cellcolor[rgb]{0.7646484375,0.9990234375,0.75} 1.06039x& \cellcolor[rgb]{0.9990234375,0.9951171875,0.75} 2.12199x& \cellcolor[rgb]{0.9990234375,0.974609375,0.75} 2.78336x& \cellcolor[rgb]{0.9990234375,0.9736328125,0.75} 2.79956x\\
+67108859 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 185779& \cellcolor[rgb]{0.7685546875,0.9990234375,0.75} 1.07588x& \cellcolor[rgb]{0.76953125,0.9990234375,0.75} 1.07923x& \cellcolor[rgb]{0.7705078125,0.9990234375,0.75} 1.08373x& \cellcolor[rgb]{0.76953125,0.9990234375,0.75} 1.08043x& \cellcolor[rgb]{0.76953125,0.9990234375,0.75} 1.07895x& \cellcolor[rgb]{0.9990234375,0.9833984375,0.75} 2.49571x& \cellcolor[rgb]{0.9990234375,0.955078125,0.75} 3.38089x& \cellcolor[rgb]{0.9990234375,0.955078125,0.75} 3.40121x\\
+134217689 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 197302& \cellcolor[rgb]{0.7802734375,0.9990234375,0.75} 1.12173x& \cellcolor[rgb]{0.775390625,0.9990234375,0.75} 1.10225x& \cellcolor[rgb]{0.779296875,0.9990234375,0.75} 1.12120x& \cellcolor[rgb]{0.779296875,0.9990234375,0.75} 1.12080x& \cellcolor[rgb]{0.779296875,0.9990234375,0.75} 1.11794x& \cellcolor[rgb]{0.9990234375,0.9677734375,0.75} 3.00248x& \cellcolor[rgb]{0.9990234375,0.9306640625,0.75} 4.17628x& \cellcolor[rgb]{0.9990234375,0.9296875,0.75} 4.20718x\\
+268435399 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 218313& \cellcolor[rgb]{0.7861328125,0.9990234375,0.75} 1.14571x& \cellcolor[rgb]{0.7861328125,0.9990234375,0.75} 1.14655x& \cellcolor[rgb]{0.7861328125,0.9990234375,0.75} 1.14531x& \cellcolor[rgb]{0.78515625,0.9990234375,0.75} 1.14376x& \cellcolor[rgb]{0.7861328125,0.9990234375,0.75} 1.14572x& \cellcolor[rgb]{0.9990234375,0.9501953125,0.75} 3.55232x& \cellcolor[rgb]{0.9990234375,0.9033203125,0.75} 5.06141x& \cellcolor[rgb]{0.9990234375,0.90234375,0.75} 5.08687x\\
+536870909 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 245682& \cellcolor[rgb]{0.794921875,0.9990234375,0.75} 1.18408x& \cellcolor[rgb]{0.794921875,0.9990234375,0.75} 1.18299x& \cellcolor[rgb]{0.794921875,0.9990234375,0.75} 1.18404x& \cellcolor[rgb]{0.794921875,0.9990234375,0.75} 1.18194x& \cellcolor[rgb]{0.794921875,0.9990234375,0.75} 1.18307x& \cellcolor[rgb]{0.9990234375,0.9296875,0.75} 4.20818x& \cellcolor[rgb]{0.9990234375,0.8720703125,0.75} 6.06747x& \cellcolor[rgb]{0.9990234375,0.8701171875,0.75} 6.13899x\\
+1073741789 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 285913& \cellcolor[rgb]{0.8037109375,0.9990234375,0.75} 1.21667x& \cellcolor[rgb]{0.8037109375,0.9990234375,0.75} 1.21894x& \cellcolor[rgb]{0.8037109375,0.9990234375,0.75} 1.21755x& \cellcolor[rgb]{0.8037109375,0.9990234375,0.75} 1.21668x& \cellcolor[rgb]{0.8037109375,0.9990234375,0.75} 1.21706x& \cellcolor[rgb]{0.9990234375,0.908203125,0.75} 4.89647x& \cellcolor[rgb]{0.9990234375,0.8369140625,0.75} 7.20172x& \cellcolor[rgb]{0.9990234375,0.8359375,0.75} 7.23917x\\
+\hline
+\end{NiceTabular}}
+\caption{\label{tab:factori32}Benchmark of \texttt{benches/factori32.ll} output by \mintinline{bash}{dune exec bench -- -f factori32 -n 1000}}
+\end{table}
+
 \subsubsection{\lstinline!benches/phis.ll!}
@@
 \caption{\label{tab:subset}Benchmark of \texttt{benches/sieven.ll} output by \mintinline{bash}{dune exec bench -- -f fib -n 1000}}
-\end{table}
-
-\subsubsection{\lstinline!benches/factori32.ll!}
-
-Integer factorization is another interesting algorithm. Generally considered to be NP, here are the results of factorizing known primes of inreasing bit length. For \(n=1024\).
-\begin{table}[H]
-\makebox[\textwidth][c]{
-\begin{NiceTabular}{|c|ccccccccc|}
-\hline
-arg(s) & clang & simple 12 & simple 2 & briggs 12 & briggs 2 & linear 12 & linear 2 & greedy 12 & greedy 0 \\
-\hline
-16777213 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 167130& \cellcolor[rgb]{0.759765625,0.9990234375,0.75} 1.04160x& \cellcolor[rgb]{0.7509765625,0.9990234375,0.75} 1.00651x& \cellcolor[rgb]{0.7626953125,0.9990234375,0.75} 1.05411x& \cellcolor[rgb]{0.7626953125,0.9990234375,0.75} 1.05372x& \cellcolor[rgb]{0.7607421875,0.9990234375,0.75} 1.04686x& \cellcolor[rgb]{0.958984375,0.9990234375,0.75} 1.84213x& \cellcolor[rgb]{0.9990234375,0.98828125,0.75} 2.33215x& \cellcolor[rgb]{0.9990234375,0.9873046875,0.75} 2.34511x\\
-33554393 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 175409& \cellcolor[rgb]{0.7646484375,0.9990234375,0.75} 1.06219x& \cellcolor[rgb]{0.7646484375,0.9990234375,0.75} 1.06070x& \cellcolor[rgb]{0.765625,0.9990234375,0.75} 1.06415x& \cellcolor[rgb]{0.7646484375,0.9990234375,0.75} 1.06212x& \cellcolor[rgb]{0.7646484375,0.9990234375,0.75} 1.06039x& \cellcolor[rgb]{0.9990234375,0.9951171875,0.75} 2.12199x& \cellcolor[rgb]{0.9990234375,0.974609375,0.75} 2.78336x& \cellcolor[rgb]{0.9990234375,0.9736328125,0.75} 2.79956x\\
-67108859 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 185779& \cellcolor[rgb]{0.7685546875,0.9990234375,0.75} 1.07588x& \cellcolor[rgb]{0.76953125,0.9990234375,0.75} 1.07923x& \cellcolor[rgb]{0.7705078125,0.9990234375,0.75} 1.08373x& \cellcolor[rgb]{0.76953125,0.9990234375,0.75} 1.08043x& \cellcolor[rgb]{0.76953125,0.9990234375,0.75} 1.07895x& \cellcolor[rgb]{0.9990234375,0.9833984375,0.75} 2.49571x& \cellcolor[rgb]{0.9990234375,0.955078125,0.75} 3.38089x& \cellcolor[rgb]{0.9990234375,0.955078125,0.75} 3.40121x\\
-134217689 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 197302& \cellcolor[rgb]{0.7802734375,0.9990234375,0.75} 1.12173x& \cellcolor[rgb]{0.775390625,0.9990234375,0.75} 1.10225x& \cellcolor[rgb]{0.779296875,0.9990234375,0.75} 1.12120x& \cellcolor[rgb]{0.779296875,0.9990234375,0.75} 1.12080x& \cellcolor[rgb]{0.779296875,0.9990234375,0.75} 1.11794x& \cellcolor[rgb]{0.9990234375,0.9677734375,0.75} 3.00248x& \cellcolor[rgb]{0.9990234375,0.9306640625,0.75} 4.17628x& \cellcolor[rgb]{0.9990234375,0.9296875,0.75} 4.20718x\\
-268435399 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 218313& \cellcolor[rgb]{0.7861328125,0.9990234375,0.75} 1.14571x& \cellcolor[rgb]{0.7861328125,0.9990234375,0.75} 1.14655x& \cellcolor[rgb]{0.7861328125,0.9990234375,0.75} 1.14531x& \cellcolor[rgb]{0.78515625,0.9990234375,0.75} 1.14376x& \cellcolor[rgb]{0.7861328125,0.9990234375,0.75} 1.14572x& \cellcolor[rgb]{0.9990234375,0.9501953125,0.75} 3.55232x& \cellcolor[rgb]{0.9990234375,0.9033203125,0.75} 5.06141x& \cellcolor[rgb]{0.9990234375,0.90234375,0.75} 5.08687x\\
-536870909 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 245682& \cellcolor[rgb]{0.794921875,0.9990234375,0.75} 1.18408x& \cellcolor[rgb]{0.794921875,0.9990234375,0.75} 1.18299x& \cellcolor[rgb]{0.794921875,0.9990234375,0.75} 1.18404x& \cellcolor[rgb]{0.794921875,0.9990234375,0.75} 1.18194x& \cellcolor[rgb]{0.794921875,0.9990234375,0.75} 1.18307x& \cellcolor[rgb]{0.9990234375,0.9296875,0.75} 4.20818x& \cellcolor[rgb]{0.9990234375,0.8720703125,0.75} 6.06747x& \cellcolor[rgb]{0.9990234375,0.8701171875,0.75} 6.13899x\\
-1073741789 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 285913& \cellcolor[rgb]{0.8037109375,0.9990234375,0.75} 1.21667x& \cellcolor[rgb]{0.8037109375,0.9990234375,0.75} 1.21894x& \cellcolor[rgb]{0.8037109375,0.9990234375,0.75} 1.21755x& \cellcolor[rgb]{0.8037109375,0.9990234375,0.75} 1.21668x& \cellcolor[rgb]{0.8037109375,0.9990234375,0.75} 1.21706x& \cellcolor[rgb]{0.9990234375,0.908203125,0.75} 4.89647x& \cellcolor[rgb]{0.9990234375,0.8369140625,0.75} 7.20172x& \cellcolor[rgb]{0.9990234375,0.8359375,0.75} 7.23917x\\
-\hline
-\end{NiceTabular}}
-\caption{\label{tab:factori32}Benchmark of \texttt{benches/fib.ll} output by \mintinline{bash}{dune exec bench -- -f fib -n 1000}}
 \end{table}

### 2024-01-14 article.tex
idk

@@
 gep      & \textrm{\textit{see \ref{impl-gep}}} & \{\}\\
-icmp     & \textrm{\textit{see \ref{impl-cond}}} & \{\texttt{\%rFLAGS}\}\\
+icmp     & cmpx, set & \{\texttt{\%rFLAGS}\}\\
 load     & movx (\%src), \%dst & \{\}\\
@@
 sdiv     & idivx \%src, \%dst & \{\texttt{\%rax}, \texttt{\%rdx}\}\\
-select   &  & \{\}\\
+select   & cmpx, cmoveq & \{\}\\
 sext     & movx \%src, \%dst & \{\}\\
@@
 Mostly because of calling conventions, caller saved registers listed in Table \ref{tab:trans} need to be saved/restored before/after all.
-This isn't necessary in all cases, only the cases where they are actually used (which ought to be a simple lookup in the assignments table), but is not accounted for in this implementation.
+This isn't necessary in all functions, only those where the registers are  actually used (which ought to be a simple lookup in the assignments table), but is not accounted for in this implementation.
 Instead they are all saved/restored regardless of use this could also be further optimized by simply subtracting from \texttt{\%rsp} and moving them to their respective stack locations manually,  instead of using push/pop.
@@
 Furthermore, the way in which arguments are passed is also unnecessarily convoluted, but stems from a corner case in which variables were overwritten by moving to the respective register of the \(n\) function argument.
-This can be remedied by ordering the \texttt{mov}s correctly, and even in the case where there are cycles and \texttt{mov}s cannot be made without overwriting, one of the scratch registers can be used instead.
+This can be remedied by ordering the \texttt{mov}s correctly, and even in the case where there are cycles and \texttt{mov}s cannot be made without overwriting at least one, one of the scratch registers can be used instead.
 But as a workaround for this, the arguments are simply pushed from the source operands and popped to their respective argument registers (in reverse order).
@@
 typedef struct {
+  char c;
   int x;
@@
 }
-
      \end{minted}
@@
 %6 = getelementptr %struct.vector,
-            %i8* %1, i64 0, i32 1
+            %i8* %1, i64 0, i32 2
 %7 = load i32, i32* %6
 %8 = getelementptr %struct.vector,
-            %i8* %0, i64 0, i32 1
+            %i8* %0, i64 0, i32 2
 %9 = load i32, i32* %8
@@
 \end{figure}
-
-\subsubsection{Implementing \mintinline{llvm}{icmp} instructions} \label{impl-cond}
-The \mintinline{llvm}{icmp} and \mintinline{llvm}{select} instructions both rely on the \texttt{cmpx} x86 instruction for most of the heavy lifting.
+\noindent Essentially \texttt{gep} instructions allow for unbounded array lookups, in that a \texttt{getelementptr} instruction is followed by the base address and one or more scaled offsets.
+What happens in Figure \ref{fig:vector.ll} is \texttt{vector} type is reduced to an \texttt{int} array, because it essentially is: in C, struct fields are generally  padded such that their respective memory addresses are aligned on four- or eight-byte boundaries \cite{cstd}.
+As such, accessing the third field entry corresponds to accessing the third element of an \texttt{int} array.
+The \texttt{getelementptr} instruction gets the pointer of this element by multiplying the size with the index and adding that to the base address.
+Therefore, the x86 that is generated essentially amounts to an \texttt{imul} instruction as can be seen in Figure
+
+\begin{figure}[H]
+    \centering
+   \begin{minipage}{0.48\textwidth}
+     \centering
+     \begin{minted}[linenos,bgcolor=LightGray]{gas}
+movq %rdi, %rcx
+movq $0, %rax
+imulq $0, %rax
+movq $2, %rax
+imulq $8, %rax
+     \end{minted}
+     \caption{\texttt{dune exec build -{}- -t x86}}\label{fig:square0s}
+   \end{minipage}
+\end{figure}
 

### 2024-01-14 article.tex
blah

@@
 The \texttt{getelementptr} instruction gets the pointer of this element by multiplying the size with the index and adding that to the base address.
-Therefore, the x86 that is generated essentially amounts to an \texttt{imul} instruction as can be seen in Figure
+Therefore, the x86 that is generated essentially amounts to an \texttt{imul} instruction as can be seen in Figure \ref{fig:gep}.
 
@@
      \begin{minted}[linenos,bgcolor=LightGray]{gas}
+; ...
 movq %rdi, %rcx
 movq $0, %rax
-imulq $0, %rax
+addq %rax, %rcx
+imulq $8, %rax
 movq $2, %rax
-imulq $8, %rax
-     \end{minted}
-     \caption{\texttt{dune exec build -{}- -t x86}}\label{fig:square0s}
+imulq $4, %rax
+addq %rax, %rcx
+; ...
+     \end{minted}
+     \caption{\texttt{dune exec build -{}- -t x86}}\label{fig:gep}
    \end{minipage}
 \end{figure}
+
+\noindent where the base register  \texttt{\%1} has assignment \(\mathit{assign}[\texttt{\%1}] = \texttt{\%rdi}\) is moved into the scratch register \texttt{\%rcx}, and each of the offset operands are 1) moved into the other scratch register \texttt{\%rax}, 2) multiplied by the size of their respective types, and 3) added to the base register.
+
+Unfortunately, the gep wasn't implemented properly in time for handin.
+Instead the code appears to be left over from what was jumbled together in the Tiger days, so it does work in \textit{some} cases but for from all.
 
@@
 \end{NiceTabular}}
-\caption{\label{tab:dumb-fib}Benchmark of \texttt{benches/fib.ll} output by \mintinline{bash}{dune exec bench -- -f fib -n 1000}}
+\caption{\label{tab:dumb-fib}Benchmark of \texttt{benches/fib.ll} output by \mintinline{bash}{dune exec bench -- -f fib -n 1024}}
 \end{table}
@@
 \end{NiceTabular}}
-\caption{\label{tab:dumb-fib-greedy}Benchmark of \texttt{benches/fib.ll} output by \mintinline{bash}{dune exec bench -- -f fib -n 1000}}
+\caption{\label{tab:dumb-fib-greedy}Benchmark of \texttt{benches/fib.ll} output by \mintinline{bash}{dune exec bench -- -f fib -n 1024}}
 \end{table}
 
-\noindent While it's not a complete 1:1 in terms of runtime as the Tiger compiler seems to perform slightly better than Greedy 0 (by \textasciitilde 30\% or so), that does make sense as there are marginally more instructions involved with each operation, not to mention the extra care taken to ensure calling conventions are upheld.
+\noindent While it's not a complete one to one in terms of runtime as the Tiger compiler seems to perform slightly better than Greedy 0 (by \textasciitilde 30\% or so), that does make sense as there are marginally more instructions involved with each operation, not to mention the extra care taken to ensure calling conventions are upheld.
 
@@
 
-\subsubsection{\lstinline!benches/factori32.ll!}
-
-Integer factorization is another interesting algorithm. Generally considered to be NP, here are the results of factorizing known primes of inreasing bit length. For \(n=1024\).
+\subsubsection{\lstinline!benches/loopn.ll!}
+
+Another simple example is a program that only consists of a loop running for \(n\) iterations.
+There stem from the \texttt{tests/loopn.c} file but with the call to \texttt{printf} removed such that no unnecessary time is spent doing I/O.
+Depending on the level of optimization chosen, either the mutation strategy of stack allocated variables or phi nodes are used.
+For the case of using the stack, the simplicty of such a program shines through as can be seen on Table \ref{tab:loopn0}:
+
 \begin{table}[H]
@@
 \hline
-16777213 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 167130& \cellcolor[rgb]{0.759765625,0.9990234375,0.75} 1.04160x& \cellcolor[rgb]{0.7509765625,0.9990234375,0.75} 1.00651x& \cellcolor[rgb]{0.7626953125,0.9990234375,0.75} 1.05411x& \cellcolor[rgb]{0.7626953125,0.9990234375,0.75} 1.05372x& \cellcolor[rgb]{0.7607421875,0.9990234375,0.75} 1.04686x& \cellcolor[rgb]{0.958984375,0.9990234375,0.75} 1.84213x& \cellcolor[rgb]{0.9990234375,0.98828125,0.75} 2.33215x& \cellcolor[rgb]{0.9990234375,0.9873046875,0.75} 2.34511x\\
-33554393 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 175409& \cellcolor[rgb]{0.7646484375,0.9990234375,0.75} 1.06219x& \cellcolor[rgb]{0.7646484375,0.9990234375,0.75} 1.06070x& \cellcolor[rgb]{0.765625,0.9990234375,0.75} 1.06415x& \cellcolor[rgb]{0.7646484375,0.9990234375,0.75} 1.06212x& \cellcolor[rgb]{0.7646484375,0.9990234375,0.75} 1.06039x& \cellcolor[rgb]{0.9990234375,0.9951171875,0.75} 2.12199x& \cellcolor[rgb]{0.9990234375,0.974609375,0.75} 2.78336x& \cellcolor[rgb]{0.9990234375,0.9736328125,0.75} 2.79956x\\
-67108859 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 185779& \cellcolor[rgb]{0.7685546875,0.9990234375,0.75} 1.07588x& \cellcolor[rgb]{0.76953125,0.9990234375,0.75} 1.07923x& \cellcolor[rgb]{0.7705078125,0.9990234375,0.75} 1.08373x& \cellcolor[rgb]{0.76953125,0.9990234375,0.75} 1.08043x& \cellcolor[rgb]{0.76953125,0.9990234375,0.75} 1.07895x& \cellcolor[rgb]{0.9990234375,0.9833984375,0.75} 2.49571x& \cellcolor[rgb]{0.9990234375,0.955078125,0.75} 3.38089x& \cellcolor[rgb]{0.9990234375,0.955078125,0.75} 3.40121x\\
-134217689 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 197302& \cellcolor[rgb]{0.7802734375,0.9990234375,0.75} 1.12173x& \cellcolor[rgb]{0.775390625,0.9990234375,0.75} 1.10225x& \cellcolor[rgb]{0.779296875,0.9990234375,0.75} 1.12120x& \cellcolor[rgb]{0.779296875,0.9990234375,0.75} 1.12080x& \cellcolor[rgb]{0.779296875,0.9990234375,0.75} 1.11794x& \cellcolor[rgb]{0.9990234375,0.9677734375,0.75} 3.00248x& \cellcolor[rgb]{0.9990234375,0.9306640625,0.75} 4.17628x& \cellcolor[rgb]{0.9990234375,0.9296875,0.75} 4.20718x\\
-268435399 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 218313& \cellcolor[rgb]{0.7861328125,0.9990234375,0.75} 1.14571x& \cellcolor[rgb]{0.7861328125,0.9990234375,0.75} 1.14655x& \cellcolor[rgb]{0.7861328125,0.9990234375,0.75} 1.14531x& \cellcolor[rgb]{0.78515625,0.9990234375,0.75} 1.14376x& \cellcolor[rgb]{0.7861328125,0.9990234375,0.75} 1.14572x& \cellcolor[rgb]{0.9990234375,0.9501953125,0.75} 3.55232x& \cellcolor[rgb]{0.9990234375,0.9033203125,0.75} 5.06141x& \cellcolor[rgb]{0.9990234375,0.90234375,0.75} 5.08687x\\
-536870909 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 245682& \cellcolor[rgb]{0.794921875,0.9990234375,0.75} 1.18408x& \cellcolor[rgb]{0.794921875,0.9990234375,0.75} 1.18299x& \cellcolor[rgb]{0.794921875,0.9990234375,0.75} 1.18404x& \cellcolor[rgb]{0.794921875,0.9990234375,0.75} 1.18194x& \cellcolor[rgb]{0.794921875,0.9990234375,0.75} 1.18307x& \cellcolor[rgb]{0.9990234375,0.9296875,0.75} 4.20818x& \cellcolor[rgb]{0.9990234375,0.8720703125,0.75} 6.06747x& \cellcolor[rgb]{0.9990234375,0.8701171875,0.75} 6.13899x\\
-1073741789 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 285913& \cellcolor[rgb]{0.8037109375,0.9990234375,0.75} 1.21667x& \cellcolor[rgb]{0.8037109375,0.9990234375,0.75} 1.21894x& \cellcolor[rgb]{0.8037109375,0.9990234375,0.75} 1.21755x& \cellcolor[rgb]{0.8037109375,0.9990234375,0.75} 1.21668x& \cellcolor[rgb]{0.8037109375,0.9990234375,0.75} 1.21706x& \cellcolor[rgb]{0.9990234375,0.908203125,0.75} 4.89647x& \cellcolor[rgb]{0.9990234375,0.8369140625,0.75} 7.20172x& \cellcolor[rgb]{0.9990234375,0.8359375,0.75} 7.23917x\\
+1048575 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 2238539& \cellcolor[rgb]{0.9990234375,0.9814453125,0.75} 2.56041x& \cellcolor[rgb]{0.9990234375,0.978515625,0.75} 2.63938x& \cellcolor[rgb]{0.9990234375,0.9814453125,0.75} 2.56101x& \cellcolor[rgb]{0.9990234375,0.978515625,0.75} 2.63915x& \cellcolor[rgb]{0.9990234375,0.9814453125,0.75} 2.56111x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 19.26926x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 27.70065x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 36.60162x\\
+2097151 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 4333585& \cellcolor[rgb]{0.9990234375,0.9794921875,0.75} 2.61339x& \cellcolor[rgb]{0.9990234375,0.9765625,0.75} 2.69416x& \cellcolor[rgb]{0.9990234375,0.9794921875,0.75} 2.61334x& \cellcolor[rgb]{0.9990234375,0.9765625,0.75} 2.69427x& \cellcolor[rgb]{0.9990234375,0.9794921875,0.75} 2.61352x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 19.87471x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 28.58590x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 37.78097x\\
+4194303 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 8529189& \cellcolor[rgb]{0.9990234375,0.978515625,0.75} 2.63937x& \cellcolor[rgb]{0.9990234375,0.9765625,0.75} 2.72120x& \cellcolor[rgb]{0.9990234375,0.978515625,0.75} 2.63934x& \cellcolor[rgb]{0.9990234375,0.9765625,0.75} 2.72124x& \cellcolor[rgb]{0.9990234375,0.978515625,0.75} 2.63938x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 20.17961x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 29.03173x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.37560x\\
+8388607 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 16919352& \cellcolor[rgb]{0.9990234375,0.978515625,0.75} 2.65269x& \cellcolor[rgb]{0.9990234375,0.9755859375,0.75} 2.73526x& \cellcolor[rgb]{0.9990234375,0.978515625,0.75} 2.65275x& \cellcolor[rgb]{0.9990234375,0.9755859375,0.75} 2.73534x& \cellcolor[rgb]{0.9990234375,0.978515625,0.75} 2.65264x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 20.33697x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 29.26211x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.68258x\\
+16777215 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 33697415& \cellcolor[rgb]{0.9990234375,0.9775390625,0.75} 2.65966x& \cellcolor[rgb]{0.9990234375,0.9755859375,0.75} 2.74260x& \cellcolor[rgb]{0.9990234375,0.9775390625,0.75} 2.65966x& \cellcolor[rgb]{0.9990234375,0.9755859375,0.75} 2.74263x& \cellcolor[rgb]{0.9990234375,0.9775390625,0.75} 2.65964x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 20.41817x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 29.39601x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.84065x\\
+33554431 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 67254339& \cellcolor[rgb]{0.9990234375,0.9775390625,0.75} 2.66311x& \cellcolor[rgb]{0.9990234375,0.9755859375,0.75} 2.74627x& \cellcolor[rgb]{0.9990234375,0.9775390625,0.75} 2.66312x& \cellcolor[rgb]{0.9990234375,0.9755859375,0.75} 2.74628x& \cellcolor[rgb]{0.9990234375,0.9775390625,0.75} 2.66311x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 20.45872x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 29.46374x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.91968x\\
+67108863 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 134365449& \cellcolor[rgb]{0.9990234375,0.9775390625,0.75} 2.66491x& \cellcolor[rgb]{0.9990234375,0.9755859375,0.75} 2.74816x& \cellcolor[rgb]{0.9990234375,0.9775390625,0.75} 2.66488x& \cellcolor[rgb]{0.9990234375,0.9755859375,0.75} 2.74816x& \cellcolor[rgb]{0.9990234375,0.9775390625,0.75} 2.66492x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 20.47947x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 29.49845x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.96011x\\
 \hline
 \end{NiceTabular}}
-\caption{\label{tab:factori32}Benchmark of \texttt{benches/factori32.ll} output by \mintinline{bash}{dune exec bench -- -f factori32 -n 1000}}
+\caption{\label{tab:loopn0}Benchmark of \texttt{benches/factori64.ll} output by \mintinline{bash}{dune exec bench -- -f factori32 -n 1024}}
+\end{table}
+
+\noindent Despite having no optimizations enabled and using the stack to store the counter, \texttt{clang} manages to build an executable wherein the loop consists of no more than 7 instructions (see Figure \ref{fig:clang-loop}).
+In comparison, mine is 19 (see Figure \ref{fig:my-loop0}):
+
+\begin{figure}[H]
+     \centering
+     \begin{minipage}[b]{0.32\textwidth}
+\begin{minted}[bgcolor=LightGray]{gas}
+# ...
+movslq 0xc(%rsp),%rax
+cmp    0x10(%rsp),%rax
+jae    0x1199 <main+89>
+mov    0xc(%rsp),%eax
+add    $0x1,%eax
+mov    %eax,0xc(%rsp)
+jmp    0x1179 <main+57>
+# ...
+     \end{minted}
+     \caption{TODO}\label{fig:clang-loop}
+   \end{minipage}
+   \begin{minipage}[b]{0.65\textwidth}
+   \begin{minipage}[b]{0.50\textwidth}
+  \begin{minted}[bgcolor=LightGray]{gas}
+# ...
+movq	%rsi, %rax
+movl	(%rax), %eax
+movq	$0, %rdx
+movl	%eax, %edx
+movq	%rdx, %rax
+movq	$0, %rbx
+movl	%eax, %ebx
+movq	%rdi, %rax
+movq	(%rax), %rax
+movq	%rax, %rdx
+     \end{minted}
+   \end{minipage}
+   \begin{minipage}[b]{0.50\textwidth}
+  \begin{minted}[bgcolor=LightGray]{gas}
+movq	%rax, %rdx
+movq	%rbx, %rax
+movq	%rdx, %rcx
+cmpq	%rcx, %rax
+movq	$0, %rdx
+setb	%dl
+movq	%rdx, %rax
+cmpq	$0, %rax
+je	_main$20
+jmp	_main$17
+# ...
+     \end{minted}
+   \end{minipage}
+     \caption{TODO}\label{fig:my-loop0}
+   \end{minipage}
+\end{figure}
+\noindent It does make sense that my compiler nearly takes 3x as many cycles considering that nearly 3x as many instructions are being executed.
+Because od the way values are stored, there is literally no benefit from Briggs coalescing since no phi nodes are involved ever.
+I was surprised to see Greedy and Linear 2 to be that much slower, but on second thought it makes perfect sense since loading variables from the stack before every single arithmetic/logic operation is considerably slower.
+What's more surprising is the fact that Tiger is nearly 4x faster than any Greedy (see Table \ref{tab:loopn0.tig}).
+It's presumably just because all binary operations consist of  about three instructions or so.
+
+\begin{table}[H]
+\makebox[\textwidth][c]{
+\begin{NiceTabular}{|c|ccccccccc|}
+\hline
+arg(s) & clang & greedy 12 & greedy 8 & greedy 6 & greedy 4 & greedy 2 & greedy 1 & greedy 0 & tiger \\
+\hline
+1048575 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 2237215& \cellcolor[rgb]{0.9990234375,0.75,0.75} 27.71735x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 36.15447x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 36.62296x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 36.62342x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 36.62283x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 36.62360x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 36.62324x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 10.97018x\\
+2097151 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 4334414& \cellcolor[rgb]{0.9990234375,0.75,0.75} 28.58063x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 37.28996x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 37.77377x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 37.77391x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 37.77373x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 37.77353x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 37.77356x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 11.29302x\\
+4194303 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 8529224& \cellcolor[rgb]{0.9990234375,0.75,0.75} 29.03182x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 37.88371x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.37535x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.37541x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.37546x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.37544x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.37543x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 11.46122x\\
+8388607 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 16919774& \cellcolor[rgb]{0.9990234375,0.75,0.75} 29.26144x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.18576x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.68170x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.68169x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.68175x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.68169x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.68163x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 11.54653x\\
+16777215 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 33698040& \cellcolor[rgb]{0.9990234375,0.75,0.75} 29.39364x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.34210x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.83996x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.83997x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.83991x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.83995x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.83993x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 11.59127x\\
+33554431 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 67253512& \cellcolor[rgb]{0.9990234375,0.75,0.75} 29.46538x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.42125x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.92015x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.92019x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.92020x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.92016x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.92020x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 11.61419x\\
+67108863 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 134364289& \cellcolor[rgb]{0.9990234375,0.75,0.75} 29.49788x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.46111x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.96041x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.96050x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.96052x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.96043x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.96040x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 11.62584x\\
+\hline
+\end{NiceTabular}}
+\caption{\label{tab:loopn0.tig}Benchmark of \texttt{benches/loopn0.ll} output by \mintinline{bash}{dune exec bench -- -f loopn0 -n 64}}
+\end{table}
+
+\noindent Even with \texttt{-O1} where phi nodes are introduced, does the Briggs allocator seem to affect anything as can be seen in Table \ref{tab:loopn1}.
+This is because the only two variables eligible for coalescing, \texttt{\%9} and \texttt{\%10}, are in conflict as both are used at separate points.
+\begin{table}[H]
+\makebox[\textwidth][c]{
+\begin{NiceTabular}{|c|ccccccccc|}
+\hline
+arg(s) & clang & simple 12 & simple 2 & briggs 12 & briggs 2 & linear 12 & linear 2 & greedy 12 & greedy 0 \\
+\hline
+131071 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 401485& \cellcolor[rgb]{0.912109375,0.9990234375,0.75} 1.65252x& \cellcolor[rgb]{0.912109375,0.9990234375,0.75} 1.65413x& \cellcolor[rgb]{0.912109375,0.9990234375,0.75} 1.65379x& \cellcolor[rgb]{0.912109375,0.9990234375,0.75} 1.65354x& \cellcolor[rgb]{0.912109375,0.9990234375,0.75} 1.65376x& \cellcolor[rgb]{0.912109375,0.9990234375,0.75} 1.65265x& \cellcolor[rgb]{0.908203125,0.9990234375,0.75} 1.63694x& \cellcolor[rgb]{0.9990234375,0.8056640625,0.75} 8.19401x\\
+262143 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 663679& \cellcolor[rgb]{0.9462890625,0.9990234375,0.75} 1.79023x& \cellcolor[rgb]{0.9462890625,0.9990234375,0.75} 1.79079x& \cellcolor[rgb]{0.9462890625,0.9990234375,0.75} 1.78950x& \cellcolor[rgb]{0.9462890625,0.9990234375,0.75} 1.79092x& \cellcolor[rgb]{0.9462890625,0.9990234375,0.75} 1.78914x& \cellcolor[rgb]{0.9462890625,0.9990234375,0.75} 1.78887x& \cellcolor[rgb]{0.9462890625,0.9990234375,0.75} 1.79096x& \cellcolor[rgb]{0.9990234375,0.7587890625,0.75} 9.69501x\\
+524287 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1188029& \cellcolor[rgb]{0.9697265625,0.9990234375,0.75} 1.88330x& \cellcolor[rgb]{0.9697265625,0.9990234375,0.75} 1.88252x& \cellcolor[rgb]{0.96875,0.9990234375,0.75} 1.88192x& \cellcolor[rgb]{0.9697265625,0.9990234375,0.75} 1.88240x& \cellcolor[rgb]{0.9697265625,0.9990234375,0.75} 1.88242x& \cellcolor[rgb]{0.96875,0.9990234375,0.75} 1.88211x& \cellcolor[rgb]{0.9697265625,0.9990234375,0.75} 1.88284x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 10.79175x\\
+1048575 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 2237256& \cellcolor[rgb]{0.9833984375,0.9990234375,0.75} 1.93727x& \cellcolor[rgb]{0.9833984375,0.9990234375,0.75} 1.93740x& \cellcolor[rgb]{0.982421875,0.9990234375,0.75} 1.93674x& \cellcolor[rgb]{0.982421875,0.9990234375,0.75} 1.93715x& \cellcolor[rgb]{0.9833984375,0.9990234375,0.75} 1.93728x& \cellcolor[rgb]{0.9833984375,0.9990234375,0.75} 1.93739x& \cellcolor[rgb]{0.982421875,0.9990234375,0.75} 1.93690x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 11.42118x\\
+2097151 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 4334546& \cellcolor[rgb]{0.990234375,0.9990234375,0.75} 1.96774x& \cellcolor[rgb]{0.990234375,0.9990234375,0.75} 1.96724x& \cellcolor[rgb]{0.990234375,0.9990234375,0.75} 1.96767x& \cellcolor[rgb]{0.990234375,0.9990234375,0.75} 1.96746x& \cellcolor[rgb]{0.990234375,0.9990234375,0.75} 1.96749x& \cellcolor[rgb]{0.990234375,0.9990234375,0.75} 1.96756x& \cellcolor[rgb]{0.990234375,0.9990234375,0.75} 1.96753x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 11.81912x\\
+4194303 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 8529268& \cellcolor[rgb]{0.994140625,0.9990234375,0.75} 1.98350x& \cellcolor[rgb]{0.994140625,0.9990234375,0.75} 1.98348x& \cellcolor[rgb]{0.994140625,0.9990234375,0.75} 1.98353x& \cellcolor[rgb]{0.994140625,0.9990234375,0.75} 1.98339x& \cellcolor[rgb]{0.994140625,0.9990234375,0.75} 1.98339x& \cellcolor[rgb]{0.994140625,0.9990234375,0.75} 1.98358x& \cellcolor[rgb]{0.994140625,0.9990234375,0.75} 1.98355x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 12.04037x\\
+8388607 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 16918996& \cellcolor[rgb]{0.99609375,0.9990234375,0.75} 1.99159x& \cellcolor[rgb]{0.99609375,0.9990234375,0.75} 1.99166x& \cellcolor[rgb]{0.99609375,0.9990234375,0.75} 1.99166x& \cellcolor[rgb]{0.99609375,0.9990234375,0.75} 1.99160x& \cellcolor[rgb]{0.99609375,0.9990234375,0.75} 1.99166x& \cellcolor[rgb]{0.99609375,0.9990234375,0.75} 1.99165x& \cellcolor[rgb]{0.99609375,0.9990234375,0.75} 1.99162x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 12.13501x\\
+\hline
+\end{NiceTabular}}
+\caption{\label{tab:loopn1}Benchmark of \texttt{benches/loopn1.ll} output by \mintinline{bash}{dune exec bench -- -f loopn1 -n 64}}
+\end{table}
+
+\subsubsection{\lstinline!benches/phis.ll!}
+To actually see the benefit of coalescing in action I have introduced some generated tests of \texttt{benches/phis\{0-7\}.ll}.
+Generated with \texttt{benches/phis.py <n> <m>} that produces a \texttt{.ll} file that consists of a loop hardcoded for \(n\) repetitions with \(m\) phi nodes.
+
+\subsubsection{\lstinline!benches/factori32.ll!}
+
+Integer factorization is another interesting algorithm. Generally considered to be NP, here are the results of factorizing known primes of inreasing bit length:
+\begin{table}[H]
+\makebox[\textwidth][c]{
+\begin{NiceTabular}{|c|ccccccccc|}
+\hline
+arg(s) & clang & simple 12 & simple 2 & briggs 12 & briggs 2 & linear 12 & linear 2 & greedy 12 & greedy 0 \\
+\hline
+16777213 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 168400& \cellcolor[rgb]{0.76171875,0.9990234375,0.75} 1.04991x& \cellcolor[rgb]{0.7568359375,0.9990234375,0.75} 1.02776x& \cellcolor[rgb]{0.76171875,0.9990234375,0.75} 1.04829x& \cellcolor[rgb]{0.7607421875,0.9990234375,0.75} 1.04435x& \cellcolor[rgb]{0.7607421875,0.9990234375,0.75} 1.04465x& \cellcolor[rgb]{0.9560546875,0.9990234375,0.75} 1.82795x& \cellcolor[rgb]{0.9990234375,0.98828125,0.75} 2.31432x& \cellcolor[rgb]{0.9990234375,0.98828125,0.75} 2.32746x\\
+33554393 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 175723& \cellcolor[rgb]{0.7646484375,0.9990234375,0.75} 1.06090x& \cellcolor[rgb]{0.7587890625,0.9990234375,0.75} 1.03777x& \cellcolor[rgb]{0.765625,0.9990234375,0.75} 1.06480x& \cellcolor[rgb]{0.7607421875,0.9990234375,0.75} 1.04465x& \cellcolor[rgb]{0.7587890625,0.9990234375,0.75} 1.03702x& \cellcolor[rgb]{0.9990234375,0.9951171875,0.75} 2.12225x& \cellcolor[rgb]{0.9990234375,0.974609375,0.75} 2.78015x& \cellcolor[rgb]{0.9990234375,0.9736328125,0.75} 2.79323x\\
+67108859 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 185913& \cellcolor[rgb]{0.763671875,0.9990234375,0.75} 1.05642x& \cellcolor[rgb]{0.76953125,0.9990234375,0.75} 1.08067x& \cellcolor[rgb]{0.7705078125,0.9990234375,0.75} 1.08352x& \cellcolor[rgb]{0.76953125,0.9990234375,0.75} 1.07946x& \cellcolor[rgb]{0.7705078125,0.9990234375,0.75} 1.08349x& \cellcolor[rgb]{0.9990234375,0.9833984375,0.75} 2.50046x& \cellcolor[rgb]{0.9990234375,0.955078125,0.75} 3.38055x& \cellcolor[rgb]{0.9990234375,0.955078125,0.75} 3.39737x\\
+134217689 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 199433& \cellcolor[rgb]{0.77734375,0.9990234375,0.75} 1.11084x& \cellcolor[rgb]{0.7783203125,0.9990234375,0.75} 1.11393x& \cellcolor[rgb]{0.771484375,0.9990234375,0.75} 1.08952x& \cellcolor[rgb]{0.77734375,0.9990234375,0.75} 1.11192x& \cellcolor[rgb]{0.775390625,0.9990234375,0.75} 1.10441x& \cellcolor[rgb]{0.9990234375,0.96875,0.75} 2.97084x& \cellcolor[rgb]{0.9990234375,0.931640625,0.75} 4.13960x& \cellcolor[rgb]{0.9990234375,0.9326171875,0.75} 4.12783x\\
+268435399 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 217520& \cellcolor[rgb]{0.78515625,0.9990234375,0.75} 1.14321x& \cellcolor[rgb]{0.787109375,0.9990234375,0.75} 1.15268x& \cellcolor[rgb]{0.7841796875,0.9990234375,0.75} 1.13781x& \cellcolor[rgb]{0.787109375,0.9990234375,0.75} 1.15059x& \cellcolor[rgb]{0.787109375,0.9990234375,0.75} 1.15174x& \cellcolor[rgb]{0.9990234375,0.9501953125,0.75} 3.56567x& \cellcolor[rgb]{0.9990234375,0.9033203125,0.75} 5.06899x& \cellcolor[rgb]{0.9990234375,0.9013671875,0.75} 5.10944x\\
+536870909 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 242651& \cellcolor[rgb]{0.798828125,0.9990234375,0.75} 1.19742x& \cellcolor[rgb]{0.798828125,0.9990234375,0.75} 1.19637x& \cellcolor[rgb]{0.796875,0.9990234375,0.75} 1.18878x& \cellcolor[rgb]{0.798828125,0.9990234375,0.75} 1.19987x& \cellcolor[rgb]{0.796875,0.9990234375,0.75} 1.19055x& \cellcolor[rgb]{0.9990234375,0.9287109375,0.75} 4.25606x& \cellcolor[rgb]{0.9990234375,0.8681640625,0.75} 6.18250x& \cellcolor[rgb]{0.9990234375,0.8671875,0.75} 6.21937x\\
+1073741789 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 282883& \cellcolor[rgb]{0.8076171875,0.9990234375,0.75} 1.23359x& \cellcolor[rgb]{0.8076171875,0.9990234375,0.75} 1.23140x& \cellcolor[rgb]{0.806640625,0.9990234375,0.75} 1.23124x& \cellcolor[rgb]{0.802734375,0.9990234375,0.75} 1.21433x& \cellcolor[rgb]{0.8076171875,0.9990234375,0.75} 1.23208x& \cellcolor[rgb]{0.9990234375,0.9072265625,0.75} 4.93689x& \cellcolor[rgb]{0.9990234375,0.8359375,0.75} 7.23489x& \cellcolor[rgb]{0.9990234375,0.8330078125,0.75} 7.31718x\\
+\hline
+\end{NiceTabular}}
+\caption{\label{tab:factori64}Benchmark of \texttt{benches/factori64.ll} output by \mintinline{bash}{dune exec bench -- -f factori32 -n 1024}}
+\end{table}
+
+\begin{table}[H]
+\makebox[\textwidth][c]{
+\begin{NiceTabular}{|c|ccccccccc|}
+\hline
+arg(s) & clang & simple 12 & simple 2 & briggs 12 & briggs 2 & linear 12 & linear 2 & greedy 12 & greedy 0 \\
+\hline
+268435399 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 218872& \cellcolor[rgb]{0.765625,0.9990234375,0.75} 1.06367x& \cellcolor[rgb]{0.7763671875,0.9990234375,0.75} 1.10706x& \cellcolor[rgb]{0.7666015625,0.9990234375,0.75} 1.06818x& \cellcolor[rgb]{0.775390625,0.9990234375,0.75} 1.10527x& \cellcolor[rgb]{0.775390625,0.9990234375,0.75} 1.10276x& \cellcolor[rgb]{0.9072265625,0.9990234375,0.75} 1.63497x& \cellcolor[rgb]{0.9990234375,0.986328125,0.75} 2.38428x& \cellcolor[rgb]{0.9990234375,0.986328125,0.75} 2.38607x\\
+536870909 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 246634& \cellcolor[rgb]{0.771484375,0.9990234375,0.75} 1.08777x& \cellcolor[rgb]{0.78125,0.9990234375,0.75} 1.12660x& \cellcolor[rgb]{0.76953125,0.9990234375,0.75} 1.08040x& \cellcolor[rgb]{0.783203125,0.9990234375,0.75} 1.13337x& \cellcolor[rgb]{0.783203125,0.9990234375,0.75} 1.13368x& \cellcolor[rgb]{0.9482421875,0.9990234375,0.75} 1.79734x& \cellcolor[rgb]{0.9990234375,0.9755859375,0.75} 2.73281x& \cellcolor[rgb]{0.9990234375,0.9755859375,0.75} 2.74118x\\
+1073741789 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 285516& \cellcolor[rgb]{0.7744140625,0.9990234375,0.75} 1.09813x& \cellcolor[rgb]{0.7900390625,0.9990234375,0.75} 1.16160x& \cellcolor[rgb]{0.7744140625,0.9990234375,0.75} 1.10120x& \cellcolor[rgb]{0.7900390625,0.9990234375,0.75} 1.16364x& \cellcolor[rgb]{0.791015625,0.9990234375,0.75} 1.16724x& \cellcolor[rgb]{0.9912109375,0.9990234375,0.75} 1.97253x& \cellcolor[rgb]{0.9990234375,0.9638671875,0.75} 3.11594x& \cellcolor[rgb]{0.9990234375,0.9638671875,0.75} 3.12868x\\
+2147483647 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 340871& \cellcolor[rgb]{0.78125,0.9990234375,0.75} 1.12733x& \cellcolor[rgb]{0.7978515625,0.9990234375,0.75} 1.19431x& \cellcolor[rgb]{0.7802734375,0.9990234375,0.75} 1.12500x& \cellcolor[rgb]{0.7978515625,0.9990234375,0.75} 1.19348x& \cellcolor[rgb]{0.7958984375,0.9990234375,0.75} 1.18683x& \cellcolor[rgb]{0.9990234375,0.994140625,0.75} 2.15246x& \cellcolor[rgb]{0.9990234375,0.9521484375,0.75} 3.49642x& \cellcolor[rgb]{0.9990234375,0.951171875,0.75} 3.51766x\\
+4294967291 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 419568& \cellcolor[rgb]{0.78515625,0.9990234375,0.75} 1.14194x& \cellcolor[rgb]{0.802734375,0.9990234375,0.75} 1.21527x& \cellcolor[rgb]{0.78515625,0.9990234375,0.75} 1.14432x& \cellcolor[rgb]{0.8037109375,0.9990234375,0.75} 1.21632x& \cellcolor[rgb]{0.8046875,0.9990234375,0.75} 1.22207x& \cellcolor[rgb]{0.9990234375,0.98828125,0.75} 2.32057x& \cellcolor[rgb]{0.9990234375,0.9404296875,0.75} 3.87806x& \cellcolor[rgb]{0.9990234375,0.939453125,0.75} 3.89161x\\
+8589934583 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 529755& \cellcolor[rgb]{0.7900390625,0.9990234375,0.75} 1.16337x& \cellcolor[rgb]{0.810546875,0.9990234375,0.75} 1.24534x& \cellcolor[rgb]{0.7900390625,0.9990234375,0.75} 1.16161x& \cellcolor[rgb]{0.8115234375,0.9990234375,0.75} 1.24781x& \cellcolor[rgb]{0.8115234375,0.9990234375,0.75} 1.25015x& \cellcolor[rgb]{0.9990234375,0.9833984375,0.75} 2.47971x& \cellcolor[rgb]{0.9990234375,0.9296875,0.75} 4.22521x& \cellcolor[rgb]{0.9990234375,0.9287109375,0.75} 4.22809x\\
+17179869143 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 682438& \cellcolor[rgb]{0.794921875,0.9990234375,0.75} 1.18197x& \cellcolor[rgb]{0.8193359375,0.9990234375,0.75} 1.28018x& \cellcolor[rgb]{0.794921875,0.9990234375,0.75} 1.18192x& \cellcolor[rgb]{0.818359375,0.9990234375,0.75} 1.27765x& \cellcolor[rgb]{0.8193359375,0.9990234375,0.75} 1.27933x& \cellcolor[rgb]{0.9990234375,0.978515625,0.75} 2.63014x& \cellcolor[rgb]{0.9990234375,0.9189453125,0.75} 4.54643x& \cellcolor[rgb]{0.9990234375,0.9189453125,0.75} 4.56048x\\
+34359738337 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 907751& \cellcolor[rgb]{0.796875,0.9990234375,0.75} 1.18869x& \cellcolor[rgb]{0.8212890625,0.9990234375,0.75} 1.28923x& \cellcolor[rgb]{0.796875,0.9990234375,0.75} 1.18944x& \cellcolor[rgb]{0.822265625,0.9990234375,0.75} 1.29097x& \cellcolor[rgb]{0.8212890625,0.9990234375,0.75} 1.28925x& \cellcolor[rgb]{0.9990234375,0.9755859375,0.75} 2.72756x& \cellcolor[rgb]{0.9990234375,0.912109375,0.75} 4.76366x& \cellcolor[rgb]{0.9990234375,0.912109375,0.75} 4.77551x\\
+\hline
+\end{NiceTabular}}
+\caption{\label{tab:factori32}Benchmark of \texttt{benches/factori32.ll} output by \mintinline{bash}{dune exec bench -- -f factori32 -n 1024}}
+\end{table}
+
+\subsubsection{\lstinline!benches/sieven.ll!}
+
+\begin{table}[H]
+\makebox[\textwidth][c]{
+\begin{NiceTabular}{|c|ccccccccc|}
+\hline
+arg(s) & clang & simple 12 & simple 2 & briggs 12 & briggs 2 & linear 12 & linear 2 & greedy 12 & greedy 0 \\
+\hline
+128 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 157871& \cellcolor[rgb]{0.7509765625,0.9990234375,0.75} 1.00777x& \cellcolor[rgb]{0.751953125,0.9990234375,0.75} 1.01034x& \cellcolor[rgb]{0.7509765625,0.9990234375,0.75} 1.00764x& \cellcolor[rgb]{0.751953125,0.9990234375,0.75} 1.01089x& \cellcolor[rgb]{0.763671875,0.9990234375,0.75} 1.05785x& \cellcolor[rgb]{0.7890625,0.9990234375,0.75} 1.16044x& \cellcolor[rgb]{0.810546875,0.9990234375,0.75} 1.24422x& \cellcolor[rgb]{0.810546875,0.9990234375,0.75} 1.24352x\\
+256 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 162153& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.98003x& \cellcolor[rgb]{0.7548828125,0.9990234375,0.75} 1.02047x& \cellcolor[rgb]{0.751953125,0.9990234375,0.75} 1.01174x& \cellcolor[rgb]{0.7548828125,0.9990234375,0.75} 1.02203x& \cellcolor[rgb]{0.77734375,0.9990234375,0.75} 1.11154x& \cellcolor[rgb]{0.8291015625,0.9990234375,0.75} 1.31965x& \cellcolor[rgb]{0.873046875,0.9990234375,0.75} 1.49736x& \cellcolor[rgb]{0.8759765625,0.9990234375,0.75} 1.50711x\\
+512 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 168660& \cellcolor[rgb]{0.755859375,0.9990234375,0.75} 1.02724x& \cellcolor[rgb]{0.759765625,0.9990234375,0.75} 1.04284x& \cellcolor[rgb]{0.7578125,0.9990234375,0.75} 1.03418x& \cellcolor[rgb]{0.7607421875,0.9990234375,0.75} 1.04433x& \cellcolor[rgb]{0.8046875,0.9990234375,0.75} 1.22076x& \cellcolor[rgb]{0.90234375,0.9990234375,0.75} 1.61297x& \cellcolor[rgb]{0.9921875,0.9990234375,0.75} 1.97437x& \cellcolor[rgb]{0.9990234375,0.9970703125,0.75} 2.03639x\\
+1024 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 186641& \cellcolor[rgb]{0.7568359375,0.9990234375,0.75} 1.02816x& \cellcolor[rgb]{0.7626953125,0.9990234375,0.75} 1.05324x& \cellcolor[rgb]{0.759765625,0.9990234375,0.75} 1.04300x& \cellcolor[rgb]{0.763671875,0.9990234375,0.75} 1.05647x& \cellcolor[rgb]{0.8447265625,0.9990234375,0.75} 1.38383x& \cellcolor[rgb]{0.9990234375,0.994140625,0.75} 2.15658x& \cellcolor[rgb]{0.9990234375,0.974609375,0.75} 2.77549x& \cellcolor[rgb]{0.9990234375,0.970703125,0.75} 2.89126x\\
+2048 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 219043& \cellcolor[rgb]{0.7685546875,0.9990234375,0.75} 1.07644x& \cellcolor[rgb]{0.775390625,0.9990234375,0.75} 1.10227x& \cellcolor[rgb]{0.7685546875,0.9990234375,0.75} 1.07577x& \cellcolor[rgb]{0.7744140625,0.9990234375,0.75} 1.10024x& \cellcolor[rgb]{0.9130859375,0.9990234375,0.75} 1.65644x& \cellcolor[rgb]{0.9990234375,0.9658203125,0.75} 3.04685x& \cellcolor[rgb]{0.9990234375,0.93359375,0.75} 4.07178x& \cellcolor[rgb]{0.9990234375,0.9287109375,0.75} 4.24534x\\
+4096 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 286355& \cellcolor[rgb]{0.779296875,0.9990234375,0.75} 1.11926x& \cellcolor[rgb]{0.7880859375,0.9990234375,0.75} 1.15395x& \cellcolor[rgb]{0.7783203125,0.9990234375,0.75} 1.11675x& \cellcolor[rgb]{0.7880859375,0.9990234375,0.75} 1.15387x& \cellcolor[rgb]{0.9990234375,0.998046875,0.75} 2.00116x& \cellcolor[rgb]{0.9990234375,0.9306640625,0.75} 4.19594x& \cellcolor[rgb]{0.9990234375,0.8818359375,0.75} 5.76251x& \cellcolor[rgb]{0.9990234375,0.873046875,0.75} 6.03484x\\
+8192 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 421739& \cellcolor[rgb]{0.7890625,0.9990234375,0.75} 1.15807x& \cellcolor[rgb]{0.8017578125,0.9990234375,0.75} 1.20870x& \cellcolor[rgb]{0.7890625,0.9990234375,0.75} 1.15821x& \cellcolor[rgb]{0.8017578125,0.9990234375,0.75} 1.20849x& \cellcolor[rgb]{0.9990234375,0.9873046875,0.75} 2.36226x& \cellcolor[rgb]{0.9990234375,0.8916015625,0.75} 5.43577x& \cellcolor[rgb]{0.9990234375,0.8251953125,0.75} 7.57708x& \cellcolor[rgb]{0.9990234375,0.814453125,0.75} 7.91676x\\
+16384 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 692242& \cellcolor[rgb]{0.7978515625,0.9990234375,0.75} 1.19241x& \cellcolor[rgb]{0.8125,0.9990234375,0.75} 1.25460x& \cellcolor[rgb]{0.7978515625,0.9990234375,0.75} 1.19587x& \cellcolor[rgb]{0.8125,0.9990234375,0.75} 1.25348x& \cellcolor[rgb]{0.9990234375,0.9775390625,0.75} 2.66022x& \cellcolor[rgb]{0.9990234375,0.8583984375,0.75} 6.50741x& \cellcolor[rgb]{0.9990234375,0.775390625,0.75} 9.15450x& \cellcolor[rgb]{0.9990234375,0.763671875,0.75} 9.54856x\\
+\hline
+\end{NiceTabular}}
+\caption{\label{tab:sieven}Benchmark of \texttt{benches/sieven.ll} output by \mintinline{bash}{dune exec bench -- -f fib -n 1000}}
 \end{table}
@@
 \caption{\label{tab:dumb-fib-greedy}Benchmark of \texttt{benches/fib.ll} output by \mintinline{bash}{dune exec bench -- -f fib -n 1000}}
-\end{table}
-
-\subsubsection{\lstinline!benches/subset.ll!}
-A suitable example for nested loop iteration is \texttt{benches/subset.ll}.
-It's a very intuitive program that simply prints all \(2^n\) of a set of \(n\) elements, given as command line arguments.
-This is done by incrementing \(i\) by 1 from 0 to \(2^n-1\) and then printing the \(j\)th entry of the arguments array iff the \(j\)th bit of \(i\) is set.
-This leads to \(2^n\) repetitions of the outer loop and \(n\) of the inner, leading to \(2^n n\) repetitions.
-Since
-\begin{table}[H]
-\makebox[\textwidth][c]{
-\begin{NiceTabular}{|c|cccccccc|}
-\hline
-n & clang & simple 12 & simple 2 & briggs 12 & briggs 2 & linear & greedy 12 & greedy 0 \\
-\hline
-14 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.771484375,0.9990234375,0.75} 1.08770x& \cellcolor[rgb]{0.787109375,0.9990234375,0.75} 1.15188x& \cellcolor[rgb]{0.771484375,0.9990234375,0.75} 1.08832x& \cellcolor[rgb]{0.7890625,0.9990234375,0.75} 1.15813x& \cellcolor[rgb]{0.98046875,0.9990234375,0.75} 1.92564x& \cellcolor[rgb]{0.9990234375,0.98046875,0.75} 2.59504x& \cellcolor[rgb]{0.9990234375,0.96875,0.75} 2.95798x\\
-15 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.7705078125,0.9990234375,0.75} 1.08304x& \cellcolor[rgb]{0.78515625,0.9990234375,0.75} 1.14485x& \cellcolor[rgb]{0.7705078125,0.9990234375,0.75} 1.08536x& \cellcolor[rgb]{0.787109375,0.9990234375,0.75} 1.14973x& \cellcolor[rgb]{0.9833984375,0.9990234375,0.75} 1.93885x& \cellcolor[rgb]{0.9990234375,0.9794921875,0.75} 2.60300x& \cellcolor[rgb]{0.9990234375,0.96875,0.75} 2.96001x\\
-16 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.7705078125,0.9990234375,0.75} 1.08622x& \cellcolor[rgb]{0.78515625,0.9990234375,0.75} 1.14309x& \cellcolor[rgb]{0.7705078125,0.9990234375,0.75} 1.08345x& \cellcolor[rgb]{0.78515625,0.9990234375,0.75} 1.14388x& \cellcolor[rgb]{0.9853515625,0.9990234375,0.75} 1.94770x& \cellcolor[rgb]{0.9990234375,0.978515625,0.75} 2.62809x& \cellcolor[rgb]{0.9990234375,0.9677734375,0.75} 2.99463x\\
-17 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.7705078125,0.9990234375,0.75} 1.08533x& \cellcolor[rgb]{0.78515625,0.9990234375,0.75} 1.14335x& \cellcolor[rgb]{0.771484375,0.9990234375,0.75} 1.08652x& \cellcolor[rgb]{0.7861328125,0.9990234375,0.75} 1.14605x& \cellcolor[rgb]{0.986328125,0.9990234375,0.75} 1.94922x& \cellcolor[rgb]{0.9990234375,0.978515625,0.75} 2.64844x& \cellcolor[rgb]{0.9990234375,0.966796875,0.75} 3.01363x\\
-\hline
-\end{NiceTabular}}
-\caption{\label{tab:subset}Benchmark of \texttt{benches/sieven.ll} output by \mintinline{bash}{dune exec bench -- -f fib -n 1000}}
 \end{table}
@@
 
-\subsection{Comparison to other work and ideas for future work}
-
-Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
-Tellus at urna condimentum mattis.
-Enim facilisis gravida neque convallis.
-Enim facilisis gravida neque convallis a cras semper auctor.
-Vitae proin sagittis nisl rhoncus mattis rhoncus urna neque.
-Justo nec ultrices dui sapien eget mi proin.
-Id aliquet risus feugiat in ante.
-A cras semper auctor neque vitae tempus quam pellentesque nec.
-Nulla porttitor massa id neque aliquam vestibulum morbi.
-Facilisi cras fermentum odio eu.
-Odio pellentesque diam volutpat commodo sed egestas egestas fringilla.
-Amet consectetur adipiscing elit duis tristique sollicitudin nibh sit amet.
-Adipiscing at in tellus integer feugiat.
-A pellentesque sit amet porttitor.
-Sed elementum tempus egestas sed sed risus pretium quam.
-Quam viverra orci sagittis eu volutpat odio facilisis mauris sit.
-Arcu dictum varius duis at.
-Sociis natoque penatibus et magnis dis parturient montes.
-Varius sit amet mattis vulputate enim nulla aliquet.
-Ac ut consequat semper viverra.
-Scelerisque eleifend donec pretium vulputate sapien nec.
-Orci ac auctor augue mauris augue.
-Pulvinar mattis nunc sed blandit.
-Semper auctor neque vitae tempus quam pellentesque nec nam.
-Risus pretium quam vulputate dignissim.
-Dui faucibus in ornare quam viverra orci sagittis eu volutpat.
-Ac tortor vitae purus faucibus.
-Eu mi bibendum neque egestas congue quisque egestas diam in.
-Curabitur vitae nunc sed velit dignissim sodales ut eu sem.
-Arcu odio ut sem nulla.
+\subsection{Future work}
+
+It goes without saying that there are countless optimnizations waiting to be made, but as the goal of this project was to explore the different approaches not many regards were had for the instruction selection itself.
 

### 2024-01-14 article.tex
Add phis final

@@
 
+\begin{table}[H]
+\makebox[\textwidth][c]{
+\begin{NiceTabular}{|c|ccccccccc|}
+\hline
+phis & clang & simple 12 & simple 2 & briggs 12 & briggs 2 & linear 12 & linear 2 & greedy 12 & greedy 0 \\
+\hline
+ 0 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1073928603& \cellcolor[rgb]{0.9990234375,0.9833984375,0.75} 2.49981x& \cellcolor[rgb]{0.9990234375,0.9833984375,0.75} 2.49981x& \cellcolor[rgb]{0.8115234375,0.9990234375,0.75} 1.24997x& \cellcolor[rgb]{0.8115234375,0.9990234375,0.75} 1.24997x& \cellcolor[rgb]{0.9990234375,0.9833984375,0.75} 2.49981x& \cellcolor[rgb]{0.9990234375,0.9833984375,0.75} 2.49982x& \cellcolor[rgb]{0.9990234375,0.9833984375,0.75} 2.49981x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 11.68360x\\
+ 1 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1073928952& \cellcolor[rgb]{0.9990234375,0.9814453125,0.75} 2.54982x& \cellcolor[rgb]{0.9990234375,0.9794921875,0.75} 2.60344x& \cellcolor[rgb]{0.8740234375,0.9990234375,0.75} 1.49994x& \cellcolor[rgb]{0.8740234375,0.9990234375,0.75} 1.49994x& \cellcolor[rgb]{0.9990234375,0.9833984375,0.75} 2.49981x& \cellcolor[rgb]{0.9990234375,0.9833984375,0.75} 2.49983x& \cellcolor[rgb]{0.9990234375,0.9814453125,0.75} 2.54981x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 11.58368x\\
+ 2 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1610830251& \cellcolor[rgb]{0.9384765625,0.9990234375,0.75} 1.75897x& \cellcolor[rgb]{0.998046875,0.9990234375,0.75} 1.99991x& \cellcolor[rgb]{0.791015625,0.9990234375,0.75} 1.16665x& \cellcolor[rgb]{0.83203125,0.9990234375,0.75} 1.33330x& \cellcolor[rgb]{0.9150390625,0.9990234375,0.75} 1.66660x& \cellcolor[rgb]{0.9150390625,0.9990234375,0.75} 1.66663x& \cellcolor[rgb]{0.9375,0.9990234375,0.75} 1.75644x& \cellcolor[rgb]{0.9990234375,0.798828125,0.75} 8.40916x\\
+ 3 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 2147723548& \cellcolor[rgb]{0.9365234375,0.9990234375,0.75} 1.74997x& \cellcolor[rgb]{0.9677734375,0.9990234375,0.75} 1.87495x& \cellcolor[rgb]{0.76171875,0.9990234375,0.75} 1.05002x& \cellcolor[rgb]{0.7646484375,0.9990234375,0.75} 1.06250x& \cellcolor[rgb]{0.8115234375,0.9990234375,0.75} 1.24999x& \cellcolor[rgb]{0.8740234375,0.9990234375,0.75} 1.49997x& \cellcolor[rgb]{0.8828125,0.9990234375,0.75} 1.53576x& \cellcolor[rgb]{0.9990234375,0.853515625,0.75} 6.64460x\\
+ 4 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 2684621503& \cellcolor[rgb]{0.849609375,0.9990234375,0.75} 1.40001x& \cellcolor[rgb]{0.9482421875,0.9990234375,0.75} 1.79998x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.82421875,0.9990234375,0.75} 1.29999x& \cellcolor[rgb]{0.9990234375,0.8916015625,0.75} 5.43928x& \cellcolor[rgb]{0.9990234375,0.8740234375,0.75} 5.99128x\\
+ 5 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 3221532192& \cellcolor[rgb]{0.8740234375,0.9990234375,0.75} 1.49997x& \cellcolor[rgb]{0.95703125,0.9990234375,0.75} 1.83329x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00000x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99999x& \cellcolor[rgb]{0.8115234375,0.9990234375,0.75} 1.24999x& \cellcolor[rgb]{0.9990234375,0.9169921875,0.75} 4.60720x& \cellcolor[rgb]{0.9990234375,0.8798828125,0.75} 5.82084x\\
+ 6 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 3758430778& \cellcolor[rgb]{0.8388671875,0.9990234375,0.75} 1.35714x& \cellcolor[rgb]{0.9462890625,0.9990234375,0.75} 1.78839x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.88572x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.90478x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.85715x& \cellcolor[rgb]{0.802734375,0.9990234375,0.75} 1.21428x& \cellcolor[rgb]{0.9990234375,0.9326171875,0.75} 4.10980x& \cellcolor[rgb]{0.9990234375,0.8876953125,0.75} 5.54983x\\
+ 7 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 4295303796& \cellcolor[rgb]{0.8740234375,0.9990234375,0.75} 1.50000x& \cellcolor[rgb]{0.9521484375,0.9990234375,0.75} 1.81439x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.87500x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.87502x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.87500x& \cellcolor[rgb]{0.7958984375,0.9990234375,0.75} 1.18750x& \cellcolor[rgb]{0.9990234375,0.9482421875,0.75} 3.62761x& \cellcolor[rgb]{0.9990234375,0.91015625,0.75} 4.84205x\\
+\hline
+\end{NiceTabular}}
+\caption{\label{tab:phis}Benchmark of \texttt{benches/phis\{0..7\}.ll} output by \mintinline{bash}{dune exec bench -- -f phis -n 16}}
+\end{table}
+
 \subsubsection{\lstinline!benches/factori32.ll!}
@@
 
-\subsubsection{\lstinline!benches/phis.ll!}
-Coalescing with the Briggs allocator is evidently quite beneficial, as the it does out perform most benchmarks.
+%\subsubsection{\lstinline!benches/phis.ll!}
+%Coalescing with the Briggs allocator is evidently quite beneficial, as the it does out perform most benchmarks.
 This is obviously because eliminating movs stand to gain.
@@
 While in theory a phi node has the potential to eliminate one \texttt{mov} instruction, in practice this might be more beneficial due to the implementation.
-The \texttt{benches/phis.py <n> <m>} script will generate a \texttt{.ll} program consisting of \(m\) phi nodes over \(n\) iterations.
+%The \texttt{benches/phis.py <n> <m>} script will generate a \texttt{.ll} program consisting of \(m\) phi nodes over \(n\) iterations.
 %Parameterized over how many phi moves need to be made per iteration.
@@
 
-\begin{NiceTabular}{|c|ccccccccc|}
-\hline
-arg(s) & clang & simple 12 & simple 2 & briggs 12 & briggs 2 & linear 2 & linear 12 & greedy 12 & greedy 0 \\
-\hline
- & \cellcolor[rgb]{0.75,0.9990234375,0.75} 33699586& \cellcolor[rgb]{0.9990234375,0.9833984375,0.75} 2.49352x& \cellcolor[rgb]{0.9990234375,0.9833984375,0.75} 2.49358x& \cellcolor[rgb]{0.8115234375,0.9990234375,0.75} 1.24889x& \cellcolor[rgb]{0.8115234375,0.9990234375,0.75} 1.24890x& \cellcolor[rgb]{0.9990234375,0.9833984375,0.75} 2.49352x& \cellcolor[rgb]{0.9990234375,0.9833984375,0.75} 2.49355x& \cellcolor[rgb]{0.9990234375,0.9833984375,0.75} 2.49351x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 11.61953x\\
- & \cellcolor[rgb]{0.75,0.9990234375,0.75} 33696431& \cellcolor[rgb]{0.9990234375,0.9814453125,0.75} 2.54359x& \cellcolor[rgb]{0.9990234375,0.9814453125,0.75} 2.55986x& \cellcolor[rgb]{0.8740234375,0.9990234375,0.75} 1.49846x& \cellcolor[rgb]{0.873046875,0.9990234375,0.75} 1.49796x& \cellcolor[rgb]{0.9990234375,0.9833984375,0.75} 2.49373x& \cellcolor[rgb]{0.9990234375,0.9833984375,0.75} 2.49378x& \cellcolor[rgb]{0.9990234375,0.9814453125,0.75} 2.54356x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 11.52326x\\
- & \cellcolor[rgb]{0.75,0.9990234375,0.75} 50475128& \cellcolor[rgb]{0.9453125,0.9990234375,0.75} 1.78694x& \cellcolor[rgb]{0.998046875,0.9990234375,0.75} 1.99722x& \cellcolor[rgb]{0.791015625,0.9990234375,0.75} 1.16621x& \cellcolor[rgb]{0.83203125,0.9990234375,0.75} 1.33250x& \cellcolor[rgb]{0.9150390625,0.9990234375,0.75} 1.66480x& \cellcolor[rgb]{0.9150390625,0.9990234375,0.75} 1.66483x& \cellcolor[rgb]{0.94140625,0.9990234375,0.75} 1.77143x& \cellcolor[rgb]{0.9990234375,0.7939453125,0.75} 8.57567x\\
- & \cellcolor[rgb]{0.75,0.9990234375,0.75} 67274018& \cellcolor[rgb]{0.935546875,0.9990234375,0.75} 1.74795x& \cellcolor[rgb]{0.966796875,0.9990234375,0.75} 1.87263x& \cellcolor[rgb]{0.76171875,0.9990234375,0.75} 1.04956x& \cellcolor[rgb]{0.7646484375,0.9990234375,0.75} 1.06208x& \cellcolor[rgb]{0.8740234375,0.9990234375,0.75} 1.49853x& \cellcolor[rgb]{0.8115234375,0.9990234375,0.75} 1.24908x& \cellcolor[rgb]{0.87890625,0.9990234375,0.75} 1.52063x& \cellcolor[rgb]{0.9990234375,0.8564453125,0.75} 6.57981x\\
- & \cellcolor[rgb]{0.75,0.9990234375,0.75} 84029983& \cellcolor[rgb]{0.8486328125,0.9990234375,0.75} 1.39938x& \cellcolor[rgb]{0.9482421875,0.9990234375,0.75} 1.79869x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00002x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00003x& \cellcolor[rgb]{0.82421875,0.9990234375,0.75} 1.29955x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00004x& \cellcolor[rgb]{0.9990234375,0.8681640625,0.75} 6.19846x& \cellcolor[rgb]{0.9990234375,0.875,0.75} 5.98194x\\
- & \cellcolor[rgb]{0.75,0.9990234375,0.75} 100810530& \cellcolor[rgb]{0.8740234375,0.9990234375,0.75} 1.49925x& \cellcolor[rgb]{0.95703125,0.9990234375,0.75} 1.83217x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99997x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99996x& \cellcolor[rgb]{0.8115234375,0.9990234375,0.75} 1.24967x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99998x& \cellcolor[rgb]{0.9990234375,0.9169921875,0.75} 4.60493x& \cellcolor[rgb]{0.9990234375,0.8798828125,0.75} 5.80149x\\
- & \cellcolor[rgb]{0.75,0.9990234375,0.75} 117587019& \cellcolor[rgb]{0.837890625,0.9990234375,0.75} 1.35669x& \cellcolor[rgb]{0.9462890625,0.9990234375,0.75} 1.78907x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.88585x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.90488x& \cellcolor[rgb]{0.802734375,0.9990234375,0.75} 1.21403x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.85730x& \cellcolor[rgb]{0.9990234375,0.9326171875,0.75} 4.10513x& \cellcolor[rgb]{0.9990234375,0.888671875,0.75} 5.51438x\\
- & \cellcolor[rgb]{0.75,0.9990234375,0.75} 134365885& \cellcolor[rgb]{0.8740234375,0.9990234375,0.75} 1.49947x& \cellcolor[rgb]{0.9521484375,0.9990234375,0.75} 1.81229x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.87512x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.87515x& \cellcolor[rgb]{0.7958984375,0.9990234375,0.75} 1.18732x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.87514x& \cellcolor[rgb]{0.9990234375,0.947265625,0.75} 3.63380x& \cellcolor[rgb]{0.9990234375,0.9111328125,0.75} 4.81382x\\
-\hline
-\end{NiceTabular}
-
-1074877208
+%\begin{NiceTabular}{|c|ccccccccc|}
+%\hline
+%arg(s) & clang & simple 12 & simple 2 & briggs 12 & briggs 2 & linear 2 & linear 12 & greedy 12 & greedy 0 \\
+%\hline
+ %& \cellcolor[rgb]{0.75,0.9990234375,0.75} 33699586& \cellcolor[rgb]{0.9990234375,0.9833984375,0.75} 2.49352x& \cellcolor[rgb]{0.9990234375,0.9833984375,0.75} 2.49358x& \cellcolor[rgb]{0.8115234375,0.9990234375,0.75} 1.24889x& \cellcolor[rgb]{0.8115234375,0.9990234375,0.75} 1.24890x& \cellcolor[rgb]{0.9990234375,0.9833984375,0.75} 2.49352x& \cellcolor[rgb]{0.9990234375,0.9833984375,0.75} 2.49355x& \cellcolor[rgb]{0.9990234375,0.9833984375,0.75} 2.49351x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 11.61953x\\
+ %& \cellcolor[rgb]{0.75,0.9990234375,0.75} 33696431& \cellcolor[rgb]{0.9990234375,0.9814453125,0.75} 2.54359x& \cellcolor[rgb]{0.9990234375,0.9814453125,0.75} 2.55986x& \cellcolor[rgb]{0.8740234375,0.9990234375,0.75} 1.49846x& \cellcolor[rgb]{0.873046875,0.9990234375,0.75} 1.49796x& \cellcolor[rgb]{0.9990234375,0.9833984375,0.75} 2.49373x& \cellcolor[rgb]{0.9990234375,0.9833984375,0.75} 2.49378x& \cellcolor[rgb]{0.9990234375,0.9814453125,0.75} 2.54356x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 11.52326x\\
+ %& \cellcolor[rgb]{0.75,0.9990234375,0.75} 50475128& \cellcolor[rgb]{0.9453125,0.9990234375,0.75} 1.78694x& \cellcolor[rgb]{0.998046875,0.9990234375,0.75} 1.99722x& \cellcolor[rgb]{0.791015625,0.9990234375,0.75} 1.16621x& \cellcolor[rgb]{0.83203125,0.9990234375,0.75} 1.33250x& \cellcolor[rgb]{0.9150390625,0.9990234375,0.75} 1.66480x& \cellcolor[rgb]{0.9150390625,0.9990234375,0.75} 1.66483x& \cellcolor[rgb]{0.94140625,0.9990234375,0.75} 1.77143x& \cellcolor[rgb]{0.9990234375,0.7939453125,0.75} 8.57567x\\
+ %& \cellcolor[rgb]{0.75,0.9990234375,0.75} 67274018& \cellcolor[rgb]{0.935546875,0.9990234375,0.75} 1.74795x& \cellcolor[rgb]{0.966796875,0.9990234375,0.75} 1.87263x& \cellcolor[rgb]{0.76171875,0.9990234375,0.75} 1.04956x& \cellcolor[rgb]{0.7646484375,0.9990234375,0.75} 1.06208x& \cellcolor[rgb]{0.8740234375,0.9990234375,0.75} 1.49853x& \cellcolor[rgb]{0.8115234375,0.9990234375,0.75} 1.24908x& \cellcolor[rgb]{0.87890625,0.9990234375,0.75} 1.52063x& \cellcolor[rgb]{0.9990234375,0.8564453125,0.75} 6.57981x\\
+ %& \cellcolor[rgb]{0.75,0.9990234375,0.75} 84029983& \cellcolor[rgb]{0.8486328125,0.9990234375,0.75} 1.39938x& \cellcolor[rgb]{0.9482421875,0.9990234375,0.75} 1.79869x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00002x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00003x& \cellcolor[rgb]{0.82421875,0.9990234375,0.75} 1.29955x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00004x& \cellcolor[rgb]{0.9990234375,0.8681640625,0.75} 6.19846x& \cellcolor[rgb]{0.9990234375,0.875,0.75} 5.98194x\\
+ %& \cellcolor[rgb]{0.75,0.9990234375,0.75} 100810530& \cellcolor[rgb]{0.8740234375,0.9990234375,0.75} 1.49925x& \cellcolor[rgb]{0.95703125,0.9990234375,0.75} 1.83217x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99997x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99996x& \cellcolor[rgb]{0.8115234375,0.9990234375,0.75} 1.24967x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99998x& \cellcolor[rgb]{0.9990234375,0.9169921875,0.75} 4.60493x& \cellcolor[rgb]{0.9990234375,0.8798828125,0.75} 5.80149x\\
+ %& \cellcolor[rgb]{0.75,0.9990234375,0.75} 117587019& \cellcolor[rgb]{0.837890625,0.9990234375,0.75} 1.35669x& \cellcolor[rgb]{0.9462890625,0.9990234375,0.75} 1.78907x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.88585x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.90488x& \cellcolor[rgb]{0.802734375,0.9990234375,0.75} 1.21403x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.85730x& \cellcolor[rgb]{0.9990234375,0.9326171875,0.75} 4.10513x& \cellcolor[rgb]{0.9990234375,0.888671875,0.75} 5.51438x\\
+ %& \cellcolor[rgb]{0.75,0.9990234375,0.75} 134365885& \cellcolor[rgb]{0.8740234375,0.9990234375,0.75} 1.49947x& \cellcolor[rgb]{0.9521484375,0.9990234375,0.75} 1.81229x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.87512x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.87515x& \cellcolor[rgb]{0.7958984375,0.9990234375,0.75} 1.18732x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.87514x& \cellcolor[rgb]{0.9990234375,0.947265625,0.75} 3.63380x& \cellcolor[rgb]{0.9990234375,0.9111328125,0.75} 4.81382x\\
+%\hline
+%\end{NiceTabular}
+
+%1074877208
 % -O0
-\begin{table}[H]
-\makebox[\textwidth][c]{
-\begin{NiceTabular}{|c|ccccccccc|}
-\hline
-phis & clang & simple 12 & simple 2 & briggs 12 & briggs 2 & linear 2 & linear 12 & greedy 12 & greedy 0 \\
-\hline
- 0 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 150762& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99922x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00386x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00104x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99880x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00325x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99096x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00077x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99764x\\
- 1 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 263112& \cellcolor[rgb]{0.9453125,0.9990234375,0.75} 1.78765x& \cellcolor[rgb]{0.9580078125,0.9990234375,0.75} 1.83810x& \cellcolor[rgb]{0.8154296875,0.9990234375,0.75} 1.26448x& \cellcolor[rgb]{0.9365234375,0.9990234375,0.75} 1.75223x& \cellcolor[rgb]{0.9384765625,0.9990234375,0.75} 1.75782x& \cellcolor[rgb]{0.939453125,0.9990234375,0.75} 1.76278x& \cellcolor[rgb]{0.9453125,0.9990234375,0.75} 1.78716x& \cellcolor[rgb]{0.9990234375,0.865234375,0.75} 6.27904x\\
- 2 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 333724& \cellcolor[rgb]{0.8603515625,0.9990234375,0.75} 1.44531x& \cellcolor[rgb]{0.8955078125,0.9990234375,0.75} 1.58670x& \cellcolor[rgb]{0.7734375,0.9990234375,0.75} 1.09621x& \cellcolor[rgb]{0.7978515625,0.9990234375,0.75} 1.19483x& \cellcolor[rgb]{0.8466796875,0.9990234375,0.75} 1.39069x& \cellcolor[rgb]{0.8466796875,0.9990234375,0.75} 1.39096x& \cellcolor[rgb]{0.8642578125,0.9990234375,0.75} 1.45969x& \cellcolor[rgb]{0.9990234375,0.896484375,0.75} 5.28416x\\
- 3 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 399042& \cellcolor[rgb]{0.8720703125,0.9990234375,0.75} 1.49027x& \cellcolor[rgb]{0.892578125,0.9990234375,0.75} 1.57311x& \cellcolor[rgb]{0.7568359375,0.9990234375,0.75} 1.03136x& \cellcolor[rgb]{0.7587890625,0.9990234375,0.75} 1.03867x& \cellcolor[rgb]{0.8310546875,0.9990234375,0.75} 1.32618x& \cellcolor[rgb]{0.7900390625,0.9990234375,0.75} 1.16354x& \cellcolor[rgb]{0.8310546875,0.9990234375,0.75} 1.32695x& \cellcolor[rgb]{0.9990234375,0.9169921875,0.75} 4.61992x\\
- 4 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 464640& \cellcolor[rgb]{0.8193359375,0.9990234375,0.75} 1.28048x& \cellcolor[rgb]{0.8896484375,0.9990234375,0.75} 1.56253x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00015x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99836x& \cellcolor[rgb]{0.8017578125,0.9990234375,0.75} 1.21081x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99861x& \cellcolor[rgb]{0.9990234375,0.9375,0.75} 3.96512x& \cellcolor[rgb]{0.9990234375,0.9208984375,0.75} 4.48737x\\
- 5 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 529535& \cellcolor[rgb]{0.841796875,0.9990234375,0.75} 1.37214x& \cellcolor[rgb]{0.9033203125,0.9990234375,0.75} 1.61569x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99931x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99959x& \cellcolor[rgb]{0.7939453125,0.9990234375,0.75} 1.17952x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99948x& \cellcolor[rgb]{0.9990234375,0.94921875,0.75} 3.58774x& \cellcolor[rgb]{0.9990234375,0.921875,0.75} 4.47259x\\
- 6 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 594553& \cellcolor[rgb]{0.818359375,0.9990234375,0.75} 1.27726x& \cellcolor[rgb]{0.9013671875,0.9990234375,0.75} 1.61016x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.91261x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.92786x& \cellcolor[rgb]{0.7900390625,0.9990234375,0.75} 1.16424x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.89141x& \cellcolor[rgb]{0.9990234375,0.955078125,0.75} 3.38554x& \cellcolor[rgb]{0.9990234375,0.921875,0.75} 4.46924x\\
- 7 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 659461& \cellcolor[rgb]{0.8486328125,0.9990234375,0.75} 1.39945x& \cellcolor[rgb]{0.90234375,0.9990234375,0.75} 1.61499x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.90111x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.90219x& \cellcolor[rgb]{0.787109375,0.9990234375,0.75} 1.15128x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.90144x& \cellcolor[rgb]{0.9990234375,0.96484375,0.75} 3.08855x& \cellcolor[rgb]{0.9990234375,0.935546875,0.75} 4.03288x\\
-\end{NiceTabular}}
-\caption{\label{tab:dumb-fib-greedy}Benchmark of \texttt{benches/fib.ll} output by \mintinline{bash}{dune exec bench -- -f fib -n 1000}}
-\end{table}
+%\begin{table}[H]
+%\makebox[\textwidth][c]{
+%\begin{NiceTabular}{|c|ccccccccc|}
+%\hline
+%phis & clang & simple 12 & simple 2 & briggs 12 & briggs 2 & linear 2 & linear 12 & greedy 12 & greedy 0 \\
+%\hline
+ %0 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 150762& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99922x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00386x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00104x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99880x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00325x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99096x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00077x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99764x\\
+ %1 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 263112& \cellcolor[rgb]{0.9453125,0.9990234375,0.75} 1.78765x& \cellcolor[rgb]{0.9580078125,0.9990234375,0.75} 1.83810x& \cellcolor[rgb]{0.8154296875,0.9990234375,0.75} 1.26448x& \cellcolor[rgb]{0.9365234375,0.9990234375,0.75} 1.75223x& \cellcolor[rgb]{0.9384765625,0.9990234375,0.75} 1.75782x& \cellcolor[rgb]{0.939453125,0.9990234375,0.75} 1.76278x& \cellcolor[rgb]{0.9453125,0.9990234375,0.75} 1.78716x& \cellcolor[rgb]{0.9990234375,0.865234375,0.75} 6.27904x\\
+ %2 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 333724& \cellcolor[rgb]{0.8603515625,0.9990234375,0.75} 1.44531x& \cellcolor[rgb]{0.8955078125,0.9990234375,0.75} 1.58670x& \cellcolor[rgb]{0.7734375,0.9990234375,0.75} 1.09621x& \cellcolor[rgb]{0.7978515625,0.9990234375,0.75} 1.19483x& \cellcolor[rgb]{0.8466796875,0.9990234375,0.75} 1.39069x& \cellcolor[rgb]{0.8466796875,0.9990234375,0.75} 1.39096x& \cellcolor[rgb]{0.8642578125,0.9990234375,0.75} 1.45969x& \cellcolor[rgb]{0.9990234375,0.896484375,0.75} 5.28416x\\
+ %3 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 399042& \cellcolor[rgb]{0.8720703125,0.9990234375,0.75} 1.49027x& \cellcolor[rgb]{0.892578125,0.9990234375,0.75} 1.57311x& \cellcolor[rgb]{0.7568359375,0.9990234375,0.75} 1.03136x& \cellcolor[rgb]{0.7587890625,0.9990234375,0.75} 1.03867x& \cellcolor[rgb]{0.8310546875,0.9990234375,0.75} 1.32618x& \cellcolor[rgb]{0.7900390625,0.9990234375,0.75} 1.16354x& \cellcolor[rgb]{0.8310546875,0.9990234375,0.75} 1.32695x& \cellcolor[rgb]{0.9990234375,0.9169921875,0.75} 4.61992x\\
+ %4 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 464640& \cellcolor[rgb]{0.8193359375,0.9990234375,0.75} 1.28048x& \cellcolor[rgb]{0.8896484375,0.9990234375,0.75} 1.56253x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00015x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99836x& \cellcolor[rgb]{0.8017578125,0.9990234375,0.75} 1.21081x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99861x& \cellcolor[rgb]{0.9990234375,0.9375,0.75} 3.96512x& \cellcolor[rgb]{0.9990234375,0.9208984375,0.75} 4.48737x\\
+ %5 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 529535& \cellcolor[rgb]{0.841796875,0.9990234375,0.75} 1.37214x& \cellcolor[rgb]{0.9033203125,0.9990234375,0.75} 1.61569x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99931x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99959x& \cellcolor[rgb]{0.7939453125,0.9990234375,0.75} 1.17952x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99948x& \cellcolor[rgb]{0.9990234375,0.94921875,0.75} 3.58774x& \cellcolor[rgb]{0.9990234375,0.921875,0.75} 4.47259x\\
+ %6 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 594553& \cellcolor[rgb]{0.818359375,0.9990234375,0.75} 1.27726x& \cellcolor[rgb]{0.9013671875,0.9990234375,0.75} 1.61016x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.91261x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.92786x& \cellcolor[rgb]{0.7900390625,0.9990234375,0.75} 1.16424x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.89141x& \cellcolor[rgb]{0.9990234375,0.955078125,0.75} 3.38554x& \cellcolor[rgb]{0.9990234375,0.921875,0.75} 4.46924x\\
+ %7 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 659461& \cellcolor[rgb]{0.8486328125,0.9990234375,0.75} 1.39945x& \cellcolor[rgb]{0.90234375,0.9990234375,0.75} 1.61499x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.90111x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.90219x& \cellcolor[rgb]{0.787109375,0.9990234375,0.75} 1.15128x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.90144x& \cellcolor[rgb]{0.9990234375,0.96484375,0.75} 3.08855x& \cellcolor[rgb]{0.9990234375,0.935546875,0.75} 4.03288x\\
+%\end{NiceTabular}}
+%\caption{\label{tab:dumb-fib-greedy}Benchmark of \texttt{benches/fib.ll} output by \mintinline{bash}{dune exec bench -- -f fib -n 1000}}
+%\end{table}
+%
 
 % -O1
-\begin{table}[H]
-\makebox[\textwidth][c]{
-\begin{NiceTabular}{|c|ccccccccc|}
-\hline
-phis & clang & simple 12 & simple 2 & briggs 12 & briggs 2 & linear 2 & linear 12 & greedy 12 & greedy 0 \\
-\hline
- & \cellcolor[rgb]{0.75,0.9990234375,0.75} 150756& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00269x& \cellcolor[rgb]{0.7509765625,0.9990234375,0.75} 1.00742x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00356x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00237x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99991x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00269x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99794x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00298x\\
- & \cellcolor[rgb]{0.75,0.9990234375,0.75} 266509& \cellcolor[rgb]{0.94140625,0.9990234375,0.75} 1.76877x& \cellcolor[rgb]{0.9521484375,0.9990234375,0.75} 1.81332x& \cellcolor[rgb]{0.8115234375,0.9990234375,0.75} 1.24997x& \cellcolor[rgb]{0.93359375,0.9990234375,0.75} 1.74009x& \cellcolor[rgb]{0.93359375,0.9990234375,0.75} 1.73912x& \cellcolor[rgb]{0.9345703125,0.9990234375,0.75} 1.74314x& \cellcolor[rgb]{0.9404296875,0.9990234375,0.75} 1.76571x& \cellcolor[rgb]{0.9990234375,0.8681640625,0.75} 6.19913x\\
- & \cellcolor[rgb]{0.75,0.9990234375,0.75} 331770& \cellcolor[rgb]{0.8720703125,0.9990234375,0.75} 1.49255x& \cellcolor[rgb]{0.8974609375,0.9990234375,0.75} 1.59420x& \cellcolor[rgb]{0.775390625,0.9990234375,0.75} 1.10326x& \cellcolor[rgb]{0.7998046875,0.9990234375,0.75} 1.20125x& \cellcolor[rgb]{0.8486328125,0.9990234375,0.75} 1.39740x& \cellcolor[rgb]{0.8486328125,0.9990234375,0.75} 1.39905x& \cellcolor[rgb]{0.8701171875,0.9990234375,0.75} 1.48485x& \cellcolor[rgb]{0.9990234375,0.8955078125,0.75} 5.31939x\\
- & \cellcolor[rgb]{0.75,0.9990234375,0.75} 398255& \cellcolor[rgb]{0.8720703125,0.9990234375,0.75} 1.49334x& \cellcolor[rgb]{0.892578125,0.9990234375,0.75} 1.57569x& \cellcolor[rgb]{0.7578125,0.9990234375,0.75} 1.03280x& \cellcolor[rgb]{0.759765625,0.9990234375,0.75} 1.04240x& \cellcolor[rgb]{0.83203125,0.9990234375,0.75} 1.32999x& \cellcolor[rgb]{0.7900390625,0.9990234375,0.75} 1.16465x& \cellcolor[rgb]{0.83203125,0.9990234375,0.75} 1.33295x& \cellcolor[rgb]{0.9990234375,0.9169921875,0.75} 4.62699x\\
- & \cellcolor[rgb]{0.75,0.9990234375,0.75} 463887& \cellcolor[rgb]{0.8203125,0.9990234375,0.75} 1.28291x& \cellcolor[rgb]{0.8876953125,0.9990234375,0.75} 1.55488x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00064x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00080x& \cellcolor[rgb]{0.8017578125,0.9990234375,0.75} 1.21103x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99887x& \cellcolor[rgb]{0.9990234375,0.9375,0.75} 3.97028x& \cellcolor[rgb]{0.9990234375,0.9208984375,0.75} 4.49752x\\
- & \cellcolor[rgb]{0.75,0.9990234375,0.75} 529682& \cellcolor[rgb]{0.841796875,0.9990234375,0.75} 1.37111x& \cellcolor[rgb]{0.9033203125,0.9990234375,0.75} 1.61877x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99918x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00048x& \cellcolor[rgb]{0.7958984375,0.9990234375,0.75} 1.18536x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99971x& \cellcolor[rgb]{0.9990234375,0.94921875,0.75} 3.58938x& \cellcolor[rgb]{0.9990234375,0.921875,0.75} 4.44715x\\
-& \cellcolor[rgb]{0.75,0.9990234375,0.75} 594307& \cellcolor[rgb]{0.8173828125,0.9990234375,0.75} 1.27157x& \cellcolor[rgb]{0.9013671875,0.9990234375,0.75} 1.60988x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.91407x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.92725x& \cellcolor[rgb]{0.791015625,0.9990234375,0.75} 1.16608x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.89112x& \cellcolor[rgb]{0.9990234375,0.955078125,0.75} 3.38938x& \cellcolor[rgb]{0.9990234375,0.921875,0.75} 4.46031x\\
- & \cellcolor[rgb]{0.75,0.9990234375,0.75} 659352& \cellcolor[rgb]{0.8486328125,0.9990234375,0.75} 1.39923x& \cellcolor[rgb]{0.9033203125,0.9990234375,0.75} 1.61575x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.90244x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.90256x& \cellcolor[rgb]{0.787109375,0.9990234375,0.75} 1.15120x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.90131x& \cellcolor[rgb]{0.9990234375,0.96484375,0.75} 3.08998x& \cellcolor[rgb]{0.9990234375,0.935546875,0.75} 4.03296x\\
-\end{NiceTabular}}
-\caption{\label{tab:dumb-fib-greedy}Benchmark of \texttt{benches/fib.ll} output by \mintinline{bash}{dune exec bench -- -f fib -n 1000}}
-\end{table}
+%\begin{table}[H]
+%\makebox[\textwidth][c]{
+%\begin{NiceTabular}{|c|ccccccccc|}
+%\hline
+%phis & clang & simple 12 & simple 2 & briggs 12 & briggs 2 & linear 2 & linear 12 & greedy 12 & greedy 0 \\
+%\hline
+ %& \cellcolor[rgb]{0.75,0.9990234375,0.75} 150756& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00269x& \cellcolor[rgb]{0.7509765625,0.9990234375,0.75} 1.00742x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00356x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00237x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99991x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00269x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99794x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00298x\\
+ %& \cellcolor[rgb]{0.75,0.9990234375,0.75} 266509& \cellcolor[rgb]{0.94140625,0.9990234375,0.75} 1.76877x& \cellcolor[rgb]{0.9521484375,0.9990234375,0.75} 1.81332x& \cellcolor[rgb]{0.8115234375,0.9990234375,0.75} 1.24997x& \cellcolor[rgb]{0.93359375,0.9990234375,0.75} 1.74009x& \cellcolor[rgb]{0.93359375,0.9990234375,0.75} 1.73912x& \cellcolor[rgb]{0.9345703125,0.9990234375,0.75} 1.74314x& \cellcolor[rgb]{0.9404296875,0.9990234375,0.75} 1.76571x& \cellcolor[rgb]{0.9990234375,0.8681640625,0.75} 6.19913x\\
+ %& \cellcolor[rgb]{0.75,0.9990234375,0.75} 331770& \cellcolor[rgb]{0.8720703125,0.9990234375,0.75} 1.49255x& \cellcolor[rgb]{0.8974609375,0.9990234375,0.75} 1.59420x& \cellcolor[rgb]{0.775390625,0.9990234375,0.75} 1.10326x& \cellcolor[rgb]{0.7998046875,0.9990234375,0.75} 1.20125x& \cellcolor[rgb]{0.8486328125,0.9990234375,0.75} 1.39740x& \cellcolor[rgb]{0.8486328125,0.9990234375,0.75} 1.39905x& \cellcolor[rgb]{0.8701171875,0.9990234375,0.75} 1.48485x& \cellcolor[rgb]{0.9990234375,0.8955078125,0.75} 5.31939x\\
+ %& \cellcolor[rgb]{0.75,0.9990234375,0.75} 398255& \cellcolor[rgb]{0.8720703125,0.9990234375,0.75} 1.49334x& \cellcolor[rgb]{0.892578125,0.9990234375,0.75} 1.57569x& \cellcolor[rgb]{0.7578125,0.9990234375,0.75} 1.03280x& \cellcolor[rgb]{0.759765625,0.9990234375,0.75} 1.04240x& \cellcolor[rgb]{0.83203125,0.9990234375,0.75} 1.32999x& \cellcolor[rgb]{0.7900390625,0.9990234375,0.75} 1.16465x& \cellcolor[rgb]{0.83203125,0.9990234375,0.75} 1.33295x& \cellcolor[rgb]{0.9990234375,0.9169921875,0.75} 4.62699x\\
+ %& \cellcolor[rgb]{0.75,0.9990234375,0.75} 463887& \cellcolor[rgb]{0.8203125,0.9990234375,0.75} 1.28291x& \cellcolor[rgb]{0.8876953125,0.9990234375,0.75} 1.55488x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00064x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00080x& \cellcolor[rgb]{0.8017578125,0.9990234375,0.75} 1.21103x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99887x& \cellcolor[rgb]{0.9990234375,0.9375,0.75} 3.97028x& \cellcolor[rgb]{0.9990234375,0.9208984375,0.75} 4.49752x\\
+ %& \cellcolor[rgb]{0.75,0.9990234375,0.75} 529682& \cellcolor[rgb]{0.841796875,0.9990234375,0.75} 1.37111x& \cellcolor[rgb]{0.9033203125,0.9990234375,0.75} 1.61877x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99918x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00048x& \cellcolor[rgb]{0.7958984375,0.9990234375,0.75} 1.18536x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99971x& \cellcolor[rgb]{0.9990234375,0.94921875,0.75} 3.58938x& \cellcolor[rgb]{0.9990234375,0.921875,0.75} 4.44715x\\
+%& \cellcolor[rgb]{0.75,0.9990234375,0.75} 594307& \cellcolor[rgb]{0.8173828125,0.9990234375,0.75} 1.27157x& \cellcolor[rgb]{0.9013671875,0.9990234375,0.75} 1.60988x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.91407x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.92725x& \cellcolor[rgb]{0.791015625,0.9990234375,0.75} 1.16608x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.89112x& \cellcolor[rgb]{0.9990234375,0.955078125,0.75} 3.38938x& \cellcolor[rgb]{0.9990234375,0.921875,0.75} 4.46031x\\
+ %& \cellcolor[rgb]{0.75,0.9990234375,0.75} 659352& \cellcolor[rgb]{0.8486328125,0.9990234375,0.75} 1.39923x& \cellcolor[rgb]{0.9033203125,0.9990234375,0.75} 1.61575x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.90244x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.90256x& \cellcolor[rgb]{0.787109375,0.9990234375,0.75} 1.15120x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.90131x& \cellcolor[rgb]{0.9990234375,0.96484375,0.75} 3.08998x& \cellcolor[rgb]{0.9990234375,0.935546875,0.75} 4.03296x\\
+%\end{NiceTabular}}
+%\caption{\label{tab:dumb-fib-greedy}Benchmark of \texttt{benches/fib.ll} output by \mintinline{bash}{dune exec bench -- -f fib -n 1000}}
+%\end{table}
 
@@
 It goes without saying that there are countless optimnizations waiting to be made, but as the goal of this project was to explore the different approaches not many regards were had for the instruction selection itself.
+In fact, much more concern was taken for making sure each instruction was translated such that no other live variables were affected.
+The call instruction alone is probably a whole order of magnitude slower tha nenceesary.
+Likewise, the function pre-/epilogue are also overly cautious.
 

### 2024-01-14 article.tex
Cleanup

@@
   \begin{minted}[bgcolor=LightGray]{gas}
-movq	%rax, %rdx
+# continued...
 movq	%rbx, %rax
@@
 
-Integer factorization is another interesting algorithm. Generally considered to be NP, here are the results of factorizing known primes of inreasing bit length:
+Integer factorization is another interesting algorithm. Generally considered to be NP, here are the results of factorizing known primes of inreasing bit size seen in Table \ref{tab:factori32}
 \begin{table}[H]
@@
 \end{NiceTabular}}
-\caption{\label{tab:factori64}Benchmark of \texttt{benches/factori64.ll} output by \mintinline{bash}{dune exec bench -- -f factori32 -n 1024}}
+\caption{\label{tab:factori32}Benchmark of \texttt{benches/factori32.ll} output by \mintinline{bash}{dune exec bench -- -f factori32 -n 1024}}
 \end{table}
+
+\noindent What's more interesting is that increasing the input size actually reduces the discrepancy with the Greedy and Linear 2 allocators, while all others seem to rise at the same pace as can be seen in Table \ref{tab:factori64}.
 
@@
 \end{NiceTabular}}
-\caption{\label{tab:factori32}Benchmark of \texttt{benches/factori32.ll} output by \mintinline{bash}{dune exec bench -- -f factori32 -n 1024}}
+\caption{\label{tab:factori64}Benchmark of \texttt{benches/factori64.ll} output by \mintinline{bash}{dune exec bench -- -f factori64 -n 1024}}
 \end{table}
@@
 
-\begin{table}[H]
-\makebox[\textwidth][c]{
-\begin{NiceTabular}{|c|ccccccccc|}
-\hline
-arg(s) & clang & simple 12 & simple 2 & briggs 12 & briggs 2 & linear 2 & linear 12 & greedy 12 & greedy 0 \\
-\hline
- & \cellcolor[rgb]{0.75,0.9990234375,0.75} 150790& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99873x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00204x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00055x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00202x& \cellcolor[rgb]{0.7509765625,0.9990234375,0.75} 1.00475x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99925x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99983x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00036x\\
- & \cellcolor[rgb]{0.75,0.9990234375,0.75} 266858& \cellcolor[rgb]{0.939453125,0.9990234375,0.75} 1.76227x& \cellcolor[rgb]{0.951171875,0.9990234375,0.75} 1.81025x& \cellcolor[rgb]{0.8095703125,0.9990234375,0.75} 1.24026x& \cellcolor[rgb]{0.93359375,0.9990234375,0.75} 1.73967x& \cellcolor[rgb]{0.93359375,0.9990234375,0.75} 1.73845x& \cellcolor[rgb]{0.9326171875,0.9990234375,0.75} 1.73599x& \cellcolor[rgb]{0.939453125,0.9990234375,0.75} 1.76279x& \cellcolor[rgb]{0.9990234375,0.8681640625,0.75} 6.18927x\\
- & \cellcolor[rgb]{0.75,0.9990234375,0.75} 332932& \cellcolor[rgb]{0.8681640625,0.9990234375,0.75} 1.47463x& \cellcolor[rgb]{0.896484375,0.9990234375,0.75} 1.58958x& \cellcolor[rgb]{0.7734375,0.9990234375,0.75} 1.09780x& \cellcolor[rgb]{0.7978515625,0.9990234375,0.75} 1.19567x& \cellcolor[rgb]{0.8466796875,0.9990234375,0.75} 1.39187x& \cellcolor[rgb]{0.84765625,0.9990234375,0.75} 1.39397x& \cellcolor[rgb]{0.86328125,0.9990234375,0.75} 1.45798x& \cellcolor[rgb]{0.9990234375,0.8955078125,0.75} 5.30079x\\
- & \cellcolor[rgb]{0.75,0.9990234375,0.75} 396964& \cellcolor[rgb]{0.873046875,0.9990234375,0.75} 1.49802x& \cellcolor[rgb]{0.89453125,0.9990234375,0.75} 1.58318x& \cellcolor[rgb]{0.7578125,0.9990234375,0.75} 1.03435x& \cellcolor[rgb]{0.7607421875,0.9990234375,0.75} 1.04439x& \cellcolor[rgb]{0.8310546875,0.9990234375,0.75} 1.32655x& \cellcolor[rgb]{0.7919921875,0.9990234375,0.75} 1.16965x& \cellcolor[rgb]{0.8330078125,0.9990234375,0.75} 1.33411x& \cellcolor[rgb]{0.9990234375,0.916015625,0.75} 4.64216x\\
- & \cellcolor[rgb]{0.75,0.9990234375,0.75} 463143& \cellcolor[rgb]{0.8193359375,0.9990234375,0.75} 1.27994x& \cellcolor[rgb]{0.8916015625,0.9990234375,0.75} 1.56894x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00193x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00219x& \cellcolor[rgb]{0.802734375,0.9990234375,0.75} 1.21525x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00036x& \cellcolor[rgb]{0.9990234375,0.9375,0.75} 3.97536x& \cellcolor[rgb]{0.9990234375,0.9208984375,0.75} 4.50366x\\
- & \cellcolor[rgb]{0.75,0.9990234375,0.75} 528926& \cellcolor[rgb]{0.841796875,0.9990234375,0.75} 1.37202x& \cellcolor[rgb]{0.904296875,0.9990234375,0.75} 1.62053x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00265x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00090x& \cellcolor[rgb]{0.7958984375,0.9990234375,0.75} 1.18442x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00260x& \cellcolor[rgb]{0.9990234375,0.94921875,0.75} 3.59263x& \cellcolor[rgb]{0.9990234375,0.921875,0.75} 4.46336x\\
- & \cellcolor[rgb]{0.75,0.9990234375,0.75} 594806& \cellcolor[rgb]{0.818359375,0.9990234375,0.75} 1.27651x& \cellcolor[rgb]{0.9013671875,0.9990234375,0.75} 1.60931x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.91222x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.92581x& \cellcolor[rgb]{0.791015625,0.9990234375,0.75} 1.16478x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.89116x& \cellcolor[rgb]{0.9990234375,0.955078125,0.75} 3.38589x& \cellcolor[rgb]{0.9990234375,0.921875,0.75} 4.46530x\\
- & \cellcolor[rgb]{0.75,0.9990234375,0.75} 660398& \cellcolor[rgb]{0.84765625,0.9990234375,0.75} 1.39599x& \cellcolor[rgb]{0.9013671875,0.9990234375,0.75} 1.60987x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.89423x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.90015x& \cellcolor[rgb]{0.7861328125,0.9990234375,0.75} 1.14754x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.90032x& \cellcolor[rgb]{0.9990234375,0.96484375,0.75} 3.08470x& \cellcolor[rgb]{0.9990234375,0.935546875,0.75} 4.02611x\\
-\hline
-\end{NiceTabular}}
-\caption{\label{tab:dumb-fib-greedy}Benchmark of \texttt{benches/fib.ll} output by \mintinline{bash}{dune exec bench -- -f fib -n 1000}}
-\end{table}
-
-\subsubsection{\lstinline!benches/factori64.ll!}
-\begin{table}[H]
-\makebox[\textwidth][c]{
-\begin{NiceTabular}{|c|ccccccccc|}
-\hline
-arg(s) & clang & simple 12 & simple 2 & briggs 12 & briggs 2 & linear 12 & linear 2 & greedy 12 & greedy 0 \\
-\hline
-268435399 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 218996& \cellcolor[rgb]{0.7666015625,0.9990234375,0.75} 1.07020x& \cellcolor[rgb]{0.7763671875,0.9990234375,0.75} 1.10754x& \cellcolor[rgb]{0.767578125,0.9990234375,0.75} 1.07114x& \cellcolor[rgb]{0.7763671875,0.9990234375,0.75} 1.10950x& \cellcolor[rgb]{0.7744140625,0.9990234375,0.75} 1.09908x& \cellcolor[rgb]{0.90234375,0.9990234375,0.75} 1.61452x& \cellcolor[rgb]{0.9990234375,0.986328125,0.75} 2.38305x& \cellcolor[rgb]{0.9990234375,0.986328125,0.75} 2.38509x\\
-536870909 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 247101& \cellcolor[rgb]{0.7705078125,0.9990234375,0.75} 1.08257x& \cellcolor[rgb]{0.7822265625,0.9990234375,0.75} 1.13159x& \cellcolor[rgb]{0.7705078125,0.9990234375,0.75} 1.08546x& \cellcolor[rgb]{0.7822265625,0.9990234375,0.75} 1.13164x& \cellcolor[rgb]{0.783203125,0.9990234375,0.75} 1.13364x& \cellcolor[rgb]{0.947265625,0.9990234375,0.75} 1.79438x& \cellcolor[rgb]{0.9990234375,0.9755859375,0.75} 2.73111x& \cellcolor[rgb]{0.9990234375,0.9755859375,0.75} 2.73643x\\
-1073741789 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 285989& \cellcolor[rgb]{0.775390625,0.9990234375,0.75} 1.10503x& \cellcolor[rgb]{0.7900390625,0.9990234375,0.75} 1.16197x& \cellcolor[rgb]{0.775390625,0.9990234375,0.75} 1.10359x& \cellcolor[rgb]{0.7900390625,0.9990234375,0.75} 1.16233x& \cellcolor[rgb]{0.7900390625,0.9990234375,0.75} 1.16392x& \cellcolor[rgb]{0.9912109375,0.9990234375,0.75} 1.97079x& \cellcolor[rgb]{0.9990234375,0.9638671875,0.75} 3.11293x& \cellcolor[rgb]{0.9990234375,0.9638671875,0.75} 3.12110x\\
-2147483647 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 341802& \cellcolor[rgb]{0.779296875,0.9990234375,0.75} 1.12053x& \cellcolor[rgb]{0.7958984375,0.9990234375,0.75} 1.18807x& \cellcolor[rgb]{0.7802734375,0.9990234375,0.75} 1.12297x& \cellcolor[rgb]{0.7978515625,0.9990234375,0.75} 1.19256x& \cellcolor[rgb]{0.796875,0.9990234375,0.75} 1.19163x& \cellcolor[rgb]{0.9990234375,0.994140625,0.75} 2.14407x& \cellcolor[rgb]{0.9990234375,0.9521484375,0.75} 3.49973x& \cellcolor[rgb]{0.9990234375,0.951171875,0.75} 3.50847x\\
-4294967291 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 419941& \cellcolor[rgb]{0.78515625,0.9990234375,0.75} 1.14339x& \cellcolor[rgb]{0.8046875,0.9990234375,0.75} 1.22187x& \cellcolor[rgb]{0.78515625,0.9990234375,0.75} 1.14366x& \cellcolor[rgb]{0.8046875,0.9990234375,0.75} 1.22078x& \cellcolor[rgb]{0.8046875,0.9990234375,0.75} 1.22093x& \cellcolor[rgb]{0.9990234375,0.98828125,0.75} 2.31858x& \cellcolor[rgb]{0.9990234375,0.9404296875,0.75} 3.87528x& \cellcolor[rgb]{0.9990234375,0.939453125,0.75} 3.88825x\\
-8589934583 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 530002& \cellcolor[rgb]{0.7900390625,0.9990234375,0.75} 1.16154x& \cellcolor[rgb]{0.8115234375,0.9990234375,0.75} 1.24776x& \cellcolor[rgb]{0.7900390625,0.9990234375,0.75} 1.16318x& \cellcolor[rgb]{0.810546875,0.9990234375,0.75} 1.24673x& \cellcolor[rgb]{0.8115234375,0.9990234375,0.75} 1.24866x& \cellcolor[rgb]{0.9990234375,0.9833984375,0.75} 2.47972x& \cellcolor[rgb]{0.9990234375,0.9296875,0.75} 4.22355x& \cellcolor[rgb]{0.9990234375,0.9287109375,0.75} 4.23487x\\
-17179869143 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 685889& \cellcolor[rgb]{0.7939453125,0.9990234375,0.75} 1.17835x& \cellcolor[rgb]{0.8173828125,0.9990234375,0.75} 1.27310x& \cellcolor[rgb]{0.7939453125,0.9990234375,0.75} 1.17878x& \cellcolor[rgb]{0.8173828125,0.9990234375,0.75} 1.27365x& \cellcolor[rgb]{0.8173828125,0.9990234375,0.75} 1.27361x& \cellcolor[rgb]{0.9990234375,0.9794921875,0.75} 2.61753x& \cellcolor[rgb]{0.9990234375,0.919921875,0.75} 4.52244x& \cellcolor[rgb]{0.9990234375,0.919921875,0.75} 4.53682x\\
-34359738337 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 905070& \cellcolor[rgb]{0.796875,0.9990234375,0.75} 1.19204x& \cellcolor[rgb]{0.8232421875,0.9990234375,0.75} 1.29498x& \cellcolor[rgb]{0.7978515625,0.9990234375,0.75} 1.19218x& \cellcolor[rgb]{0.8232421875,0.9990234375,0.75} 1.29535x& \cellcolor[rgb]{0.8232421875,0.9990234375,0.75} 1.29450x& \cellcolor[rgb]{0.9990234375,0.9755859375,0.75} 2.73632x& \cellcolor[rgb]{0.9990234375,0.912109375,0.75} 4.77726x& \cellcolor[rgb]{0.9990234375,0.912109375,0.75} 4.79190x\\
-\hline
-\end{NiceTabular}}
-\caption{\label{tab:factori64-table}Benchmark of \texttt{benches/fib.ll} output by \mintinline{bash}{dune exec bench -- -f fib -n 1000}}
-\end{table}
-
-\subsubsection{\lstinline!benches/sieven.ll!}
-\begin{table}[H]
-\makebox[\textwidth][c]{
-\begin{NiceTabular}{|c|cccccccc|}
-\hline
-arg(s) & clang & simple 12 & simple 2 & briggs 12 & briggs 2 & linear & greedy 12 & greedy 0 \\
-\hline
-128 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 159130& \cellcolor[rgb]{0.7509765625,0.9990234375,0.75} 1.00777x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00293x& \cellcolor[rgb]{0.7509765625,0.9990234375,0.75} 1.00644x& \cellcolor[rgb]{0.7529296875,0.9990234375,0.75} 1.01562x& \cellcolor[rgb]{0.8173828125,0.9990234375,0.75} 1.27104x& \cellcolor[rgb]{0.82421875,0.9990234375,0.75} 1.29856x& \cellcolor[rgb]{0.82421875,0.9990234375,0.75} 1.29894x\\
-256 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 163800& \cellcolor[rgb]{0.7509765625,0.9990234375,0.75} 1.00726x& \cellcolor[rgb]{0.7529296875,0.9990234375,0.75} 1.01200x& \cellcolor[rgb]{0.751953125,0.9990234375,0.75} 1.00852x& \cellcolor[rgb]{0.7529296875,0.9990234375,0.75} 1.01440x& \cellcolor[rgb]{0.8818359375,0.9990234375,0.75} 1.53225x& \cellcolor[rgb]{0.89453125,0.9990234375,0.75} 1.58048x& \cellcolor[rgb]{0.8974609375,0.9990234375,0.75} 1.59431x\\
-512 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 172096& \cellcolor[rgb]{0.75390625,0.9990234375,0.75} 1.01902x& \cellcolor[rgb]{0.755859375,0.9990234375,0.75} 1.02488x& \cellcolor[rgb]{0.75390625,0.9990234375,0.75} 1.01789x& \cellcolor[rgb]{0.75390625,0.9990234375,0.75} 1.01876x& \cellcolor[rgb]{0.9990234375,0.9970703125,0.75} 2.03762x& \cellcolor[rgb]{0.9990234375,0.994140625,0.75} 2.13731x& \cellcolor[rgb]{0.9990234375,0.9931640625,0.75} 2.15792x\\
-1024 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 188591& \cellcolor[rgb]{0.7587890625,0.9990234375,0.75} 1.03687x& \cellcolor[rgb]{0.7587890625,0.9990234375,0.75} 1.03706x& \cellcolor[rgb]{0.7568359375,0.9990234375,0.75} 1.02964x& \cellcolor[rgb]{0.7587890625,0.9990234375,0.75} 1.03689x& \cellcolor[rgb]{0.9990234375,0.9697265625,0.75} 2.93413x& \cellcolor[rgb]{0.9990234375,0.9638671875,0.75} 3.09867x& \cellcolor[rgb]{0.9990234375,0.962890625,0.75} 3.15326x\\
-2048 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 221875& \cellcolor[rgb]{0.76171875,0.9990234375,0.75} 1.04971x& \cellcolor[rgb]{0.765625,0.9990234375,0.75} 1.06586x& \cellcolor[rgb]{0.76171875,0.9990234375,0.75} 1.04806x& \cellcolor[rgb]{0.765625,0.9990234375,0.75} 1.06306x& \cellcolor[rgb]{0.9990234375,0.92578125,0.75} 4.34620x& \cellcolor[rgb]{0.9990234375,0.916015625,0.75} 4.64145x& \cellcolor[rgb]{0.9990234375,0.9140625,0.75} 4.72466x\\
-4096 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 288132& \cellcolor[rgb]{0.771484375,0.9990234375,0.75} 1.08649x& \cellcolor[rgb]{0.7744140625,0.9990234375,0.75} 1.09864x& \cellcolor[rgb]{0.76953125,0.9990234375,0.75} 1.08082x& \cellcolor[rgb]{0.7744140625,0.9990234375,0.75} 1.10033x& \cellcolor[rgb]{0.9990234375,0.8662109375,0.75} 6.25886x& \cellcolor[rgb]{0.9990234375,0.8525390625,0.75} 6.69614x& \cellcolor[rgb]{0.9990234375,0.84765625,0.75} 6.84597x\\
-8192 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 424179& \cellcolor[rgb]{0.7763671875,0.9990234375,0.75} 1.10601x& \cellcolor[rgb]{0.78125,0.9990234375,0.75} 1.12721x& \cellcolor[rgb]{0.7763671875,0.9990234375,0.75} 1.10683x& \cellcolor[rgb]{0.78125,0.9990234375,0.75} 1.12700x& \cellcolor[rgb]{0.9990234375,0.8037109375,0.75} 8.26470x& \cellcolor[rgb]{0.9990234375,0.78515625,0.75} 8.85856x& \cellcolor[rgb]{0.9990234375,0.7783203125,0.75} 9.07318x\\
-16384 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 697378& \cellcolor[rgb]{0.78125,0.9990234375,0.75} 1.12603x& \cellcolor[rgb]{0.787109375,0.9990234375,0.75} 1.15191x& \cellcolor[rgb]{0.78125,0.9990234375,0.75} 1.12581x& \cellcolor[rgb]{0.787109375,0.9990234375,0.75} 1.15269x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 10.00521x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 10.73828x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 10.98249x\\
-\hline
-\end{NiceTabular}}
-\caption{\label{tab:sieven}Benchmark of \texttt{benches/sieven.ll} output by \mintinline{bash}{dune exec bench -- -f fib -n 1000}}
-\end{table}
+%\begin{table}[H]
+%\makebox[\textwidth][c]{
+%\begin{NiceTabular}{|c|ccccccccc|}
+%\hline
+%arg(s) & clang & simple 12 & simple 2 & briggs 12 & briggs 2 & linear 2 & linear 12 & greedy 12 & greedy 0 \\
+%\hline
+ %& \cellcolor[rgb]{0.75,0.9990234375,0.75} 150790& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99873x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00204x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00055x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00202x& \cellcolor[rgb]{0.7509765625,0.9990234375,0.75} 1.00475x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99925x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.99983x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00036x\\
+ %& \cellcolor[rgb]{0.75,0.9990234375,0.75} 266858& \cellcolor[rgb]{0.939453125,0.9990234375,0.75} 1.76227x& \cellcolor[rgb]{0.951171875,0.9990234375,0.75} 1.81025x& \cellcolor[rgb]{0.8095703125,0.9990234375,0.75} 1.24026x& \cellcolor[rgb]{0.93359375,0.9990234375,0.75} 1.73967x& \cellcolor[rgb]{0.93359375,0.9990234375,0.75} 1.73845x& \cellcolor[rgb]{0.9326171875,0.9990234375,0.75} 1.73599x& \cellcolor[rgb]{0.939453125,0.9990234375,0.75} 1.76279x& \cellcolor[rgb]{0.9990234375,0.8681640625,0.75} 6.18927x\\
+ %& \cellcolor[rgb]{0.75,0.9990234375,0.75} 332932& \cellcolor[rgb]{0.8681640625,0.9990234375,0.75} 1.47463x& \cellcolor[rgb]{0.896484375,0.9990234375,0.75} 1.58958x& \cellcolor[rgb]{0.7734375,0.9990234375,0.75} 1.09780x& \cellcolor[rgb]{0.7978515625,0.9990234375,0.75} 1.19567x& \cellcolor[rgb]{0.8466796875,0.9990234375,0.75} 1.39187x& \cellcolor[rgb]{0.84765625,0.9990234375,0.75} 1.39397x& \cellcolor[rgb]{0.86328125,0.9990234375,0.75} 1.45798x& \cellcolor[rgb]{0.9990234375,0.8955078125,0.75} 5.30079x\\
+ %& \cellcolor[rgb]{0.75,0.9990234375,0.75} 396964& \cellcolor[rgb]{0.873046875,0.9990234375,0.75} 1.49802x& \cellcolor[rgb]{0.89453125,0.9990234375,0.75} 1.58318x& \cellcolor[rgb]{0.7578125,0.9990234375,0.75} 1.03435x& \cellcolor[rgb]{0.7607421875,0.9990234375,0.75} 1.04439x& \cellcolor[rgb]{0.8310546875,0.9990234375,0.75} 1.32655x& \cellcolor[rgb]{0.7919921875,0.9990234375,0.75} 1.16965x& \cellcolor[rgb]{0.8330078125,0.9990234375,0.75} 1.33411x& \cellcolor[rgb]{0.9990234375,0.916015625,0.75} 4.64216x\\
+ %& \cellcolor[rgb]{0.75,0.9990234375,0.75} 463143& \cellcolor[rgb]{0.8193359375,0.9990234375,0.75} 1.27994x& \cellcolor[rgb]{0.8916015625,0.9990234375,0.75} 1.56894x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00193x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00219x& \cellcolor[rgb]{0.802734375,0.9990234375,0.75} 1.21525x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00036x& \cellcolor[rgb]{0.9990234375,0.9375,0.75} 3.97536x& \cellcolor[rgb]{0.9990234375,0.9208984375,0.75} 4.50366x\\
+ %& \cellcolor[rgb]{0.75,0.9990234375,0.75} 528926& \cellcolor[rgb]{0.841796875,0.9990234375,0.75} 1.37202x& \cellcolor[rgb]{0.904296875,0.9990234375,0.75} 1.62053x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00265x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00090x& \cellcolor[rgb]{0.7958984375,0.9990234375,0.75} 1.18442x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 1.00260x& \cellcolor[rgb]{0.9990234375,0.94921875,0.75} 3.59263x& \cellcolor[rgb]{0.9990234375,0.921875,0.75} 4.46336x\\
+ %& \cellcolor[rgb]{0.75,0.9990234375,0.75} 594806& \cellcolor[rgb]{0.818359375,0.9990234375,0.75} 1.27651x& \cellcolor[rgb]{0.9013671875,0.9990234375,0.75} 1.60931x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.91222x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.92581x& \cellcolor[rgb]{0.791015625,0.9990234375,0.75} 1.16478x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.89116x& \cellcolor[rgb]{0.9990234375,0.955078125,0.75} 3.38589x& \cellcolor[rgb]{0.9990234375,0.921875,0.75} 4.46530x\\
+ %& \cellcolor[rgb]{0.75,0.9990234375,0.75} 660398& \cellcolor[rgb]{0.84765625,0.9990234375,0.75} 1.39599x& \cellcolor[rgb]{0.9013671875,0.9990234375,0.75} 1.60987x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.89423x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.90015x& \cellcolor[rgb]{0.7861328125,0.9990234375,0.75} 1.14754x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.90032x& \cellcolor[rgb]{0.9990234375,0.96484375,0.75} 3.08470x& \cellcolor[rgb]{0.9990234375,0.935546875,0.75} 4.02611x\\
+%\hline
+%\end{NiceTabular}}
+%\caption{\label{tab:dumb-fib-greedy}Benchmark of \texttt{benches/fib.ll} output by \mintinline{bash}{dune exec bench -- -f fib -n 1000}}
+%\end{table}
 
@@
 
-\begin{table}[H]
-\makebox[\textwidth][c]{
-\begin{NiceTabular}{|c|ccccccccc|}
-\hline
-arg(s) & clang & simple 12 & simple 2 & briggs 12 & briggs 2 & linear 12 & linear 2 & greedy 12 & greedy 0 \\
-\hline
-4 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 171300& \cellcolor[rgb]{0.751953125,0.9990234375,0.75} 1.01148x& \cellcolor[rgb]{0.7568359375,0.9990234375,0.75} 1.02792x& \cellcolor[rgb]{0.755859375,0.9990234375,0.75} 1.02449x& \cellcolor[rgb]{0.755859375,0.9990234375,0.75} 1.02731x& \cellcolor[rgb]{0.7880859375,0.9990234375,0.75} 1.15660x& \cellcolor[rgb]{0.7841796875,0.9990234375,0.75} 1.13751x& \cellcolor[rgb]{0.8173828125,0.9990234375,0.75} 1.27163x& \cellcolor[rgb]{0.8037109375,0.9990234375,0.75} 1.21946x\\
-5 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 189030& \cellcolor[rgb]{0.76171875,0.9990234375,0.75} 1.04909x& \cellcolor[rgb]{0.771484375,0.9990234375,0.75} 1.08923x& \cellcolor[rgb]{0.7626953125,0.9990234375,0.75} 1.05266x& \cellcolor[rgb]{0.7705078125,0.9990234375,0.75} 1.08530x& \cellcolor[rgb]{0.8486328125,0.9990234375,0.75} 1.39953x& \cellcolor[rgb]{0.85546875,0.9990234375,0.75} 1.42370x& \cellcolor[rgb]{0.9033203125,0.9990234375,0.75} 1.61577x& \cellcolor[rgb]{0.908203125,0.9990234375,0.75} 1.63543x\\
-6 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 305772& \cellcolor[rgb]{0.8115234375,0.9990234375,0.75} 1.24795x& \cellcolor[rgb]{0.830078125,0.9990234375,0.75} 1.32393x& \cellcolor[rgb]{0.8125,0.9990234375,0.75} 1.25272x& \cellcolor[rgb]{0.83203125,0.9990234375,0.75} 1.33146x& \cellcolor[rgb]{0.9990234375,0.9873046875,0.75} 2.34850x& \cellcolor[rgb]{0.9990234375,0.9873046875,0.75} 2.37422x& \cellcolor[rgb]{0.9990234375,0.9609375,0.75} 3.21562x& \cellcolor[rgb]{0.9990234375,0.9599609375,0.75} 3.23381x\\
-7 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1331652& \cellcolor[rgb]{0.86328125,0.9990234375,0.75} 1.45633x& \cellcolor[rgb]{0.90625,0.9990234375,0.75} 1.62799x& \cellcolor[rgb]{0.8740234375,0.9990234375,0.75} 1.50098x& \cellcolor[rgb]{0.9091796875,0.9990234375,0.75} 1.64250x& \cellcolor[rgb]{0.9990234375,0.9580078125,0.75} 3.31242x& \cellcolor[rgb]{0.9990234375,0.955078125,0.75} 3.39020x& \cellcolor[rgb]{0.9990234375,0.904296875,0.75} 5.03373x& \cellcolor[rgb]{0.9990234375,0.9033203125,0.75} 5.06248x\\
-8 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 11197605& \cellcolor[rgb]{0.904296875,0.9990234375,0.75} 1.62006x& \cellcolor[rgb]{0.9375,0.9990234375,0.75} 1.75454x& \cellcolor[rgb]{0.89453125,0.9990234375,0.75} 1.58268x& \cellcolor[rgb]{0.939453125,0.9990234375,0.75} 1.76405x& \cellcolor[rgb]{0.9990234375,0.9482421875,0.75} 3.61197x& \cellcolor[rgb]{0.9990234375,0.9482421875,0.75} 3.62996x& \cellcolor[rgb]{0.9990234375,0.8876953125,0.75} 5.56708x& \cellcolor[rgb]{0.9990234375,0.888671875,0.75} 5.53880x\\
-9 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 112923594& \cellcolor[rgb]{0.9140625,0.9990234375,0.75} 1.66209x& \cellcolor[rgb]{0.9541015625,0.9990234375,0.75} 1.82157x& \cellcolor[rgb]{0.904296875,0.9990234375,0.75} 1.62219x& \cellcolor[rgb]{0.9453125,0.9990234375,0.75} 1.78657x& \cellcolor[rgb]{0.9990234375,0.947265625,0.75} 3.65915x& \cellcolor[rgb]{0.9990234375,0.94140625,0.75} 3.82090x& \cellcolor[rgb]{0.9990234375,0.880859375,0.75} 5.79421x& \cellcolor[rgb]{0.9990234375,0.8818359375,0.75} 5.75327x\\
-10 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1348654063& \cellcolor[rgb]{0.884765625,0.9990234375,0.75} 1.54471x& \cellcolor[rgb]{0.9306640625,0.9990234375,0.75} 1.72799x& \cellcolor[rgb]{0.896484375,0.9990234375,0.75} 1.59112x& \cellcolor[rgb]{0.943359375,0.9990234375,0.75} 1.77928x& \cellcolor[rgb]{0.9990234375,0.9501953125,0.75} 3.55946x& \cellcolor[rgb]{0.9990234375,0.9462890625,0.75} 3.68201x& \cellcolor[rgb]{0.9990234375,0.8828125,0.75} 5.71345x& \cellcolor[rgb]{0.9990234375,0.8837890625,0.75} 5.67565x\\
-\hline
-\end{NiceTabular}}
-\caption{\label{tab:fannkuch}Benchmark of \texttt{benches/sieven.ll} output by \mintinline{bash}{dune exec bench -- -f fib -n 1000}}
-\end{table}
-
 % TODO: comment that this is basically writing an llvm backend, except redoing all the register allocation stuff.
@@
 Likewise, the function pre-/epilogue are also overly cautious.
+
+Certainly the selection of benchmarks could also have been broader, with the cast majority just showing how Greedy is considerably slower.
+Not to mention more benchmarks with support for the Tiger compiler that was written the year prior.
+Likewise, some more detailed profiling of the entire compilation would have been interesting.
+Some
 

### 2024-01-14 article.tex
Add debugger harness section again

@@
 
-\subsection{Test Suite and  Framework}
+\subsection{Test suite/framework}
 
@@
 Except the \~ 20\% of the LLVM test suite that were never working.
+
+\subsubsection{Debug Harness}
+
+A more elaborate way of asserting correctness of translation is to assert that there are no effects except for those intended.
+One way of achieving this is to check if a sequence of x86 instructions representing a single LLVM-{}- operation change any values except those expected.
+An approach to this is to insert a breakpoint before every LLVM-{}- instruction, then attach a debugger that breaks on these,  and ensure that only the intended registers are altered.
+
+To achieve this, only a few changes need to be made to the codebase.
+A debug flag is introduced, which inserts a label before every instruction by adding a \lstinline!Breakpoint!
+variant of the \lstinline!ins!
+type before the assembly emitted of that specific operation.
+A unique identifier as well as a list of changes expected to be made is encoded.
+This includes a bitmask of GPRs that can possibly be changed (aside from the scratch registers of course) as well as which parts of the stack may be modified. % TODO: implement the stack checker
+In addition to this, each of the labels are output with a \lstinline!.globl! directive in the file header. A file is built with these extra symbols using the \lstinline!-d! flag.
+
+Unfortunately, this was never made to work in practice, but the \texttt{harness.py} script is attached regardless. It was made and tested with \texttt{clang16} in mind on Linux.
 
@@
 Fannkuch-redux involves a permutation algorithm that has a non-trivial algorithmic complexity.
-It requires both recursion and iterative processes, making it a good test for a language's ability to handle complex computations efficiently.
+%It requires both recursion and iterative processes, making it a good test for a language's ability to handle complex computations efficiently.
+It stems from the Benchmarks game 
+It was made for the Debian benchmarks game 
 The fannkuch algorithm is

### 2024-01-14 article.tex
fsdf

@@
 It could be chosen at random but that could be costly, instead some spilling heuristic is applied.
+
+Total time spent during compilation is somewhere around those listed in Figure \ref{fig:build-time}.
+
+\begin{figure}[H]
+  \centering
+  \begin{minipage}[b]{0.69\textwidth}
+    \begin{minted}[bgcolor=LightGray]{bash}
+$ time dune exec build -- tests/sha256.ll -a greedy
+38,72s user 0,07s system 99% cpu 38,833 total
+$ time dune exec build -- tests/sha256.ll -a linear
+38,72s user 0,04s system 99% cpu 38,813 total
+$ time dune exec build -- tests/sha256.ll -a simple
+120,55s user 0,07s system 99% cpu 2:00,74 total
+    \end{minted}
+    \caption{Runtime over single runs of compilation}\label{fig:build-time}
+  \end{minipage}
+\end{figure}
 
@@
 
-It goes without saying that there are countless optimnizations waiting to be made, but as the goal of this project was to explore the different approaches not many regards were had for the instruction selection itself.
+It goes without saying that there are countless optimizations waiting to be made, but as the goal of this project was to explore the different approaches not many regards were had for the instruction selection itself.
 In fact, much more concern was taken for making sure each instruction was translated such that no other live variables were affected.
@@
 Likewise, some more detailed profiling of the entire compilation would have been interesting.
-Some
+Some short descriptions of time taking during certain phases are included (see Figure \ref{fig:lva-time} and \ref{fig:build-time}) although a far more detailed report could've been made with instrumentation frameworks like Landmarks \cite{landmarks} as the time spent during compile time is quite significant, prticularly for JIT.
+
+The complexity of the liveness analysis phase could also be significantly improved, as
 

### 2024-01-14 article.tex
sdf

@@
 120,55s user 0,07s system 99% cpu 2:00,74 total
+$ time dune exec build -- tests/sha256.ll -a briggs
+119,86s user 0,05s system 99% cpu 2:00,03 total
     \end{minted}
@@
 Instructions for which the size varies, the suffix is denoted with \texttt{x} instead.
-This is then replaced with the appropriate size suffix depending on the size denoted in the LLVM-{}- instruction.
+This is then replaced with the appropriate size suffix depending on the type noted in the LLVM-{}- instruction.
 For example, most binary operations move the left-hand operand into \texttt{\%rax} and the right-hand operand into  \texttt{\%rcx}, after which the associated instruction from the table is performed.
@@
 sdiv     & idivx \%src, \%dst & \{\texttt{\%rax}, \texttt{\%rdx}\}\\
-select   & cmpx, cmoveq & \{\}\\
+select   & cmpx, cmoveq & \{\texttt{\%rFLAGS}\}\\
 sext     & movx \%src, \%dst & \{\}\\
@@
 % FIXME: check if above is sound
+
+\subsubsection{Translating terminators} \label{impl-term}
+
+As for implementing terminators, the simplest is \texttt{unreachable}, as this literally doesn't evaluate to anything.
+Next is \texttt{ret}, which is more involved, in that it needs to concern itself with calling conventions.
+Although these are fairly straight forward, expecially since functions generally apply the same operations in the prologue by pushing all callee-saved registers to the stack and then, occasionally, subctracting the amount of spilled variables from \texttt{\%rsp}.
+The function epilogue doesn't have to concern itself with the number of spiller variables though, as the stack pointer is saved to \texttt{\%rbp} and simply restored from that.
+The return value (if specified) is moved to the return register which is \texttt{\%rax} according to the System V AMD64 ABI \cite[26]{sysv}.
 
@@
 Likewise, some more detailed profiling of the entire compilation would have been interesting.
-Some short descriptions of time taking during certain phases are included (see Figure \ref{fig:lva-time} and \ref{fig:build-time}) although a far more detailed report could've been made with instrumentation frameworks like Landmarks \cite{landmarks} as the time spent during compile time is quite significant, prticularly for JIT.
+Some short descriptions of time taking during certain phases are included (see Figure \ref{fig:lva-time} and \ref{fig:build-time}) although a far more detailed report could've been made with instrumentation frameworks like Landmarks \cite{landmarks} as the time spent during compile time is quite significant, prticularly for runtime environments that depend on JIT, but also for AOT compilers as long compile times can negatively affect productivity.
 

### 2024-01-14 article.tex
Fix code offset

@@
      \begin{minted}[linenos,bgcolor=LightGray]{ocaml}
-let rec linearscan insns spills ?(v = 0)
-    (avail, active, assign, incstart, incend) =
+let rec linearscan insns spills (avail, active, assign, incstart, incend) =
   match IT.min_binding_opt incstart with
@@
                   S.ST.add e (Reg reg) assign,
+     \end{minted}
+\end{figure}
+
+\begin{figure}[H]
+     \centering
+     \begin{minted}[linenos,bgcolor=LightGray,firstnumber=18]{ocaml}
                   incstart',
@@
             in
-     \end{minted}
-\end{figure}
-
-\begin{figure}[H]
-     \centering
-     \begin{minted}[linenos,bgcolor=LightGray]{ocaml}
             ( spills + 1,

### 2024-01-15 article.tex
more

@@
 \section*{Abstract}
-Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
-Tellus at urna condimentum mattis.
-Enim facilisis gravida neque convallis.
-Enim facilisis gravida neque convallis a cras semper auctor.
-Vitae proin sagittis nisl rhoncus mattis rhoncus urna neque.
-Justo nec ultrices dui sapien eget mi proin.
-Id aliquet risus feugiat in ante.
-A cras semper auctor neque vitae tempus quam pellentesque nec.
-Nulla porttitor massa id neque aliquam vestibulum morbi.
-Facilisi cras fermentum odio eu.
-Odio pellentesque diam volutpat commodo sed egestas egestas fringilla.
-Amet consectetur adipiscing elit duis tristique sollicitudin nibh sit amet.
-Adipiscing at in tellus integer feugiat.
-A pellentesque sit amet porttitor.
-Sed elementum tempus egestas sed sed risus pretium quam.
-Quam viverra orci sagittis eu volutpat odio facilisis mauris sit.
-Arcu dictum varius duis at.
-Sociis natoque penatibus et magnis dis parturient montes.
-Varius sit amet mattis vulputate enim nulla aliquet.
-Ac ut consequat semper viverra.
-Scelerisque eleifend donec pretium vulputate sapien nec.
-Orci ac auctor augue mauris augue.
-Pulvinar mattis nunc sed blandit.
-Semper auctor neque vitae tempus quam pellentesque nec nam.
-Risus pretium quam vulputate dignissim.
-Dui faucibus in ornare quam viverra orci sagittis eu volutpat.
-Ac tortor vitae purus faucibus.
-Eu mi bibendum neque egestas congue quisque egestas diam in.
-Curabitur vitae nunc sed velit dignissim sodales ut eu sem.
-Arcu odio ut sem nulla.
+In this day and age compilers are used in more places than one would think of at a glance.
+While programmers heavily rely on them to get their work done in a reasonable amount of time, %, as the ability to manually construct machine code is a ridiculous expectation to have.
+ quite a significant amount of time that could otherwise have been spent being productive
+ is spent waiting on the compiler to finish, hence the productivity in the field of software development stands to benefit from improved efficiency.
+ Likewise, programming languages that used to be run by interpreting source text text chunks at a time are now being  translated to raw machine instructions immediately prior to execution just to squeeze out fractions of performance benefits on computers as Moore's law seems to nearing it's natural limit as dictated by the laws of our universe.
 
@@
 First of all, I can hardly begin to express my gratitude towards my advisors Aslan and Amin for the weekly meetings where most if not all concerns were settled.
+I'm also grateful for any fellow students who gave me inspiration and some insight along the way, naturally including some additional test cases that did manage to eek  out some bugs.
+Additionally, all of the program snippets I have found online around the place have been quite helpful and saved me a lot of time, including of course the source of some benchmarks.
 %Firstly, we would like to thank Bas Spitters, our supervisor, for his guidance and expertise on the subject, as
@@
 
-The complexity of the liveness analysis phase could also be significantly improved, as
+As for the liveness analysis, research has been conducted allowing for liveness analysis to be conducted over IR of strict SSA form in only two passes \cite{BrandnerFlorian2011CLSf} shadowing  the current approach that depends on iterative dataflow analysis.
+Similarly, linear scan can be aware of 'holes' in live ranges such that variables don't need to live across widely disjointed basic blocks.
+
+%The complexity of the liveness analysis phase could also be significantly improved, as
 
@@
 The intended goal of this project has been to implement both the graph coloring and linear scan algorithms and evaluate them both in terms of benefits and drawbacks when used to compile LLVM-{}- IR to x86 assembly.
+At first, control flow analysis was implemented to generate a control-flow graph. As this finalized,
+dataflow analysis was applied to derive live ranges for each variable, which was eventually used to conduct both the graph coloring by simplifcation and linear scan algoritms.
+All the way throughout development a test framework has been evolving with the phases to validate correct- and soundness to some degree.
 
